### 2015-08-23 paper.tex
add document skeleton

\documentclass[preprint]{aastex}
%\documentclass{emulateapj}

% has to be before amssymb it seems
%\usepackage{color,hyperref}
%\definecolor{linkcolor}{rgb}{0,0,0.5}
%\hypersetup{colorlinks=true,linkcolor=linkcolor,citecolor=linkcolor,
%            filecolor=linkcolor,urlcolor=linkcolor}
%\usepackage{amssymb,amsmath}

\usepackage{color}
\usepackage{url}
\usepackage{graphicx}
\graphicspath{{figures/}}

% For Python code
\usepackage{listings}
\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}
\lstset{language=Python,
        basicstyle=\footnotesize\ttfamily,
        showspaces=false,
        showstringspaces=false,
        tabsize=2,
        breaklines=false,
        breakatwhitespace=true,
        identifierstyle=\ttfamily,
        keywordstyle=\bfseries\color[rgb]{0.133,0.545,0.133},
        commentstyle=\color[rgb]{0.133,0.545,0.133},
        stringstyle=\color[rgb]{0.627,0.126,0.941},
    }

% Draft watermark:
%\usepackage{draftwatermark}
%\SetWatermarkLightness{0.9}
%\SetWatermarkScale{4}

% Some macros
\newcommand{\todo}[1]{{\color{red} [TODO: #1]}}
\newcommand{\foreign}[1]{{\it #1}}

\newcommand{\apriori}{\foreign{a priori}}
\newcommand{\adhoc}{\foreign{ad hoc}}
\newcommand{\etal}{\foreign{et\,al.}}
\newcommand{\etc}{\foreign{etc.}}

\newcommand{\Fig}[1]{Figure~\ref{fig:#1}}
\newcommand{\fig}[1]{\Fig{#1}}
\newcommand{\figlabel}[1]{\label{fig:#1}}
\newcommand{\Eq}[1]{Equation~(\ref{eq:#1})}
\newcommand{\eq}[1]{\Eq{#1}}
\newcommand{\eqs}[2]{Equations~(\ref{eq:#1})-(\ref{eq:#2})}
\newcommand{\eqlabel}[1]{\label{eq:#1}}
\newcommand{\Sect}[1]{Section~\ref{sect:#1}}
\newcommand{\sect}[1]{\Sect{#1}}
\newcommand{\sects}[1]{Sections~#1}
\newcommand{\App}[1]{Appendix~\ref{sect:#1}}
\newcommand{\app}[1]{\App{#1}}
\newcommand{\sectlabel}[1]{\label{sect:#1}}

\usepackage[normalem]{ulem}
\newcommand{\new}[1]{{\color{red} #1}}
\newcommand{\old}[1]{{\sout{#1}}}

\begin{document}

\title{Practical Considerations for Lomb-Scargle Periodic Analysis}

\newcommand{\escience}{1}
\newcommand{\uwastro}{2}
\author{Jacob T. VanderPlas\altaffilmark{\escience}}
\altaffiltext{\escience}{eScience Institute, University of Washington}

\begin{abstract}
An introduction to practical aspects of the use of Lomb-Scargle-type algorithms for periodic analysis.
\end{abstract}

\keywords{
    methods: data analysis ---
    methods: statistical
}

\section{Introduction}
\sectlabel{introduction}

\todo{Quick motivating example with code \& figures}

\section{Measuring Periods in Continuous Signals}

Continuous signal $g(t)$:

\begin{equation}
  \hat{g}(\omega) = \int_{-\infty}^\infty g(t) e^{-2\pi i \nu t} dt
\end{equation}

We can denote this

\begin{equation}
  \hat{g} = \mathcal{F}\left\{g\right\}
\end{equation}

\subsection{Important Properties of the CFT}

\subsubsection{Linearity}
The Fourier Transform is a linear operation, meaning that for two functions $f(t)$ and $g(t)$ and a scalar value $A$,

\begin{eqnarray}
  \mathcal{F}\{g + h\} &=& \mathcal{F}\{g\} + \mathcal{F}\{h\}\\
  \mathcal{F}\{A g\} &=& A\mathcal{F}\{g\}
\end{eqnarray}

These properties follow straightforwardly from the linearity of the Fourier transform definition.

\subsubsection{The Convolution Theorem}

The Fourier transform is central to an identity known as {\it the convolution theorem}.
The convolution of two functions $g(t)$ and $h(t)$ is itself a function of $t$, denoted $\{g \ast h\}(t)$, and is defined as

\begin{equation}
  \{g \ast h\}(t) = \int_{-\infty}^\infty g(t - \tau) h(\tau) d\tau
\end{equation}

The convolution theorem states that the Fourier transform of a convolution is equal to the product of Fourier transforms; that is,

\begin{equation}
  \mathcal{F}\{g \ast h\} = \mathcal{F}\{g\}\mathcal{F}\{h\}
\end{equation}

This property can also be shown (relatively) straightforwardly from the convolution and Fourier transform definitions.

A particularly special function in the complex exponential, $e^{2\pi i\nu_0 t}$, which has the Fourier transform $\mathcal{F}\{e^{2\pi i\nu_0 t}\} = \delta(\nu - \nu_0)$, which follows directly from the definition of the Dirac delta function:

\begin{equation}
  \delta(\nu) \equiv \int_{-\infty}^\infty e^{-2\pi i\nu t}dt
\end{equation}

This is the sense in which a Fourier power is an ideal measure of periodicity: the Fourier transform of a pure periodic signal is a single localized spike identifying its frequency.

\subsection{Fourier Power}

\todo{talk about offsets and phase-shifts?}

With this in mind, we can define the Fourier power,

\begin{equation}
  P(\nu) = \left|\hat{g}(\nu)\right|^2,
\end{equation}

which measures the (real) squared-magnitude of the (complex) Fourier transform.

\todo{Some examples of functions, their transforms, and their power.}

\todo{convolution function and windowing}

\subsection{Evenly-Spaced Discrete Data: the Schuster Periodogram}

\todo{Evenly-spaced delta functions $\to$ evenly-spaced delta in frequency, separated by nyquist}

\subsection{Unevenly-Spaced Discrete Data: the Lomb-Scargle Periodogram}

\todo{Unevenly-spaced delta functions: no cancelations, and so window is generally {\it infinitely} wide.}

\todo{Least-squares equivalence of Lomb-Scargle}

\section{What Frequency Grid to Use?}

\todo{Follow discussion from gatspy documentation}

\section{Reporting errors in Frequency: Uncertainty vs. Precision}

\todo{Talk about peak width and False Alarm Probability; recommend bootstrap (show example of when analytic FAP fails?)}

\section{Algorithmic Considerations}

\todo{Talk about Press \& Rybicki method; show some benchmarks.}

\section{Generalizations \& Challenges}

\todo{Multiterm makes the true period fit better, but also bumps the background noise.}

\citet{VanderPlas2015}

\bibliographystyle{apj}
\bibliography{paper}

\end{document}

### 2015-08-24 paper.tex
more filling-in of outline

@@
 
-\todo{Quick motivating example with code \& figures}
+[Quick motivating example with figures: RR Lyrae star from LINEAR?]
 
 \section{Measuring Periods in Continuous Signals}
+
+[Perhaps put this info in an appendix?]
 
@@
 
-A particularly special function in the complex exponential, $e^{2\pi i\nu_0 t}$, which has the Fourier transform $\mathcal{F}\{e^{2\pi i\nu_0 t}\} = \delta(\nu - \nu_0)$, which follows directly from the definition of the Dirac delta function:
+\subsubsection{Fourier Transform and Periodic Functions}
+
+A particularly special function in the complex exponential, $e^{2\pi i\nu_0 t} \equiv \cos(2\pi \nu_0 t) + i\sin(2\pi\nu_0 t)$, which has the Fourier transform $\mathcal{F}\{e^{2\pi i\nu_0 t}\} = \delta(\nu - \nu_0)$, which follows directly from the definition of the Dirac delta function:
 
@@
 
-This is the sense in which a Fourier power is an ideal measure of periodicity: the Fourier transform of a pure periodic signal is a single localized spike identifying its frequency.
+Similarly, if we have a signal composed of a sum of complex exponentials, the linearity of the Fourier transform assures that the transform is simply a sum of delta functions:
+\begin{equation}
+  \mathcal{F}\left\{ \sum_{n=1}^N e^{-2\pi i \nu_n t} \right\} = \sum_{n=1}^N \delta(\nu - \nu_n)
+\end{equation}
+This result shows the sense in which a Fourier transform is an ideal measure of periodicity: the Fourier transform pulls-out and identifies the individual periodic components of any input.
 
@@
 
-\todo{talk about offsets and phase-shifts?}
+Using the above definitions, it is straightforward to show that a time-offset in a function leads to a multiplicative phase in the Fourier transform; that is, if
+\begin{equation}
+  g_\tau(t) \equiv g(t - \tau)
+\end{equation}
+then 
+\begin{equation}
+  \mathcal{F}\{g_\tau\} = e^{-2\pi i \nu \tau}\mathcal{F}\{g\}
+\end{equation}
+Because the phase merely represents the chosen coordinate system, we'd like to remove it for spectral analysis.
+One way to do this is to compute the power spectral density, which is proportional to \todo{const of proportionality?} the squared modulus of the Fourier Transform:
+\begin{equation}
+  P\{g\} \propto \left| \mathcal{F}\{g\} \right|^2.
+\end{equation}
+Because the modulus of a phase is equal to unity, the resulting measure is unchanged under time-translations in $g(t)$.
 
-With this in mind, we can define the Fourier power,
+\begin{itemize}
+\item Some examples of functions, their transforms, and their power.
+\end{itemize}
 
-\begin{equation}
-  P(\nu) = \left|\hat{g}(\nu)\right|^2,
-\end{equation}
+\subsection{Effect of Window Functions}
 
-which measures the (real) squared-magnitude of the (complex) Fourier transform.
-
-\todo{Some examples of functions, their transforms, and their power.}
-
-\todo{convolution function and windowing}
+\begin{itemize}
+\item convolution function and windowing
+\end{itemize}
 
@@
 
-\todo{Evenly-spaced delta functions $\to$ evenly-spaced delta in frequency, separated by nyquist}
+\begin{itemize}
+\item Evenly-spaced delta functions $\to$ evenly-spaced delta in frequency, separated by nyquist
+\end{itemize}
 
@@
 
-\todo{Unevenly-spaced delta functions: no cancelations, and so window is generally {\it infinitely} wide.}
-
-\todo{Least-squares equivalence of Lomb-Scargle}
+\begin{itemize}
+\item Unevenly-spaced delta functions: no cancelations, and so window is generally {\it infinitely} wide.
+\item Least-squares equivalence of Lomb-Scargle
+\end{itemize}
 
@@
 
-\todo{Follow discussion from gatspy documentation}
+\begin{itemize}
+\item Follow discussion from gatspy documentation
+\end{itemize}
 
@@
 
-\todo{Talk about peak width and False Alarm Probability; recommend bootstrap (show example of when analytic FAP fails?)}
+\begin{itemize}
+\item Talk about peak width and False Alarm Probability; recommend bootstrap (show example of when analytic FAP fails?)
+\end{itemize}
 
@@
 
-\todo{Talk about Press \& Rybicki method; show some benchmarks.}
+\begin{itemize}
+\item Talk about Press \& Rybicki method; show some benchmarks.
+\end{itemize}
 
@@
 
-\todo{Multiterm makes the true period fit better, but also bumps the background noise.}
+\begin{itemize}
+\item Multiterm makes the true period fit better, but also bumps the background noise.
+\end{itemize}
 

### 2016-11-15 paper.tex
some tweaks to the paper

@@
 
-\subsubsection{Fourier Transform and Periodic Functions}
+\subsubsection{$\mathcal{F}$ of a Gaussian is a Gaussian}
+One of the special functions under a Fourier transform is the Gaussian distribution, which is proportional to
+
+\subsubsection{$\mathcal{F}$ of a Periodic Function is a Delta Function}
 
@@
 \end{equation}
-Because the modulus of a phase is equal to unity, the resulting measure is unchanged under time-translations in $g(t)$.
+Because the modulus of a phase is equal to unity, the resulting measure is unchanged under time-translations in $g(t)$, and $P\{g\} = P\{g_\tau\}$ for any time-shift $\tau$.
 
@@
 
+\section{From Idealistic to Realistic}
+
+All of the above theory is beautiful and elegant, but when the theory meets the real world things get a bit more messy.
+In particular, we rarely (if ever) are able to measure continuous functions like $g(t)$, but instead measure discrete realizations of the underlying function at a finite number of particular times $t_n$.
+These discrete observations are effectively a window function over our data, and as we saw above, the presence of such a window function will propagate through our computation of the Fourier power!
+The quantitative result of this windowning depends on the exact nature of the observation pattern; we will first consider the somewhat simpler case of evenly-spaced discrete data.
+
 \subsection{Evenly-Spaced Discrete Data: the Schuster Periodogram}
+
+Let's first consider the case of a large series of evenly-spaced observations.
+Discrete samples can
 

### 2016-11-15 paper.tex
Add Figures 1 and 2

@@
 
-[Quick motivating example with figures: RR Lyrae star from LINEAR?]
+The Lomb-Scargle periodogram is a classic algorithm for detecting periodicity
+in unevenly-sampled time-series, particularly within the astronomy community.
+For example, consider the data showing in \Fig{LINEAR_data}.
+This is an irregularly-sampled timeseries showing a single object from the
+LINEAR survey, with magnitude measured 280 times over the course of five and
+a half years.
 
-\section{Measuring Periods in Continuous Signals}
+It is clear that this object varies with an amplitude of around 0.6 magnitudes.
+Computing the Lomb-Scargle periodogram for the data gives us a measure of the
+power as a function of period of oscillation (\fig{LINEAR_power}, left), from
+which we can determine the period of oscillation of approximately 2.58 hours.
+The right panel of \fig{LINEAR_power} shows a folded visualization of
+the same data as \fig{LINEAR_data} -- i.e.{} plotted as a function of phase
+rather than time.
+
+Often this is exactly how the Lomb-Scargle periodogram is presented: as a clean
+procedure to detect the periodic component in an unevenly-sampled dataset.
+In practice, however, there are a number of subtle issues that must be considered
+when applying a Lomb-Scargle analysis to real-world datasets.
+This paper contains a practical guide to the effective use of the Lomb-Scargle
+periodogram for analyzing real-world time-series.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\textwidth]{fig01_LINEAR_data}
+\caption{Observed light curve from LINEAR object ID 11375941
+    \figlabel{LINEAR_data}.
+}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\textwidth]{fig02_LINEAR_PSD}
+\caption{{\it Left panel:} the Lomb-Scargle periodogram computed from the data
+    in \fig{LINEAR_data}, with an inset detailing the interesting region.
+    {\it Right panel:} the input data in \fig{LINEAR_data}, folded over the
+    2.58-hour period to show the coherent periodic variability.
+    \figlabel{LINEAR_power}
+}
+\end{figure}
+
+\section{Background: Periods in Continuous Signals}
+
+for a continuously defined signal $g(t)$, the Fourier transform is $\hat{g}(\nu)$,
+
+\begin{equation}
+    \mathcal{F}\left\{g\right\}(f) = \int_{-\infty}^\infty g(t) e^{-2\pi i f t} dt
+\end{equation}
 
@@
 \begin{equation}
-  \hat{g}(\omega) = \int_{-\infty}^\infty g(t) e^{-2\pi i \nu t} dt
+  \mathcal{F}\left\{g\right\}(f) = \int_{-\infty}^\infty g(t) e^{-2\pi i f t} dt
+\end{equation}
+
+The inverse of this operation is
+
+\begin{equation}
+  \mathcal{F}^{-1}\left\{g\right\}(f) = \int_{-\infty}^\infty g(t) e^{+2\pi i f t} dt
 \end{equation}
@@
 \begin{equation}
-  \{g \ast h\}(t) = \int_{-\infty}^\infty g(t - \tau) h(\tau) d\tau
+  \{g \ast h\}(t) \equiv \int_{-\infty}^\infty g(t - \tau) h(\tau) d\tau
 \end{equation}
@@
 \subsubsection{$\mathcal{F}$ of a Gaussian is a Gaussian}
-One of the special functions under a Fourier transform is the Gaussian distribution, which is proportional to
+One of the special functions under a Fourier transform is the Gaussian distribution:
+
+\begin{equation}
+    g(t; \mu, \sigma) = \frac{1}{\sqrt{2\pi\sigma^2}} \exp\left[\frac{- (x - \mu)^2}{2\sigma^2}\right]
+\end{equation}, which is proportional to
 

### 2016-11-15 paper.tex
update continuous discussion

@@
 
-\newcommand{\escience}{1}
-\newcommand{\uwastro}{2}
+\newcommand{\escience}{*}
+\newcommand{\uwastro}{+}
 \author{Jacob T. VanderPlas\altaffilmark{\escience}}
@@
 
-The Lomb-Scargle periodogram is a classic algorithm for detecting periodicity
-in unevenly-sampled time-series, particularly within the astronomy community.
-For example, consider the data showing in \Fig{LINEAR_data}.
-This is an irregularly-sampled timeseries showing a single object from the
-LINEAR survey, with magnitude measured 280 times over the course of five and
-a half years.
-
-It is clear that this object varies with an amplitude of around 0.6 magnitudes.
-Computing the Lomb-Scargle periodogram for the data gives us a measure of the
-power as a function of period of oscillation (\fig{LINEAR_power}, left), from
-which we can determine the period of oscillation of approximately 2.58 hours.
-The right panel of \fig{LINEAR_power} shows a folded visualization of
-the same data as \fig{LINEAR_data} -- i.e.{} plotted as a function of phase
-rather than time.
-
-Often this is exactly how the Lomb-Scargle periodogram is presented: as a clean
-procedure to detect the periodic component in an unevenly-sampled dataset.
-In practice, however, there are a number of subtle issues that must be considered
-when applying a Lomb-Scargle analysis to real-world datasets.
-This paper contains a practical guide to the effective use of the Lomb-Scargle
-periodogram for analyzing real-world time-series.
-
 \begin{figure}[ht]
@@
 \caption{{\it Left panel:} the Lomb-Scargle periodogram computed from the data
-    in \fig{LINEAR_data}, with an inset detailing the interesting region.
+    shown in \fig{LINEAR_data}, with an inset detailing the interesting region.
     {\it Right panel:} the input data in \fig{LINEAR_data}, folded over the
@@
 
-\section{Background: Periods in Continuous Signals}
+The Lomb-Scargle periodogram \citep{Lomb76, Scargle82}
+is a well-known algorithm for detecting periodicity
+in unevenly-sampled time-series, particularly within the astronomy community.
+For example, consider the data showing in \Fig{LINEAR_data}.
+This is an irregularly-sampled timeseries showing a single object from the
+LINEAR survey, with magnitude measured 280 times over the course of five and
+a half years.
 
-for a continuously defined signal $g(t)$, the Fourier transform is $\hat{g}(\nu)$,
+It is clear that this object varies with an amplitude of around 0.6 magnitudes.
+Computing the Lomb-Scargle periodogram for the data gives us a measure of the
+power as a function of period of oscillation (\fig{LINEAR_power}, left)\footnote{
+This periodogram was computed using tools in the AstroPy project \citep{Astropy2013};
+code is available at \url{http://github.com/jakevdp/PracticalLombScargle/}
+}, from
+which we can determine the period of oscillation of approximately 2.58 hours.
+The right panel of \fig{LINEAR_power} shows a folded visualization of
+the same data as \fig{LINEAR_data} -- i.e.{} plotted as a function of phase
+rather than time.
 
+Often this is exactly how the Lomb-Scargle periodogram is presented: as a clean,
+well-defined procedure to detect the periodic component in an unevenly-sampled dataset.
+In practice, however, there are a number of subtle issues that must be considered
+when applying a Lomb-Scargle analysis to real-world datasets, and I have found
+that these are rarely presented together at an introductory level.
+This paper seeks to fill that gap, and provide a practical, semi-technical
+guide to the effective use of the Lomb-Scargle method of periodic analysis.
+
+\section{Background: Continuous Fourier Transform}
+
+To understand what the Lomb-Scargle periodogram is doing, we will first step
+back and review the subject of Fourier analysis of continuous signals.
+If we have a continuously-defined signal $g(t)$, we can compute its Fourier
+transform $\hat{g}(f)$ using the following integral:
 \begin{equation}
-    \mathcal{F}\left\{g\right\}(f) = \int_{-\infty}^\infty g(t) e^{-2\pi i f t} dt
+    \hat{g}(f) \equiv \int_{-\infty}^\infty g(t) e^{-2\pi i f t} dt
+    \eqlabel{FT-def}
 \end{equation}
-
-[Perhaps put this info in an appendix?]
-
-Continuous signal $g(t)$:
-
+The inverse relationship is given by:
 \begin{equation}
-  \mathcal{F}\left\{g\right\}(f) = \int_{-\infty}^\infty g(t) e^{-2\pi i f t} dt
+    g(t) \equiv \int_{-\infty}^\infty \hat{g}(f) e^{+2\pi i f t} df
+    \eqlabel{IFT-def}
 \end{equation}
-
-The inverse of this operation is
-
-\begin{equation}
-  \mathcal{F}^{-1}\left\{g\right\}(f) = \int_{-\infty}^\infty g(t) e^{+2\pi i f t} dt
-\end{equation}
-
-We can denote this
-
-\begin{equation}
-  \hat{g} = \mathcal{F}\left\{g\right\}
-\end{equation}
-
-\subsection{Important Properties of the CFT}
-
-\subsubsection{Linearity}
-The Fourier Transform is a linear operation, meaning that for two functions $f(t)$ and $g(t)$ and a scalar value $A$,
-
+This can be straightforwardly confirmed using the linearity of integration and
+the integral definition of the Dirac delta function.
+For convenience below, we will define the Fourier transform operator $\mathcal{F}$,
+such that
 \begin{eqnarray}
-  \mathcal{F}\{g + h\} &=& \mathcal{F}\{g\} + \mathcal{F}\{h\}\\
-  \mathcal{F}\{A g\} &=& A\mathcal{F}\{g\}
+    \mathcal{F}\{g\} &=& \hat{g} \\
+    \mathcal{F}^{-1}\{\hat{g}\} &=& g
 \end{eqnarray}
 
-These properties follow straightforwardly from the linearity of the Fourier transform definition.
+\subsection{Motivating the Power Spectral Density}
+
+If you explore the properties of the Fourier transform, you will discover
+a couple important properties:
+
+\begin{description}
+\item[The Fourier Transform of a Sine wave is a delta function at its frequency.]
+    That is, if $g(t) = \sin(2\pi f_0 t)$, then
+    \begin{equation}
+        \hat{g}(f) \propto \delta(f - f_0) - \delta(f + f_0)
+    \end{equation}
+\item[The Fourier Transform is a Linear Operation.]
+    This means that for functions $a$ and $b$ and scalar $A$,
+    \begin{eqnarray}
+        \mathcal{F}(a + b) &=& \mathcal{F}(a) + \mathcal{F}(b) \\
+        \mathcal{F}(A \cdot a) &=& A\cdot\mathcal{F}(a).
+    \end{eqnarray}
+    These properties follow straightforwardly from the linearity of integration.
+\item[The Fourier Transform of a time-shifted function is unchanged up to a phase.]
+    That is, if $g_\tau(t) = g(t - \tau)$, then
+    \begin{equation}
+        \hat{g}_\tau(f) = e^{-2\pi i f \tau}\hat{g}(f)
+    \end{equation}
+    This can be confirmed via a change of variables in \eq{FT-def}.
+\end{description}
+
+Taken together, what this tells us is that if we define the spectral density
+\begin{equation}
+    \mathcal{P}(g) \equiv \left|\mathcal{F}(g)\right|^2
+\end{equation}
+we have a quantity that is {\it insensitive to time-shifts} and consists of a
+{\it delta function for each sinusoidal frequency within the data}. In other
+words, the spectral density {\it P} measures ...
+
+Because of the first two properties, we see that the Fourier transform can essentially
+extract sinusoidal components within a signal:
+
+\begin{description}
+
+\item[Convolution Theorem.]
+    The convolution theorem states that the
+    Fourier transform of a convolution is a product of the Fourier transforms:
+    \begin{equation}
+      \mathcal{F}\{g \ast h\} = \mathcal{F}\{g\}\mathcal{F}\{h\}
+    \end{equation}
+    where the convlution operator $\ast$ is defined by
+    \begin{equation}
+      \{g \ast h\}(t) \equiv \int_{-\infty}^\infty g(t - \tau) h(\tau) d\tau.
+    \end{equation}
+    This can be confirmed via the linearity of integration and the integral
+    definition of the Dirac delta function.
+
+\item[The Fourier Transform of a Gaussian is a Gaussian.]
+    The Gaussian function is defined as
+    \begin{equation}
+    G(t;\sigma) \propto \exp\left[\frac{-t^2}{2\sigma^2}\right].
+    \end{equation}
+    The Fourier transform of a Gaussian is another Gaussian:
+    \begin{equation}
+    \mathcal{F}\{G(t;\sigma)\} \propto G(f; \sigma_f)
+    \end{equation}
+    where $\sigma_f = (2\pi\sigma)^{-1}$. This can be shown by completing the
+    square within the integrand.
+\end{description}
 
@@
 
-\begin{equation}
-  \{g \ast h\}(t) \equiv \int_{-\infty}^\infty g(t - \tau) h(\tau) d\tau
-\end{equation}
-
 The convolution theorem states that the Fourier transform of a convolution is equal to the product of Fourier transforms; that is,
-
-\begin{equation}
-  \mathcal{F}\{g \ast h\} = \mathcal{F}\{g\}\mathcal{F}\{h\}
-\end{equation}
 

### 2016-12-07 paper.tex
Add some figures & rearrange the outline

@@
 in unevenly-sampled time-series, particularly within the astronomy community.
-For example, consider the data showing in \Fig{LINEAR_data}.
+For example, consider the data shown in \fig{LINEAR_data}.
 This is an irregularly-sampled timeseries showing a single object from the
@@
 
-It is clear that this object varies with an amplitude of around 0.6 magnitudes.
+By eye, it is clear that the brightness of the object varies in time with a range spanning approximately 0.8 magnitudes, but what is not immediately clear is that this variation is periodic in time.
+The Lomb-Scargle periodogram is a method that allows efficient computation of a Fourier-like power spectrum from such unevenly-sampled data, resulting in an intuitive means of determining the period of oscillation.
+
+The left panel of \fig{LINEAR_power} shows the Lomb-Scargle periodogram computed for the data in \fig{LINEAR_data}
 Computing the Lomb-Scargle periodogram for the data gives us a measure of the
-power as a function of period of oscillation (\fig{LINEAR_power}, left)\footnote{
-This periodogram was computed using tools in the AstroPy project \citep{Astropy2013};
-code is available at \url{http://github.com/jakevdp/PracticalLombScargle/}
-}, from
+power as a function of period of oscillation (\fig{LINEAR_power}, left), from
 which we can determine the period of oscillation of approximately 2.58 hours.
@@
 the same data as \fig{LINEAR_data} -- i.e.{} plotted as a function of phase
-rather than time.
+rather than time.\footnote{
+    The periodogram in \fig{LINEAR_power}was computed using
+    {\tt astropy.stats.LombScargle}
+    from the AstroPy project \citep{Astropy2013};
+    Python code to reproduce this figure is available at
+    \url{http://github.com/jakevdp/PracticalLombScargle/}
+}
 
-Often this is exactly how the Lomb-Scargle periodogram is presented: as a clean,
-well-defined procedure to detect the periodic component in an unevenly-sampled dataset.
-In practice, however, there are a number of subtle issues that must be considered
-when applying a Lomb-Scargle analysis to real-world datasets, and I have found
-that these are rarely presented together at an introductory level.
-This paper seeks to fill that gap, and provide a practical, semi-technical
-guide to the effective use of the Lomb-Scargle method of periodic analysis.
+Often this is exactly how the Lomb-Scargle periodogram is presented: as a clean, well-defined procedure to detect the periodic component in an unevenly-sampled dataset.
+In practice, however, there are a number of subtle issues that must be considered when applying a Lomb-Scargle analysis to real-world datasets, and I have found that these are rarely presented together at an introductory level.
+This paper seeks to fill that gap, and provide a practical, semi-technical guide to the effective use of the Lomb-Scargle method of periodic analysis.
+This paper does not seek a complete or rigorous treatment of the mathematics involved, but rather seeks to develop the intuition of {\it how} to think about the issues involved.
 
-\section{Background: Continuous Fourier Transform}
+\section{Background: the Continuous Fourier Transform}
 
-To understand what the Lomb-Scargle periodogram is doing, we will first step
-back and review the subject of Fourier analysis of continuous signals.
-If we have a continuously-defined signal $g(t)$, we can compute its Fourier
-transform $\hat{g}(f)$ using the following integral:
+In order to understand how we should interpret the Lomb-Scargle periodogram, we will first briefly step back and review the subject of Fourier analysis of continuous signals.
+Consider a continuously-defined signal $g(t)$.
+Its Fourier transform is given by the following integral:
 \begin{equation}
@@
 \end{equation}
-This can be straightforwardly confirmed using the linearity of integration and
-the integral definition of the Dirac delta function.
-For convenience below, we will define the Fourier transform operator $\mathcal{F}$,
-such that
+For convenience we will also define the Fourier transform operator
+$\mathcal{F}$, such that
 \begin{eqnarray}
@@
 \end{eqnarray}
+
+The continuous Fourier transform has a number of useful properties that we will make use of in our discussion.
+
+\begin{description}
+   \item[The Fourier transform is a linear operation.]
+     That is, for any constant $A$ and any functions $f$ and $g$, we can write
+     \begin{eqnarray}
+       \mathcal{F}\{f(t) + g(t)\} &=& \mathcal{F}\{f(t)\} + \mathcal{F}\{g(t)\}\nonumber\\
+       \mathcal{F}\{A f(t)\} &=& A\mathcal{F}\{f(t)\},
+     \end{eqnarray}
+     which follows from the linearity of the Fourier integral.
+
+   \item[The Fourier transform of sinusoid with frequency $f_0$ is a sum of delta functions at $\pm f_0$.]
+     Given the integral definition of the Dirac delta function, it is straightforward to show that
+     \begin{equation}
+       \mathcal{F}\{e^{2\pi f_0 t}\} = \delta(f - f_0).
+       \eqlabel{delta_FT}
+     \end{equation}
+     Euler's formula shows how such a complex exponential
+     can be expressed in terms of sines and cosines:
+     \begin{equation}
+       e^{2\pi f_0 t} = \cos(2\pi f_0 t) + i\sin(2\pi f_0 t).
+       \eqlabel{Euler_formula}
+     \end{equation}
+     Using the linearity of the Fourier transform, we can take these identites
+     and derive the following:
+     \begin{eqnarray}
+       \mathcal{F}\{\cos(2\pi f_0 t)\} &=& \frac{1}{2}\left[\delta(f - f_0) + \delta(f + f_0)\right]\\
+       \mathcal{F}\{\sin(2\pi f_0 t)\} &=& \frac{1}{2i}\left[\delta(f - f_0) - \delta(f + f_0)\right].
+     \end{eqnarray}
+     In other words, a sinusoidal signal with frequency $f_0$ has a Fourier transform consisting of a weighted sum of delta functions at $\pm f_0$.
+
+   \item[A time-shift imparts a phase in the Fourier transform.]
+     That is for any function $g(t)$,
+     \begin{equation}
+       \mathcal{F}\{g(t - t_0)\} = e^{-2\pi i ft_0}\mathcal{F}\{g(t)\}
+     \end{equation}
+     The key observation here is that the time-shift does not change the
+     amplitude of the result, but only the phase.
+\end{description}
+
+There are a couple useful consequences of these properties.
+
+...
+
+\subsection{Some Useful Fourier Pairs}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\textwidth]{fig03_Fourier_pairs}
+\caption{Visualization of important Fourier pairs.}
+\end{figure}
+
+Dirac comb
+\begin{equation}
+    C(t, \Delta t) = \sum_{k=-\infty}^\infty \delta(t - k\Delta t)
+\end{equation}
+Fourier transform is
+\begin{equation}
+    \mathcal{F}\{C(t, \Delta t)\} = \frac{1}{\Delta t} C(f, 1/\Delta t)
+\end{equation}
+
+Tophat and Sinc
+\begin{equation}
+  tophat(t; T) \equiv \left\{
+  \begin{array}{ll}
+    1 / T & {\rm if}\ |t| < T / 2 \\
+    0     & {\rm otherwise}
+  \end{array}
+  \right.
+\end{equation}
+Fourier transform is
+\begin{equation}
+  sinc(t) \equiv \frac{\sin(\pi t)}{\pi t}
+\end{equation}
+
+Gaussian
+
+\begin{itemize}
+  \item Chart of Fourier pairs
+\end{itemize}
+
+\subsection{Windowing and Convolution}
+
+\begin{itemize}
+  \item mention convolution theorem; show how it leads to windowing
+  \item figure: show how windowing works
+\end{itemize}
+
+\section{From Continuous to Discrete Signals}
+
+\begin{itemize}
+  \item discrete signal as a Dirac comb window
+  \item derive the Nyquist limit from the Dirac comb FT
+  \item derive the Schuster periodogram
+  \item mention the FFT
+\end{itemize}
+
+\section{Non-uniform sampling}
+
+\begin{itemize}
+  \item nonuniform signal as sum-of-deltas window function
+  \item Nyquist limit does not exist!
+  \item 
+\end{itemize}
+
+\section{The Lomb-Scargle Periodogram}
+
+\begin{itemize}
+  \item Show relation to Schuster periodogram
+  \item Show relation to least squares fitting of sinusoid
+  \item Key: LS measures {\it windowed} power spectrum
+  \item Show effects of truely random sampling; of Kepler-like sampling; of ground-based nightly sampling
+\end{itemize}
+
+\section{Using Lomb-Scargle Effectively}
+
+\begin{itemize}
+  \item Frequency grid
+  \item Peak width
+\end{itemize}
+
+\subsection{Uncertainties in Periods}
+
+\begin{itemize}
+  \item Peak Width
+  \item False Alarm Probability and why it fails
+  \item Baluev method and why it's not enough
+  \item Boostrap
+
+\end{itemize}
+
+\section{Algorithmic Considerations}
+
+\begin{itemize}
+\item Press \& Rybicki method; show some benchmarks.
+\item NFFT method
+\end{itemize}
+
+\section{Generalizations \& Challenges}
+
+\begin{itemize}
+\item Bayesian view doesn't make sense
+\item Multiterm makes the true period fit better, but also bumps the background noise.
+\end{itemize}
+
+\newpage
+
+\section{Draft stuff below...}
+
+\begin{itemize}
+    \item transform and inverse
+    \item properties (delta function $\to$ sinusoid, time shift $\to$ phase, convolution theorem)
+    \item motivating the power spectrum
+    \item some important Fourier pairs (sinusoid$\to$delta, dirac comb$\to$dirac comb, Gaussian$\to$gaussian, top-hat$\to$sinc)
+\end{itemize}
 
@@
 \begin{description}
-\item[The Fourier Transform of a Sine wave is a delta function at its frequency.]
-    That is, if $g(t) = \sin(2\pi f_0 t)$, then
-    \begin{equation}
-        \hat{g}(f) \propto \delta(f - f_0) - \delta(f + f_0)
-    \end{equation}
 \item[The Fourier Transform is a Linear Operation.]
@@
 These properties follow straightforwardly from the linearity of integration.
+\item[The Fourier Transform of a Sinusoid with frequency $f_0$ is a pair of delta
+      functions at $\pm f_0$.]
+    That is, if $g(t) = \sin(2\pi f_0 t)$, then
+    \begin{equation}
+        \hat{g}(f) = \frac{i}{2} \bigg(\delta(f - f_0) - \delta(f + f_0) \bigg)
+    \end{equation}
+    This can be shown straightforwardly using Euler's formula and the integral
+    definition of the Dirac delta function.
 \item[The Fourier Transform of a time-shifted function is unchanged up to a phase.]
@@
     \begin{equation}
-        \hat{g}_\tau(f) = e^{-2\pi i f \tau}\hat{g}(f)
+        \hat{g}_\tau(f) = \phi(\tau)\hat{g}(f)
     \end{equation}
+    where $\phi$ is a phase that depends on the time-shift $\tau$.
     This can be confirmed via a change of variables in \eq{FT-def}.
 \end{description}
-
-Taken together, what this tells us is that if we define the spectral density
+These properties are central to the well-known ability of the Fourier transform
+to measure frequency information from a continuous signal: a signal
+with a periodic component at frequency $f_0$ will have a Fourier transform
+reflecting that frequency. Moreover, we can remove the effect of the phase by
+taking the squared magnitude of the Fourier transform; this is the power
+spectral density:
 \begin{equation}
@@
 \end{equation}
-we have a quantity that is {\it insensitive to time-shifts} and consists of a
-{\it delta function for each sinusoidal frequency within the data}. In other
-words, the spectral density {\it P} measures ...
+Because this operation removes the phase, we now have a quantity that is
+{\it insensitive to time-shifts} and consists of a
+{\it delta function for each sinusoidal frequency within the data}.
+If we want to detect periodic components in a continuous signal,
+this will be a very useful quantity.
 
-Because of the first two properties, we see that the Fourier transform can essentially
-extract sinusoidal components within a signal:
+\subsection{Other Identities}
 
 \begin{description}
-
 \item[Convolution Theorem.]
@@
     definition of the Dirac delta function.
-
 \item[The Fourier Transform of a Gaussian is a Gaussian.]
@@
 \end{description}
-
-\subsubsection{The Convolution Theorem}
-
-The Fourier transform is central to an identity known as {\it the convolution theorem}.
-The convolution of two functions $g(t)$ and $h(t)$ is itself a function of $t$, denoted $\{g \ast h\}(t)$, and is defined as
-
-The convolution theorem states that the Fourier transform of a convolution is equal to the product of Fourier transforms; that is,
-
-This property can also be shown (relatively) straightforwardly from the convolution and Fourier transform definitions.
-
-\subsubsection{$\mathcal{F}$ of a Gaussian is a Gaussian}
-One of the special functions under a Fourier transform is the Gaussian distribution:
-
-\begin{equation}
-    g(t; \mu, \sigma) = \frac{1}{\sqrt{2\pi\sigma^2}} \exp\left[\frac{- (x - \mu)^2}{2\sigma^2}\right]
-\end{equation}, which is proportional to
-
-\subsubsection{$\mathcal{F}$ of a Periodic Function is a Delta Function}
-
-A particularly special function in the complex exponential, $e^{2\pi i\nu_0 t} \equiv \cos(2\pi \nu_0 t) + i\sin(2\pi\nu_0 t)$, which has the Fourier transform $\mathcal{F}\{e^{2\pi i\nu_0 t}\} = \delta(\nu - \nu_0)$, which follows directly from the definition of the Dirac delta function:
-
-\begin{equation}
-  \delta(\nu) \equiv \int_{-\infty}^\infty e^{-2\pi i\nu t}dt
-\end{equation}
-
-Similarly, if we have a signal composed of a sum of complex exponentials, the linearity of the Fourier transform assures that the transform is simply a sum of delta functions:
-\begin{equation}
-  \mathcal{F}\left\{ \sum_{n=1}^N e^{-2\pi i \nu_n t} \right\} = \sum_{n=1}^N \delta(\nu - \nu_n)
-\end{equation}
-This result shows the sense in which a Fourier transform is an ideal measure of periodicity: the Fourier transform pulls-out and identifies the individual periodic components of any input.
 
@@
 \begin{itemize}
-\item convolution function and windowing
+\item convolution theorem and windowing
+\item infinite, regularly-sampled observations
+\item finite, regularly-sampled observations
+\item finite, irregularly-sampled observations
 \end{itemize}
@@
 \begin{itemize}
-\item Talk about peak width and False Alarm Probability; recommend bootstrap (show example of when analytic FAP fails?)
+\item Peak width
+\item False Alarm Probability and why it fails
+\item Baluev method and why it's not enough
+\item Boostrap
 \end{itemize}
@@
 \item Talk about Press \& Rybicki method; show some benchmarks.
+\item NFFT method
 \end{itemize}

### 2016-12-08 paper.tex
Finish discussion of Fourier pairs

@@
 %\SetWatermarkScale{4}
+
+\usepackage{amsmath}
+\DeclareMathOperator{\sinc}{sinc}
+\DeclareMathOperator{\III}{III}
 
@@
    \item[A time-shift imparts a phase in the Fourier transform.]
-     That is for any function $g(t)$,
+     In other words, for any well-behaved function $g(t)$ we can write
      \begin{equation}
-       \mathcal{F}\{g(t - t_0)\} = e^{-2\pi i ft_0}\mathcal{F}\{g(t)\}
+       \mathcal{F}\{g(t - t_0)\} = \mathcal{F}\{g(t)\} e^{-2\pi i ft_0}
      \end{equation}
      The key observation here is that the time-shift does not change the
-     amplitude of the result, but only the phase.
+     amplitude of the resulting transform, but only the phase.
 \end{description}
 
-There are a couple useful consequences of these properties.
+These properties taken together make the Fourier transform quite useful for the study of periodic signals.
+The linearity of the transform means that any signal made up of a sum of periodic components will have a Fourier transform consisting of a sum of delta functions at those frequencies: that is, the Fourier transform is a {\it direct} measure of the periodic content of a function.
 
-...
+Further, if we compute the squared amplitude of the resulting transform, we can both do away with complex components and remove the phase imparted by any time offsets; this squared amplitude is usually known as the {\it Power Spectrum}:
+\begin{equation}
+  \mathcal{P}\{g\} \equiv \left|\mathcal{F}\{g\}\right|^2
+\end{equation}
+The power spectrum of a function is a positive real-valued function of $f$ that quantifies the contribution of each frequency $f$ to the total signal.
 
@@
 \includegraphics[width=\textwidth]{fig03_Fourier_pairs}
-\caption{Visualization of important Fourier pairs.}
+\caption{Visualization of important Fourier pairs.\figlabel{fourier_pairs}}
 \end{figure}
 
-Dirac comb
+We have already discussed that the Fourier transform of a complex exponential is a single delta function.
+This is just one of many ``Fourier Pairs'' to keep in mind as we progress to understanding the Lomb-Scargle Periodogram.
+We list a few more important pairs here (see \Fig{fourier_pairs} for a visual representation of the following pairs):
+
+\begin{description}
+  \item[The Fourier transform of a sinusoid is a pair of Delta functions.]
     \begin{equation}
-    C(t, \Delta t) = \sum_{k=-\infty}^\infty \delta(t - k\Delta t)
+      \mathcal{F}\{\cos(2\pi f_0 t)\} = \frac{1}{2}\left[\delta(f-f_0) + \delta(f+f_0)\right]
     \end{equation}
-Fourier transform is
+    We saw this above, but repeat it here for completeness.
+
+   \item[The Fourier transform of a Gaussian is a Gaussian.]
      \begin{equation}
-    \mathcal{F}\{C(t, \Delta t)\} = \frac{1}{\Delta t} C(f, 1/\Delta t)
+       \mathcal{F}\{{\rm N}(t; \sigma)\} = \frac{1}{\sqrt{2\pi\sigma^2}}{\rm N}\left(f;\frac{1}{2\pi\sigma}\right)
+     \end{equation}
+     The Gaussian function ${\rm N}(t, \sigma)$ is given by
+     \begin{equation}
+       {\rm N}(t; \sigma) \equiv \frac{1}{\sqrt{2\pi\sigma^2}}e^{-t^2 / (2\sigma^2)}
      \end{equation}
 
-Tophat and Sinc
+  \item[The Fourier transform of a tophat function is a sinc function.]
     \begin{equation}
-  tophat(t; T) \equiv \left\{
+      \mathcal{F}\{\Pi_T(t)\} = \sinc(f T)
+    \end{equation}
+    The tophat function, $\Pi_T(t)$ is a normalized symmetric function that
+    is uniform within a range given by $T$, and zero elsewhere:
+    \begin{equation}
+      \Pi_T(t)  \equiv \left\{
       \begin{array}{ll}
-    1 / T & {\rm if}\ |t| < T / 2 \\
-    0     & {\rm otherwise}
+        1 / T, & |t| \le T / 2 \\
+        0,     & |t| > T / 2
       \end{array}
@@
     \end{equation}
-Fourier transform is
+    The sinc function is given by the standard definition:
     \begin{equation}
-  sinc(t) \equiv \frac{\sin(\pi t)}{\pi t}
+      \sinc(x) \equiv \frac{\sin(\pi x)}{\pi x}
     \end{equation}
 
-Gaussian
+  \item[The Fourier transform of a Dirac comb is a Dirac comb.]
+    \begin{equation}
+      \mathcal{F}\{\III_T(t)\} = \frac{1}{T}\III_{1/T}(f)
+    \end{equation}
+    The Dirac comb $\III_T(t)$ is an infinite sequence of Dirac delta functions
+    placed at even intervals of size $T$:
+    \begin{equation}
+      \III_T(t) \equiv \sum_{k=-\infty}^\infty \delta(t - kT)
+    \end{equation}
+\end{description}
 
-\begin{itemize}
-  \item Chart of Fourier pairs
-\end{itemize}
+The key observation in each of these is the reciprocity of scales across the
+Fourier transform: a function with a characteristic scale $T$ will generally
+have a Fourier transform with characteristic scale proportional to $1/T$.
+This feature will turn out to be very important as we push further in
+understanding the Lomb-Scargle Periodogram.
 

### 2016-12-08 paper.tex
Text tweaks; remove some outdated content

@@
 \begin{abstract}
-An introduction to practical aspects of the use of Lomb-Scargle-type algorithms for periodic analysis.
+This paper gives an introduction to the practical aspects of the use of Lomb-Scargle periodograms and related algorithms to detect periodic signals.
+Rather than a rigorous mathematical treatment, the goal of this paper is to build intuition about what assumptions are baked into any periodogram computation.
+We discuss... and ... and ...
 \end{abstract}
@@
 \caption{{\it Left panel:} the Lomb-Scargle periodogram computed from the data
-    shown in \fig{LINEAR_data}, with an inset detailing the interesting region.
+    shown in \fig{LINEAR_data}, with an inset detailing the region near the peak.
     {\it Right panel:} the input data in \fig{LINEAR_data}, folded over the
-    2.58-hour period to show the coherent periodic variability.
+    detected 2.58-hour period to show the coherent periodic variability.
     \figlabel{LINEAR_power}
@@
 This is an irregularly-sampled timeseries showing a single object from the
-LINEAR survey, with magnitude measured 280 times over the course of five and
+LINEAR survey \citep{LINEAR}, with magnitude measured 280 times over the course of five and
 a half years.
@@
 
-The left panel of \fig{LINEAR_power} shows the Lomb-Scargle periodogram computed for the data in \fig{LINEAR_data}
+The left panel of \fig{LINEAR_power} shows the Lomb-Scargle periodogram computed for the data in \fig{LINEAR_data}.
 Computing the Lomb-Scargle periodogram for the data gives us a measure of the
@@
 rather than time.\footnote{
-    The periodogram in \fig{LINEAR_power}was computed using
+    The periodogram in \fig{LINEAR_power} was computed using
     {\tt astropy.stats.LombScargle}
@@
      Using the linearity of the Fourier transform, we can take these identites
-     and derive the following:
+     and derive the following identities:
      \begin{eqnarray}
-       \mathcal{F}\{\cos(2\pi f_0 t)\} &=& \frac{1}{2}\left[\delta(f - f_0) + \delta(f + f_0)\right]\\
+       \mathcal{F}\{\cos(2\pi f_0 t)\} &=& \frac{1}{2}\left[\delta(f - f_0) + \delta(f + f_0)\right]\nonumber\\
        \mathcal{F}\{\sin(2\pi f_0 t)\} &=& \frac{1}{2i}\left[\delta(f - f_0) - \delta(f + f_0)\right].
@@
    \item[A time-shift imparts a phase in the Fourier transform.]
-     In other words, for any well-behaved function $g(t)$ we can write
+     Given a well-behaved function $g(t)$ we can use a transformation of
+     variables to derive the following identity:
      \begin{equation}
@@
 \end{description}
-
 These properties taken together make the Fourier transform quite useful for the study of periodic signals.
@@
   \item[The Fourier transform of a sinusoid is a pair of Delta functions.]
+    (See \fig{fourier_pairs}{\it a})
     \begin{equation}
@@
    \item[The Fourier transform of a Gaussian is a Gaussian.]
+    (See \fig{fourier_pairs}{\it b})
      \begin{equation}
@@
      \begin{equation}
-       {\rm N}(t; \sigma) \equiv \frac{1}{\sqrt{2\pi\sigma^2}}e^{-t^2 / (2\sigma^2)}
+       {\rm N}(t; \sigma) \equiv \frac{1}{\sqrt{2\pi\sigma^2}}e^{-t^2/(2\sigma^2)}
      \end{equation}
@@
   \item[The Fourier transform of a tophat function is a sinc function.]
+    (See \fig{fourier_pairs}{\it c})
     \begin{equation}
@@
   \item[The Fourier transform of a Dirac comb is a Dirac comb.]
+    (See \fig{fourier_pairs}{\it d})
     \begin{equation}
@@
   \item nonuniform signal as sum-of-deltas window function
-  \item Nyquist limit does not exist!
-  \item 
+  \item Nyquist limit is meaningless!
+  \item Some use notions of ``average Nyquist limit'', but this can be misleading (reference LINEAR data in \fig{LINEAR_data})
 \end{itemize}
@@
 \begin{itemize}
-  \item Show relation to Schuster periodogram
+  \item Show relation to Shuster periodogram
   \item Show relation to least squares fitting of sinusoid
@@
 \begin{itemize}
-  \item Frequency grid
-  \item Peak width
+  \item Expected peak width
+  \item How to choose the frequency grid
+  \item Normalization and interpretation
 \end{itemize}
@@
 \begin{itemize}
-  \item Peak Width
+  \item Peak Width and why it's a poor proxy
   \item False Alarm Probability and why it fails
   \item Baluev method and why it's not enough
-  \item Boostrap
+  \item Motivation of Bootstrap
 
@@
 
-\section{Draft stuff below...}
-
-\begin{itemize}
-    \item transform and inverse
-    \item properties (delta function $\to$ sinusoid, time shift $\to$ phase, convolution theorem)
-    \item motivating the power spectrum
-    \item some important Fourier pairs (sinusoid$\to$delta, dirac comb$\to$dirac comb, Gaussian$\to$gaussian, top-hat$\to$sinc)
-\end{itemize}
-
-\subsection{Motivating the Power Spectral Density}
-
-If you explore the properties of the Fourier transform, you will discover
-a couple important properties:
-
-\begin{description}
-\item[The Fourier Transform is a Linear Operation.]
-This means that for functions $a$ and $b$ and scalar $A$,
-\begin{eqnarray}
-    \mathcal{F}(a + b) &=& \mathcal{F}(a) + \mathcal{F}(b) \\
-    \mathcal{F}(A \cdot a) &=& A\cdot\mathcal{F}(a).
-\end{eqnarray}
-These properties follow straightforwardly from the linearity of integration.
-\item[The Fourier Transform of a Sinusoid with frequency $f_0$ is a pair of delta
-      functions at $\pm f_0$.]
-    That is, if $g(t) = \sin(2\pi f_0 t)$, then
-    \begin{equation}
-        \hat{g}(f) = \frac{i}{2} \bigg(\delta(f - f_0) - \delta(f + f_0) \bigg)
-    \end{equation}
-    This can be shown straightforwardly using Euler's formula and the integral
-    definition of the Dirac delta function.
-\item[The Fourier Transform of a time-shifted function is unchanged up to a phase.]
-    That is, if $g_\tau(t) = g(t - \tau)$, then
-    \begin{equation}
-        \hat{g}_\tau(f) = \phi(\tau)\hat{g}(f)
-    \end{equation}
-    where $\phi$ is a phase that depends on the time-shift $\tau$.
-    This can be confirmed via a change of variables in \eq{FT-def}.
-\end{description}
-These properties are central to the well-known ability of the Fourier transform
-to measure frequency information from a continuous signal: a signal
-with a periodic component at frequency $f_0$ will have a Fourier transform
-reflecting that frequency. Moreover, we can remove the effect of the phase by
-taking the squared magnitude of the Fourier transform; this is the power
-spectral density:
-\begin{equation}
-    \mathcal{P}(g) \equiv \left|\mathcal{F}(g)\right|^2
-\end{equation}
-Because this operation removes the phase, we now have a quantity that is
-{\it insensitive to time-shifts} and consists of a
-{\it delta function for each sinusoidal frequency within the data}.
-If we want to detect periodic components in a continuous signal,
-this will be a very useful quantity.
+\section{Draft Stuff}
 
@@
 
-\citet{VanderPlas2015}
-
 \bibliographystyle{apj}

### 2016-12-09 paper.tex
Add convolution diagram

@@
     \end{equation}
-    The tophat function, $\Pi_T(t)$ is a normalized symmetric function that
+    The tophat function, $\Pi_T(t)$, is a normalized symmetric function that
     is uniform within a range given by $T$, and zero elsewhere:
@@
     \end{equation}
-    The Dirac comb $\III_T(t)$ is an infinite sequence of Dirac delta functions
-    placed at even intervals of size $T$:
+    The Dirac comb, $\III_T(t)$, is an infinite sequence of Dirac delta functions placed at even intervals of size $T$:
     \begin{equation}
-      \III_T(t) \equiv \sum_{k=-\infty}^\infty \delta(t - kT)
+      \III_T(t) \equiv \sum_{n=-\infty}^\infty \delta(t - nT)
     \end{equation}
 \end{description}
-
-The key observation in each of these is the reciprocity of scales across the
-Fourier transform: a function with a characteristic scale $T$ will generally
-have a Fourier transform with characteristic scale proportional to $1/T$.
+The key observation in each of these Fourier pairs is the reciprocity
+of scales between a function and its Fourier transform:
+that is, a function with a characteristic scale $T$ will in general
+have a Fourier transform with characteristic scale of $1/T$.
 This feature will turn out to be very important as we push further in
@@
 \begin{itemize}
-  \item mention convolution theorem; show how it leads to windowing
-  \item figure: show how windowing works
+  \item Convolution theorem; \fig{convolution}
+  \item Windowing: Convolution theorem in reverse (fig?)
 \end{itemize}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig04_Convolution_Diagram}
+  \caption{Visualization of the Convolution Theorem\figlabel{convolution}}
+\end{figure}
 

### 2016-12-10 paper.tex
More window func diagrams

@@
 
-\begin{itemize}
-  \item Convolution theorem; \fig{convolution}
-  \item Windowing: Convolution theorem in reverse (fig?)
-\end{itemize}
+A final property of the Fourier transform that we will discuss here is its
+ability to convert convolutions into point-wise products.
+A convolution of two functions, usually denoted with a $\ast$, is defined
+as follows:
+\begin{equation}
+  [f \ast g](t) = \int_{-\infty}^\infty f(\tau)g(t - \tau) d\tau
+  \eqlabel{convolution_definition}
+\end{equation}
+From the definition, it is clear that a convolution amounts to ``sliding'' one
+function past the other.
+Such an operation is commonly used, for example, in smoothing a function,
+as visualized in \Fig{convolution} for a tophat-shaped smoothing window.
 
@@
   \includegraphics[width=\textwidth]{fig04_Convolution_Diagram}
-  \caption{Visualization of the Convolution Theorem\figlabel{convolution}}
+  \caption{Visualization of a convolution as often used for kernel smoothing.
+    The normalized tophat-shaped window function slides across the domain (upper panel),
+    such that at each point the mean of the values within the window are
+    used to compute the smoothed function (lower panel).
+    \figlabel{convolution}}
 \end{figure}
+
+Given this definition of a convolution, it can be shown that the Fourier transform of a convolution is the point-wise product of the Fourier transforms:
+\begin{equation}
+  \mathcal{F}\{f \ast g\} = \mathcal{F}\{f\} \cdot \mathcal{F}\{g\}
+  \eqlabel{convolution_theorem}
+\end{equation}
+In practice, this is a much more efficient means of computing the convolution
+than to directly solve at each time $t$ the integral over $\tau$ that appears
+in \eq{convolution_definition}.
+This result in \eq{convolution_theorem} is known as the
+{\it convolution theorem}, and is illustrated in \fig{convolution_theorem}.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig05_Convolution_Theorem}
+  \caption{Visualization of a convolution used for kernel smoothing
+    \figlabel{convolution_theorem}}
+\end{figure}
+
+Likewise, the Fourier transform of a product is a convolution of the transforms:
+\begin{equation}
+  \mathcal{F}\{f \cdot g\} = \mathcal{F}\{f\} \ast \mathcal{F}\{g\}
+\end{equation}
+This becomes
 

### 2016-12-11 paper.tex
Finish convolution diagram

@@
   \includegraphics[width=\textwidth]{fig04_Convolution_Diagram}
-  \caption{Visualization of a convolution as often used for kernel smoothing.
+  \caption{Visualization of a convolution between a continuous signal and a tophat smoothing kernel.
     The normalized tophat-shaped window function slides across the domain (upper panel),
@@
   \includegraphics[width=\textwidth]{fig05_Convolution_Theorem}
-  \caption{Visualization of a convolution used for kernel smoothing
+  \caption{Visualization of the Convolution theorem as used to compute the
+    result in \fig{convolution}. The key is that the Fourier transform of
+    a convolution is the pointwise product of the two Fourier transforms.
     \figlabel{convolution_theorem}}

### 2016-12-12 paper.tex
Update windowing figures

@@
 This is an irregularly-sampled timeseries showing a single object from the
-LINEAR survey \citep{LINEAR}, with magnitude measured 280 times over the course of five and
-a half years.
+LINEAR survey \citep{LINEAR}, with magnitude measured 280 times over the
+course of five and a half years\footnote{
+  Python code to reproduce \fig{LINEAR_data}, as well as all other figures
+  in this manuscript, is available at 
+  \url{http://github.com/jakevdp/PracticalLombScargle/}}.
 
@@
     {\tt astropy.stats.LombScargle}
-    from the AstroPy project \citep{Astropy2013};
-    Python code to reproduce this figure is available at
-    \url{http://github.com/jakevdp/PracticalLombScargle/}
+    from the AstroPy project; see \url{http://astropy.org/}.
 }
@@
 In practice, however, there are a number of subtle issues that must be considered when applying a Lomb-Scargle analysis to real-world datasets, and I have found that these are rarely presented together at an introductory level.
+Here are a few questions in particular that we might wish to ask about the results in \fig{LINEAR_power}:
+\begin{enumerate}
+  \item What is the source of the multiple peaks revealed by the Lomb-Scargle Periodogram?
+  \item What is a suitable range of frequencies to search for a given dataset?
+  \item What is the Nyquist frequency for this type of analysis?
+  \item How should we choose the spacing of the frequency grid for our periodogram?
+  \item How should we understand the uncertainty of the computed frequency?
+  \item What assumptions is the Lomb-Scargle periodogram making about the signal?
+\end{enumerate}
+Questions like these are discussed in various textbooks and review papers, but I have not come across any concise reference that discusses such questions in depth.
 This paper seeks to fill that gap, and provide a practical, semi-technical guide to the effective use of the Lomb-Scargle method of periodic analysis.
@@
 \end{eqnarray}
+As a further shorthand, we'll sometimes use a double arrow to denote a Fourier
+pair; for example
+\begin{equation}
+  g(x) \Longleftrightarrow \hat{g}(f).
+\end{equation}
+
+\subsection{Properties of the Fourier Transform}
 
@@
 
-\subsection{Windowing and Convolution}
+\subsection{The Convolution Theorem}
 
@@
 {\it convolution theorem}, and is illustrated in \fig{convolution_theorem}.
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig05_Convolution_Theorem}
+  \caption{Visualization of the Convolution theorem (\eq{convolution_theorem}).
+    The key is that the Fourier transform of
+    a convolution is the pointwise product of the two Fourier transforms.
+    \figlabel{convolution_theorem}}
+\end{figure}
+When thinking about frequency components of time-domain measurements, this
+property of the Fourier transform becomes essential, as we will see.
+
+An important corrollary is that the Fourier transform of a product is a convolution of the two transforms:
+\begin{equation}
+  \mathcal{F}\{f \cdot g\} = \mathcal{F}\{f\} \ast \mathcal{F}\{g\}
+  \eqlabel{convolution_theorem_inverse}
+\end{equation}
+We will explore the impact of this in the next section.
+
+\section{Window Functions: From Continuous to Discrete Signals}
+In a typical observational setting, we will be observing some small part of
+a larger time-varying signal.
+For example, if you have a very long signal that you measure for a finite
+amount of time, your measurement is a point-wise product between the true
+signal and a suitable tophat function describing your measurement process.
+If you have a continuous signal that you measure at regular intervals, your
+measurement is a point-wise product between the true signal and an appropriate
+Dirac comb describing your measurement process.
+From what we learned about the convolution theorem in the previous section, it
+is clear that {\it the measured Fourier transform will be the true transform
+convolved with the transform of the obverving window}.
+This has some interesting consequences, as we shall see.
+
+\subsection{Effect of a Tophat Window}
+First, let's consider the case of observing a continuous periodic signal over
+a limited span of time, as illustrated in \Fig{tophat_window}.
+Here we consider a periodic function made up of three frequency components, observed within a 10-unit time window.
+The observed signal in this case can be thought of as a pointwise product of an infinite periodic signal, and a tophat window function.
+Using \eq{convolution_theorem_inverse}, we see that the Fourier transform of
+this function is equivalent to the convolution of the Fourier transforms of each component, which are a set of delta functions and a sinc function, respectively.
+For a purely periodic signal, this convolution has the effect of broadening the delta functions representing the signal, turning each of them into a shifted sinc function.
+Because of the inverse relationship between the width of the window and the width of its transform (see \Fig{fourier_pairs}), it follows that a wider observing window leads to proportionally less spread in the Fourier transform of the observed function.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{fig05_Convolution_Theorem}
-  \caption{Visualization of the Convolution theorem as used to compute the
-    result in \fig{convolution}. The key is that the Fourier transform of
-    a convolution is the pointwise product of the two Fourier transforms.
-    \figlabel{convolution_theorem}}
+  \includegraphics[width=\textwidth]{fig06_Tophat_Window}
+  \caption{Visualization of the effect on the Fourier transform of a
+    tophat-shaped observing window (i.e., a continuous signal observed
+    fully, but only within a finite range of time). The observed Fourier
+    transform is a convolution of the true transform (here a series of Delta
+    functions) and the window transform (here a narrow sinc function).
+    \figlabel{tophat_window}}
 \end{figure}
 
-Likewise, the Fourier transform of a product is a convolution of the transforms:
-\begin{equation}
-  \mathcal{F}\{f \cdot g\} = \mathcal{F}\{f\} \ast \mathcal{F}\{g\}
-\end{equation}
-This becomes
+\subsection{Effect of a Dirac Comb}
 
-\section{From Continuous to Discrete Signals}
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig07_comb_window_1}
+  \caption{Visualization of the effect on the Fourier transform of a
+    Dirac Comb observing window (i.e., a long string of evenly-spaced
+    discrete observations). The observed Fourier
+    transform is a convolution of the true transform (here a localized
+    Gaussian) and the window transform (here another Dirac comb).
+    \figlabel{comb_window}}
+\end{figure}
+
+\subsection{Understanding the Nyquist Frequency}
 
@@
 
-\section{Non-uniform sampling}
+\section{Non-uniform Sampling Windows}
 
@@
 
+\section{Acknowledgements}
+
+\citet{Astropy2013}
+
 \bibliographystyle{apj}

### 2016-12-13 paper.tex
Add second dirac comb window figure

@@
 \subsection{Effect of a Dirac Comb}
+Another window function that commonly arises is when a continuous signal is
+sampled instantaneously at regular intervals.
+Such an observation is, in effect, a point-wise product between the true
+underlying signal and a Dirac comb with the $T$ parameter matching the spacing
+of the observations; this is illustrated in \fig{comb_window_1}.
 
@@
     Gaussian) and the window transform (here another Dirac comb).
-    \figlabel{comb_window}}
+    \figlabel{comb_window_1}}
 \end{figure}
 
-\subsection{Understanding the Nyquist Frequency}
+Interestingly, because the Fourier transform of a Dirac comb is another Dirac
+comb, the effect of such an observing window is to create a long sequence
+of aliases with a spacing of $1/T$.
+With this in mind, we can be assured that evaluating the transform in the
+range $[-(2T)^{-1}, (2T)^{-1}]$ is sufficient to capture all the relevant
+frequency information: everything else will simply be an alias of what appears
+within that range.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig08_comb_window_2}
+  \caption{Visualization of the effect on the Fourier transform of a
+    Dirac Comb observing window (i.e., a long string of evenly-spaced
+    discrete observations). The observed Fourier
+    transform is a convolution of the true transform (here a localized
+    Gaussian) and the window transform (here another Dirac comb).
+    \figlabel{comb_window_2}}
+\end{figure}
 
@@
 
-\subsection{Uncertainties in Periods}
+\section{Uncertainties in Periods}
 
@@
 
-\newpage
-
-\section{Draft Stuff}
-
-\subsection{Other Identities}
-
-\begin{description}
-\item[Convolution Theorem.]
-    The convolution theorem states that the
-    Fourier transform of a convolution is a product of the Fourier transforms:
-    \begin{equation}
-      \mathcal{F}\{g \ast h\} = \mathcal{F}\{g\}\mathcal{F}\{h\}
-    \end{equation}
-    where the convlution operator $\ast$ is defined by
-    \begin{equation}
-      \{g \ast h\}(t) \equiv \int_{-\infty}^\infty g(t - \tau) h(\tau) d\tau.
-    \end{equation}
-    This can be confirmed via the linearity of integration and the integral
-    definition of the Dirac delta function.
-\item[The Fourier Transform of a Gaussian is a Gaussian.]
-    The Gaussian function is defined as
-    \begin{equation}
-    G(t;\sigma) \propto \exp\left[\frac{-t^2}{2\sigma^2}\right].
-    \end{equation}
-    The Fourier transform of a Gaussian is another Gaussian:
-    \begin{equation}
-    \mathcal{F}\{G(t;\sigma)\} \propto G(f; \sigma_f)
-    \end{equation}
-    where $\sigma_f = (2\pi\sigma)^{-1}$. This can be shown by completing the
-    square within the integrand.
-\end{description}
-
-\subsection{Fourier Power}
-
-Using the above definitions, it is straightforward to show that a time-offset in a function leads to a multiplicative phase in the Fourier transform; that is, if
-\begin{equation}
-  g_\tau(t) \equiv g(t - \tau)
-\end{equation}
-then
-\begin{equation}
-  \mathcal{F}\{g_\tau\} = e^{-2\pi i \nu \tau}\mathcal{F}\{g\}
-\end{equation}
-Because the phase merely represents the chosen coordinate system, we'd like to remove it for spectral analysis.
-One way to do this is to compute the power spectral density, which is proportional to \todo{const of proportionality?} the squared modulus of the Fourier Transform:
-\begin{equation}
-  P\{g\} \propto \left| \mathcal{F}\{g\} \right|^2.
-\end{equation}
-Because the modulus of a phase is equal to unity, the resulting measure is unchanged under time-translations in $g(t)$, and $P\{g\} = P\{g_\tau\}$ for any time-shift $\tau$.
-
-\begin{itemize}
-\item Some examples of functions, their transforms, and their power.
-\end{itemize}
-
-\subsection{Effect of Window Functions}
-
-\begin{itemize}
-\item convolution theorem and windowing
-\item infinite, regularly-sampled observations
-\item finite, regularly-sampled observations
-\item finite, irregularly-sampled observations
-\end{itemize}
-
-\section{From Idealistic to Realistic}
-
-All of the above theory is beautiful and elegant, but when the theory meets the real world things get a bit more messy.
-In particular, we rarely (if ever) are able to measure continuous functions like $g(t)$, but instead measure discrete realizations of the underlying function at a finite number of particular times $t_n$.
-These discrete observations are effectively a window function over our data, and as we saw above, the presence of such a window function will propagate through our computation of the Fourier power!
-The quantitative result of this windowning depends on the exact nature of the observation pattern; we will first consider the somewhat simpler case of evenly-spaced discrete data.
-
-\subsection{Evenly-Spaced Discrete Data: the Schuster Periodogram}
-
-Let's first consider the case of a large series of evenly-spaced observations.
-Discrete samples can
-
-\begin{itemize}
-\item Evenly-spaced delta functions $\to$ evenly-spaced delta in frequency, separated by nyquist
-\end{itemize}
-
-\subsection{Unevenly-Spaced Discrete Data: the Lomb-Scargle Periodogram}
-
-\begin{itemize}
-\item Unevenly-spaced delta functions: no cancelations, and so window is generally {\it infinitely} wide.
-\item Least-squares equivalence of Lomb-Scargle
-\end{itemize}
-
-\section{What Frequency Grid to Use?}
-
-\begin{itemize}
-\item Follow discussion from gatspy documentation
-\end{itemize}
-
-\section{Reporting errors in Frequency: Uncertainty vs. Precision}
-
-\begin{itemize}
-\item Peak width
-\item False Alarm Probability and why it fails
-\item Baluev method and why it's not enough
-\item Boostrap
-\end{itemize}
-
-\section{Algorithmic Considerations}
-
-\begin{itemize}
-\item Talk about Press \& Rybicki method; show some benchmarks.
-\item NFFT method
-\end{itemize}
-
-\section{Generalizations \& Challenges}
-
-\begin{itemize}
-\item Multiterm makes the true period fit better, but also bumps the background noise.
-\end{itemize}
-
 \section{Acknowledgements}

### 2016-12-13 paper.tex
Add Nyquist discussion

@@
 
-\subsection{Effect of a Dirac Comb}
+\subsection{The Dirac Comb and the Nyquist Limit}
 Another window function that commonly arises is when a continuous signal is
@@
 of aliases with a spacing of $1/T$.
-With this in mind, we can be assured that evaluating the transform in the
-range $[-(2T)^{-1}, (2T)^{-1}]$ is sufficient to capture all the relevant
-frequency information: everything else will simply be an alias of what appears
-within that range.
+With this in mind, we can be assured in this case that evaluating the
+transform in the range $-(2T)^{-1} \le f < (2T)^{-1}$ is sufficient to capture
+all the relevant frequency information: the signal outside that range is a
+string of duplicates of what lies in that range.
+
+The example in \fig{comb_window_1} is somewhat of a best-case scenario, because
+the true Fourier transform lies entirely in the range noted above.
+If we change the sampling rate so that this is no longer the case, we will
+have a situation similar to that in \fig{comb_window_2}.
+With a lower sampling rate, the signal transform no longer fits within the
+frequency window, and the true Fourier transform {\it cannot be recovered}
+from the observed transform.
+
+This implies that if we have a regularly-sampled function with a sampling
+rate of of $f_0 = 1/T$, we can only fully recover the freuqency information
+if the signal is {\it band-limited} between frequencies $\pm f_0/2$.
+This is one way to motivate the famous Nyquist sampling limit, which says
+that to fully represent the frequency content of a ``band-limited signal''
+with all Fourier power in the range $\pm B$, we must sample the data with a
+rate of at least $2B$.
 

### 2016-12-13 paper.tex
Add discussion of Discrete FT

@@
 This is an irregularly-sampled timeseries showing a single object from the
-LINEAR survey \citep{LINEAR}, with magnitude measured 280 times over the
+LINEAR survey \citep{LINEAR1, LINEAR3}, with magnitude measured 280 times over the
 course of five and a half years\footnote{
@@
        \mathcal{F}\{f(t) + g(t)\} &=& \mathcal{F}\{f(t)\} + \mathcal{F}\{g(t)\}\nonumber\\
-       \mathcal{F}\{A f(t)\} &=& A\mathcal{F}\{f(t)\},
+       \mathcal{F}\{A f(t)\} &=& A\mathcal{F}\{f(t)\}
      \end{eqnarray}
@@
       \III_T(t) \equiv \sum_{n=-\infty}^\infty \delta(t - nT)
+      \eqlabel{dirac_comb}
     \end{equation}
@@
 
+\subsubsection{The Nyquist Limit}
+
 The example in \fig{comb_window_1} is somewhat of a best-case scenario, because
@@
   \includegraphics[width=\textwidth]{fig08_comb_window_2}
-  \caption{Visualization of the effect on the Fourier transform of a
-    Dirac Comb observing window (i.e., a long string of evenly-spaced
-    discrete observations). The observed Fourier
-    transform is a convolution of the true transform (here a localized
-    Gaussian) and the window transform (here another Dirac comb).
+  \caption{Repeating the visualization from \fig{comb_window_1}, but here with
+    a lower sampling rate. The result is that the Fourier transform of the
+    window function (lower left) has spacing narrower than the Fourier transform
+    of the signal (middle left), meaning the observed Fourier transform has
+    aliasing of signals, such that not all frequency information can be
+    recovered.
     \figlabel{comb_window_2}}
@@
 
-\begin{itemize}
-  \item discrete signal as a Dirac comb window
-  \item derive the Nyquist limit from the Dirac comb FT
-  \item derive the Schuster periodogram
-  \item mention the FFT
-\end{itemize}
+\subsubsection{The Discrete Fourier Transform}
+For regularly-spaced observations defined by a Dirac comb, the delta functions
+serve to collapse the Fourier integral into a sum over periodic components.
+if $g^{(T)}(t) \equiv g(t) \III_T(t)$, then its Fourier transform is
+\begin{equation}
+  \hat{g}^{(T)}(f) = \sum_{n=-\infty}^\infty g(nT) e^{-2\pi i f n T},
+\end{equation}
+which follows directly from \eq{FT-def} and \eq{dirac_comb}.
+In the real world, we will not have an infinite number of samples, but rather
+a finite number of samples $N$, and we can choose the coordinate system
+appropriately and define $g(nT) \equiv g_n$ to write
+\begin{equation}
+  \hat{g}^{(N,T)}(f) = \sum_{n=0}^N g_n e^{-2\pi i f n T}
+\end{equation}
+From the arguments around Nyquist aliasing, we know that the only relevant
+frequency range is from $0 \le f \le 1/T$, and so we can define $N$
+evenly-spaced frequencies with $\Delta f = 1 / (NT)$ covering this range.
+Denoting the sampled transform as
+$\hat{g}_k \equiv \hat{g}^{(N, T)}(k\Delta f)$, we can write
+\begin{equation}
+  \hat{g}_k = \sum_{n=0}^N g_n e^{-2\pi i k n / N}
+\end{equation}
+which you may recognize as the standard discrete Fourier transform.
+
+But you may have noticed we glossed over something: the effect of switching
+from an infinite number of samples to a finite number of samples.
+In moving from $\hat{g}^{(T)}$ to $\hat{g}^{(N,T)}$, we have effectively applied
+to our data a rectangular window function of width $NT$.
+From the discussion accompanying \fig{tophat_window}, we know what this does:
+it gives us a Fourier transform convolved with a sinc function of width
+$1 / (NT)$, resulting in the ``smearing'' of the Fourier transform signal
+with this width.
+Roughly speaking, then, any two Fourier transform values at frequencies within
+$1/(NT)$ of each other will not be independent, and so we should space our
+evaluations of the frequency with $\Delta f \ge 1/(NT)$.
+Comparing to above, we see that this is {\it exactly the frequency spacing}
+we arrived at from Nyquist-frequency arguments.
+
+What this indicates is that the frequency spacing of the discrete Fourier
+transform is optimal in terms of both the Nyquist sampling limit
+{\it and} the effect of the finite observing window!
+Now, this argument has admittedly been a bit hand-wavy, but there do exist more
+mathematically rigorous approaches to showing these
+properties of the discrete Fourier transform
+\citep[see, e.g.][]{FoundationsOfSignalProcessing}.
+Nevertheless, I find this to be a helpful approach
+in developing intuition regarding the relationship between the
+continuous and discrete Fourier transforms.
+
+\subsubsection{The Schuster Periodogram}
+
+...
 

### 2016-12-16 paper.tex
Add random window figures

@@
 \includegraphics[width=\textwidth]{fig01_LINEAR_data}
-\caption{Observed light curve from LINEAR object ID 11375941
+\caption{Observed light curve from LINEAR object ID 11375941. Uncertainties
+  are indicated by the gray errorbars on each point.
   \figlabel{LINEAR_data}.
@@
   \mathcal{P}\{g\} \equiv \left|\mathcal{F}\{g\}\right|^2
+  \eqlabel{power-spectrum}
 \end{equation}
@@
 
-\section{Window Functions: From Continuous to Discrete Signals}
+\section{Window Functions: From Idealized to Real-world signals}
 In a typical observational setting, we will be observing some small part of
@@
 
-\subsection{The Dirac Comb and the Nyquist Limit}
+\subsection{The Dirac Comb and the Discrete Fourier Transform}
 Another window function that commonly arises is when a continuous signal is
@@
 With this in mind, we can be assured in this case that evaluating the
-transform in the range $-(2T)^{-1} \le f < (2T)^{-1}$ is sufficient to capture
-all the relevant frequency information: the signal outside that range is a
-string of duplicates of what lies in that range.
+transform in the range $(2T)^{-1} \le f < (2T)^{-1}$ is sufficient to capture
+all the available frequency information:
+the signal outside that range is a sequence of identical aliases of
+what lies in that range.
+In fact, due to the nature of the aliasing, it is in fact sufficient to
+compute the transform in {\it any} range of width $1/T$, and it is often
+convenient to instead use the range $0 \le f < 1/T$.
 
@@
 The example in \fig{comb_window_1} is somewhat of a best-case scenario, because
-the true Fourier transform lies entirely in the range noted above.
+the true Fourier transform values lie entirely in a range of width $1/T$.
 If we change the sampling rate so that this is no longer the case, we will
@@
 serve to collapse the Fourier integral into a sum over periodic components.
-if $g^{(T)}(t) \equiv g(t) \III_T(t)$, then its Fourier transform is
+Suppose we have a true (infinitely long and continuous) signal $g(t)$, but
+we observe it only at a regular grid with spacing $T$. In this case, our
+observed signal is $g_{obs} = g(t) \III_T(t)$ and its Fourier transform is
 \begin{equation}
-  \hat{g}^{(T)}(f) = \sum_{n=-\infty}^\infty g(nT) e^{-2\pi i f n T},
+  \hat{g}_{obs}(f) = \sum_{n=-\infty}^\infty g(nT) e^{-2\pi i f n T},
 \end{equation}
 which follows directly from \eq{FT-def} and \eq{dirac_comb}.
-In the real world, we will not have an infinite number of samples, but rather
-a finite number of samples $N$, and we can choose the coordinate system
-appropriately and define $g(nT) \equiv g_n$ to write
+
+In the real world, however, we will not have an infinite number of observations,
+but rather a finite number of samples $N$.
+We can choose the coordinate system appropriately and define
+$g_n \equiv g(nT)$ to write
 \begin{equation}
-  \hat{g}^{(N,T)}(f) = \sum_{n=0}^N g_n e^{-2\pi i f n T}
+  \hat{g}_{obs}(f) = \sum_{n=0}^N g_n e^{-2\pi i f n T}
+  \eqlabel{DFT_f}
 \end{equation}
@@
 Denoting the sampled transform as
-$\hat{g}_k \equiv \hat{g}^{(N, T)}(k\Delta f)$, we can write
+$\hat{g}_k \equiv \hat{g}_{obs}(k\Delta f)$, we can write
 \begin{equation}
   \hat{g}_k = \sum_{n=0}^N g_n e^{-2\pi i k n / N}
+  \eqlabel{DFT}
 \end{equation}
-which you may recognize as the standard discrete Fourier transform.
+which you might recognize as the standard form of the discrete Fourier
+transform.
 
-But you may have noticed we glossed over something: the effect of switching
+You may have noticed we glossed over something here: the effect of switching
 from an infinite number of samples to a finite number of samples.
@@
 {\it and} the effect of the finite observing window!
-Now, this argument has admittedly been a bit hand-wavy, but there do exist more
-mathematically rigorous approaches to showing these
-properties of the discrete Fourier transform
+Now, this argument has admittedly been a bit hand-wavy, but there do exist
+mathematically rigorous approaches to proving that the discrete Fourier
+transform in \eq{DFT} captures all of the available frequency information
+for a uniformly-sampled function $g_n$
 \citep[see, e.g.][]{FoundationsOfSignalProcessing}.
-Nevertheless, I find this to be a helpful approach
+Despite the semi-qualitative nature of our discussion,
+I find it to be a helpful approach
 in developing intuition regarding the relationship between the
@@
 
-...
+With the discrete Fourier transform defined in \eqs{DFT_f}{DFT}, we can
+apply the definition of the Fourier power spectrum from \eq{power-spectrum}
+to compute the {\it Schuster Periodogram}, first proposed by \citet{Schuster98}:
+\begin{equation}
+  P_S(f) = \frac{1}{N}\left|\sum_{n=1}^N g_n e^{-2\pi i f t_n}\right|^2
+\end{equation}
+Apart from the $1/N$ scaling, this is precisely the Fourier power defined
+in \eq{power-spectrum} for the expression in \eq{DFT_f}, and it follows that
+for a uniformly-sampled function $g_n$, the Schuster periodogram captures
+all of the relevant frequency information present in the data.
+
+\todo{Add simulated RR Lyrae data and show the Shuster periodogram}
 
 \section{Non-uniform Sampling Windows}
+
+In the real world, particularly in fields like Astronomy where observations are
+subject to interruptions from weather and diurnal cycles, the sampling rate
+is generally far from uniform.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig10_random_window}
+  \caption{
+    \figlabel{random_window}}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig11_random_window_2}
+  \caption{
+    \figlabel{random_window_2}}
+\end{figure}
 

### 2016-12-16 paper.tex
Update nonuniform window discussion

@@
 is generally far from uniform.
+Using the same approach as we used to explore uniform sampling in the previous
+section, we can now explore non-uniform sampling here.
+
+In the general non-uniform case, we measure some signal at a set of $N$ times
+which we will denote $\{t_n\}$.
+Our observing window function in this case is still a sum of delta functions,
+but at non-uniform locations.
+Mathematically, the observing window will look like this:
+
+\begin{equation}
+W(t; \{t_n\}) = \sum_{n=1}^{N} \delta(t - t_n)
+\end{equation}
+
+Applying this window to our true underlying signal $g(t)$, we find an observed
+signal of the form:
+
+\begin{eqnarray}
+  g_{obs}(t) &=& g(t) W(t; \{t_n\}) \nonumber\\
+             &=& \sum_{n=1}^{N} g(t_n)\delta(t - t_n)
+\end{eqnarray}
+
+Just as in the uniform case, the observed signal is a convolution of the true
+signal transform and the window transform; unlike in the uniform case, the
+window transform will generally {\it not} be a nice clean function.
+\Fig{random_window} shows the effect on the Fourier transform of a
+non-uniform observing cadence with a sampling rate identical to that
+in \fig{comb_window_1}.
+
+A few things stand-out in this figure. In particular, the Fourier transform of
+the non-uniformly spaced delta functions looks like random noise, and in some
+senses it is: the Fourier space representation reflects the frequency of
+observations, and so non-structured spacing of observations will lead to
+a non-structured Fourier representation of the window function.
+This non-structured window transform, when convolved with the Fourier transform
+of the true signal, results in an observed Fourier transform with the same
+lack of useful structure.
+Unlike the uniform result in \Fig{comb_window_1}, there is no direct aliasing
+of the true signal, and no way to exactly recover any portion of the true
+Fourier transform of this data.
 
@@
   \includegraphics[width=\textwidth]{fig10_random_window}
-  \caption{
+  \caption{The effect of non-uniform sampling on the observed Fourier transform.
+    These samples have the same average spacing as those in \fig{comb_window_1},
+    but the lack of structure in observing window translates to a lack of
+    structure in its transform, causing the observed transform to be ``noisy''.
     \figlabel{random_window}}
 \end{figure}
+
+One might hope that sampling the signal more densely might limit these problems,
+and it does, but only to a degree.
+\Fig{random_window_2} increases the density of observations by a factor of 10,
+such that there are 200 total observations over the length-10 observing window.
+The observed Fourier transform in this case is much more reflective of the
+underlying signal, but still contains a degree of ``noise'' rooted in the
+unstructured nature of the spacing between observations.
 
@@
   \includegraphics[width=\textwidth]{fig11_random_window_2}
-  \caption{
+  \caption{The effect of non-uniform sampling on the observed Fourier transform,
+    with a factor of 10 more samples than \fig{random_window}.
+    Even with very dense sampling of the function, the Fourier transform
+    cannot be exactly recovered due to the 
+    These samples have the same average spacing as those in \fig{comb_window_1},
+    but the lack of structure in observing window translates to a lack of
+    structure in its transform, causing the observed transform to be a
+    ``noisy'' representation of the true transform
     \figlabel{random_window_2}}
@@
 
+Since the observed transform is a well-defined convolution of the (unknown) true
+transform and the (known) window transform, you might hope that we could
+perform some sort of deconvolution operation to recover the result, but this
+turns out not to be straightforward, as deconvolution (particularly in this
+case where delta functions are involved) is an ill-posed problem without a
+unique solution.
+
+\subsection{Non-Uniform Periodogram}
+
+\subsection{A Non-uniform Nyquist Limit?}
+
+NO!!!
+
+\todo{Reference LINEAR data avg frequency vs computed period}
+
+\subsection{Semi-structured Observing Windows}
+
 \begin{itemize}
-  \item nonuniform signal as sum-of-deltas window function
-  \item Nyquist limit is meaningless!
-  \item Some use notions of ``average Nyquist limit'', but this can be misleading (reference LINEAR data in \fig{LINEAR_data})
+  \item Nightly (e.g. ground-based imaging)
+  \item Uniform, but with missing data (e.g. Kepler)
 \end{itemize}
@@
 \begin{itemize}
-  \item Show relation to Shuster periodogram
-  \item Show relation to least squares fitting of sinusoid
-  \item Key: LS measures {\it windowed} power spectrum
-  \item Show effects of truely random sampling; of Kepler-like sampling; of ground-based nightly sampling
+  \item ... is like a Schuster periodogram.
+  \item ... is also like a least-squares sinusoid fit.
+  \item ... reminder that it is a {\it windowed} power spectrum
 \end{itemize}

### 2016-12-18 PracticalLombScargle.tex
rename paper
Add some citations to pseudo-Nyquist discussions

--- 2016-12-16 paper.tex
@@
 
-\title{Practical Considerations for Lomb-Scargle Periodic Analysis}
+\title{Understanding the Lomb-Scargle Periodogram}
 
@@
   are indicated by the gray errorbars on each point.
-  \figlabel{LINEAR_data}.
+  \figlabel{LINEAR_data}
 }
@@
   \includegraphics[width=\textwidth]{fig05_Convolution_Theorem}
-  \caption{Visualization of the Convolution theorem (\eq{convolution_theorem}).
+  \caption{Visualization of the convolution theorem (\eq{convolution_theorem}).
     The key is that the Fourier transform of
     a convolution is the pointwise product of the two Fourier transforms.
+    In the right panels, the black and gray lines represent the real and
+    imaginary part of the transform, respectively.
     \figlabel{convolution_theorem}}
@@
   P_S(f) = \frac{1}{N}\left|\sum_{n=1}^N g_n e^{-2\pi i f t_n}\right|^2
+  \eqlabel{schuster-periodogram}
 \end{equation}
@@
              &=& \sum_{n=1}^{N} g(t_n)\delta(t - t_n)
+  \eqlabel{g-nonuniform}
 \end{eqnarray}
@@
 
+From this non-uniform Fourier transform, it is straightforward to extend the
+definition of the Schuster periodogram to the non-uniform case.
+Plugging \eq{g-nonuniform} into the Fourier transform of \eq{FT-def}, in analogy to \eq{schuster-periodogram}, gives the following form of the power spectrum:
+\begin{equation}
+  P(f) = \frac{1}{N}\left|\sum_{n=1}^N g(t_n)e^{-2\pi i t_n} \right|^2
+\end{equation}
+Becuase it is derived from the Fourier transform of a non-uniformly sampled
+function, this periodogram is subject to all the caveats mentioned here: in
+particular, the periodogram measures the periodic content of the
+{\it windowed signal} and thus reflects a convolution of the true signal
+with a non-structured window function.
+The resulting periodogram, then, can only be used as a qualitative approximation
+of the true periodogram, and will only approach the true periodogram as the
+number of observed samples gets very large.
+
 \subsection{A Non-uniform Nyquist Limit?}
@@
 
-\todo{Reference LINEAR data avg frequency vs computed period}
+Examples: average \citep{NumRec, Horne86}, minimum \citep{Scargle82, Roberts87}, average of inverse time intervals \citep{Debosscher07}.
+
+time interval of measurement \citep{ICVG2014}, common multiple of observing time \citep{Eyer99}
+
+\todo{Reference Figure 1 LINEAR data avg frequency vs computed period}
+
+\citet{Scargle82}: ``...''
+
+\citet{NumRec}: ``...''
+
+\citet{Roberts87}: ``for arbitrary $\{t_r\}$,
+the sampling theorem tells us nothing.
+If the data samples are otherwise equally spaced but with missing
+points, the theorem says thatthe data completely determine
+a function whose FT is zero for $|v| > 1/(2\Delta_{max})$, where $\Delta_{max}$
+is the {\it largest completely sampled} data spacing. However,
+there are smaller spacings, and these certainly carry information
+about frequencies greater than $1/(2\Delta_{max} )$; some information
+is available about frequencies as high as
+$1/(2\Delta_{min} )$, where $\Delta_{min}$ is the smallest spacing between data
+points. Furthermore, if the $\{t_r\}$ are more or less randomly
+distributed, so that a wide range ofspacings are present and
+there is little redundancy in the spacing between various
+points, tests have shown (Paper II) that significant information
+is available on frequencies greater than $l/(2\Delta_{min})$.
+
+``Nonetheless, in the present paper we will restrict ourselves
+to frequencies obeying
+$\nu < \nu_{max} = 1/(2\Delta_{min})$.''
+
+\citet{Horne86}: ``The largest frequency we calculated was $\pi N-)/T$ which
+is the traditional Nyquist frequency for evenly-spaced data. The Nyquist
+frequency is not well defined for unevenly spaced signals, but it can serve as
+a reasonable upper limit for the calculation.''
+
+\citet{Debosscher07}: ``For the
+highest frequency, we used the average of the inverse time intervals
+between the measurements: $f_N = 0.5(1/\Delta T)$ as a pseudo
+Nyquist frequency. Note that $f_N$ is equal to the Nyquist frequency
+in the case of equidistant sampling. For particular cases,
+an even higher upper limit can be used \citep[see][]{Eyer99}.
+Our upper limit should be seen as a compromise between
+the required resolution to allow a good fitting, and computation
+time.''
 
@@
 \begin{itemize}
-\item Bayesian view doesn't make sense
+\item Bayesian view is (almost always) not useful
 \item Multiterm makes the true period fit better, but also bumps the background noise.
@@
 \bibliographystyle{apj}
-\bibliography{paper}
+\bibliography{PracticalLombScargle}
 

### 2016-12-19 PracticalLombScargle.tex
Add more pseudo nyquist quotes

@@
 
-Examples: average \citep{NumRec, Horne86}, minimum \citep{Scargle82, Roberts87}, average of inverse time intervals \citep{Debosscher07}.
-
+Examples: average \citep{NumRec, Horne86, Scargle82}, minimum \citep{Roberts87, Press89}, average of inverse time intervals \citep{Debosscher07},
 time interval of measurement \citep{ICVG2014}, common multiple of observing time \citep{Eyer99}
@@
 
-\citet{Scargle82}: ``...''
+\citet{Scargle82} (citing the following: Beutler 1966, 1970; Masry \& Lui 1975; Higgins 1976; Wiley 1978; Gaster \& Roberts 1975, 1977; Kar, Hornkohl \& Farmer 1981; Ludeman 1981):
+\begin{quote}
+Error-free recovery of a band-limited signal [i.e.~reproduction of the entire function $X(t)$ from the samples $X(t_i)$] can be achieved with irregular sampling as long as the mean sampling rate exceeds the Nyquist rate (i.e.
+the average number of samples per unit time must exceed twice the highest frequency component in the signal).
+\end{quote}
 
-\citet{NumRec}: ``...''
+\citet{NumRec}: 
+\begin{quote}
+One guide to choosing $f_{hi}$ is to compare it with the Nyquist frequency $f_c$ which would obtain if the $N$ data points were evenly spaced over the same span $T$, that is $f_c = N/(2T)$.
+The accompanying program includes an input parameter {\tt hifac}, defined as $f_{hi}/f_{c}$.
+\end{quote}
 
-\citet{Roberts87}: ``for arbitrary $\{t_r\}$,
-the sampling theorem tells us nothing.
+\citet{Press89}:
+\begin{quote}
+It is often meaningful to examine frequencies significantly higher than the
+Nyquist frequency that would obtain if the same number of data points were
+evenly spaced in the same total length of time. Some spectral information is
+obtainable for frequencies all the way up to something like half the inverse
+spacing of the {\it closest} spaced points.
+\end{quote}
+
+\citet{Roberts87}:
+\begin{quote}
+...for arbitrary $\{t_r\}$, the sampling theorem tells us nothing.
 If the data samples are otherwise equally spaced but with missing
-points, the theorem says thatthe data completely determine
+points, the theorem says that the data completely determine
 a function whose FT is zero for $|v| > 1/(2\Delta_{max})$, where $\Delta_{max}$
@@
 
-``Nonetheless, in the present paper we will restrict ourselves
+Nonetheless, in the present paper we will restrict ourselves
 to frequencies obeying
-$\nu < \nu_{max} = 1/(2\Delta_{min})$.''
+$\nu < \nu_{max} = 1/(2\Delta_{min})$.
+\end{quote}
 
-\citet{Horne86}: ``The largest frequency we calculated was $\pi N-)/T$ which
+\citet{Horne86}:
+\begin{quote}
+The largest frequency we calculated was $\pi N-)/T$ which
 is the traditional Nyquist frequency for evenly-spaced data. The Nyquist
 frequency is not well defined for unevenly spaced signals, but it can serve as
-a reasonable upper limit for the calculation.''
+a reasonable upper limit for the calculation.
+\end{quote}
 
-\citet{Debosscher07}: ``For the
-highest frequency, we used the average of the inverse time intervals
+\citet{Debosscher07}:
+\begin{quote}
+For the highest frequency, we used the average of the inverse time intervals
 between the measurements: $f_N = 0.5(1/\Delta T)$ as a pseudo
@@
 the required resolution to allow a good fitting, and computation
-time.''
+time.
+\end{quote}
+
+\todo{Demonstrate recovery of a signal well beyond these limits (Is LINEAR sufficient?}
 

### 2016-12-20 PracticalLombScargle.tex
update pseudo-nyquist references

@@
 
-NO!!!
+From \citet{Eyer99}: let $p$ be the largest value such that each $t_i$ can be written $t_i = t_0 + n_i p$, for integers $n_i$. The Nyquist Frequency then is:
+\begin{equation}
+  \nu_{Ny} = 1 / (2p)
+\end{equation}
 
-Examples: average \citep{NumRec, Horne86, Scargle82}, minimum \citep{Roberts87, Press89}, average of inverse time intervals \citep{Debosscher07},
+Examples of wrong approaches: average \citep{NumRec, Horne86, Scargle82}, minimum \citep{Roberts87, Press89}, average of inverse time intervals \citep{Debosscher07},
 time interval of measurement \citep{ICVG2014}, common multiple of observing time \citep{Eyer99}
@@
 the average number of samples per unit time must exceed twice the highest frequency component in the signal).
+\end{quote}
+
+\citet{Percy86}:
+\begin{quote}
+... the Nyquist frequency is not well defined for input data obtained at unequally spaced time intervals, the usual case in astronomy \citep{Scargle82}.
+Theoretically such a data set contains informatin on the periodicities down to
+$\Delta t = \min(t_i - t{i-1})$.
+In practice, however, a pseudo-Nyquist frequency may be defined by averaging $\Delta t = (t_i - t_{i-1})$, where large, uncharacteristic temporal gaps are avoided.
+Alternatively, the harmonic mean of all $\Delta t$ may be used.
+The result is that a useful pseudo-Nyquist frequency may be defined by $\nu_{Ny} = (2 \langle\Delta t \rangle)^{-1}$.
 \end{quote}
@@
 
+\citet{ICVG2014}:
+\begin{quote}
+As a good choice for the maximum search frequency, a pseudo-Nyquist frequency $\omega_{max} = \pi/\Delta t$, where $1/\Delta t$ is the median of the inverse time interval between data points, was proposed by \citet{Debosscher07} (in the case of even sampling, $\omega_{max}$ is equal to the Nyquist frequency).
+In practice, this choice may be a gross underestimate because unevenly sampled data can detect periodicity with frequencies even higher than $2\pi / \Delta t_{min}$ \citep[see][]{Eyer99}.
+An appropriate choice of $\omega_{max}$ thus depends on sampling (the phase coverage at a given frequency is the relevant quantity) and needs to be carefully chosen: a hard limit on maximum detectable frequency is of course given by the time interval over which individual measurements are performed, such as imaging exposure time.
+\end{quote}
+
+\citet{Hilditch01}:
+\begin{quote}
+The highest frequency, for equally spaced data, is formally given by the {\it Nyquist frequency}, $f_N = 1/(2\Delta t)$, where $\Delta t$ is the {\it sampling interval} in the data string.
+For unequally spaced data, it is common practice to estimate a pseudo-Nyquist frequency from the minimum value of $\Delta t$.
+\end{quote}
+
 \todo{Demonstrate recovery of a signal well beyond these limits (Is LINEAR sufficient?}

### 2016-12-20 PracticalLombScargle.tex
More material for pseudo-nyquist section

@@
 \subsection{A Non-uniform Nyquist Limit?}
+\sectlabel{pseudo-nyquist}
 
@@
 
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig12_pseudo_nyquist}
+  \caption{An example of data for which the various poorly-motivated
+    ``pseudo-Nyquist'' approaches outlined in \sect{pseudo-nyquist} fail
+    spectacularly. The upper panels show the data, a noisy sinusoid with
+    a frequency of 100 (i.e. a period of 0.01).
+    The lower left panel shows a histogram of spacings between observations:
+    the minimum spacing is 2.5, meaning that the signal has
+    {\it over 250 full cycles} between the minimum spacing.
+    Nevertheless, the periodogram (lower right) clearly identifies the correct
+    period, though it's orders of magnitude larger than Nyquist estimates based
+    on average or minimum sampling rate.
+    \figlabel{pseudo_nyquist}}
+\end{figure}
+
 \todo{Reference Figure 1 LINEAR data avg frequency vs computed period}
-
-\citet{Scargle82} (citing the following: Beutler 1966, 1970; Masry \& Lui 1975; Higgins 1976; Wiley 1978; Gaster \& Roberts 1975, 1977; Kar, Hornkohl \& Farmer 1981; Ludeman 1981):
-\begin{quote}
-Error-free recovery of a band-limited signal [i.e.~reproduction of the entire function $X(t)$ from the samples $X(t_i)$] can be achieved with irregular sampling as long as the mean sampling rate exceeds the Nyquist rate (i.e.
-the average number of samples per unit time must exceed twice the highest frequency component in the signal).
-\end{quote}
-
-\citet{Percy86}:
-\begin{quote}
-... the Nyquist frequency is not well defined for input data obtained at unequally spaced time intervals, the usual case in astronomy \citep{Scargle82}.
-Theoretically such a data set contains informatin on the periodicities down to
-$\Delta t = \min(t_i - t{i-1})$.
-In practice, however, a pseudo-Nyquist frequency may be defined by averaging $\Delta t = (t_i - t_{i-1})$, where large, uncharacteristic temporal gaps are avoided.
-Alternatively, the harmonic mean of all $\Delta t$ may be used.
-The result is that a useful pseudo-Nyquist frequency may be defined by $\nu_{Ny} = (2 \langle\Delta t \rangle)^{-1}$.
-\end{quote}
-
-\citet{NumRec}: 
-\begin{quote}
-One guide to choosing $f_{hi}$ is to compare it with the Nyquist frequency $f_c$ which would obtain if the $N$ data points were evenly spaced over the same span $T$, that is $f_c = N/(2T)$.
-The accompanying program includes an input parameter {\tt hifac}, defined as $f_{hi}/f_{c}$.
-\end{quote}
-
-\citet{Press89}:
-\begin{quote}
-It is often meaningful to examine frequencies significantly higher than the
-Nyquist frequency that would obtain if the same number of data points were
-evenly spaced in the same total length of time. Some spectral information is
-obtainable for frequencies all the way up to something like half the inverse
-spacing of the {\it closest} spaced points.
-\end{quote}
-
-\citet{Roberts87}:
-\begin{quote}
-...for arbitrary $\{t_r\}$, the sampling theorem tells us nothing.
-If the data samples are otherwise equally spaced but with missing
-points, the theorem says that the data completely determine
-a function whose FT is zero for $|v| > 1/(2\Delta_{max})$, where $\Delta_{max}$
-is the {\it largest completely sampled} data spacing. However,
-there are smaller spacings, and these certainly carry information
-about frequencies greater than $1/(2\Delta_{max} )$; some information
-is available about frequencies as high as
-$1/(2\Delta_{min} )$, where $\Delta_{min}$ is the smallest spacing between data
-points. Furthermore, if the $\{t_r\}$ are more or less randomly
-distributed, so that a wide range ofspacings are present and
-there is little redundancy in the spacing between various
-points, tests have shown (Paper II) that significant information
-is available on frequencies greater than $l/(2\Delta_{min})$.
-
-Nonetheless, in the present paper we will restrict ourselves
-to frequencies obeying
-$\nu < \nu_{max} = 1/(2\Delta_{min})$.
-\end{quote}
-
-\citet{Horne86}:
-\begin{quote}
-The largest frequency we calculated was $\pi N-)/T$ which
-is the traditional Nyquist frequency for evenly-spaced data. The Nyquist
-frequency is not well defined for unevenly spaced signals, but it can serve as
-a reasonable upper limit for the calculation.
-\end{quote}
-
-\citet{Debosscher07}:
-\begin{quote}
-For the highest frequency, we used the average of the inverse time intervals
-between the measurements: $f_N = 0.5(1/\Delta T)$ as a pseudo
-Nyquist frequency. Note that $f_N$ is equal to the Nyquist frequency
-in the case of equidistant sampling. For particular cases,
-an even higher upper limit can be used \citep[see][]{Eyer99}.
-Our upper limit should be seen as a compromise between
-the required resolution to allow a good fitting, and computation
-time.
-\end{quote}
-
-\citet{ICVG2014}:
-\begin{quote}
-As a good choice for the maximum search frequency, a pseudo-Nyquist frequency $\omega_{max} = \pi/\Delta t$, where $1/\Delta t$ is the median of the inverse time interval between data points, was proposed by \citet{Debosscher07} (in the case of even sampling, $\omega_{max}$ is equal to the Nyquist frequency).
-In practice, this choice may be a gross underestimate because unevenly sampled data can detect periodicity with frequencies even higher than $2\pi / \Delta t_{min}$ \citep[see][]{Eyer99}.
-An appropriate choice of $\omega_{max}$ thus depends on sampling (the phase coverage at a given frequency is the relevant quantity) and needs to be carefully chosen: a hard limit on maximum detectable frequency is of course given by the time interval over which individual measurements are performed, such as imaging exposure time.
-\end{quote}
-
-\citet{Hilditch01}:
-\begin{quote}
-The highest frequency, for equally spaced data, is formally given by the {\it Nyquist frequency}, $f_N = 1/(2\Delta t)$, where $\Delta t$ is the {\it sampling interval} in the data string.
-For unequally spaced data, it is common practice to estimate a pseudo-Nyquist frequency from the minimum value of $\Delta t$.
-\end{quote}
 
@@
 
+\appendix
+
+\section{Non-uniform Nyquist}
+A selection of quotations showing approaches to a non-uniform Nyquist limit:
+
+\begin{description}
+
+\item[Average Sampling Rate]
+
+\citet{Scargle82} (citing the following: Beutler 1966, 1970; Masry \& Lui 1975; Higgins 1976; Wiley 1978; Gaster \& Roberts 1975, 1977; Kar, Hornkohl \& Farmer 1981; Ludeman 1981):
+\begin{quote}
+Error-free recovery of a band-limited signal [i.e.~reproduction of the entire function $X(t)$ from the samples $X(t_i)$] can be achieved with irregular sampling as long as the mean sampling rate exceeds the Nyquist rate (i.e.
+the average number of samples per unit time must exceed twice the highest frequency component in the signal).
+\end{quote}
+
+\citet{NumRec}: 
+\begin{quote}
+One guide to choosing $f_{hi}$ is to compare it with the Nyquist frequency $f_c$ which would obtain if the $N$ data points were evenly spaced over the same span $T$, that is $f_c = N/(2T)$.
+The accompanying program includes an input parameter {\tt hifac}, defined as $f_{hi}/f_{c}$.
+\end{quote}
+
+\citet{Horne86}:
+\begin{quote}
+The largest frequency we calculated was $\pi N/T$ which
+is the traditional Nyquist frequency for evenly-spaced data. The Nyquist
+frequency is not well defined for unevenly spaced signals, but it can serve as
+a reasonable upper limit for the calculation.
+\end{quote}
+
+\item[Harmonic Average of Sampling Rate]
+
+\citet{Debosscher07}:
+\begin{quote}
+For the highest frequency, we used the average of the inverse time intervals
+between the measurements: $f_N = 0.5(1/\Delta T)$ as a pseudo
+Nyquist frequency. Note that $f_N$ is equal to the Nyquist frequency
+in the case of equidistant sampling. For particular cases,
+an even higher upper limit can be used \citep[see][]{Eyer99}.
+Our upper limit should be seen as a compromise between
+the required resolution to allow a good fitting, and computation
+time.
+\end{quote}
+
+\item[Minimum Sample Spacing]
+
+\citet{Percy86}:
+\begin{quote}
+... the Nyquist frequency is not well defined for input data obtained at unequally spaced time intervals, the usual case in astronomy \citep{Scargle82}.
+Theoretically such a data set contains informatin on the periodicities down to
+$\Delta t = \min(t_i - t{i-1})$.
+In practice, however, a pseudo-Nyquist frequency may be defined by averaging $\Delta t = (t_i - t_{i-1})$, where large, uncharacteristic temporal gaps are avoided.
+Alternatively, the harmonic mean of all $\Delta t$ may be used.
+The result is that a useful pseudo-Nyquist frequency may be defined by $\nu_{Ny} = (2 \langle\Delta t \rangle)^{-1}$.
+\end{quote}
+
+\citet{Press89}:
+\begin{quote}
+It is often meaningful to examine frequencies significantly higher than the
+Nyquist frequency that would obtain if the same number of data points were
+evenly spaced in the same total length of time. Some spectral information is
+obtainable for frequencies all the way up to something like half the inverse
+spacing of the {\it closest} spaced points.
+\end{quote}
+
+\citet{Roberts87}:
+\begin{quote}
+...for arbitrary $\{t_r\}$, the sampling theorem tells us nothing.
+If the data samples are otherwise equally spaced but with missing
+points, the theorem says that the data completely determine
+a function whose FT is zero for $|v| > 1/(2\Delta_{max})$, where $\Delta_{max}$
+is the {\it largest completely sampled} data spacing. However,
+there are smaller spacings, and these certainly carry information
+about frequencies greater than $1/(2\Delta_{max} )$; some information
+is available about frequencies as high as
+$1/(2\Delta_{min} )$, where $\Delta_{min}$ is the smallest spacing between data
+points. Furthermore, if the $\{t_r\}$ are more or less randomly
+distributed, so that a wide range ofspacings are present and
+there is little redundancy in the spacing between various
+points, tests have shown (Paper II) that significant information
+is available on frequencies greater than $l/(2\Delta_{min})$.
+
+Nonetheless, in the present paper we will restrict ourselves
+to frequencies obeying
+$\nu < \nu_{max} = 1/(2\Delta_{min})$.
+\end{quote}
+
+\citet{Hilditch01}:
+\begin{quote}
+The highest frequency, for equally spaced data, is formally given by the {\it Nyquist frequency}, $f_N = 1/(2\Delta t)$, where $\Delta t$ is the {\it sampling interval} in the data string.
+For unequally spaced data, it is common practice to estimate a pseudo-Nyquist frequency from the minimum value of $\Delta t$.
+\end{quote}
+
+\item[Getting it Right]
+
+\citet{ICVG2014}:
+\begin{quote}
+As a good choice for the maximum search frequency, a pseudo-Nyquist frequency $\omega_{max} = \pi/\Delta t$, where $1/\Delta t$ is the median of the inverse time interval between data points, was proposed by \citet{Debosscher07} (in the case of even sampling, $\omega_{max}$ is equal to the Nyquist frequency).
+In practice, this choice may be a gross underestimate because unevenly sampled data can detect periodicity with frequencies even higher than $2\pi / \Delta t_{min}$ \citep[see][]{Eyer99}.
+An appropriate choice of $\omega_{max}$ thus depends on sampling (the phase coverage at a given frequency is the relevant quantity) and needs to be carefully chosen: a hard limit on maximum detectable frequency is of course given by the time interval over which individual measurements are performed, such as imaging exposure time.
+\end{quote}
+
+\citet{Eyer99}
+
+\end{description}
+
 \end{document}

### 2016-12-23 PracticalLombScargle.tex
Finish pseudo-nyquist discussion

@@
 \subsubsection{The Nyquist Limit}
+\sectlabel{nyquist}
 
@@
 
-From \citet{Eyer99}: let $p$ be the largest value such that each $t_i$ can be written $t_i = t_0 + n_i p$, for integers $n_i$. The Nyquist Frequency then is:
-\begin{equation}
-  \nu_{Ny} = 1 / (2p)
-\end{equation}
+As we have seen in \sect{nyquist}, the Nyquist limit is a direct consequence
+of the window function created by evenly-sampled data,
+and uneven sampling destroys the symmetry that underlies its definition.
+Nevertheless, the idea of the ``Nyquist frequency'' seems to have taken hold
+in the scientific psyche to the extent that it's often mis-applied in areas
+where it is utterly irrelevant.
+For unevenly-sampled data, the ``Nyquist limit'' might or might not exist,
+and even in cases where it does exist it tends to be far less relevant than
+in the evenly-sampled case.
 
-Examples of wrong approaches: average \citep{NumRec, Horne86, Scargle82}, minimum \citep{Roberts87, Press89}, average of inverse time intervals \citep{Debosscher07},
-time interval of measurement \citep{ICVG2014}, common multiple of observing time \citep{Eyer99}
+\subsubsection{Incorrect Limits in the Literature}
+
+In the scientific literature, there are countless examples of attempts to
+define a ``pseudo-Nyquist'' frequency to apply in the uneven sampling case.
+A few common approaches include the mean of the sampling intervals
+\citep[e.g.][]{Scargle82, Horne86, NumRec},
+the harmonic mean of the sampling intervals \citep[e.g.][]{Debosscher07},
+or the minimum sample spacing \citep[e.g.][]{Percy86, Roberts87, Press89, Hilditch01}.
+All of these are tempting criteria in that they recover the classical Nyquist 
+frequency in the limit of evenly-spaced data, but this does not make up for the
+fact that they are wrong.
+Several of these cases do parenthetically note that signals at frequencies
+beyond the proposed limit are observable in principle, but they nevertheless
+proceed under the apparent impression that such pseudo-Nyquist reasoning
+is relevant.
+
+As a simple example of where such logic fails,
+consider the data from \fig{LINEAR_data}: though the mean sample
+spacing is one observation every seven days, in \fig{LINEAR_power}
+we were nevertheless able to quite clearly identify a period of 2.58 hours---
+an order of magnitude shorter than the mean-based pseudo-Nyquist limit
+would indicate as possible.
+For the data in \fig{LINEAR_data}, the minimum sample spacing is just under 10
+seconds, but it it irresponsible to pretend that this single pair of
+observations itself defines some limit beyond which frequency information is
+unattainable.
 
@@
     The lower left panel shows a histogram of spacings between observations:
-    the minimum spacing is 2.5, meaning that the signal has
-    {\it over 250 full cycles} between the minimum spacing.
+    the minimum spacing is 2.55, meaning that the signal has
+    {\it over 250 full cycles} between the closest pair of observations.
     Nevertheless, the periodogram (lower right) clearly identifies the correct
-    period, though it's orders of magnitude larger than Nyquist estimates based
-    on average or minimum sampling rate.
+    period, though it's orders of magnitude larger than pseudo-Nyquist
+    estimates based on average or minimum sampling rate.
     \figlabel{pseudo_nyquist}}
@@
 
-\todo{Reference Figure 1 LINEAR data avg frequency vs computed period}
+As a more extreme example, consider the data shown in \fig{pseudo_nyquist}.
+This consists of noisy samples from a sinusoid with a period of 0.01 units,
+with sample spacings ranging between 2 and 18 units: needless to say, any
+pseudo-Nyquist definition based on an average or minimum sample spacings
+will be {\it far} smaller than the true frequency of 100; still, the
+the Lomb-Scargle periodogram in the lower right panel shows quite cleanly
+the true frequency.
 
-\todo{Demonstrate recovery of a signal well beyond these limits (Is LINEAR sufficient?}
+\subsubsection{The Correct Non-uniform Nyquist Limit}
+
+While Nyquist arguments based on average or minimum sampling fail spectacularly,
+there is a sense in which the Nyquist limit can be applied to unevenly-spaced
+data. \citet{Eyer99} explores this issue in some detail, and in particular
+proves the following:
+\begin{quote}
+Let $p$ be the largest value such that each $t_i$ can be written $t_i = t_0 + n_i p$, for integers $n_i$. The Nyquist Frequency then is $f_{Ny} = 1 / (2p)$.
+\end{quote}
+In other words, computing the Nyquist rate for unevenly-spaced data requires
+finding some common factor, such that each spacing $\Delta t_i$ is {\it exactly}
+an integer multiple of this factor.
+This can be proven formally, but can be understood by thinking of such data as
+a windowed version of uniformly sampled data with spacing $p$.
+Such uniform data has a classical Nyquist limit of $1/(2p)$, and a window function applied to that data cannot change that fact.
+
+\Fig{nyquist_eyer99} shows an example of such a Nyquist frequency.
+The data are non-uniformly sampled at times $t_i = n_i * p$, with $p=0.01$.
+This results in a Nyquist frequency $f_{Ny} = 50$, and we see the expected
+behavior beyond this frequency: the signal at $f > f_{Ny}$ consists of repeated
+aliases of the signal at $f < f_{Ny}$.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig13_nyquist_eyer99}
+  \caption{A visualization of the \citep{Eyer99} definition of the Nyquist
+    frequency. Data are non-uniformly sampled at times $t_i = n_i p$, for
+    integer $n_i$ and $p=0.01$.
+    This results in a Nyquist frequency of $f_{Ny}= (2p)^{-1} = 50$:
+    the periodogram outside the range $0 \le f < f_{Ny}$ is a series of
+    repetitions of the signal within that range.
+    \figlabel{nyquist_eyer99}}
+\end{figure}
+
+We should keep in mind one consequence of this Nyquist definition:
+if you have any pair of observation spacings
+whose ratio is irrational, {\it the Nyquist limit does not exist!}.
+To realize this situation in practice, however, 
+requires infinitely precise measurements of the
+times $t_i$; more often in the real-world, the Nyquist frequency will be
+defined by the precision of the time measurements.
+That is, if times are recorded to $D$ decimal places, the Nyquist frequency
+will be (at most) $f_{Ny} = \frac{1}{2} 10^D$, as in \fig{nyquist_eyer99}.
+In other words, the Nyquist frequency for irregularly-sampled data is most
+typically defined by {\it the precision of the time measurements}.
+
+\subsubsection{Limit due to Windowing}
+
+In contexts where observations are not instantaneous, but rather consist of
+short-duration integrations of a continuous signal, a qualitatively different
+kind of frequency limit exists.
+This is typical in, e.g., optical astronomy, where observations take place over
+a finite duration $\delta t$.
+As noted by \citet{ICVG2014}, this time-scale of integration represents another
+kind of limiting frequency for irregularly-sampled data.
+Such a situation means that the measurement is effectively a convolution of
+the underlying signal with a rectangular window function of width $\delta t$,
+in a manner analogous to \fig{convolution_theorem}.
+By the convolution theorem, the observed transform will be a point-wise
+product between the true transform and the transform of the window, which
+will generally have a width proportional to $1/\delta t$.
+This means that---absent other more constraining window effects---the
+frequency limit is $f_{max} \propto 1/(2\delta t)$, with the constant of
+proportionality dependent on the shape of the effective window describing
+individual observations.
+
+Keep in mind that the windowing limit $f_{max}$ is quite different than a
+Nyquist limit: the Nyquist limit is the frequency beyond which all signal
+is aliased into the Nyquist range; the windowing limit is the frequency
+beyond which all signal is zero.
 
@@
 
-\citet{Scargle82} (citing the following: Beutler 1966, 1970; Masry \& Lui 1975; Higgins 1976; Wiley 1978; Gaster \& Roberts 1975, 1977; Kar, Hornkohl \& Farmer 1981; Ludeman 1981):
+\citet{Scargle82} (citing the following: Beutler 1966, 1970; Masry \& Lui 1975; Higgins 1976; Wiley 1978; Gaster \& Roberts 1975, 1977; Kar, Hornkohl \& Farmer 1981; Ludeman 1981 \todo{read and summarize these!}):
 \begin{quote}
@@
 
-\citet{Eyer99}
-
 \end{description}

### 2016-12-24 PracticalLombScargle.tex
Add initial discussion of structured windows

@@
 
+We have seen that for uniform data, the perfect aliasing beyond the Nyquist
+frequency is a direct consequence of the symmetry of the Dirac-comb window
+function.
+For non-uniform observations, such symmetry does not exist, but {\it structure}
+in the observing window can lead to partial aliasing of signals in the
+data.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig14_LINEAR_window}
+  \caption{The power spectrum of the observing window for the data shown
+    in \fig{LINEAR_data}. Notice the strong spike in power at a period of
+    1 day, and related aliases at $1/n$ days for integer $n$.
+    There is also a strong spike at $365$ days, and a noticeable spike at
+    $\sim 14$ days. Each of these indicate time intervals that appear often
+    in the data.
+    \figlabel{LINEAR_window}}
+\end{figure}
+
+Let's again consider the data shown in \fig{LINEAR_data}. The window power
+spectrum for this in \fig{LINEAR_window} shows some quite distinct features,
+and these features have an intuitive interpretation.
+Namely, if the window power shows a spike at a period of $p$ days, this means
+that an observation at time $t_0$ is likely to be followed by another
+observation near a time $t_0 + np$ for integer $n$.
+
+Thus the strong spike at a period of 1 day indicate that observations are
+taken near the same time of day: this is typical of a ground-based survey
+which observes only at night.
+The additional spikes at periods of $1/n$ days (for integer $n$) are aliases
+of this same effect.
+
+This window also shows a strong spike at a period of 1 year, indicating that
+there was a strong seasonal pattern in the observations.
+Finally, there is a noticeable spike at approximately 14 days that is likely
+related to patterns in scheduling of observing runs.
+
+Recall that a Fourier spectrum observed through a particular window will reflect
+a convolution of the true spectrum and the window spectrum
+(cf.\ \fig{random_window}), and so we would expect this structure to be
+imprinted on the power spectrum measured from the data.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig15_LINEAR_window_effect}
+  \caption{The effect of the window function in \fig{LINEAR_window} on the
+    power spectrum in \fig{LINEAR_power}.
+    The top panel shows the window power spectrum, and the bottom panel shows
+    the observed signal power spectrum.
+    Both are plotted as a function of frequency (we earlier saw both of these
+    as a function of period; see \fig{LINEAR_window} and \fig{LINEAR_power},
+    respectively).
+    Viewing these as a function of frequency makes it clear that the structure
+    in the window function is imprinted on the observed spectrum: both the
+    diurnal structure in the main panel, and the annual structure in the inset
+    are apparent in the observed spectrum.
+    \figlabel{LINEAR_window_effect}}
+\end{figure}
+
+This behavior is illustrated in \fig{LINEAR_window_effect}.
+The upper panel shows the window power spectrum as a function of frequency, while lower panel shows the observed signal power spectrum as a function of frequency.
+As noted in the discussion of \fig{LINEAR_window}, the window function shows a strong signal at frequencies corresponding to 1 day and 1 year; the latter can be seen in the inset plot.
+The power at both these scales is clearly imprinted on the observed power spectrum, as shown in the lower panel and its inset.
+This imperfect aliasing is similar to the perfect aliasing seen in regularly-sampled data at the Nyquist frequency; however, in this case the aliasing is
+imperfect and the effect fades as you move further from the frequency driving
+the signal.
+
 \begin{itemize}
-  \item Nightly (e.g. ground-based imaging)
   \item Uniform, but with missing data (e.g. Kepler)
@@
   \item Motivation of Bootstrap
-
 \end{itemize}

### 2016-12-25 PracticalLombScargle.tex
Add Kepler window example

@@
 
+\subsubsection{A Ground-based Observing Window: LINEAR}
+
 Let's again consider the data shown in \fig{LINEAR_data}. The window power
@@
 The power at both these scales is clearly imprinted on the observed power spectrum, as shown in the lower panel and its inset.
-This imperfect aliasing is similar to the perfect aliasing seen in regularly-sampled data at the Nyquist frequency; however, in this case the aliasing is
-imperfect and the effect fades as you move further from the frequency driving
-the signal.
+This imperfect aliasing is similar to the perfect aliasing seen in regularly-sampled data at the Nyquist frequency; however, in this case the magnitude of the
+aliased signal fades further from the frequency driving the signal.
 
-\begin{itemize}
-  \item Uniform, but with missing data (e.g. Kepler)
-\end{itemize}
+\subsubsection{A Space-based Observing Window: Kepler}
 
-\section{The Lomb-Scargle Periodogram}
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig16_kepler_data}
+  \caption{An RR Lyrae variable observed by the Kepler project \citep[see][]{Kolenberg2010}.
+    {\it upper left:} the 4083 observed fluxes over roughly three months.
+    {\it upper right:} the window power spectrum, which is quite close to that of regularly-spaced data with a cadence of 29.4 minutes.
+    {\it lower left:} the time between observations.
+other than missing measurements, the spacing between observations is nearly uniform.
+the lower panel gives a closer look at the majority of the spacings, which are not exactly the same but rather span a range of $\pm 50$ms around the mean.
+    {\it lower right:} the data power spectrum, showing approximate aliasing and clearly indicating a peak near a period of 13.6 hours, along with higher-order components at multiples of this value.
+    \figlabel{kepler-data}}
+\end{figure}
+
+Space-based missions often have quite different observing windows.
+For example, \fig{kepler-data} shows observations of an RR-Lyrae variable star
+from the Kepler survey \citep[For discussion of this object, see][]{Kolenberg2010}.
+
+blah blah blah...
+
+\section{From Schuster to Lomb-Scargle}
 

### 2016-12-26 PracticalLombScargle.tex
Finish Kepler discussion

@@
 \newcommand{\Fig}[1]{Figure~\ref{fig:#1}}
-\newcommand{\fig}[1]{\Fig{#1}}
+\newcommand{\fig}[1]{Figure~\ref{fig:#1}}
+\newcommand{\figs}[2]{Figures~\ref{fig:#1}-\ref{fig:#2}}
 \newcommand{\figlabel}[1]{\label{fig:#1}}
@@
     detected 2.58-hour period to show the coherent periodic variability.
+    For more discussion of this data, see \citep{LINEAR3}.
     \figlabel{LINEAR_power}
@@
 This is an irregularly-sampled timeseries showing a single object from the
-LINEAR survey \citep{LINEAR1, LINEAR3}, with magnitude measured 280 times over the
-course of five and a half years\footnote{
+LINEAR survey \citep{LINEAR1, LINEAR3}, with magnitude measured 280 times over
+the course of five and a half years\footnote{
   Python code to reproduce \fig{LINEAR_data}, as well as all other figures
@@
 Space-based missions often have quite different observing windows.
-For example, \fig{kepler-data} shows observations of an RR-Lyrae variable star
-from the Kepler survey \citep[For discussion of this object, see][]{Kolenberg2010}.
-
-blah blah blah...
+For example, the upper-left panel of \fig{kepler-data}
+shows observations of an RR-Lyrae variable star from the Kepler survey,
+measured over 4000 times over a period of three months,
+with an irregular observing cadence of just under 30 minutes
+ \citep[For deeper discussion of these observations, see][]{Kolenberg2010}.
+The Kepler observations are very nearly uniformly-spaced, and this is reflected
+in the window power spectrum, shown in the upper-right panel of
+\fig{kepler-data}.
+The window function is a series of very narrow evenly-spaced spikes,
+reminiscent of the Dirac comb shown in \figs{comb_window_1}{comb_window_2},
+which we used to motivate the idea of the Nyquist limit.
+Thus, in this case we can treat $f_{Ny} = 0.5/29.4$ minutes$^{-1}$ as the
+effective Nyquist limit for the data, though aliasing is imperfect due to
+the uneven spacing of samples, shown in the lower-left panels of
+\fig{kepler-data}.
+
+The lower-right panel of \fig{kepler-data} shows the power spectrum of the
+observations, with gray shading indicating the ``Aliased'' region of the
+spectrum.
+The period of 13.6 hours is strongly apparent, along with smaller spikes at
+integer multiples of this frequency that indicate higher-order periodic
+components in the signal.
+
+The window functions for ground-based and space-based observations, reflected
+by LINEAR data in \fig{LINEAR_window_effect} and Kepler data in
+\fig{kepler-data}, are quite different, but in both cases essential features
+of the observed power spectra can be understood by recognizing it reflects
+not the power spectrum of the underlying signal, but a power specrtrum from
+the convolution of the true signal transform and the Fourier transform of
+the window function.
 

### 2016-12-26 PracticalLombScargle.tex
Start Schuster->Lomb Scargle discussion

@@
   are indicated by the gray errorbars on each point.
-  \figlabel{LINEAR_data}
+  \figlabel{LINEAR-data}
 }
@@
 \caption{{\it Left panel:} the Lomb-Scargle periodogram computed from the data
-    shown in \fig{LINEAR_data}, with an inset detailing the region near the peak.
-    {\it Right panel:} the input data in \fig{LINEAR_data}, folded over the
+    shown in \fig{LINEAR-data}, with an inset detailing the region near the peak.
+    {\it Right panel:} the input data in \fig{LINEAR-data}, folded over the
     detected 2.58-hour period to show the coherent periodic variability.
     For more discussion of this data, see \citep{LINEAR3}.
-    \figlabel{LINEAR_power}
+    \figlabel{LINEAR-power}
 }
@@
 in unevenly-sampled time-series, particularly within the astronomy community.
-For example, consider the data shown in \fig{LINEAR_data}.
+For example, consider the data shown in \fig{LINEAR-data}.
 This is an irregularly-sampled timeseries showing a single object from the
@@
 the course of five and a half years\footnote{
-  Python code to reproduce \fig{LINEAR_data}, as well as all other figures
+  Python code to reproduce \fig{LINEAR-data}, as well as all other figures
   in this manuscript, is available at 
@@
 
-The left panel of \fig{LINEAR_power} shows the Lomb-Scargle periodogram computed for the data in \fig{LINEAR_data}.
+The left panel of \fig{LINEAR-power} shows the Lomb-Scargle periodogram computed for the data in \fig{LINEAR-data}.
 Computing the Lomb-Scargle periodogram for the data gives us a measure of the
-power as a function of period of oscillation (\fig{LINEAR_power}, left), from
+power as a function of period of oscillation (\fig{LINEAR-power}, left), from
 which we can determine the period of oscillation of approximately 2.58 hours.
-The right panel of \fig{LINEAR_power} shows a folded visualization of
-the same data as \fig{LINEAR_data} -- i.e.{} plotted as a function of phase
+The right panel of \fig{LINEAR-power} shows a folded visualization of
+the same data as \fig{LINEAR-data} -- i.e.{} plotted as a function of phase
 rather than time.\footnote{
-    The periodogram in \fig{LINEAR_power} was computed using
+    The periodogram in \fig{LINEAR-power} was computed using
     {\tt astropy.stats.LombScargle}
@@
 In practice, however, there are a number of subtle issues that must be considered when applying a Lomb-Scargle analysis to real-world datasets, and I have found that these are rarely presented together at an introductory level.
-Here are a few questions in particular that we might wish to ask about the results in \fig{LINEAR_power}:
+Here are a few questions in particular that we might wish to ask about the results in \fig{LINEAR-power}:
 \begin{enumerate}
@@
        \mathcal{F}\{e^{2\pi f_0 t}\} = \delta(f - f_0).
-       \eqlabel{delta_FT}
+       \eqlabel{delta-FT}
      \end{equation}
@@
        e^{2\pi f_0 t} = \cos(2\pi f_0 t) + i\sin(2\pi f_0 t).
-       \eqlabel{Euler_formula}
+       \eqlabel{Euler-formula}
      \end{equation}
@@
 \includegraphics[width=\textwidth]{fig03_Fourier_pairs}
-\caption{Visualization of important Fourier pairs.\figlabel{fourier_pairs}}
+\caption{Visualization of important Fourier pairs.\figlabel{fourier-pairs}}
 \end{figure}
@@
 This is just one of many ``Fourier Pairs'' to keep in mind as we progress to understanding the Lomb-Scargle Periodogram.
-We list a few more important pairs here (see \Fig{fourier_pairs} for a visual representation of the following pairs):
+We list a few more important pairs here (see \Fig{fourier-pairs} for a visual representation of the following pairs):
 
@@
   \item[The Fourier transform of a sinusoid is a pair of Delta functions.]
-    (See \fig{fourier_pairs}{\it a})
+    (See \fig{fourier-pairs}{\it a})
     \begin{equation}
@@
    \item[The Fourier transform of a Gaussian is a Gaussian.]
-    (See \fig{fourier_pairs}{\it b})
+    (See \fig{fourier-pairs}{\it b})
      \begin{equation}
@@
   \item[The Fourier transform of a tophat function is a sinc function.]
-    (See \fig{fourier_pairs}{\it c})
+    (See \fig{fourier-pairs}{\it c})
     \begin{equation}
@@
   \item[The Fourier transform of a Dirac comb is a Dirac comb.]
-    (See \fig{fourier_pairs}{\it d})
+    (See \fig{fourier-pairs}{\it d})
     \begin{equation}
@@
       \III_T(t) \equiv \sum_{n=-\infty}^\infty \delta(t - nT)
-      \eqlabel{dirac_comb}
+      \eqlabel{dirac-comb}
     \end{equation}
@@
   [f \ast g](t) = \int_{-\infty}^\infty f(\tau)g(t - \tau) d\tau
-  \eqlabel{convolution_definition}
+  \eqlabel{convolution-definition}
 \end{equation}
@@
   \mathcal{F}\{f \ast g\} = \mathcal{F}\{f\} \cdot \mathcal{F}\{g\}
-  \eqlabel{convolution_theorem}
+  \eqlabel{convolution-theorem}
 \end{equation}
@@
 than to directly solve at each time $t$ the integral over $\tau$ that appears
-in \eq{convolution_definition}.
-This result in \eq{convolution_theorem} is known as the
-{\it convolution theorem}, and is illustrated in \fig{convolution_theorem}.
+in \eq{convolution-definition}.
+This result in \eq{convolution-theorem} is known as the
+{\it convolution theorem}, and is illustrated in \fig{convolution-theorem}.
 \begin{figure}[ht]
@@
   \includegraphics[width=\textwidth]{fig05_Convolution_Theorem}
-  \caption{Visualization of the convolution theorem (\eq{convolution_theorem}).
+  \caption{Visualization of the convolution theorem (\eq{convolution-theorem}).
     The key is that the Fourier transform of
@@
     imaginary part of the transform, respectively.
-    \figlabel{convolution_theorem}}
+    \figlabel{convolution-theorem}}
 \end{figure}
@@
   \mathcal{F}\{f \cdot g\} = \mathcal{F}\{f\} \ast \mathcal{F}\{g\}
-  \eqlabel{convolution_theorem_inverse}
+  \eqlabel{convolution-theorem-inverse}
 \end{equation}
@@
 First, let's consider the case of observing a continuous periodic signal over
-a limited span of time, as illustrated in \Fig{tophat_window}.
+a limited span of time, as illustrated in \Fig{tophat-window}.
 Here we consider a periodic function made up of three frequency components, observed within a 10-unit time window.
 The observed signal in this case can be thought of as a pointwise product of an infinite periodic signal, and a tophat window function.
-Using \eq{convolution_theorem_inverse}, we see that the Fourier transform of
+Using \eq{convolution-theorem-inverse}, we see that the Fourier transform of
 this function is equivalent to the convolution of the Fourier transforms of each component, which are a set of delta functions and a sinc function, respectively.
 For a purely periodic signal, this convolution has the effect of broadening the delta functions representing the signal, turning each of them into a shifted sinc function.
-Because of the inverse relationship between the width of the window and the width of its transform (see \Fig{fourier_pairs}), it follows that a wider observing window leads to proportionally less spread in the Fourier transform of the observed function.
+Because of the inverse relationship between the width of the window and the width of its transform (see \Fig{fourier-pairs}), it follows that a wider observing window leads to proportionally less spread in the Fourier transform of the observed function.
 
@@
     functions) and the window transform (here a narrow sinc function).
-    \figlabel{tophat_window}}
+    \figlabel{tophat-window}}
 \end{figure}
@@
 underlying signal and a Dirac comb with the $T$ parameter matching the spacing
-of the observations; this is illustrated in \fig{comb_window_1}.
+of the observations; this is illustrated in \fig{comb-window-1}.
 
@@
     Gaussian) and the window transform (here another Dirac comb).
-    \figlabel{comb_window_1}}
+    \figlabel{comb-window-1}}
 \end{figure}
@@
 
-The example in \fig{comb_window_1} is somewhat of a best-case scenario, because
+The example in \fig{comb-window-1} is somewhat of a best-case scenario, because
 the true Fourier transform values lie entirely in a range of width $1/T$.
 If we change the sampling rate so that this is no longer the case, we will
-have a situation similar to that in \fig{comb_window_2}.
+have a situation similar to that in \fig{comb-window-2}.
 With a lower sampling rate, the signal transform no longer fits within the
@@
   \includegraphics[width=\textwidth]{fig08_comb_window_2}
-  \caption{Repeating the visualization from \fig{comb_window_1}, but here with
+  \caption{Repeating the visualization from \fig{comb-window-1}, but here with
     a lower sampling rate. The result is that the Fourier transform of the
@@
     recovered.
-    \figlabel{comb_window_2}}
+    \figlabel{comb-window-2}}
 \end{figure}
@@
 \end{equation}
-which follows directly from \eq{FT-def} and \eq{dirac_comb}.
+which follows directly from \eq{FT-def} and \eq{dirac-comb}.
 
@@
   \hat{g}_{obs}(f) = \sum_{n=0}^N g_n e^{-2\pi i f n T}
-  \eqlabel{DFT_f}
+  \eqlabel{DFT-f}
 \end{equation}
@@
 to our data a rectangular window function of width $NT$.
-From the discussion accompanying \fig{tophat_window}, we know what this does:
+From the discussion accompanying \fig{tophat-window}, we know what this does:
 it gives us a Fourier transform convolved with a sinc function of width
@@
 
-With the discrete Fourier transform defined in \eqs{DFT_f}{DFT}, we can
+With the discrete Fourier transform defined in \eqs{DFT-f}{DFT}, we can
 apply the definition of the Fourier power spectrum from \eq{power-spectrum}
@@
 Apart from the $1/N$ scaling, this is precisely the Fourier power defined
-in \eq{power-spectrum} for the expression in \eq{DFT_f}, and it follows that
+in \eq{power-spectrum} for the expression in \eq{DFT-f}, and it follows that
 for a uniformly-sampled function $g_n$, the Schuster periodogram captures
@@
 window transform will generally {\it not} be a nice clean function.
-\Fig{random_window} shows the effect on the Fourier transform of a
+\Fig{random-window} shows the effect on the Fourier transform of a
 non-uniform observing cadence with a sampling rate identical to that
-in \fig{comb_window_1}.
+in \fig{comb-window-1}.
 
@@
 lack of useful structure.
-Unlike the uniform result in \Fig{comb_window_1}, there is no direct aliasing
+Unlike the uniform result in \Fig{comb-window-1}, there is no direct aliasing
 of the true signal, and no way to exactly recover any portion of the true
@@
   \caption{The effect of non-uniform sampling on the observed Fourier transform.
-    These samples have the same average spacing as those in \fig{comb_window_1},
+    These samples have the same average spacing as those in \fig{comb-window-1},
     but the lack of structure in observing window translates to a lack of
     structure in its transform, causing the observed transform to be ``noisy''.
-    \figlabel{random_window}}
+    \figlabel{random-window}}
 \end{figure}
@@
 and it does, but only to a degree.
-\Fig{random_window_2} increases the density of observations by a factor of 10,
+\Fig{random-window-2} increases the density of observations by a factor of 10,
 such that there are 200 total observations over the length-10 observing window.
@@
   \caption{The effect of non-uniform sampling on the observed Fourier transform,
-    with a factor of 10 more samples than \fig{random_window}.
+    with a factor of 10 more samples than \fig{random-window}.
     Even with very dense sampling of the function, the Fourier transform
     cannot be exactly recovered due to the 
-    These samples have the same average spacing as those in \fig{comb_window_1},
+    These samples have the same average spacing as those in \fig{comb-window-1},
     but the lack of structure in observing window translates to a lack of
@@
     ``noisy'' representation of the true transform
-    \figlabel{random_window_2}}
+    \figlabel{random-window-2}}
 \end{figure}
@@
 As a simple example of where such logic fails,
-consider the data from \fig{LINEAR_data}: though the mean sample
-spacing is one observation every seven days, in \fig{LINEAR_power}
+consider the data from \fig{LINEAR-data}: though the mean sample
+spacing is one observation every seven days, in \fig{LINEAR-power}
 we were nevertheless able to quite clearly identify a period of 2.58 hours---
@@
 would indicate as possible.
-For the data in \fig{LINEAR_data}, the minimum sample spacing is just under 10
+For the data in \fig{LINEAR-data}, the minimum sample spacing is just under 10
 seconds, but it it irresponsible to pretend that this single pair of
@@
     estimates based on average or minimum sampling rate.
-    \figlabel{pseudo_nyquist}}
-\end{figure}
-
-As a more extreme example, consider the data shown in \fig{pseudo_nyquist}.
+    \figlabel{pseudo-nyquist}}
+\end{figure}
+
+As a more extreme example, consider the data shown in \fig{pseudo-nyquist}.
 This consists of noisy samples from a sinusoid with a period of 0.01 units,
@@
 
-\Fig{nyquist_eyer99} shows an example of such a Nyquist frequency.
+\Fig{nyquist-eyer99} shows an example of such a Nyquist frequency.
 The data are non-uniformly sampled at times $t_i = n_i * p$, with $p=0.01$.
@@
     repetitions of the signal within that range.
-    \figlabel{nyquist_eyer99}}
+    \figlabel{nyquist-eyer99}}
 \end{figure}
@@
 That is, if times are recorded to $D$ decimal places, the Nyquist frequency
-will be (at most) $f_{Ny} = \frac{1}{2} 10^D$, as in \fig{nyquist_eyer99}.
+will be (at most) $f_{Ny} = \frac{1}{2} 10^D$, as in \fig{nyquist-eyer99}.
 In other words, the Nyquist frequency for irregularly-sampled data is most
@@
 the underlying signal with a rectangular window function of width $\delta t$,
-in a manner analogous to \fig{convolution_theorem}.
+in a manner analogous to \fig{convolution-theorem}.
 By the convolution theorem, the observed transform will be a point-wise
@@
 in the observing window can lead to partial aliasing of signals in the
-data.
+data \citep[see, e.g.][]{Deeming75}.
 
@@
   \caption{The power spectrum of the observing window for the data shown
-    in \fig{LINEAR_data}. Notice the strong spike in power at a period of
+    in \fig{LINEAR-data}. Notice the strong spike in power at a period of
     1 day, and related aliases at $1/n$ days for integer $n$.
@@
     in the data.
-    \figlabel{LINEAR_window}}
+    \figlabel{LINEAR-window}}
 \end{figure}
@@
 
-Let's again consider the data shown in \fig{LINEAR_data}. The window power
-spectrum for this in \fig{LINEAR_window} shows some quite distinct features,
+Let's again consider the data shown in \fig{LINEAR-data}. The window power
+spectrum for this in \fig{LINEAR-window} shows some quite distinct features,
 and these features have an intuitive interpretation.
@@
 a convolution of the true spectrum and the window spectrum
-(cf.\ \fig{random_window}), and so we would expect this structure to be
+(cf.\ \fig{random-window}), and so we would expect this structure to be
 imprinted on the power spectrum measured from the data.
@@
   \includegraphics[width=\textwidth]{fig15_LINEAR_window_effect}
-  \caption{The effect of the window function in \fig{LINEAR_window} on the
-    power spectrum in \fig{LINEAR_power}.
+  \caption{The effect of the window function in \fig{LINEAR-window} on the
+    power spectrum in \fig{LINEAR-power}.
     The top panel shows the window power spectrum, and the bottom panel shows
@@
     Both are plotted as a function of frequency (we earlier saw both of these
-    as a function of period; see \fig{LINEAR_window} and \fig{LINEAR_power},
+    as a function of period; see \fig{LINEAR-window} and \fig{LINEAR-power},
     respectively).
@@
     are apparent in the observed spectrum.
-    \figlabel{LINEAR_window_effect}}
-\end{figure}
-
-This behavior is illustrated in \fig{LINEAR_window_effect}.
+    \figlabel{LINEAR-window-effect}}
+\end{figure}
+
+This behavior is illustrated in \fig{LINEAR-window-effect}.
 The upper panel shows the window power spectrum as a function of frequency, while lower panel shows the observed signal power spectrum as a function of frequency.
-As noted in the discussion of \fig{LINEAR_window}, the window function shows a strong signal at frequencies corresponding to 1 day and 1 year; the latter can be seen in the inset plot.
+As noted in the discussion of \fig{LINEAR-window}, the window function shows a strong signal at frequencies corresponding to 1 day and 1 year; the latter can be seen in the inset plot.
 The power at both these scales is clearly imprinted on the observed power spectrum, as shown in the lower panel and its inset.
@@
 The window function is a series of very narrow evenly-spaced spikes,
-reminiscent of the Dirac comb shown in \figs{comb_window_1}{comb_window_2},
+reminiscent of the Dirac comb shown in \figs{comb-window-1}{comb-window-2},
 which we used to motivate the idea of the Nyquist limit.
@@
 The window functions for ground-based and space-based observations, reflected
-by LINEAR data in \fig{LINEAR_window_effect} and Kepler data in
+by LINEAR data in \fig{LINEAR-window-effect} and Kepler data in
 \fig{kepler-data}, are quite different, but in both cases essential features
@@
 the window function.
+For another example of this type of observing window analysis, see
+\citep{Deeming75}
 
@@
 
+Up until now, we've been mainly discussing the direct extension of the Schuster
+periodogram in \eq{schuster-periodogram} to non-uniform data.
+Returning to this definition, we can rewrite the expression in a more
+suggestive way:
+\begin{eqnarray}
+  P(f)
+  &=& \frac{1}{N}\left|\sum_{n=1}^N g(t_n)e^{-2\pi i t_n} \right|^2 \nonumber\\
+  &=& \frac{1}{N}\left[
+    \left(\sum_n g_n \cos(2\pi f t_n)\right)^2
+    + \left(\sum_n g_n \sin(2\pi f t_n)\right)^2
+    \right]
+  \eqlabel{classical-periodogram}
+\end{eqnarray}
+Although this form of the non-uniform periodogram can be useful for identifying
+periodic signals, its statistical properties are not as straightforward as in
+the uniform case.
+In this case, when we say an estimator with ``good statistical properties'',
+we mean that when it is applied to data with straightforward, uncorrelated
+noise, the result is a periodogram estimate in which the noise is also
+straightforward and uncorrelated.
+
+\citet{Scargle82} shows that the classical periodogram of
+\eq{classical-periodogram} has such desireable statistical properties in the
+case of uniform sampling, but not in the case of non-uniform sampling.
+Such desireable properties can be recovered in the non-uniform case by
+applying a slight modification to the standard periodogram:
+\begin{equation}
+  P(f)
+  = \frac{1}{2}\left[
+    \frac{\left[\sum_n g_n \cos(2\pi f [t_n-\tau])\right]^2}
+         {\sum_n g_n \cos^2(2\pi f [t_n-\tau])}
+  + \frac{\left[\sum_n g_n \sin(2\pi f [t_n-\tau])\right]^2}
+         {\sum_n g_n \sin^2(2\pi f [t_n-\tau])}
+  \right]
+  \eqlabel{lomb-scargle-periodogram}
+\end{equation}
+where $\tau$ is given by
+\begin{equation}
+  \tan(4\pi f \tau) = \frac{\sum_n 4\pi f \tau}{\sum_n 4\pi f\tau}.
+  \eqlabel{tau-def}
+\end{equation}
+This form has since become commonly referred to as the ``Lomb-Scargle Periodogram'', after \citet{Lomb76} and \citet{Scargle82} who proposed this alternate
+periodogram and explored its statistical properties.
+
+\todo{properties of the periodogram}
 \begin{itemize}
-  \item ... is like a Schuster periodogram.
-  \item ... is also like a least-squares sinusoid fit.
-  \item ... reminder that it is a {\it windowed} power spectrum
+  \item Reduces to standard form for even spacing
+  \item Identical to the $\chi^2$ for a simple sinusoidal fit
+    at each frequency
+  \item Is qualitatively similar to the classical periodogram, meaning all
+    the above discussion is still applicable
 \end{itemize}
+
+\todo{show classical vs. lomb-scargle periodogram for LINEAR data}
 

### 2016-12-27 PracticalLombScargle.tex
Add discussion of Classical->Lomb Scargle connection

@@
     the periodogram outside the range $0 \le f < f_{Ny}$ is a series of
-    repetitions of the signal within that range.
+    perfect aliases of the signal within that range.
     \figlabel{nyquist-eyer99}}
@@
 
-\citet{Scargle82} shows that the classical periodogram of
-\eq{classical-periodogram} has such desireable statistical properties in the
-case of uniform sampling, but not in the case of non-uniform sampling.
-Such desireable properties can be recovered in the non-uniform case by
-applying a slight modification to the standard periodogram:
-\begin{equation}
-  P(f)
-  = \frac{1}{2}\left[
-    \frac{\left[\sum_n g_n \cos(2\pi f [t_n-\tau])\right]^2}
-         {\sum_n g_n \cos^2(2\pi f [t_n-\tau])}
-  + \frac{\left[\sum_n g_n \sin(2\pi f [t_n-\tau])\right]^2}
-         {\sum_n g_n \sin^2(2\pi f [t_n-\tau])}
-  \right]
+\citet{Scargle82} showed that while the classical periodogram of
+\eq{classical-periodogram} has these desireable statistical properties
+in the case of uniform sampling, this does not hold for
+\eq{classical-periodogram} when the sampling becomes non-uniform.
+To address this, he considered a generalized form of the periodogram,
+\begin{equation}
+  P(f) = \frac{A^2}{2}\left(\sum_n g_n \cos(2\pi f t_n)\right)^2
+       + \frac{B^2}{2} \left(\sum_n g_n \sin(2\pi f t_n)\right)^2,
+\end{equation}
+where $A$ and $B$ are arbitrary functions of the frequency $f$ and
+observing times $\{t_i\}$ (but not the values $\{g_n\}$, and showed
+that you can choose one particular form of $A$ and $B$ such that
+\begin{itemize}
+  \item The periodogram reduces to the classical form in the case of equally-spaced data,
+  \item The periodogram has desireable statistical properties, in the sense discussed above,
+  \item The periodogram is insensitive to global time-shifts in the data.
+\end{itemize}
+The values of $A$ and $B$ leading to these properties lead to the following
+modification of the classical expression for the periodogram:
+\begin{eqnarray}
+  P(f) =
+  \frac{1}{2} \Bigg\{ &
+  \bigg(\sum_n g_n \cos(2\pi f [t_n-\tau])\bigg)^2 \bigg/
+  \sum_n \cos^2(2\pi f [t_n-\tau]) &\nonumber\\
+  & + ~ \bigg(\sum_n g_n \sin(2\pi f [t_n-\tau])\bigg)^2 \bigg/
+  \sum_n \sin^2(2\pi f [t_n-\tau]) & \Bigg\}
   \eqlabel{lomb-scargle-periodogram}
-\end{equation}
-where $\tau$ is given by
-\begin{equation}
-  \tan(4\pi f \tau) = \frac{\sum_n 4\pi f \tau}{\sum_n 4\pi f\tau}.
+\end{eqnarray}
+Here $\tau$ is defined for each $f$ to ensure time-shift invariance of $P(f)$:
+\begin{equation}
+  \tau = \frac{1}{4\pi f}\tan^{-1}\Bigg(
+  \frac{\sum_n \sin(4\pi f t_n)}{\sum_n \cos(4\pi f t_n)}\Bigg).
   \eqlabel{tau-def}
 \end{equation}
-This form has since become commonly referred to as the ``Lomb-Scargle Periodogram'', after \citet{Lomb76} and \citet{Scargle82} who proposed this alternate
-periodogram and explored its statistical properties.
+A remarkable feature of this result is that it is {\it identical} to the result
+obtained by fitting a simple sinusoid to the data at each frequency $f$, and
+constructing a ``periodogram'' from the $\chi^2$ goodness-of-fit at each
+frequency--an estimator which was considered in some depth by \citet{Lomb76}.
+Partly due to this deep connection between Fourier analysis and least-squares
+analysis, the modified periodogram in \eq{lomb-scargle-periodogram}
+has since become commonly referred to as the ``Lomb-Scargle Periodogram''.
+
+Often, the Lomb-Scargle periodogram only differs slightly from the
+classical/Schuster periodogram; an example of this is seen in
+\fig{ls-comparison}.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig17_ls_comparison}
+  \caption{A comparison of the Classical periodogram
+    (\eq{classical-periodogram}) and the Lomb-Scargle periodogram
+    (\eq{lomb-scargle-periodogram}) for 30 noisy points drawn from a sinusoid.
+    Though the two differ quantitatively, the essential qualitative
+    features---namely the position of significant peaks---generally remains
+    the same.
+    \figlabel{ls-comparison}}
+\end{figure}
+
+\subsection{Properties of the Lomb-Scargle Periodogram}
 
@@
 \begin{itemize}
+\item Uncertainties in Data
 \item Bayesian view is (almost always) not useful

### 2016-12-27 PracticalLombScargle.tex
add more discussion of LS denominator

@@
 that you can choose one particular form of $A$ and $B$ such that
-\begin{itemize}
+\begin{enumerate}
   \item The periodogram reduces to the classical form in the case of equally-spaced data,
@@
   \item The periodogram is insensitive to global time-shifts in the data.
-\end{itemize}
+\end{enumerate}
 The values of $A$ and $B$ leading to these properties lead to the following
@@
 frequency--an estimator which was considered in some depth by \citet{Lomb76}.
+In this view, the $\tau$ shift defined in \eq{tau-def} serves to orthogonalize
+the normal equations used in the least squares analysis.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig17_ls_comparison}
+  \caption{{\it upper panel:} A comparison of the Classical periodogram
+    (\eq{classical-periodogram}) and the Lomb-Scargle periodogram
+    (\eq{lomb-scargle-periodogram}) for 30 noisy points drawn from a sinusoid.
+    Though the two periodogram estimates differ quantitatively, the essential
+    qualitative features---namely the position of significant peaks---generally
+    remain the same.
+    {\it lower panel:} the denominators in \eq{lomb-scargle-periodogram}.
+    The difference between the Lomb-Scargle periodogram and the classical
+    periodogram stems from the difference between these quantities
+    and $N/2 = 15$.
+    \figlabel{ls-comparison}}
+\end{figure}
+
 Partly due to this deep connection between Fourier analysis and least-squares
 analysis, the modified periodogram in \eq{lomb-scargle-periodogram}
-has since become commonly referred to as the ``Lomb-Scargle Periodogram''.
-
-Often, the Lomb-Scargle periodogram only differs slightly from the
-classical/Schuster periodogram; an example of this is seen in
-\fig{ls-comparison}.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{fig17_ls_comparison}
-  \caption{A comparison of the Classical periodogram
-    (\eq{classical-periodogram}) and the Lomb-Scargle periodogram
-    (\eq{lomb-scargle-periodogram}) for 30 noisy points drawn from a sinusoid.
-    Though the two differ quantitatively, the essential qualitative
-    features---namely the position of significant peaks---generally remains
-    the same.
-    \figlabel{ls-comparison}}
-\end{figure}
-
-\subsection{Properties of the Lomb-Scargle Periodogram}
-
-\todo{properties of the periodogram}
-\begin{itemize}
-  \item Reduces to standard form for even spacing
-  \item Identical to the $\chi^2$ for a simple sinusoidal fit
-    at each frequency
-  \item Is qualitatively similar to the classical periodogram, meaning all
-    the above discussion is still applicable
-\end{itemize}
-
-\todo{show classical vs. lomb-scargle periodogram for LINEAR data}
+has since become commonly referred to as the {\it Lomb-Scargle Periodogram}.
+The Lomb-Scargle periodogram differs from the classical periodogram only to
+the extent that the denominators $\sum_n \sin^2(2\pi f t_n)$ and
+$\sum_n \cos^2(2\pi f t_n)$ differ from $N/2$, which is the expected value of
+each of these quantities in the limit of very dense sampling.
+Thus, in many cases of interest the Lomb-Scargle periodogram only differs
+slightly from the classical/Schuster periodogram; an example of this is seen
+in \fig{ls-comparison}.
+
+Because of this similarity between the classical and Lomb-Scargle periodograms,
+our discussion of windowing effects applies--at least qualitatively--to results
+computed with the Lomb-Scargle periodogram.
+In particular, reasoning about the effect of window functions on the observed
+Lomb-Scargle power spectrum remains useful even if it is true in only a
+qualitative sense.
 

### 2016-12-27 PracticalLombScargle.tex
Add LS model figure

@@
 the normal equations used in the least squares analysis.
+In other words, the Lomb-Scargle periodogram essentially {\it assumes a
+sinusoidal model} for the data, which is visualized in \fig{ls-model} for
+the data we had seen in \fig{LINEAR-data}.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig18_ls_model}
+  \caption{The sinusoidal model implied by the Lomb-Scargle periodogram for
+    the LINEAR data seen in \fig{LINEAR-data}.
+    Although a sinusoid does not perfectly fit the data, it is close enough
+    that it serves to find the correct period without a problem.
+    \figlabel{ls-model}}
+\end{figure}
+
+Partly due to the deep connection between Fourier analysis and least-squares
+analysis, the modified periodogram in \eq{lomb-scargle-periodogram}
+has since become commonly referred to as the {\it Lomb-Scargle Periodogram}.
+The Lomb-Scargle periodogram differs from the classical periodogram only to
+the extent that the denominators $\sum_n \sin^2(2\pi f t_n)$ and
+$\sum_n \cos^2(2\pi f t_n)$ differ from $N/2$, which is the expected value of
+each of these quantities in the limit of very dense sampling.
+Thus, in many cases of interest the Lomb-Scargle periodogram only differs
+slightly from the classical/Schuster periodogram; an example of this is seen
+in \fig{ls-comparison}.
 
@@
 
-Partly due to this deep connection between Fourier analysis and least-squares
-analysis, the modified periodogram in \eq{lomb-scargle-periodogram}
-has since become commonly referred to as the {\it Lomb-Scargle Periodogram}.
-The Lomb-Scargle periodogram differs from the classical periodogram only to
-the extent that the denominators $\sum_n \sin^2(2\pi f t_n)$ and
-$\sum_n \cos^2(2\pi f t_n)$ differ from $N/2$, which is the expected value of
-each of these quantities in the limit of very dense sampling.
-Thus, in many cases of interest the Lomb-Scargle periodogram only differs
-slightly from the classical/Schuster periodogram; an example of this is seen
-in \fig{ls-comparison}.
-
 Because of this similarity between the classical and Lomb-Scargle periodograms,
-our discussion of windowing effects applies--at least qualitatively--to results
-computed with the Lomb-Scargle periodogram.
+our discussion of windowing effects applies---at least qualitatively---to
+results computed with the Lomb-Scargle periodogram.
 In particular, reasoning about the effect of window functions on the observed
@@
 
-\section{Using Lomb-Scargle Effectively}
-
-\begin{itemize}
-  \item Expected peak width
-  \item How to choose the frequency grid
-  \item Normalization and interpretation
-\end{itemize}
+\section{Practical Considerations when using Lomb-Scargle Periodograms}
+
+\subsection{Choosing a Frequency Grid}
+
+\subsection{Normalization of the Periodogram}
+
+\subsection{Algorithmic Considerations}
 
@@
 
-\section{Algorithmic Considerations}
+\section{Generalizations \& Challenges}
 
 \begin{itemize}
-\item Press \& Rybicki method; show some benchmarks.
-\item NFFT method
-\end{itemize}
-
-\section{Generalizations \& Challenges}
-
-\begin{itemize}
+\item floating-mean
 \item Uncertainties in Data
@@
 
-\appendix
-
-\section{Non-uniform Nyquist}
-A selection of quotations showing approaches to a non-uniform Nyquist limit:
-
-\begin{description}
-
-\item[Average Sampling Rate]
-
-\citet{Scargle82} (citing the following: Beutler 1966, 1970; Masry \& Lui 1975; Higgins 1976; Wiley 1978; Gaster \& Roberts 1975, 1977; Kar, Hornkohl \& Farmer 1981; Ludeman 1981 \todo{read and summarize these!}):
-\begin{quote}
-Error-free recovery of a band-limited signal [i.e.~reproduction of the entire function $X(t)$ from the samples $X(t_i)$] can be achieved with irregular sampling as long as the mean sampling rate exceeds the Nyquist rate (i.e.
-the average number of samples per unit time must exceed twice the highest frequency component in the signal).
-\end{quote}
-
-\citet{NumRec}: 
-\begin{quote}
-One guide to choosing $f_{hi}$ is to compare it with the Nyquist frequency $f_c$ which would obtain if the $N$ data points were evenly spaced over the same span $T$, that is $f_c = N/(2T)$.
-The accompanying program includes an input parameter {\tt hifac}, defined as $f_{hi}/f_{c}$.
-\end{quote}
-
-\citet{Horne86}:
-\begin{quote}
-The largest frequency we calculated was $\pi N/T$ which
-is the traditional Nyquist frequency for evenly-spaced data. The Nyquist
-frequency is not well defined for unevenly spaced signals, but it can serve as
-a reasonable upper limit for the calculation.
-\end{quote}
-
-\item[Harmonic Average of Sampling Rate]
-
-\citet{Debosscher07}:
-\begin{quote}
-For the highest frequency, we used the average of the inverse time intervals
-between the measurements: $f_N = 0.5(1/\Delta T)$ as a pseudo
-Nyquist frequency. Note that $f_N$ is equal to the Nyquist frequency
-in the case of equidistant sampling. For particular cases,
-an even higher upper limit can be used \citep[see][]{Eyer99}.
-Our upper limit should be seen as a compromise between
-the required resolution to allow a good fitting, and computation
-time.
-\end{quote}
-
-\item[Minimum Sample Spacing]
-
-\citet{Percy86}:
-\begin{quote}
-... the Nyquist frequency is not well defined for input data obtained at unequally spaced time intervals, the usual case in astronomy \citep{Scargle82}.
-Theoretically such a data set contains informatin on the periodicities down to
-$\Delta t = \min(t_i - t{i-1})$.
-In practice, however, a pseudo-Nyquist frequency may be defined by averaging $\Delta t = (t_i - t_{i-1})$, where large, uncharacteristic temporal gaps are avoided.
-Alternatively, the harmonic mean of all $\Delta t$ may be used.
-The result is that a useful pseudo-Nyquist frequency may be defined by $\nu_{Ny} = (2 \langle\Delta t \rangle)^{-1}$.
-\end{quote}
-
-\citet{Press89}:
-\begin{quote}
-It is often meaningful to examine frequencies significantly higher than the
-Nyquist frequency that would obtain if the same number of data points were
-evenly spaced in the same total length of time. Some spectral information is
-obtainable for frequencies all the way up to something like half the inverse
-spacing of the {\it closest} spaced points.
-\end{quote}
-
-\citet{Roberts87}:
-\begin{quote}
-...for arbitrary $\{t_r\}$, the sampling theorem tells us nothing.
-If the data samples are otherwise equally spaced but with missing
-points, the theorem says that the data completely determine
-a function whose FT is zero for $|v| > 1/(2\Delta_{max})$, where $\Delta_{max}$
-is the {\it largest completely sampled} data spacing. However,
-there are smaller spacings, and these certainly carry information
-about frequencies greater than $1/(2\Delta_{max} )$; some information
-is available about frequencies as high as
-$1/(2\Delta_{min} )$, where $\Delta_{min}$ is the smallest spacing between data
-points. Furthermore, if the $\{t_r\}$ are more or less randomly
-distributed, so that a wide range ofspacings are present and
-there is little redundancy in the spacing between various
-points, tests have shown (Paper II) that significant information
-is available on frequencies greater than $l/(2\Delta_{min})$.
-
-Nonetheless, in the present paper we will restrict ourselves
-to frequencies obeying
-$\nu < \nu_{max} = 1/(2\Delta_{min})$.
-\end{quote}
-
-\citet{Hilditch01}:
-\begin{quote}
-The highest frequency, for equally spaced data, is formally given by the {\it Nyquist frequency}, $f_N = 1/(2\Delta t)$, where $\Delta t$ is the {\it sampling interval} in the data string.
-For unequally spaced data, it is common practice to estimate a pseudo-Nyquist frequency from the minimum value of $\Delta t$.
-\end{quote}
-
-\item[Getting it Right]
-
-\citet{ICVG2014}:
-\begin{quote}
-As a good choice for the maximum search frequency, a pseudo-Nyquist frequency $\omega_{max} = \pi/\Delta t$, where $1/\Delta t$ is the median of the inverse time interval between data points, was proposed by \citet{Debosscher07} (in the case of even sampling, $\omega_{max}$ is equal to the Nyquist frequency).
-In practice, this choice may be a gross underestimate because unevenly sampled data can detect periodicity with frequencies even higher than $2\pi / \Delta t_{min}$ \citep[see][]{Eyer99}.
-An appropriate choice of $\omega_{max}$ thus depends on sampling (the phase coverage at a given frequency is the relevant quantity) and needs to be carefully chosen: a hard limit on maximum detectable frequency is of course given by the time interval over which individual measurements are performed, such as imaging exposure time.
-\end{quote}
-
-\end{description}
-
 \end{document}

### 2016-12-27 notes.tex
Add LS model figure

\section{Non-uniform Nyquist}
A selection of quotations showing approaches to a non-uniform Nyquist limit:

\begin{description}

\item[Average Sampling Rate]

\citet{Scargle82} (citing the following: Beutler 1966, 1970; Masry \& Lui 1975; Higgins 1976; Wiley 1978; Gaster \& Roberts 1975, 1977; Kar, Hornkohl \& Farmer 1981; Ludeman 1981 \todo{read and summarize these!}):
\begin{quote}
Error-free recovery of a band-limited signal [i.e.~reproduction of the entire function $X(t)$ from the samples $X(t_i)$] can be achieved with irregular sampling as long as the mean sampling rate exceeds the Nyquist rate (i.e. the average number of samples per unit time must exceed twice the highest frequency component in the signal).
\end{quote}

\citet{NumRec}: 
\begin{quote}
One guide to choosing $f_{hi}$ is to compare it with the Nyquist frequency $f_c$ which would obtain if the $N$ data points were evenly spaced over the same span $T$, that is $f_c = N/(2T)$. The accompanying program includes an input parameter {\tt hifac}, defined as $f_{hi}/f_{c}$.
\end{quote}

\citet{Horne86}:
\begin{quote}
The largest frequency we calculated was $\pi N/T$ which
is the traditional Nyquist frequency for evenly-spaced data. The Nyquist
frequency is not well defined for unevenly spaced signals, but it can serve as
a reasonable upper limit for the calculation.
\end{quote}

\item[Harmonic Average of Sampling Rate]

\citet{Debosscher07}:
\begin{quote}
For the highest frequency, we used the average of the inverse time intervals
between the measurements: $f_N = 0.5(1/\Delta T)$ as a pseudo
Nyquist frequency. Note that $f_N$ is equal to the Nyquist frequency
in the case of equidistant sampling. For particular cases,
an even higher upper limit can be used \citep[see][]{Eyer99}.
Our upper limit should be seen as a compromise between
the required resolution to allow a good fitting, and computation
time.
\end{quote}

\item[Minimum Sample Spacing]

\citet{Percy86}:
\begin{quote}
... the Nyquist frequency is not well defined for input data obtained at unequally spaced time intervals, the usual case in astronomy \citep{Scargle82}.
Theoretically such a data set contains informatin on the periodicities down to
$\Delta t = \min(t_i - t{i-1})$. In practice, however, a pseudo-Nyquist frequency may be defined by averaging $\Delta t = (t_i - t_{i-1})$, where large, uncharacteristic temporal gaps are avoided. Alternatively, the harmonic mean of all $\Delta t$ may be used. The result is that a useful pseudo-Nyquist frequency may be defined by $\nu_{Ny} = (2 \langle\Delta t \rangle)^{-1}$.
\end{quote}

\citet{Press89}:
\begin{quote}
It is often meaningful to examine frequencies significantly higher than the
Nyquist frequency that would obtain if the same number of data points were
evenly spaced in the same total length of time. Some spectral information is
obtainable for frequencies all the way up to something like half the inverse
spacing of the {\it closest} spaced points.
\end{quote}

\citet{Roberts87}:
\begin{quote}
...for arbitrary $\{t_r\}$, the sampling theorem tells us nothing.
If the data samples are otherwise equally spaced but with missing
points, the theorem says that the data completely determine
a function whose FT is zero for $|v| > 1/(2\Delta_{max})$, where $\Delta_{max}$
is the {\it largest completely sampled} data spacing. However,
there are smaller spacings, and these certainly carry information
about frequencies greater than $1/(2\Delta_{max} )$; some information
is available about frequencies as high as
$1/(2\Delta_{min} )$, where $\Delta_{min}$ is the smallest spacing between data
points. Furthermore, if the $\{t_r\}$ are more or less randomly
distributed, so that a wide range ofspacings are present and
there is little redundancy in the spacing between various
points, tests have shown (Paper II) that significant information
is available on frequencies greater than $l/(2\Delta_{min})$.

Nonetheless, in the present paper we will restrict ourselves
to frequencies obeying
$\nu < \nu_{max} = 1/(2\Delta_{min})$.
\end{quote}

\citet{Hilditch01}:
\begin{quote}
The highest frequency, for equally spaced data, is formally given by the {\it Nyquist frequency}, $f_N = 1/(2\Delta t)$, where $\Delta t$ is the {\it sampling interval} in the data string.
For unequally spaced data, it is common practice to estimate a pseudo-Nyquist frequency from the minimum value of $\Delta t$.
\end{quote}

\item[Getting it Right]

\citet{ICVG2014}:
\begin{quote}
As a good choice for the maximum search frequency, a pseudo-Nyquist frequency $\omega_{max} = \pi/\Delta t$, where $1/\Delta t$ is the median of the inverse time interval between data points, was proposed by \citet{Debosscher07} (in the case of even sampling, $\omega_{max}$ is equal to the Nyquist frequency). In practice, this choice may be a gross underestimate because unevenly sampled data can detect periodicity with frequencies even higher than $2\pi / \Delta t_{min}$ \citep[see][]{Eyer99}. An appropriate choice of $\omega_{max}$ thus depends on sampling (the phase coverage at a given frequency is the relevant quantity) and needs to be carefully chosen: a hard limit on maximum detectable frequency is of course given by the time interval over which individual measurements are performed, such as imaging exposure time.
\end{quote}

\end{description}

### 2016-12-30 PracticalLombScargle.tex
Add discussion of grid spacing

@@
 \newcommand{\figlabel}[1]{\label{fig:#1}}
-\newcommand{\Eq}[1]{Equation~(\ref{eq:#1})}
+\newcommand{\Eq}[1]{Equation~\ref{eq:#1}}
 \newcommand{\eq}[1]{\Eq{#1}}
-\newcommand{\eqs}[2]{Equations~(\ref{eq:#1})-(\ref{eq:#2})}
+\newcommand{\eqs}[2]{Equations~\ref{eq:#1}-\ref{eq:#2}}
 \newcommand{\eqlabel}[1]{\label{eq:#1}}
@@
 Rather than a rigorous mathematical treatment, the goal of this paper is to build intuition about what assumptions are baked into any periodogram computation.
-We discuss... and ... and ...
+etc. etc....
 \end{abstract}
@@
   \item What is the source of the multiple peaks revealed by the Lomb-Scargle Periodogram?
-  \item What is a suitable range of frequencies to search for a given dataset?
-  \item What is the Nyquist frequency for this type of analysis?
+  \item What is the Nyquist frequency for unevenly-sampled data?
   \item How should we choose the spacing of the frequency grid for our periodogram?
+  \item What assumptions is the Lomb-Scargle periodogram making about the signal?
   \item How should we understand the uncertainty of the computed frequency?
-  \item What assumptions is the Lomb-Scargle periodogram making about the signal?
 \end{enumerate}
-Questions like these are discussed in various textbooks and review papers, but I have not come across any concise reference that discusses such questions in depth.
-This paper seeks to fill that gap, and provide a practical, semi-technical guide to the effective use of the Lomb-Scargle method of periodic analysis.
+The answers to these sorts of questions are discussed in various textbooks and review papers, but I have not come across any single concise reference that gives a good intuition for how to think about such questions.
+This paper seeks to fill that gap, and provide a practical, just-technical-enough guide to the effective use of the Lomb-Scargle method for periodic analysis.
 This paper does not seek a complete or rigorous treatment of the mathematics involved, but rather seeks to develop the intuition of {\it how} to think about the issues involved.
 
+The paper is organized as follows:
+In \sect{continuous-fourier-transform}, we review the continous Fourier
+transform and some of its useful properties, including defining the notion of
+a power spectrum for measuring periodic content in a signal.
+In \sect{window-functions}, we use these properties to understand how different
+observation patterns (i.e. window functions) transform the power spectrum, and
+discuss a conceptual understanding of the well-known Nyquist limit.
+In \sect{non-uniform-sampling}, we consider non-uniformly sampled signals, and
+show that the Nyquist limit for this case is quite different than what many
+practitioners in the field seem to assume.
+In \sect{schuster-to-lomb-scargle}, we discuss the Lomb-Scargle periodogram, a
+modified version of the classical periodogram for unevenly-sampled data, as
+well as the motivation behind these modifications.
+In \sect{practical-considerations}, we discuss important practical
+considerations for the use of the Lomb-Scargle periodogram, including
+the choice of frequency grid and the various normalizations of the periodogram.
+In \sect{uncertainties}, we discuss how to quantify uncertainties in periods
+derived from the periodogram, including various notions of the
+``False Alarm Probability''.
+In \sect{generalizations}, we consider and critique some generalizations of
+the Lomb-Scargle model, including multiterm periodograms and Bayesian
+approaches.
+
+All figures in this work were produced using Python, and in particular the
+Numpy \citep{numpy},
+Pandas \citep{pandas},
+Astropy \citep{Astropy2013},
+and Matplotlib \citep{matplotlib} packages.
+Code to reproduce these figures is available in IPython notebooks at \url{http://github.com/jakevdp/PracticalLombScargle/}.
+
 \section{Background: the Continuous Fourier Transform}
+\sectlabel{continuous-fourier-transform}
 
@@
 \section{Window Functions: From Idealized to Real-world signals}
+\sectlabel{window-functions}
+
 In a typical observational setting, we will be observing some small part of
@@
 \subsection{Effect of a Tophat Window}
+
 First, let's consider the case of observing a continuous periodic signal over
@@
 if the signal is {\it band-limited} between frequencies $\pm f_0/2$.
-This is one way to motivate the famous Nyquist sampling limit, which says
+This is one way to motivate the famous Nyquist sampling limit, which
+approaches the question the other way around and says
 that to fully represent the frequency content of a ``band-limited signal''
-with all Fourier power in the range $\pm B$, we must sample the data with a
-rate of at least $2B$.
+whose Fourier transform is zero outside the range $\pm B$,
+we must sample the data with a rate of at least $2B$.
 
@@
 
-\section{Non-uniform Sampling Windows}
+\section{Non-uniform Sampling}
+\sectlabel{non-uniform-sampling}
 
@@
 
-In the scientific literature, there are countless examples of attempts to
-define a ``pseudo-Nyquist'' frequency to apply in the uneven sampling case.
-A few common approaches include the mean of the sampling intervals
+In the scientific literature it is quite common to come across various proposals
+for a Nyquist-like or pseudo-Nyquist limit in the case of irregular sampling.
+A few typical approaches include using the mean of the sampling intervals
 \citep[e.g.][]{Scargle82, Horne86, NumRec},
@@
 All of these are tempting criteria in that they recover the classical Nyquist 
-frequency in the limit of evenly-spaced data, but this does not make up for the
-fact that they are wrong.
-Several of these cases do parenthetically note that signals at frequencies
-beyond the proposed limit are observable in principle, but they nevertheless
-proceed under the apparent impression that such pseudo-Nyquist reasoning
-is relevant.
+frequency in the limit of evenly-spaced data.
+Unfortunately, none of these approaches is correct: in general,
+unevenly-sampled data can probe frequencies far larger than any of these
+supposed limits (a fact that some of these citations do hint at
+parenthetically).
 
@@
 we were nevertheless able to quite clearly identify a period of 2.58 hours---
-an order of magnitude shorter than the mean-based pseudo-Nyquist limit
+an order of magnitude shorter than the average-based pseudo-Nyquist limit
 would indicate as possible.
 For the data in \fig{LINEAR-data}, the minimum sample spacing is just under 10
-seconds, but it it irresponsible to pretend that this single pair of
+seconds, but it would be irresponsible to claim that this single pair of
 observations itself defines some limit beyond which frequency information is
@@
 
-\subsubsection{The Correct Non-uniform Nyquist Limit}
+\subsubsection{The Non-uniform Nyquist Limit}
 
@@
 a windowed version of uniformly sampled data with spacing $p$.
-Such uniform data has a classical Nyquist limit of $1/(2p)$, and a window function applied to that data cannot change that fact.
+Such uniform data has a classical Nyquist limit of $1/(2p)$, and a window function applied to that data does not change that fact.
 
@@
 That is, if times are recorded to $D$ decimal places, the Nyquist frequency
-will be (at most) $f_{Ny} = \frac{1}{2} 10^D$, as in \fig{nyquist-eyer99}.
-In other words, the Nyquist frequency for irregularly-sampled data is most
-typically defined by {\it the precision of the time measurements}.
+will be, at most
+\begin{equation}
+  f_{Ny} = \frac{1}{2} 10^D,
+  \eqlabel{nonuniform-nyquist}
+\end{equation}
+as in \fig{nyquist-eyer99}.
+In other words, absent other relevant patterns in the observations,
+the Nyquist frequency for irregularly-sampled data is most
+typically set by {\it the precision of the time measurements}.
 
@@
 kind of frequency limit exists.
-This is typical in, e.g., optical astronomy, where observations take place over
-a finite duration $\delta t$.
+This is typical in, e.g., optical astronomy, where a single observation typically consists of an integration of observed flux over a finite duration $\delta t$.
 As noted by \citet{ICVG2014}, this time-scale of integration represents another
@@
 other than missing measurements, the spacing between observations is nearly uniform.
-the lower panel gives a closer look at the majority of the spacings, which are not exactly the same but rather span a range of $\pm 50$ms around the mean.
+the lower panel gives a closer look at the majority of the spacings, which are not exactly the same but rather span a range of $\pm 50$ms around the main period observed in the window function.
     {\it lower right:} the data power spectrum, showing approximate aliasing and clearly indicating a peak near a period of 13.6 hours, along with higher-order components at multiples of this value.
@@
 \section{From Schuster to Lomb-Scargle}
+\sectlabel{schuster-to-lomb-scargle}
 
@@
 \begin{equation}
-  P(f) = \frac{A^2}{2}\left(\sum_n g_n \cos(2\pi f t_n)\right)^2
-       + \frac{B^2}{2} \left(\sum_n g_n \sin(2\pi f t_n)\right)^2,
+  P(f) = \frac{A^2}{2}\left(\sum_n g_n \cos(2\pi f [t_n-\tau])\right)^2
+       + \frac{B^2}{2} \left(\sum_n g_n \sin(2\pi f [t_n-\tau])\right)^2,
 \end{equation}
@@
 \end{eqnarray}
-Here $\tau$ is defined for each $f$ to ensure time-shift invariance of $P(f)$:
+where $\tau$ is specified for each $f$ to ensure time-shift invariance:
 \begin{equation}
@@
 \end{equation}
-A remarkable feature of this result is that it is {\it identical} to the result
+This modified periodogram differs from the classical periodogram only to
+the extent that the denominators $\sum_n \sin^2(2\pi f t_n)$ and
+$\sum_n \cos^2(2\pi f t_n)$ differ from $N/2$, which is the expected value of
+each of these quantities in the limit of complete phase sampling at each
+frequency.
+Thus, in many cases of interest the Lomb-Scargle periodogram only differs
+slightly from the classical/Schuster periodogram; an example of this is seen
+in \fig{ls-comparison}.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig17_ls_comparison}
+  \caption{{\it upper panel:} A comparison of the Classical periodogram
+    (\eq{classical-periodogram}) and the Lomb-Scargle periodogram
+    (\eq{lomb-scargle-periodogram}) for 30 noisy points drawn from a sinusoid.
+    Though the two periodogram estimates differ quantitatively, the essential
+    qualitative features---namely the position of significant peaks---generally
+    remain the same.
+    {\it lower panel:} the values of the denominators in
+    \eq{lomb-scargle-periodogram}.
+    The difference between the Lomb-Scargle periodogram and the classical
+    periodogram stems from the difference between these quantities
+    and $N/2 = 15$ (the dotted line).
+    \figlabel{ls-comparison}}
+\end{figure}
+
+A remarkable feature of Scargle's modified periodogram is that it is
+{\it identical} to the result
 obtained by fitting a simple sinusoid to the data at each frequency $f$, and
@@
 has since become commonly referred to as the {\it Lomb-Scargle Periodogram}.
-The Lomb-Scargle periodogram differs from the classical periodogram only to
-the extent that the denominators $\sum_n \sin^2(2\pi f t_n)$ and
-$\sum_n \cos^2(2\pi f t_n)$ differ from $N/2$, which is the expected value of
-each of these quantities in the limit of very dense sampling.
-Thus, in many cases of interest the Lomb-Scargle periodogram only differs
-slightly from the classical/Schuster periodogram; an example of this is seen
-in \fig{ls-comparison}.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{fig17_ls_comparison}
-  \caption{{\it upper panel:} A comparison of the Classical periodogram
-    (\eq{classical-periodogram}) and the Lomb-Scargle periodogram
-    (\eq{lomb-scargle-periodogram}) for 30 noisy points drawn from a sinusoid.
-    Though the two periodogram estimates differ quantitatively, the essential
-    qualitative features---namely the position of significant peaks---generally
-    remain the same.
-    {\it lower panel:} the denominators in \eq{lomb-scargle-periodogram}.
-    The difference between the Lomb-Scargle periodogram and the classical
-    periodogram stems from the difference between these quantities
-    and $N/2 = 15$.
-    \figlabel{ls-comparison}}
-\end{figure}
+The Lomb-Scargle periodogram differs from the classical periodogram
 
@@
 \section{Practical Considerations when using Lomb-Scargle Periodograms}
+\sectlabel{practical-considerations}
 
 \subsection{Choosing a Frequency Grid}
+The choice of frequency grid for a Lomb-Scargle analysis is an important one
+that is too-often glossed-over, probably because the choice is so
+straightforward in the more familiar case of uniformly-sampled data.
+For non-uniform data, there are two important choices to make: the
+{\bf frequency limits} and the {\bf grid spacing}.
+
+The frequency limit on the low end is relatively easy: for an observing window
+of length $T$, a signal with frequency $1/T$ will just complete a full cycle.
+Often, it's more convenient just to set this minimum frequency to zero, as it
+doesn't add much of a computational burden and is unlikely to add any
+significant peak to the periodogram.
+
+The high-frequency limit is more interesting, and goes back to the discussion
+of Nyquist and/or limiting frequencies from \sect{pseudo-nyquist}:
+in order to not miss relevant information, it is important to compute the
+periodogram up to some well-motivated limiting frequency $f_{max}$. This could
+be a true Nyquist limit based on the \citep{Eyer99} definition, a pseudo-Nyquist
+limit based on careful scrutiny of the window function (cf. \fig{kepler-data}),
+or a limiting frequency based on the integration time of individual
+observations.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig19_LINEAR_coarse_grid}
+  \caption{An example of a poorly-chosen frequency grid for the data in
+    \fig{LINEAR-data}. The dark curve shows a periodogram evaluated on a
+    grid of 10,000 equally-spaced periods; the light curve shows the true
+    periodogram (evaluated at ${\sim}200,000$ equally-spaced frequencies).
+    This demonstrates that using too coarse a grid can lead to a periodogram
+    that entirely misses relevant peaks.
+    \figlabel{LINEAR-coarse-grid}}
+\end{figure}
+
+With the frequency range decided, we now must determine how finely to sample
+the space between the limits.
+This turns out to be quite important: if you choose too fine a grid, it can
+lead to unnecessarily long computation times that can add up quickly in the
+case of large surveys. If you choose too coarse a grid, you
+risk entirely missing narrow peaks that fall between grid points.
+For example, \fig{LINEAR-coarse-grid} shows the true, well-sampled periodogram
+(gray line), along with a periodogram computed for 10,000 equally-spaced
+periods covering the same range (black line).
+Because the spacing of the 10,000-point grid is much larger than the width of
+the periodogram peaks, the computation {\it entirely misses} the most
+important peaks in the periodogram!
+
+To choose an appropriate periodogram grid, then, it is essential to take into
+account the expected widths of periodogram peaks, and construct a grid fine
+enough to assure that each peak will be covered by the grid.
+From our discussion of windowing in \sect{window-functions}
+(particularly \fig{tophat-window}), we know that data observed through a
+rectangular window of length $T$ will have sinc-shaped peaks of width
+$\sim 1/T$.
+To ensure that our grid sufficiently covers each peak, it is prudent to
+over-sample by some factor---say $n_o$ samples per peak---and use a grid of size
+$\Delta f = \frac{1}{n_o T}$. This pushes the total number of required
+periodogram evaluations to
+\begin{equation}
+  \eqlabel{n-eval}
+  N_{eval} = n_o T f_{max}
+\end{equation}
+The numerical result of this can vary greatly from dataset to dataset.
+
+For example, the Kepler data shown in \fig{kepler-data} has a pseudo-Nyquist
+frequency of $48.9\ {\rm days}^{-1}$ and
+an observing window of $T = 90\ {\rm days}$.
+To compute five samples per peak thus requires $N_{eval} \approx 22,000$
+evaluations of the periodogram.
+
+On the other hand, the LINEAR data shown in \fig{LINEAR-data} does not have
+any significant structure in its window function.
+In this case the maximum detectable frequency is the Nyquist limit defined
+by its temporal resolution, which is 6 digits beyond the decimal point in days.
+From \eq{nonuniform-nyquist}, we can write $f_{Ny} = 500,000\ {\rm days}^{-1}$.
+Given the observing window of $T = 1962\ {\rm days}$, we find that five
+evaluations per peak across the entire detectable frequency range
+would require $N_{eval} \approx 4.9 \times 10^9$ evaluations of the periodogram!
+Computing this large a periodogram in most cases is computationally intractable
+(see \sect{algorithmic-considerations}), and so in practice one must choose
+a smaller limiting frequency based on prior information about what kind of
+signals are expected in the data: for example, in \fig{LINEAR-power}, we chose
+a limiting frequency of $f_{max} = (1\ {\rm hour})^{-1}$ based on typical
+oscillation periods of SX Phe-type stars. This leads to a much more manageable
+$N_{eval} \approx 240,000$ periodogram evaluations.
+
+Data from the upcoming LSST survey \citep{Ivezic08LSST} will fall somewhere
+in-between: for full frequency coverage of
+the 10-year data up to a limiting frequency defined by the 30-second
+integration time for each observation will require on order 2.5 million
+periodogram evaluations per object, which for fast Lomb-Scargle implementations
+(see the next section) can be accomplished in seconds on a single core.
+
+One final note: although it can be more easily interpretable to
+visualize periodograms as a function of period rather than a function of
+frequency, the peak widths are not constant in period, and for this reason
+it is ideal to construct the grid itself in frequency space.
+
+\subsection{Algorithmic Considerations}
+\sectlabel{algorithmic-considerations}
+Given the size of the frequency grid required to fully characterize the periods
+from a given dataset, it is vital to have an efficient algorithm for evaluating
+the periodogram. Unfortunately, the periodogram itself is a relatively slow
+operation.
+
+Suppose you have a set of $N$ observations over a time $T$ with an average
+cadence of $\bar{\delta t} = N/T$.
+From \eq{n-eval} we see that the number of frequencies evaluated is
+$N_f \propto T f_{max} = N f_{max} / \bar{\delta t}$.
+Holding constant the average survey cadence $\bar{\delta t}$ and individual
+observation properties (which control $f_{max}$), we find that the number of
+frequencies is simply proportional to the number of data points.
+The computation of the Lomb-scargle periodogram in \eq{lomb-scargle-periodogram}
+requires several sums over $N$ sinusoids at each of the
+$N_f$ frequencies, and thus we see that the
+na{\"i}ve scaling of the algorithm with the size of the dataset is
+$\mathcal{O}(N^2)$, holding observing properties constant.
+Due to the trigonometric functions involved, this turns out to be a rather
+``expensive'' $\mathcal{O}(N^2)$, which makes the straightforward
+implementation impractical for even modestly-sized datasets.
+
+Fortunately, several faster implementations have been proposed to compute the
+periodogram to arbitrary precision in $\mathcal{O}(N\log N)$ time.
+The first of these is discussed in \citet{Press89}, which uses an inverse
+interpolation operation along with a Fast Fourier Transform to compute the
+trigonometric components of \eq{lomb-scargle-periodogram} very efficiently over
+a large number of frequencies.
+A qualitatively similar approach is presented by \citet{Leroy2012},
+and makes use of the Non-equispaced Fast Fourier Transform
+\citep[NFFT, see][]{Keiner2009} to compute the Lomb-Scargle periodogram about
+a factor of 10 faster than the \citet{Press89} approach.
+All periodograms in this paper were computed using the Astropy
+implementation of the \citet{Press89} algorithm, which was adapted from Python
+code released to accompany \citet{VanderPlas2015}.
 
@@
 
-\subsection{Algorithmic Considerations}
+\begin{itemize}
+  \item $\chi^2$ normalization
+  \item PSD normalization
+\end{itemize}
 
 \section{Uncertainties in Periods}
+\sectlabel{uncertainties}
 
@@
   \item Peak Width and why it's a poor proxy
-  \item False Alarm Probability and why it fails
+  \item False Alarm Probability and its pitfalls
   \item Baluev method and why it's not enough
-  \item Motivation of Bootstrap
+  \item Motivation and example of Bootstrap
 \end{itemize}
@@
 \section{Generalizations \& Challenges}
+\sectlabel{generalizations}
 
@@
 
-\section{Acknowledgements}
-
-\citet{Astropy2013}
+\section{Conclusion and Summary}
+
+(Summarize key recommendations here)
 

### 2016-12-30 PracticalLombScargle.tex
Change tophat->rectangular

@@
 
-  \item[The Fourier transform of a tophat function is a sinc function.]
+  \item[The Fourier transform of a rectangular function is a sinc function.]
     (See \fig{fourier-pairs}{\it c})
@@
     \end{equation}
-    The tophat function, $\Pi_T(t)$, is a normalized symmetric function that
+    The rectangular function, $\Pi(t)$, is a normalized symmetric function that
     is uniform within a range given by $T$, and zero elsewhere:
@@
 Such an operation is commonly used, for example, in smoothing a function,
-as visualized in \Fig{convolution} for a tophat-shaped smoothing window.
+as visualized in \Fig{convolution} for a rectangular smoothing window.
 
@@
   \includegraphics[width=\textwidth]{fig04_Convolution_Diagram}
-  \caption{Visualization of a convolution between a continuous signal and a tophat smoothing kernel.
-    The normalized tophat-shaped window function slides across the domain (upper panel),
+  \caption{Visualization of a convolution between a continuous signal and a rectangular smoothing kernel.
+    The normalized rectangular window function slides across the domain (upper panel),
     such that at each point the mean of the values within the window are
@@
 amount of time, your measurement is a point-wise product between the true
-signal and a suitable tophat function describing your measurement process.
+signal and a suitable rectangular function describing your measurement process.
 If you have a continuous signal that you measure at regular intervals, your
@@
 
-\subsection{Effect of a Tophat Window}
+\subsection{Effect of a Rectangular Window}
 
 First, let's consider the case of observing a continuous periodic signal over
-a limited span of time, as illustrated in \Fig{tophat-window}.
+a limited span of time, as illustrated in \Fig{rectangular-window}.
 Here we consider a periodic function made up of three frequency components, observed within a 10-unit time window.
-The observed signal in this case can be thought of as a pointwise product of an infinite periodic signal, and a tophat window function.
+The observed signal in this case can be thought of as a pointwise product of an infinite periodic signal, and a rectangular window function.
 Using \eq{convolution-theorem-inverse}, we see that the Fourier transform of
@@
   \centering
-  \includegraphics[width=\textwidth]{fig06_Tophat_Window}
+  \includegraphics[width=\textwidth]{fig06_Rectangular_Window}
   \caption{Visualization of the effect on the Fourier transform of a
-    tophat-shaped observing window (i.e., a continuous signal observed
-    fully, but only within a finite range of time). The observed Fourier
+    rectangular observing window (i.e., a continuous signal observed in its
+    entirety within a finite range of time). The observed Fourier
     transform is a convolution of the true transform (here a series of Delta
     functions) and the window transform (here a narrow sinc function).
-    \figlabel{tophat-window}}
+    \figlabel{rectangular-window}}
 \end{figure}
@@
 to our data a rectangular window function of width $NT$.
-From the discussion accompanying \fig{tophat-window}, we know what this does:
+From the discussion accompanying \fig{rectangular-window}, we know what this does:
 it gives us a Fourier transform convolved with a sinc function of width
@@
 From our discussion of windowing in \sect{window-functions}
-(particularly \fig{tophat-window}), we know that data observed through a
+(particularly \fig{rectangular-window}), we know that data observed through a
 rectangular window of length $T$ will have sinc-shaped peaks of width

### 2016-12-30 PracticalLombScargle.tex
Small tweaks

@@
 straightforward in the more familiar case of uniformly-sampled data.
-For non-uniform data, there are two important choices to make: the
-{\bf frequency limits} and the {\bf grid spacing}.
-
-The frequency limit on the low end is relatively easy: for an observing window
-of length $T$, a signal with frequency $1/T$ will just complete a full cycle.
+For non-uniform data, it is not so simple, and there are two important
+considerations: the {\bf frequency limits} and the {\bf grid spacing}.
+
+The frequency limit on the low end is relatively easy: for a set of observations
+spanning a length of time $T$,
+a signal with frequency $1/T$ will complete exactly one oscillation cycle,
+and so this is a suitable minimum frequency.
 Often, it's more convenient just to set this minimum frequency to zero, as it
 doesn't add much of a computational burden and is unlikely to add any
-significant peak to the periodogram.
+significant spurious peak to the periodogram.
 
@@
 
-With the frequency range decided, we now must determine how finely to sample
-the space between the limits.
+With the frequency range decided, we next must determine how finely to sample
+the frequencies between the limits.
 This turns out to be quite important: if you choose too fine a grid, it can
@@
 
-To choose an appropriate periodogram grid, then, it is essential to take into
-account the expected widths of periodogram peaks, and construct a grid fine
-enough to assure that each peak will be covered by the grid.
+This shows us that it's important to choose grid spacings smaller than the
+expected widths of the periodogram peaks.
 From our discussion of windowing in \sect{window-functions}
@@
 rectangular window of length $T$ will have sinc-shaped peaks of width
-$\sim 1/T$.
+${\sim}1/T$.
 To ensure that our grid sufficiently covers each peak, it is prudent to
-over-sample by some factor---say $n_o$ samples per peak---and use a grid of size
-$\Delta f = \frac{1}{n_o T}$. This pushes the total number of required
-periodogram evaluations to
+over-sample by some factor---say $n_{over}$ samples per peak---and
+use a grid of size $\Delta f = \frac{1}{n_{over} T}$.
+This pushes the total number of required periodogram evaluations to
 \begin{equation}
@@
 \end{equation}
-The numerical result of this can vary greatly from dataset to dataset.
-
+
+Depending on the characteristics of the dataset, this number can vary greatly.
 For example, the Kepler data shown in \fig{kepler-data} has a pseudo-Nyquist
@@
 On the other hand, the LINEAR data shown in \fig{LINEAR-data} does not have
-any significant structure in its window function.
+any noteable aliasing structure in its window function.
 In this case the maximum detectable frequency is the Nyquist limit defined
@@
 
-Data from the upcoming LSST survey \citep{Ivezic08LSST} will fall somewhere
-in-between: for full frequency coverage of
+By comparison, data from the upcoming LSST survey \citep{Ivezic08LSST}
+will fall somewhere in-between: full frequency coverage of
 the 10-year data up to a limiting frequency defined by the 30-second
-integration time for each observation will require on order 2.5 million
+exposure time for each observation will require roughly 2.5 million
 periodogram evaluations per object, which for fast Lomb-Scargle implementations
@@
 visualize periodograms as a function of period rather than a function of
-frequency, the peak widths are not constant in period, and for this reason
-it is ideal to construct the grid itself in frequency space.
+frequency, the peak widths are not constant in period.
+Thus regular period grids tend to over-sample at large periods and under-sample
+at small periods;
+for this reason it is preferable to use a regular grid in frequency.
 
@@
   \item PSD normalization
+  \item Other normalizations \citep{Baluev2008}
 \end{itemize}

### 2017-01-04 PracticalLombScargle.tex
Add modifications section

@@
 {\it identical} to the result
-obtained by fitting a simple sinusoid to the data at each frequency $f$, and
+obtained by fitting a model consisting of a simple sinusoid to the data at
+each frequency $f$ and
 constructing a ``periodogram'' from the $\chi^2$ goodness-of-fit at each
@@
 the normal equations used in the least squares analysis.
-In other words, the Lomb-Scargle periodogram essentially {\it assumes a
-sinusoidal model} for the data, which is visualized in \fig{ls-model} for
-the data we had seen in \fig{LINEAR-data}.
+
+Partly due to the deep connection between Fourier analysis and least-squares
+analysis, the modified periodogram in \eq{lomb-scargle-periodogram}
+has since become commonly referred to as the {\it Lomb-Scargle Periodogram}.
+
+Because of this similarity between the classical and Lomb-Scargle periodograms,
+our discussion of windowing effects applies---at least qualitatively---to
+results computed with the Lomb-Scargle periodogram.
+In particular, reasoning about the effect of window functions on the observed
+Lomb-Scargle power spectrum remains useful even if it is true in only a
+qualitative sense.
+
+\section{Least-Squares Extensions to Lomb-Scargle}
+\sectlabel{lomb-scargle-extensions}
 
@@
 
-Partly due to the deep connection between Fourier analysis and least-squares
-analysis, the modified periodogram in \eq{lomb-scargle-periodogram}
-has since become commonly referred to as the {\it Lomb-Scargle Periodogram}.
-The Lomb-Scargle periodogram differs from the classical periodogram
-
-Because of this similarity between the classical and Lomb-Scargle periodograms,
-our discussion of windowing effects applies---at least qualitatively---to
-results computed with the Lomb-Scargle periodogram.
-In particular, reasoning about the effect of window functions on the observed
-Lomb-Scargle power spectrum remains useful even if it is true in only a
-qualitative sense.
+The equivalence of the Fourier interpretation and least squares interpretation of the Lomb-Scargle periodogram allows for some interesting and useful extensions to the periodogram.
+In the least squares interpretation of the periodogram, a sinusoidal model is proposed at each candidate frequency $f$:
+\begin{equation}
+  y_{model}(t;f) = A_f \sin(2 \pi f (t - \phi_f))
+\end{equation}
+where the amplitude $A_f$ and phase $\phi_f$ can vary as a function of frequency.
+This is fit to the data in the standard least-squares sense, by constructing
+the $\chi^2$ statistic at each frequency:
+\begin{equation}
+  \chi^2(f) \equiv \sum_n \big(y_n - y_{model}(t_n;f)\big)^2
+  \eqlabel{chi2-simple}
+\end{equation}
+We can find the ``best'' model $\hat{y}(t;f)$ by minimizing $\chi^2(f)$ at
+each freuqnecy with respect to $A_f$ and $\phi_f$;
+we will denote this minimum value as $\hat{\chi}^2(f)$.
+\citet{Scargle82} showed that with this setup, the Lomb-Scargle
+periodogram from \eq{lomb-scargle-periodogram} can be equivalently written:
+\begin{equation}
+  P(f) = \frac{N}{2}\big[\hat{\chi}^2_0 - \hat{\chi}^2(f)\big]
+  \eqlabel{lomb-scargle-chi2}
+\end{equation}
+where $\hat{\chi}^2_0$ is the reference model that assumes no variation in the data.
+The key realization here is that the Lomb-Scargle periodogram essentially
+{\it assumes a sinusoidal model} for the data; this is visualized in
+\fig{ls-model} for the data we had seen in \fig{LINEAR-data}.
+This immediately begs the question: can we compute a ``periodogram'' based
+on more general forms of $y(t;f)$ to more effectively fit the data?
+
+\subsection{Observational Noise}
+\label{extensions-observational-noise}
+Perhaps the most immediately obvious modification of the periodogram is to
+change not the form of the model $y_{model}$, but the $\chi^2$ statistic itself
+to account for observational errors in the data.
+If there are Gaussian errors $\sigma_n$ on each observed $y_n$, we can re-write \eq{chi2-simple} in the following (standard) way:
+\begin{equation}
+  \chi^2(f) \equiv \sum_n \left(\frac{y_n - y_{model}(t_n;f)}{\sigma_n}\right)^2
+  \eqlabel{chi2-with-errors}
+\end{equation}
+The periodogram constructed from this $\chi^2$ definition will more accurately
+reflect the spectral power of noisy observations.
+The effect of this modification on \eq{lomb-scargle-periodogram} is the addition
+of a multiplicative weight $1/\sigma_n$ within each of the summations.
+Early versions of this sort of modification appear in \citet{Gilliland87}
+and \citet{Irwin89}; \citet{Zechmeister09} showed that such a modification
+does not change the desirable statistical properties of the periodogram.
+
+\subsection{Data Centering and the Floating Mean Periodogram}
+\sectlabel{centering-data}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig20_standard_vs_floatingmean}
+  \caption{A comparison of the standard and floating mean periodograms for
+    data with a frequency of 0.3 and a selection effect which removes
+    faint observations.
+    In this case the mean estimated from the observed data is not close to
+    the true mean, which leads to the failure of the standard periodogram to
+    recover the correct frequency. A floating mean model correctly recovers
+    the true frequency of 0.3.
+    \figlabel{standard-vs-floatingmean}}
+\end{figure}
+
+We have noted that the periodogram explored in \citep{Lomb76} and
+\citep{Scargle82} is equivalent to computing the $\chi^2$ goodness-of-fit
+of a sinusoid at each frequency.
+Such a fit requires that the data be pre-centered; that is,
+\begin{equation}
+  \int_0^P g(t) dt = 0
+\end{equation}fitting a sinusoid to the data at each
+frrrequires the data to be pre-centered; i.e. the signal
+must be modified so that
+
+will only work correctly if the data is pre-centered so
+that a sinusoid with no offset will fit the data.
+In the past this was primarily accomplished by pre-centering the data using the
+mean estimated from the data points themselves.
+With full phase coverage of the observed signal, this approach is generally
+robust, but due to selection effects and survey cadence, full coverage can
+not always be guaranteed.
+This can potentially lead to situations where the periodogram entirely misses
+peaks of interest; an example of this is shown in \fig{standard-vs-floatingmean}.
+This simulates noisy observations of a sinusoidal signal situation in which the faintest parts of the light curve fall below the detection threshold.
+Using the standard periodogram, where the mean is estimated from the observations, leads to a periodogram that misses the true period of 0.3 days.
+The floating-mean periodogram, on the other hand, fits the mean as part of the model and more accurately characterizes the periodic content of the partially-observed data.
+
+\subsection{Fitting Higher-order Models}
+
+\citet{Bretthorst88}
+
+\todo{Use a multi-band periodogram on Kepler Data}
+
+\todo{Show poor performance on eclipsing data}
 
@@
 
-Suppose you have a set of $N$ observations over a time $T$ with an average
-cadence of $\bar{\delta t} = N/T$.
+Suppose you have a set of $N$ observations over a time-span $T$ for an average
+cadence of $\overline{\delta t} = N/T$.
 From \eq{n-eval} we see that the number of frequencies evaluated is
-$N_f \propto T f_{max} = N f_{max} / \bar{\delta t}$.
-Holding constant the average survey cadence $\bar{\delta t}$ and individual
+$N_f \propto T f_{max} = N f_{max} / \overline{\delta t}$.
+Holding constant the average survey cadence $\overline{\delta t}$ and individual
 observation properties (which control $f_{max}$), we find that the number of
-frequencies is simply proportional to the number of data points.
+frequencies is directly proportional to the number of data points.
 The computation of the Lomb-scargle periodogram in \eq{lomb-scargle-periodogram}
-requires several sums over $N$ sinusoids at each of the
+requires sums over $N$ sinusoids at each of the
 $N_f$ frequencies, and thus we see that the
 na{\"i}ve scaling of the algorithm with the size of the dataset is
-$\mathcal{O}(N^2)$, holding observing properties constant.
+$\mathcal{O}(N^2)$, when survey properties are held constant.
 Due to the trigonometric functions involved, this turns out to be a rather

### 2017-01-04 PracticalLombScargle.tex
Add multiterm figures

@@
 
-\section{Least-Squares Extensions to Lomb-Scargle}
+\section{The Least-Square Periodogram and its Extensions}
 \sectlabel{lomb-scargle-extensions}
 
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{fig18_ls_model}
-  \caption{The sinusoidal model implied by the Lomb-Scargle periodogram for
-    the LINEAR data seen in \fig{LINEAR-data}.
-    Although a sinusoid does not perfectly fit the data, it is close enough
-    that it serves to find the correct period without a problem.
-    \figlabel{ls-model}}
-\end{figure}
-
-The equivalence of the Fourier interpretation and least squares interpretation of the Lomb-Scargle periodogram allows for some interesting and useful extensions to the periodogram.
+The equivalence of the Fourier interpretation and least squares interpretation of the Lomb-Scargle periodogram allows for some interesting and useful
+extensions, which we will explore in this section.
+First, let's look at the Least Squares periodogram itself:
+
 In the least squares interpretation of the periodogram, a sinusoidal model is proposed at each candidate frequency $f$:
@@
 \end{equation}
-where $\hat{\chi}^2_0$ is the reference model that assumes no variation in the data.
+where $\hat{\chi}^2_0$ is the non-varying reference model.
 The key realization here is that the Lomb-Scargle periodogram essentially
@@
 This immediately begs the question: can we compute a ``periodogram'' based
-on more general forms of $y(t;f)$ to more effectively fit the data?
+on more general forms of the above model to more effectively fit the data?
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig18_ls_model}
+  \caption{The sinusoidal model implied by the Lomb-Scargle periodogram for
+    the LINEAR data seen in \fig{LINEAR-data}.
+    Although a sinusoid does not perfectly fit the data, the sinusoidal model
+    is close enough that it serves to locate the correct frequency.
+    \figlabel{ls-model}}
+\end{figure}
 
@@
 \label{extensions-observational-noise}
-Perhaps the most immediately obvious modification of the periodogram is to
-change not the form of the model $y_{model}$, but the $\chi^2$ statistic itself
-to account for observational errors in the data.
+Perhaps the most important modification of the periodogram is to
+construct a periodogram that correctly handles measurement error in the data.
+This involves changing the expression for the $\chi^2$ statistic in a manner
+analogous to the change for a regression problem.
 If there are Gaussian errors $\sigma_n$ on each observed $y_n$, we can re-write \eq{chi2-simple} in the following (standard) way:
@@
 of a multiplicative weight $1/\sigma_n$ within each of the summations.
-Early versions of this sort of modification appear in \citet{Gilliland87}
+Early versions of this sort of modification appeared in \citet{Gilliland87}
 and \citet{Irwin89}; \citet{Zechmeister09} showed that such a modification
-does not change the desirable statistical properties of the periodogram.
+does not change the desirable statistical properties of the resulting
+periodogram.
 
@@
 \sectlabel{centering-data}
+
+Another commonly-applied modification of the periodogram has variously been
+called the {\it Date-compensated Discrete Fourier Transform} \citep{Ferraz-Mello81}, the {\it floating-mean periodogram} \citep{Cumming99}, and the {\it generalized Lomb-Scargle Method} \citep{Zechmeister09}, and involves adding an offset term to the sinusoidal model at each frequency:
+\begin{equation}
+  y_{model}(t;f) = y_{0, f} + A_f \sin(2 \pi f (t - \phi_f))
+  \eqlabel{floating-mean-model}
+\end{equation}
+This is important, because the standard Lomb-Scargle approach assumes that the
+data are pre-centered around the mean value of the signal.
+In the past this was primarily accomplished by pre-centering the data before
+the Lomb-Scargle fit using the mean estimated from the data points themselves.
+
+With full phase coverage of the observed signal, this approach is generally
+robust. However, due to selection effects and survey cadence, full phase
+coverage can not always be guaranteed, which can potentially lead to situations
+where the periodogram entirely misses peaks of interest.
+\Fig{standard-vs-floatingmean} shows a simulated example of this: the data
+consist of noisy observations of a sinusoidal signal in which the faintest
+observations are omitted from the dataset (due to, {e.g.} a detection threshold).
+Applying the standard Lomb-Scargle periodogram to pre-centered data leads to a
+periodogram that misses the true period of 0.3 days (lower right panel).
+Using the floating-mean model of \eq{floating-mean-model} yields a periodogram
+that detects this true period (upper right panel).
 
@@
 
-We have noted that the periodogram explored in \citep{Lomb76} and
-\citep{Scargle82} is equivalent to computing the $\chi^2$ goodness-of-fit
-of a sinusoid at each frequency.
-Such a fit requires that the data be pre-centered; that is,
-\begin{equation}
-  \int_0^P g(t) dt = 0
-\end{equation}fitting a sinusoid to the data at each
-frrrequires the data to be pre-centered; i.e. the signal
-must be modified so that
-
-will only work correctly if the data is pre-centered so
-that a sinusoid with no offset will fit the data.
-In the past this was primarily accomplished by pre-centering the data using the
-mean estimated from the data points themselves.
-With full phase coverage of the observed signal, this approach is generally
-robust, but due to selection effects and survey cadence, full coverage can
-not always be guaranteed.
-This can potentially lead to situations where the periodogram entirely misses
-peaks of interest; an example of this is shown in \fig{standard-vs-floatingmean}.
-This simulates noisy observations of a sinusoidal signal situation in which the faintest parts of the light curve fall below the detection threshold.
-Using the standard periodogram, where the mean is estimated from the observations, leads to a periodogram that misses the true period of 0.3 days.
-The floating-mean periodogram, on the other hand, fits the mean as part of the model and more accurately characterizes the periodic content of the partially-observed data.
-
-\subsection{Fitting Higher-order Models}
-
-\citet{Bretthorst88}
-
-\todo{Use a multi-band periodogram on Kepler Data}
-
-\todo{Show poor performance on eclipsing data}
+\citet{Zechmeister09} provides a detailed study of this
+generalized/floating-mean periodogram, and shows that the beneficial
+statistical properties of the standard periodogram apply here as well;
+namely, periodogram power is independent and distributed according to a
+simple, known function.
+
+\subsection{Higher-order Fourier Models}
+
+A further generalization involves multi-term Fourier models at each frequency:
+rather than fitting just a single sinusoid at each frequency, we might fit a
+partial Fourier series at each frequency:
+\begin{equation}
+  y_{model}(t;f) = y^{0}_f + \sum_{k=1}^K A_f^{(k)} \sin(2\pi k f (t - \phi_f^{(k)})).
+\end{equation}
+This type of multi-term Lomb-Scargle periodogram was studied in detail by
+\citet{Bretthorst88}.
+
+This kind of multiterm generalization is quite tempting, because it means that
+the periodogram can fit models that are more complicated than
+
+\todo{Show performance on eclipsing data}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig21_kepler_multiterm}
+  \caption{Multi-term models (left) and their corresponding periodograms (right)
+    for the kepler data shown in \fig{kepler-data}.
+    \figlabel{kepler-multiterm}}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig22_LINEAR_multiterm}
+  \caption{Multi-term models (left) and their corresponding periodograms (right)
+    for the LINEAR data shown in \fig{LINEAR-data}
+    \figlabel{LINEAR-multiterm}}
+\end{figure}
 
@@
 
+\todo{Reference fast chi-square approach}
+
 \subsection{Normalization of the Periodogram}

### 2017-01-05 PracticalLombScargle.tex
Add eclipsing binary example

@@
 rather than fitting just a single sinusoid at each frequency, we might fit a
-partial Fourier series at each frequency:
+partial Fourier series at each frequency, adding $K$ additional components at
+integer multiples of the fundamental frequency:
 \begin{equation}
@@
 
-This kind of multiterm generalization is quite tempting, because it means that
-the periodogram can fit models that are more complicated than
-
-\todo{Show performance on eclipsing data}
+This kind of Fourier series generalization to the periodogram is quite
+tempting, because it means that the periodogram can be tuned to fit models
+that are more complicated than simple sine waves.
+\citet{Bretthorst88} takes this even further, exploring applications to
+decaying signals and other more complex models.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig23_binary_multiterm}
+  \caption{1-term and 6-term Lomb-Scargle models fit to LINEAR object
+    14752041, an eclipsing binary. Notice that the standard periodogram
+    (top panels) finds an alias of the true 17.5-hour frequency,
+    because a simple sinusoidal model cannot closely fit both the primary
+    and secondary eclipse. A six-term Fourier model (bottom panels), on
+    the other hand, does find the true period.
+    \figlabel{binary-multiterm}}
+\end{figure}
 
@@
 
-\section{Generalizations \& Challenges}
-\sectlabel{generalizations}
-
-\begin{itemize}
-\item floating-mean
-\item Uncertainties in Data
-\item Bayesian view is (almost always) not useful
-\item Multiterm makes the true period fit better, but also bumps the background noise.
-\end{itemize}
-
 \section{Conclusion and Summary}

### 2017-01-06 PracticalLombScargle.tex
Add multiterm discussion

@@
 ``False Alarm Probability''.
-In \sect{generalizations}, we consider and critique some generalizations of
-the Lomb-Scargle model, including multiterm periodograms and Bayesian
-approaches.
 
@@
 \end{equation}
-This type of multi-term Lomb-Scargle periodogram was studied in detail by
-\citet{Bretthorst88}.
-
-This kind of Fourier series generalization to the periodogram is quite
-tempting, because it means that the periodogram can be tuned to fit models
-that are more complicated than simple sine waves.
-\citet{Bretthorst88} takes this even further, exploring applications to
-decaying signals and other more complex models.
+\citet{Bretthorst88} takes a comprehensive look at this type of multi-term
+extension to the periodogram, as well as extensions to decaying signals and
+other more complex models.
 
@@
 
+This kind of Fourier series generalization to the periodogram is quite
+tempting, because it means that the periodogram can be tuned to fit models
+that are more complicated than simple sine waves.
+In some cases, this can be very useful; for example, \fig{binary-multiterm}
+shows a Lomb-Scargle analysis of an eclipsing binary star, characterized by
+both a primary and secondary eclipse.
+The standard Lomb-Scargle periodogram (upper panels) is maximized at twice the true rotation
+frequency, because the simple sinusoidal model is unable to closely fit the
+primary and secondary eclipses separately.
+A six-term Fourier model (lower panels) is sufficiently flexible that it can
+detect the true 17.5-hour period, though this comes at the expense of a much
+noisier periodogram.
+
+This background noise is easy to understand: in the least-squares view of the
+periodogram, the periodogram height at any frequency is directly related to
+how well the model fits the data.
+For Linear models, adding additional terms will {\it always} provide an
+equal or better fit to data than the simpler model, and so a periodogram based
+on a more complex model will be higher {\it at all frequencies}, not just
+at frequencies of interest!
+
 \begin{figure}[ht]
@@
   \caption{Multi-term models (left) and their corresponding periodograms (right)
-    for the kepler data shown in \fig{kepler-data}.
+    for the kepler data shown in \fig{kepler-data}
     \figlabel{kepler-multiterm}}
 \end{figure}
+
+This background noise can also be understood in terms of aliasing.
+Consider \fig{kepler-multiterm}, which shows several multi-term fits to the
+data from \fig{kepler-data}.
+Given a standard periodogram peak at $f_0$, a 2-term periodogram will have an
+additional peak at $f_0 / 2$, in which the original peak is the second
+harmonic. Similarly, a 3-term periodogram will add additional peaks both at
+$f_0 / 2$ and $f_0 / 3$, due to the original peak falling in the second and
+third harmonic.
 
@@
 \end{figure}
+
+In cases where the single-term power spectrum is itself very noisy, these
+additional aliases can quickly wash-out any gains from the more complicated
+model.
+Consider \fig{LINEAR-multiterm}, in which the $n=3$ model leads to very large
+noise at the low-frequency end. In this case, spurious periodogram peaks due
+to coincidences of such aliases may become problematic for even moderately
+noisy models.
 

### 2017-01-06 PracticalLombScargle.tex
Add section on extensions

@@
 All figures in this work were produced using Python, and in particular the
-Numpy \citep{numpy},
+Numpy \citep{numpy, numpybook},
 Pandas \citep{pandas},
@@
 unevenly-sampled data can probe frequencies far larger than any of these
-supposed limits (a fact that some of these citations do hint at
+supposed limits (a fact that several of these citations do hint at
 parenthetically).
@@
 there is a sense in which the Nyquist limit can be applied to unevenly-spaced
-data. \citet{Eyer99} explores this issue in some detail, and in particular
+data. \citet{Eyer99} explore this issue in some detail, and in particular
 proves the following:
@@
 This can be proven formally, but can be understood by thinking of such data as
-a windowed version of uniformly sampled data with spacing $p$.
-Such uniform data has a classical Nyquist limit of $1/(2p)$, and a window function applied to that data does not change that fact.
+a windowed version of uniformly sampled data with spacing $p$, where the window
+is zero at some observations.
+Such uniform data has a classical Nyquist limit of $1/(2p)$, and a window
+function applied to that data does not change that fact.
 
 \Fig{nyquist-eyer99} shows an example of such a Nyquist frequency.
-The data are non-uniformly sampled at times $t_i = n_i * p$, with $p=0.01$.
-This results in a Nyquist frequency $f_{Ny} = 50$, and we see the expected
+The data are non-uniformly sampled at times $t_i = n_i * p$, with $p=0.01$
+and $n_i$ drawn randomly from positive integers less than 10,000.
+According to the \citet{Eyer99} definition,
+this results in a Nyquist frequency $f_{Ny} = 50$, and we see the expected
 behavior beyond this frequency: the signal at $f > f_{Ny}$ consists of repeated
@@
 
-Keep in mind that the windowing limit $f_{max}$ is quite different than a
+Keep in mind that the windowing limit $1/(2\delta t)$ is quite different than a
 Nyquist limit: the Nyquist limit is the frequency beyond which all signal
 is aliased into the Nyquist range; the windowing limit is the frequency
-beyond which all signal is zero.
+beyond which all signal is zero or nearly zero.
 
@@
 taken near the same time of day: this is typical of a ground-based survey
-which observes only at night.
+with observations recorded only at night.
 The additional spikes at periods of $1/n$ days (for integer $n$) are aliases
@@
 This behavior is illustrated in \fig{LINEAR-window-effect}.
-The upper panel shows the window power spectrum as a function of frequency, while lower panel shows the observed signal power spectrum as a function of frequency.
-As noted in the discussion of \fig{LINEAR-window}, the window function shows a strong signal at frequencies corresponding to 1 day and 1 year; the latter can be seen in the inset plot.
-The power at both these scales is clearly imprinted on the observed power spectrum, as shown in the lower panel and its inset.
-This imperfect aliasing is similar to the perfect aliasing seen in regularly-sampled data at the Nyquist frequency; however, in this case the magnitude of the
+The upper panel shows the window power spectrum as a function of frequency (not period, as in \fig{LINEAR-window}), while lower panel shows the observed signal power spectrum as a function of frequency.
+The upper panel and its inset show clearly the 1-day and 1-year features we noted previously.
+The lower panel shows the observed power spectrum of the data: these diurnal and annual peaks in the window function are quite clearly imprinted on the observed power spectrum at relevant scales.
+This approximate aliasing is similar to the exact aliasing seen in regularly-sampled data at the Nyquist frequency; however, in this case the magnitude of the
 aliased signal fades further from the frequency driving the signal.
@@
     {\it lower left:} the time between observations.
-other than missing measurements, the spacing between observations is nearly uniform.
-the lower panel gives a closer look at the majority of the spacings, which are not exactly the same but rather span a range of $\pm 50$ms around the main period observed in the window function.
+Missing measurements aside, the spacing between observations is nearly uniform.
+the lower panel gives a closer look at the majority of the spacings, which are not exactly the same but rather span a range of $\pm 50$ms around the 29.4-minute period observed in the window function.
     {\it lower right:} the data power spectrum, showing approximate aliasing and clearly indicating a peak near a period of 13.6 hours, along with higher-order components at multiples of this value.
@@
 
-Space-based missions often have quite different observing windows.
+Space-based surveys will generally have quite different observing windows.
 For example, the upper-left panel of \fig{kepler-data}
@@
 periodic signals, its statistical properties are not as straightforward as in
-the uniform case.
-In this case, when we say an estimator with ``good statistical properties'',
-we mean that when it is applied to data with straightforward, uncorrelated
-noise, the result is a periodogram estimate in which the noise is also
-straightforward and uncorrelated.
+the uniform case: the periodogram has noise properties that are nontrivial
+to quantify, even when applied to ideal data (data with uncorrelated
+Gaussian noise).
+
+For the sake of evaluating the significance of periodogram peaks
+(see \sect{uncertainties}), it would be preferable if the non-equispaced
+periodogram, had uncertainties that were uncorrelated, and could be
+straightforwardly computed from the uncertainties of the data.
 
 \citet{Scargle82} showed that while the classical periodogram of
-\eq{classical-periodogram} has these desireable statistical properties
+\eq{classical-periodogram} has such desireable statistical properties
 in the case of uniform sampling, this does not hold for
@@
 \subsection{Data Centering and the Floating Mean Periodogram}
-\sectlabel{centering-data}
+\sectlabel{floating-mean}
 
 Another commonly-applied modification of the periodogram has variously been
-called the {\it Date-compensated Discrete Fourier Transform} \citep{Ferraz-Mello81}, the {\it floating-mean periodogram} \citep{Cumming99}, and the {\it generalized Lomb-Scargle Method} \citep{Zechmeister09}, and involves adding an offset term to the sinusoidal model at each frequency:
+called the {\it Date-compensated Discrete Fourier Transform} \citep{Ferraz-Mello81}, the {\it floating-mean periodogram} \citep{Cumming99}, and the {\it generalized Lomb-Scargle Method} \citep{Zechmeister09}, and involves adding an offset term to the sinusoidal model at each frequency\footnote{We choose to follow \citet{VanderPlas2015} and call this a ``Floating Mean'' model, to avoid confusion surrounding the different meanings of the term ``generalized periodogram'' in, {e.g.} \citet{Bretthorst88} and \citet{Zechmeister09}.}:
 \begin{equation}
@@
 \end{equation}
+
 This is important, because the standard Lomb-Scargle approach assumes that the
@@
 \subsection{Higher-order Fourier Models}
+\sectlabel{multiterm}
 
@@
 Consider \fig{kepler-multiterm}, which shows several multi-term fits to the
-data from \fig{kepler-data}.
+Kepler data from \fig{kepler-data}.
 Given a standard periodogram peak at $f_0$, a 2-term periodogram will have an
@@
 third harmonic.
+In general, you can expect an $N$-term periodogram to contain $N$ copies of
+every feature present in the standard periodogram, in addition to any
+new or strengthened features due to the more flexible model.
 
@@
 
-In cases where the single-term power spectrum is itself very noisy, these
-additional aliases can quickly wash-out any gains from the more complicated
-model.
-Consider \fig{LINEAR-multiterm}, in which the $n=3$ model leads to very large
-noise at the low-frequency end. In this case, spurious periodogram peaks due
-to coincidences of such aliases may become problematic for even moderately
-noisy models.
+In cases where the single-term power spectrum is itself noisy and/or dominated
+by partial aliasing due to window effects, these additional aliases can
+quickly wash-out any gains from the more complicated model.
+For example, consider the periodogram in \fig{LINEAR-multiterm}: as previously
+discussed, this periodogram contains a number of aliases of the true peak due
+to the dominant 24-hour signal in the window function.
+Adding terms to this model only compounds this problem, increasing the number of
+spurious peaks at the low-frequency end.
+For data with even moderate levels of noise, coincidences of these peaks can
+easily lead to spurious signals dominating the true peak.
+
+\subsection{Additional Extensions}
+When the periodogram is viewed from the least-squares model-fitting perspective,
+there is no need to limit the analysis to sums of sinusoids.
+There have been some very interesting applications extending this type of
+analysis to more arbitrary models.
+For example, \citet{Bretthorst88} studies ``periodograms'' which measure
+decaying signals, non-stationary signals, multiple signals, and other models,
+and shows the effectiveness of such approaches in applications from
+medical imaging and astronomy.
+
+On the astronomy side, several examples of such analyses exist.
+For example, the Supersmoother approach to detecting periodicity
+involves fitting a flexible non-parametric smoothing
+function to the data at each frequency: the flexibility of the model leads
+to fewer issues of aliasing \citep{Reimann94}.
+Another approach is to use empirically-derived templates as a functional fit
+at each frequency; this has been employed effectively in \citet{Sesar2010,
+Sesar2013} and related work.
+Finally, there has been some exploration of extensions to the Lomb-Scargle
+periodogram for use with multi-band observations, using various forms of
+regularization to control model complexity \citep{VanderPlas2015, Long2016}.
+
+\subsection{A Note About ``Bayesian'' Periodograms}
 
@@
 a large number of frequencies.
-A qualitatively similar approach is presented by \citet{Leroy2012},
-and makes use of the Non-equispaced Fast Fourier Transform
+\citet{Zechmeister09} showed how this approach can be easily extended to
+the floating mean periodogram discussed in \sect{floating-mean}.
+
+A qualitatively similar approach to the \citet{Press89} algorithm
+is presented by \citet{Leroy2012}:
+It makes use of the Non-equispaced Fast Fourier Transform
 \citep[NFFT, see][]{Keiner2009} to compute the Lomb-Scargle periodogram about
 a factor of 10 faster than the \citet{Press89} approach.
+For the multi-term models discussed in \sect{multiterm}, \citet{Palmer09}
+presents an adaptation of the \citet{Press89} method that can compute the
+multi-term result in $\mathcal{O}(NK\log N)$, for $N$ data points and
+$K$ Fourier components.
+
 All periodograms in this paper were computed using the Astropy
-implementation of the \citet{Press89} algorithm, which was adapted from Python
-code released to accompany \citet{VanderPlas2015}.
-
-\todo{Reference fast chi-square approach}
+implementation of the \citet{Press89}, \citet{Zechmeister09}, and
+\citet{Palmer09} algorithms, which was adapted from Python code originally
+published by \citet{ICVG2014} and \citet{VanderPlas2015}.
+All code for this paper is available in the form of Jupyter notebooks on the
+author's GitHub repository\footnote{\url{http://github.com/jakevdp/PracticalLombScargle}}
 

### 2017-01-06 PracticalLombScargle.tex
Add initial Bayesian discussion

@@
 regularization to control model complexity \citep{VanderPlas2015, Long2016}.
+With many of these types of extensions, computational complexity quickly
+becomes an issue, because fast $\mathcal{O}(N\log N)$ approaches along the
+lines of \citet{Press89} are not available for more general functional forms
+(see \sect{algorithmic-considerations}).
 
 \subsection{A Note About ``Bayesian'' Periodograms}
+\sectlabel{bayesian-periodograms}
+The least squares view of the Lomb-Scargle periodogram creates a natural
+bridge, via maximum likelihood, to Bayesian approaches to periodograms.
+In fact, \citet{Jaynes87} showed that the standard form of the Lomb-Scargle
+periodogram can be derived directly from the principles of Bayesian Probability
+Theory outlined in his comprehensive treatment of the subject \citep{Jaynes03}.
+In the Bayesian view, the Lomb-Scargle periodogram is the optimal statistic
+for detecting a stationary sinusoidal signal in the presence of Gaussian
+noise.
+
+$P(f)$ from \eq{lomb-scargle-periodogram}
+
+\begin{equation}
+  p(f~|~D) \propto e^{-n \chi^2 / 2} \propto e^{P(f)}
+\end{equation}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig24_bayesian_periodogram}
+  \caption{Comparison of the Lomb-Scargle periodogram (upper right) and the
+    Bayesian posterior periodogram (lower right) for simulated data (left)
+    drawn from a simple sinusoid.
+    The Bayesian posterior is equal to the exponentiation of the unnormalized
+    Lomb-Scargle periodogram, and thus tends to suppress all but the largest
+    peak.
+    \figlabel{bayesian-periodogram}}
+\end{figure}
+
+Warnings:
+
+\citep{Mortier15}
+
+What works:
+
+Bayesian models with strong priors \citep{Bretthorst88}
+Flexible Bayesian models \citep{Gregory92, Kelly14}
+Gaussian process models \citep{Angus16}
 
@@
 
-\subsection{Normalization of the Periodogram}
+\subsection{Periodogram Normalization and Interpretation}
+\sectlabel{normalizations}
 

### 2017-01-09 PracticalLombScargle.tex
Add failure modes figure

@@
 
-$P(f)$ from \eq{lomb-scargle-periodogram}
-
-\begin{equation}
-  p(f~|~D) \propto e^{-n \chi^2 / 2} \propto e^{P(f)}
-\end{equation}
+For the standard model, the Bayesian periodogram is given by
+\begin{equation}
+  p(f~|~D) \propto e^{-N \chi^2 / 2} \propto e^{P_{LS}(f)}
+  \eqlabel{bayesian-periodogram}
+\end{equation}
+where $P_{LS}(f)$ is the Lomb-Scargle power from \eq{lomb-scargle-periodogram}.
+The effect of this exponentiation, as seen in \fig{bayesian-periodogram},
+is to suppress all peaks in relation to the largest one of the spectrum.
 
@@
 
+\subsection{Failure Modes}
+\sectlabel{failure-modes}
+
+When using the Lomb-Scargle periodogram in an observational pipeline, it is
+vital to keep in mind the failure modes of the periodogram approach, which
+are rooted in the aliasing and pseudo-aliasing effects discussed in
+\sect{window-functions}.
+Due to the interaction of the signal, the convolution due to the survey window,
+and noise in the data, it is quite common for the largest peak in the
+periodogram to correspond not to the true frequency, but some alias of that
+frequency.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig25_failure_modes}
+  \caption{Comparison of the true period and peak Lomb-Scargle period for
+    1000 simulated periodic light curves.
+    Each has 60 irregular observations over 180 nights,
+    with a cadence typical of ground-based surveys ({i.e.} showing a strong
+    diurnal window pattern similar to \fig{LINEAR-window}).
+    The Lomb-Scargle peak does not always coincide with the true period,
+    and there is noticeable structure among these failures.
+    Panels (a)-(d) show some of the failure modes we can expect due to this
+    window function; see the text for more discussion.
+    \figlabel{failure-modes}}
+\end{figure}
+
+\todo{Detailed discussion of \fig{failure-modes}
+\citep[see, e.g.][]{VanderPlas2015, Long2016}
+
 \subsection{Algorithmic Considerations}
@@
 All periodograms in this paper were computed using the Astropy
-implementation of the \citet{Press89}, \citet{Zechmeister09}, and
-\citet{Palmer09} algorithms, which was adapted from Python code originally
+implementations of the \citet{Press89}, \citet{Zechmeister09}, and
+\citet{Palmer09} algorithms, which were adapted from Python code originally
 published by \citet{ICVG2014} and \citet{VanderPlas2015}.
-All code for this paper is available in the form of Jupyter notebooks on the
-author's GitHub repository\footnote{\url{http://github.com/jakevdp/PracticalLombScargle}}
+All code for this paper is available in the form of Jupyter notebooks in the
+paper's GitHub repository\footnote{\url{http://github.com/jakevdp/PracticalLombScargle}}
 

### 2017-01-09 PracticalLombScargle.tex
Finish Bayesian section

@@
 This paper gives an introduction to the practical aspects of the use of Lomb-Scargle periodograms and related algorithms to detect periodic signals.
-Rather than a rigorous mathematical treatment, the goal of this paper is to build intuition about what assumptions are baked into any periodogram computation.
-etc. etc....
+Rather than a rigorous mathematical treatment, the goal of this paper is to build intuition about what assumptions are baked into any periodogram computation
+and to help develop the intuition required to interpret the periodogram
+correctly and use it effectively.
 \end{abstract}
@@
     Lomb-Scargle periodogram, and thus tends to suppress all but the largest
-    peak.
+    peak. The Bayesian approach can be useful, but is problematic if
+    not used carefully (see text for discussion).
     \figlabel{bayesian-periodogram}}
@@
 
-Warnings:
-
-\citep{Mortier15}
-
-What works:
-
-Bayesian models with strong priors \citep{Bretthorst88}
-Flexible Bayesian models \citep{Gregory92, Kelly14}
-Gaussian process models \citep{Angus16}
+The benefit of the Bayesian approach is that it allows more principled treatment
+of nuisance parameters related to noise in the data and choices of model,
+and Bayesian periodogram extensions have been used quite effectively in a
+variety of circumstances \citep[see, e.g.][]{Bretthorst88}.
+Additionally, the Bayesian result explicitly returns probability density as
+a function of frequency, and it is tempting to think of this as the
+``natural'' way to interpret the periodogram.
+
+The problem, however, is that the Bayesian periodogram does {\it not} return
+a simple probability of the data showing a particular frequency, but rather
+a probability of each frequency conditioned on a relatively strong assumption:
+that the data are drawn from a sinusoidal model.
+Now, the standard Lomb-Scargle periodogram also assumes a sinusoidal model,
+and is thus subject to the same criticism.
+But unlike the simple Bayesian approach, the standard periodogram affords
+interpretation in light of Fourier transforms, window functions, and the
+like: in most cases of interest, this turns out to be {\it vital} to the
+application and interpretation of the periodogram in most cases of interest
+(see \sect{failure-modes}).
+
+In other words, the Bayesian approach goes ``all-in'' on the least squares
+interpretation of the periodogram, exponentially suppressing the information
+that allows you to reason about the periodogram in light of its connection
+to Fourier analysis.
+In contrived cases with structureless window functions, exponential attenuation
+of side-peaks and side-lobes may seem appealing \citep[see, e.g.][]{Mortier15},
+but that appeal extends to the real world only in special cases.
+That is not to say that all Bayesian approaches to periodic analysis are
+similarly flawed; there have been many interesting studies that go beyond
+the simple sinusoidal model and use more complex and/or flexible models.
+Some examples are models based on Fourier
+extensions with strong priors \citep[e.g.][]{Bretthorst88}, flexible
+non-parametric functions \citep[e.g.][]{Gregory92}, constrained stochastic
+models \citep[e.g.][]{Kelly14}, and Gaussian process models \citep{Angus16}.
+Though these can often be more accurate and powerful than the Lomb-Scargle
+approach, they tend to be far more expensive computationally.
 
@@
 \subsection{Choosing a Frequency Grid}
+\sectlabel{frequency-grid}
 The choice of frequency grid for a Lomb-Scargle analysis is an important one
@@
 
-\todo{Detailed discussion of \fig{failure-modes}
-\citep[see, e.g.][]{VanderPlas2015, Long2016}
+\fig{failure-modes} demonstrates this for some simulated data.
+The data consist of 60 noisy observations of each object within a span of
+180 days, with each observation recorded within a few hours of midnight
+each night.
+The left panel shows the results: the peridogram peak coincides with the true
+period in under 50\% of cases, and the modes of failure lead to noticeable
+patterns in the resulting plot.
+Though these are simulated observations, the pattern seen here is typical of
+real observations: see, for example, \citet{VanderPlas2015}
+and \citet{Long2016} which show similar plots for RR Lyrae candidates from
+the Sloan Digital Sky Survey.
+
+These patterns can be understood in terms of the interaction between the
+window function and the underlying spectral power.
+As we discussed in \sect{window-functions},
+a nightly observation pattern---typical of ground-based surveys---leads to
+a window function with a strong diurnal component that causes each frequency
+signature $f_0$ to be partially aliased at $f_0 + n \delta f$, for integers
+$n$ and $\delta f = 1\ {\rm day}^{-1}$
+(Recall \figs{LINEAR-window}{LINEAR-window-effect}).
+This is the source of the failure modes depicted in panel (a) of
+\fig{failure-modes}; the periods shown are
+\begin{equation}
+  P_{obs} = \left(\frac{1}{P_{true}} + n\delta f\right)^{-1}
+\end{equation}
+and for this data, close to 36\% of the objects are mis-characterized this way.
+
+Another mode of failure is when the periodogram for an object with frequency
+$f_0$ isolates a higher harmonic of that fundamental frequency, such as $2f_0$.
+This can occur for signals that are not strictly sinusoidal, and this harmonic
+frequency peak is subject to the same windowing effects as above.
+Thus we can describe the failure modes depicted in panel (b) of \fig{failure-modes} with the following equation, for integers $m > 0$:
+\begin{equation}
+  P_{obs} = \left(\frac{m}{P_{true}} + n\delta f\right)^{-1}
+\end{equation}
+In \fig{failure-modes}, the $m=2$ harmonic and its aliases account for roughly
+15\% of the results.
+
+Panel (c) of \fig{failure-modes} shows the final pattern of failure in
+Lomb-Scarlge results.
+This is closely related to panels (a) and (b), but comes from the fact that
+the Lomb-Scargle periodogram at frequency $f_0$ contains contributions from
+signals at both $+f_0$ and $-f_0$.
+When diurnal aliasing leads to power at negative frequencies, this is reflected
+at the absolute value of that frequency, so it is more accurate to take the
+absolute value of the above formula
+\begin{equation}
+  P_{obs} = \left|\frac{m}{P_{true}} + n\delta f\right|^{-1}
+  \eqlabel{failure-modes}
+\end{equation}
+Panel (c) shows the 5\% of objects that fall along this failure mode for
+$m=1$, $n=-2$.
+After accounting for all these known sources of failure, only roughly
+1\% of points are misclassified in an ``unexplained'' way, as seen in panel (d).
+
+When applying a periodogram in practice, it is vital to take such effects into
+account, rather than blindly relying on the single periodogram peak as your
+best estimate of the period.
+Applying understanding of windowing and aliasing effects can help in detecting
+failures of the periodogram, but is no silver bullet.
+For an observed peak at $f_{peak}$ from a survey whose window has strong power
+at $\delta f$, something like the following should probably be employed:
+\begin{enumerate}
+  \item Check for a peak at $f_{peak}/m$ for at least $m \in \{2, 3\}$.
+    If a significant peak is found, then $f_{peak}$ is probably an order-$m$
+    harmonic of the true frequency.
+  \item Check for peaks at $|f_{peak} \pm n\delta f|$ for at least
+    $n \in \{1, 2\}$. If these aliases exist, then it is likely that you have
+    found a peak on the sequence of expected aliases---though keep in mind that
+    there is no way to know from the periodogram alone whether or not
+    this is the ``true'' peak!
+  \item For each of the top few of these aliases, fit a more complicated model
+    (such as a multi-term Fourier series, template-based fit, etc.) to select
+    among them.
+\end{enumerate}
+For noisy observations, this procedure cannot generally guarantee that you
+have found the correct peak, but it is far preferable than making the mistake
+of blindly trusting the highest peak in the periodogram!
+We will come back to the question of uncertainty in the periodogram result
+in \sect{uncertainties}
 
@@
 All code for this paper is available in the form of Jupyter notebooks in the
-paper's GitHub repository\footnote{\url{http://github.com/jakevdp/PracticalLombScargle}}
+paper's GitHub repository\footnote{\url{http://github.com/jakevdp/PracticalLombScargle}}.
 

### 2017-01-14 PracticalLombScargle.tex
Add some discussion of uncertainties

@@
 and Matplotlib \citep{matplotlib} packages.
-Code to reproduce these figures is available in IPython notebooks at \url{http://github.com/jakevdp/PracticalLombScargle/}.
+Code to reproduce these figures is available in Jupyter notebooks at
+\url{http://github.com/jakevdp/PracticalLombScargle/}.
 
@@
 data. \citet{Eyer99} explore this issue in some detail, and in particular
-proves the following:
+prove the following:
 \begin{quote}
@@
 an integer multiple of this factor.
-This can be proven formally, but can be understood by thinking of such data as
-a windowed version of uniformly sampled data with spacing $p$, where the window
-is zero at some observations.
+\citet{Eyer99} prove this formally, but the result can be understood
+by thinking of such data as a windowed version of uniformly sampled
+data with spacing $p$, where the window is zero at some observations.
 Such uniform data has a classical Nyquist limit of $1/(2p)$, and a window
@@
 analysis, the modified periodogram in \eq{lomb-scargle-periodogram}
-has since become commonly referred to as the {\it Lomb-Scargle Periodogram}.
+has since become commonly referred to as the {\it Lomb-Scargle Periodogram}
+\citep[though versions of this approach had been employed even earlier;
+  see, e.g.][]{Gottlieb75}.
 
@@
 \subsection{Observational Noise}
-\label{extensions-observational-noise}
+\sectlabel{extensions-observational-noise}
 Perhaps the most important modification of the periodogram is to
@@
 Another commonly-applied modification of the periodogram has variously been
-called the {\it Date-compensated Discrete Fourier Transform} \citep{Ferraz-Mello81}, the {\it floating-mean periodogram} \citep{Cumming99}, and the {\it generalized Lomb-Scargle Method} \citep{Zechmeister09}, and involves adding an offset term to the sinusoidal model at each frequency\footnote{We choose to follow \citet{VanderPlas2015} and call this a ``Floating Mean'' model, to avoid confusion surrounding the different meanings of the term ``generalized periodogram'' in, {e.g.} \citet{Bretthorst88} and \citet{Zechmeister09}.}:
-\begin{equation}
-  y_{model}(t;f) = y_{0, f} + A_f \sin(2 \pi f (t - \phi_f))
+called the {\it Date-compensated Discrete Fourier Transform} \citep{Ferraz-Mello81}, the {\it floating-mean periodogram} \citep{Cumming99}, and the {\it generalized Lomb-Scargle Method} \citep{Zechmeister09}, and involves adding an offset term to the sinusoidal model at each frequency\footnote{We choose to follow \citet{Cumming99} and \citet{VanderPlas2015} and call this a ``Floating Mean'' model, to avoid confusion surrounding the different uses of the term ``generalized periodogram'' in, {e.g.} \citet{Bretthorst88} and \citet{Zechmeister09}.}:
+\begin{equation}
+  y_{model}(t;f) = y_0(f) + A_f \sin(2 \pi f (t - \phi_f))
   \eqlabel{floating-mean-model}
 \end{equation}
-
-This is important, because the standard Lomb-Scargle approach assumes that the
-data are pre-centered around the mean value of the signal.
-In the past this was primarily accomplished by pre-centering the data before
-the Lomb-Scargle fit using the mean estimated from the data points themselves.
-
-With full phase coverage of the observed signal, this approach is generally
-robust. However, due to selection effects and survey cadence, full phase
+This turns out to be quite important in practice, because the standard
+Lomb-Scargle approach assumes that the
+data are pre-centered around the mean value of the (unknown) signal.
+In many analyses, this requirement is accomplished by pre-centering using
+the mean estimated from the data points themselves.
+
+When the data provide full phase coverage of the observed signal,
+this approach is generally robust.
+However, due to selection effects and survey cadence, full phase
 coverage can not always be guaranteed, which can potentially lead to situations
@@
 \begin{equation}
-  y_{model}(t;f) = y^{0}_f + \sum_{k=1}^K A_f^{(k)} \sin(2\pi k f (t - \phi_f^{(k)})).
+  y_{model}(t;f) = A_f^{0} + \sum_{k=1}^K A_f^{(k)} \sin(2\pi k f (t - \phi_f^{(k)})).
 \end{equation}
@@
 The least squares view of the Lomb-Scargle periodogram creates a natural
-bridge, via maximum likelihood, to Bayesian approaches to periodograms.
+bridge, via maximum likelihood, to Bayesian periodic analysis.
 In fact, \citet{Jaynes87} showed that the standard form of the Lomb-Scargle
-periodogram can be derived directly from the principles of Bayesian Probability
-Theory outlined in his comprehensive treatment of the subject \citep{Jaynes03}.
+periodogram can be derived directly from the principles of Bayesian probability
+theory outlined in his comprehensive treatment of the subject \citep{Jaynes03}.
 In the Bayesian view, the Lomb-Scargle periodogram is the optimal statistic
@@
 
-For the standard model, the Bayesian periodogram is given by
-\begin{equation}
-  p(f~|~D) \propto e^{-N \chi^2 / 2} \propto e^{P_{LS}(f)}
+For the standard model, the Bayesian periodogram is given by the posterior
+probability of frequency $f$ given the data $D$ and sinusoidal model $M$:
+\begin{equation}
+  p(f~|~D, M) \propto e^{-N \chi^2 / 2} \propto e^{P_{LS}(f)}
   \eqlabel{bayesian-periodogram}
@@
 The problem, however, is that the Bayesian periodogram does {\it not} return
-a simple probability of the data showing a particular frequency, but rather
+a simple probability that the data oscillates with a given frequency, but rather
 a probability of each frequency conditioned on a relatively strong assumption:
@@
 of side-peaks and side-lobes may seem appealing \citep[see, e.g.][]{Mortier15},
-but that appeal extends to the real world only in special cases.
-That is not to say that all Bayesian approaches to periodic analysis are
+but that appeal extends to the real world only in the most favorable of
+cases---{i.e.} high signal-to-noise measurements with a very well-behaved
+survey window.
+
+This is not to say that all Bayesian approaches to periodic analysis are
 similarly flawed; there have been many interesting studies that go beyond
@@
 Some examples are models based on Fourier
-extensions with strong priors \citep[e.g.][]{Bretthorst88}, flexible
-non-parametric functions \citep[e.g.][]{Gregory92}, constrained stochastic
-models \citep[e.g.][]{Kelly14}, and Gaussian process models \citep{Angus16}.
+extensions with strong priors \citep[e.g.][]{Bretthorst88},
+instrument-dependent parametric models \citep[e.g.][]{Angus16},
+flexible non-parametric functions \citep[e.g.][]{Gregory92}, and
+specially-designed stochastic models \citep[e.g.][]{Kelly14}.
 Though these can often be more accurate and powerful than the Lomb-Scargle
 approach, they tend to be far more expensive computationally.
+Bayesian approaches based on MCMC also tend to run into stability problems,
+particularly for multimodal or other complicated posterior distributions
+\citep[See, for example, the RR Lyrae discussion in][]{Kelly14}.
 
@@
     and there is noticeable structure among these failures.
-    Panels (a)-(d) show some of the failure modes we can expect due to this
-    window function; see the text for more discussion.
+    Panels (a)-(d) isolate some of the specific modes of failure that should
+    be expected for this kind of window function;
+    see the text for more discussion.
     \figlabel{failure-modes}}
@@
 This is the source of the failure modes depicted in panel (a) of
-\fig{failure-modes}; the periods shown are
-\begin{equation}
-  P_{obs} = \left(\frac{1}{P_{true}} + n\delta f\right)^{-1}
-\end{equation}
-and for this data, close to 36\% of the objects are mis-characterized this way.
+\fig{failure-modes}; in terms of periods the above expression becomes:
+\begin{equation}
+  P_{obs} = \left(\frac{1}{P_{true}} + n\ \delta f\right)^{-1}
+\end{equation}
+where, to be clear, $n$ is a positive or negative integer and $\delta f$ is
+a strong feature in the window function (here, 1 day).
+For this data, close to 36\% of the objects are mis-characterized along these
+failure modes.
 
@@
 $f_0$ isolates a higher harmonic of that fundamental frequency, such as $2f_0$.
-This can occur for signals that are not strictly sinusoidal, and this harmonic
+This may occur for signals that are not strictly sinusoidal, and this harmonic
 frequency peak is subject to the same windowing effects as above.
-Thus we can describe the failure modes depicted in panel (b) of \fig{failure-modes} with the following equation, for integers $m > 0$:
+Thus we can describe the failure modes depicted in panel (b) of \fig{failure-modes} with the following equation, for positive integers $m > 0$:
 \begin{equation}
@@
 Panel (c) of \fig{failure-modes} shows the final pattern of failure in
-Lomb-Scarlge results.
-This is closely related to panels (a) and (b), but comes from the fact that
+Lomb-Scarlge results, which has an opposite trend with true Period.
+This is closely related to the effects shown in panels (a) and (b),
+but comes from the fact that
 the Lomb-Scargle periodogram at frequency $f_0$ contains contributions from
@@
 After accounting for all these known sources of failure, only roughly
-1\% of points are misclassified in an ``unexplained'' way, as seen in panel (d).
+1\% of points are misclassified in an ``unexplained'' way, seen in panel (d).
 
@@
 We will come back to the question of uncertainty in the periodogram result
-in \sect{uncertainties}
+in \sect{uncertainties}.
 
@@
 from a given dataset, it is vital to have an efficient algorithm for evaluating
-the periodogram. Unfortunately, the periodogram itself is a relatively slow
-operation.
+the periodogram. We will see that the na{\"i}ve implementation of
+\eq{lomb-scargle-periodogram} scales poorly with the size of the data, but that
+fast alternatives are available.
 
@@
 observation properties (which control $f_{max}$), we find that the number of
-frequencies is directly proportional to the number of data points.
+frequencies required is directly proportional to the number of data points.
 The computation of the Lomb-scargle periodogram in \eq{lomb-scargle-periodogram}
@@
 published by \citet{ICVG2014} and \citet{VanderPlas2015}.
-All code for this paper is available in the form of Jupyter notebooks in the
+All code behind this paper is available in the form of Jupyter notebooks in the
 paper's GitHub repository\footnote{\url{http://github.com/jakevdp/PracticalLombScargle}}.
@@
 
-\begin{itemize}
-  \item $\chi^2$ normalization
-  \item PSD normalization
-  \item Other normalizations \citep{Baluev2008}
-\end{itemize}
-
-\section{Uncertainties in Periods}
+One point of confusion in periodogram implementations is the choice of
+normalization.
+Recall that there are two equally valid perspectives of what the periodogram
+measures: the Fourier view and the least-squares view.
+Each of these perspectives lends itself to a different approach to
+periodogram normalization.
+
+\subsubsection{PSD normalization}
+When considering the periodogram from the Fourier perspective,
+it is useful to normalize the periodogram such that in the special case
+of equal spaced data, it recovers the standard Fourier power spectrum.
+This is the normalization used in \eq{lomb-scargle-periodogram} and the
+equivalent least-squares expression seen in \eq{lomb-scargle-chi2}:
+\begin{equation}
+  P(f) = \frac{N}{2}\big[\hat{\chi}^2_0 - \hat{\chi}^2(f)\big]
+  \eqlabel{lomb-scargle-chi2-2}
+\end{equation}
+For equally-spaced data, this periodogram is identical to the standard
+definition based on the fast Fourier transform:
+\begin{equation}
+  P(f) = \frac{1}{N} \left| FFT(y_n) \right|^2
+\end{equation}
+in particular, this means that the units of the periodogram are $unit(y)^2$,
+and can be interpreted as squared-amplitudes of the Fourier component at
+each frequency.
+Note that this changes if data uncertainties are incorporated into
+the periodogram as in \sect{extensions-observational-noise},
+the periodogram in this normalization becomes unitless: it is essentially
+a measure of periodic content in signal-to-noise ratio
+rather than in signal itself.
+
+\subsubsection{Least-squares normalization}
+In the least-squares view of the periodogram, it is more common to normalize
+the periodogram so that the values lie between 0 and 1.
+When we express the periodogram as a function of $\chi^2$ model fits as in
+\eq{lomb-scargle-chi2-2}, it becomes clear that the periodogram has a
+mathematical maximum value: if the sinusoidal model {\it perfectly} fits
+the data at some frequency $f_0$, then $\hat{\chi}^2(f_0) = 0$ and the
+periodogram is maximized at a value of $N\hat{\chi}^2_0 / 2$.
+On the other end, it is mathematically impossible for a sinusoidal model to
+do worse than the simple constant, non-varying fit reflected in
+$\hat{\chi}^2_0$, and so the minimum value of \eq{lomb-scargle-chi2-2} is
+exactly zero.
+
+This suggests a different normalization of the periodogram, where the values
+are dimensionless and lie between 0 and 1:
+\begin{equation}
+  P_{norm}(f) = 1 - \frac{\hat{\chi}^2(f)}{\hat{\chi}^2_0}
+  \eqlabel{lomb-scargle-normalized}
+\end{equation}
+This ``normalized periodogram'' is unitless, and
+is directly proportional to the unnormalized
+(or PSD-normalized) periodogram in \eq{lomb-scargle-chi2-2}.
+
+The main practical consequence of the normalization choice is that the
+resulting periodogram has slightly different statistical properties, and
+this needs to be taken into account when computing uncertainties and false
+alarm probabilities.
+Other normalizations exist as well, but seem to be rarely used in practice.
+For a summary of these alternative normalizations and their statistical
+properties, refer to the introduction of \citet{Baluev2008}.
+
+\subsection{Uncertainties in Periodogram Results}
 \sectlabel{uncertainties}
-
-\begin{itemize}
-  \item Peak Width and why it's a poor proxy
-  \item False Alarm Probability and its pitfalls
-  \item Baluev method and why it's not enough
-  \item Motivation and example of Bootstrap
-\end{itemize}
+An important piece of reporting results from the Lomb-Scargle periodogram is
+the uncertainty of the resulting period estimate.
+In most areas of science we are used to being able to report uncertainties in
+terms of errorbars, {e.g.} ``the period is $16.3 \pm 0.6$ hours''.
+For periods derived from the Lomb-Scargle periodogram, however, uncertainties
+generally cannot be meaningfully expressed in this way:
+as we saw in the discussion of failure modes in \sect{failure-modes},
+the concern for periodograms is more often errors associated with
+false peaks and aliases, rather than with imprecision in peak location.
+
+\subsubsection{Peak Width and Frequency Precision}
+
+A periodic signal will be reflected in a periodogram by the presence of a
+peak of a certain height and width.
+The width of the peak is directly related to the precision with which that
+peak's frequency is measured; the height of the peak 
+is related to the significance of the detection.
+We have motivated from Fourier window discussions that the peak width scales
+with the observational baseline; the surprising result is that the peak width
+{\it does not depend} on either the number of observations or the noise
+in those observations!
+
+This can be proven mathematically \citep[see, e.g.][]{Bretthorst88, ICVG2014},
+but it is perhaps more intuitive to see this visually.
+\fig{peak-width-height} shows periodograms for simulated data with a fixed
+baseline $T=4$ units covering four cycles, with varying sample sizes and
+signal-to-noise ratios.
+In all cases, the width of the peaks (most clearly visible in one or another
+normalization) are essentially {\it identical}, regardless of the data
+quality or quantity!
+Evidently, peak width ({i.e.} frequency precision) is not a particularly
+useful measure of uncertainty in the detection of periods.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig26_peak_width_height}
+  \caption{The effect of the number of points $N$ and the signal-to-noise
+    ratio $S/N$ on the expected width and height of the periodogram peak.
+    Top panels show the normalized periodogram (\eq{lomb-scargle-normalized}),
+    while bottom panels show the PSD-normalized periodogram
+    (\eq{lomb-scargle-periodogram})
+    Somewhat surprisingly, neither the number of points nor the signal-to-noise
+    ratio affects the peak width; nor do they affect the peak height under
+    the appropriate normalization.
+    \figlabel{peak-width-height}}
+\end{figure}
+
+If you insist on reporting precision derived from a peak width
+in terms of an errorbar, the results can be pretty silly for observations
+with long baselines.
+For example, going by the peak width in \fig{LINEAR-power}, the
+periodogram reveals a period of $2.58024 \pm 0.00005$ hours.
+While this is an accurate characterization of the period precision
+{\it assuming that peak is the correct one}, it does not capture the
+much more relevant uncertainties demonstrated in \sect{failure-modes},
+in which we might find ourselves on the wrong peak entirely!
+This is why in general, peak widths and Gaussian error bars are not all that
+useful for reporting uncertainties in the context of periodogram analysis.
+
+\subsubsection{False Alarm Probability}
+
+A much more relevant quantity for uncertainty in periodogram results is the
+height of the peak, and in particular the height compared to the background
+noise.
+\fig{peak-width-height} indicates that this property {\it does} depend on
+both the number of observations and their signal-to-noise ratio: for the
+small-$N$/low signal-to-noise cases, the spurious peaks in the background
+become comparable to the true peak.
+In fact, as we saw in \sect{schuster-to-lomb-scargle}, the ability to
+analytically define and quantify the relationship between peak height
+and significance is one of the considerations that led to the standard
+form of the Lomb-Scargle periodogram.
+
+The typical approach to quantifying the significance of a peak is the {\it
+False Alarm Probability} (FAP), which measures the probability that a dataset
+with no signal would---due to coincidental alignment among the random
+errors---lead to a peak of a similar magnitude.
+\citet{Scargle82} showed that for Gaussian noise, the values of the unnormalized
+periodogram in \eq{lomb-scargle-periodogram} are exponentially distributed;
+that is, for {\it one single frequency} $f_0$, if $Z = P(f_0)$ is the
+periodogram value from \eq{lomb-scargle-periodogram}, then
+\begin{equation}
+  P_{single}(Z) = 1 - \exp(-Z)
+  \eqlabel{LS-prob-z}
+\end{equation}
+is the cumulative probability of observing a periodogram value less than $Z$,
+in data with no periodic signal
+(be aware that for different periodogram normalizations,
+the form of this distribution changes; see \citet{Cumming99} or
+\citet{Baluev2008} for a summary for various normalizations).
+
+\subsubsubsection{Independent Frequency Method}
+
+But we are generally not interested in the distribution of one particular
+frequency, but rather the distribution of {\it the highest peak} of the
+periodogram, which is a quite different situation.
+If this is confusing, consider rolling a standard 6-sided die.
+The probability, in a single roll, of rolling a number, say $r > 4$ is easy
+to compute: it's 2 sides out of six, or $p(r>4) = 1/3$.
+If, on the other hand, you roll 10 dice and choose the largest number among
+them, the probability that it will be greater than 4 is {\it far} larger
+than $1/3$; it is $p(max(r,10) > 4) = 1 - (1 - 1/3)^{10} \approx 0.98$.
+The case for the periodogram is analogous: the chance of getting a high peak
+at any random location (\eq{LS-prob-z}) is relatively small, but if you search
+among a number of large frequencies it is much larger.
+
+With the dice, the calculation is easy because the number of independent
+rolls is clear.
+With the periodogram, though, the value at one frequency is correlated with
+the value at other frequencies in a way that is quite difficult to analytically
+quantify---these correlations come from the convolution with the survey
+window.
+Nonetheless, one common approach to estimating the distribution has been to
+assume it can be modeled on some ``effective number'' of independent
+frequencies $N_{eff}$, so that the FAP can be estimated as
+\begin{equation}
+  FAP(z) \approx 1 - \big(P_{single}(z)\big)^{N_{eff}}
+  \eqlabel{FAP-neff}
+\end{equation}
+A very simple estimate for $N_{eff}$ is based on our arguments of the
+expected peak width, $\delta f = 1/T$.
+In this approximation, the number of independent peaks in a range
+$0 \le f \le f_{max}$ is assumed to be $N_{eff} = f_{max} T$.
+There have been various attempts to estimate this value more carefully,
+both analytically and via simulations
+\citep[see, e.g.][]{Horne86,Schwarzenberg-Czerny98,Cumming04,Fresura2008},
+but all such approaches are necessarily only approximations.
+
+\subsubsubsection{Baluev (2008) Method}
+
+A more sophisticated treatment of the problem is that of \citet{Baluev2008},
+who derived an analytic result based on the theory of stochastic processes.
+The Baluev result is not an exact measure of the FAP, but rather an upper
+limit that is shown to be quite close to the true FAP even
+with highly structured window functons.
+For the PSD-normalized periodogram, \citet{Baluev08} showed that the following
+formula for the FAP provides a close upper-limit even in the case of
+highly structured window functions:
+\begin{equation}
+  FAP(z) \approx 1 - P_{single}(z)e^{-\tau(z)}
+  \eqlabel{FAP-baluev}
+\end{equation}
+where, for the PSD-normalized periodogram (\eq{lomb-scargle-periodogram}),
+\begin{equation}
+  \tau(z) = W e^{-z}\sqrt{z}
+\end{equation}
+and $W = f_{max} \sqrt{4\pi\ {\rm var}(t)}$ measures the effective width of the
+observing window in units of the maximum frequency chosen for the analysis.
+This result is not an exact measure of the FAP, but rather an approximation
+to the upper limit that is valid for alias-free periodograms, {i.e.} cases
+where the window function has very little structure.
+Nonetheless, \citet{Baluev08} showed that this approximation holds quite well
+even in the case of realistic survey windows.
+
+\subsubsubsection{Bootstrap Method}
+In the absence of a true analytic solution to the False Alarm Probability, we
+can turn to computational methods such as the bootstrap.
+The bootstrap method is a technique in which the statistic in question is
+computed repeatedly on many random resamplings of the data in order to
+approximate the distribution of that statistic
+\citep[see][for a useful general discussion of this technique]{ICVG2014}.
+For the periodogram, in each resampling we keep the temporal coordinates
+the same, draw observations randomly with replacement from the observed
+values, and then compute the maximum of the resulting periodogram.
+For enough resamplings, the distribution of these maxima will approximate
+the true distribution for the case with no periodic signal present.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig27_FAP_bootstrap}
+  \caption{
+    \figlabel{FAP-bootstrap}}
+\end{figure}
+
+\fig{FAP-bootstrap} compares
+
+A nice description of the use of the bootstrap to estimate false alarm
+probabilities in a periodogram is gien by \citet{Suveges12}.
+
+he correlations are the result of the convolution with the su; nonetheless, a common approach has been to assume that the
+distribution can be modeled based on some effective number of independent
+
+This strong correlation between frequencies means that the simple formul
+There have been various attempts to estimate this effective number of
+frequencies and/or quantify the distribution of the maximum peak,
+both analytically and via simulations
+\citep[see, e.g.][]{Horne86,Schwarzenberg-Czerny98,Cumming04,
+  Fresura2008,Baluev2008,Suveges12}.
+Perhaps the most sophisticated approach is that of \citet{Baluev2008},
+who derived an analytic result based on the theory of stochastic processes.
+The Baluev result is not an exact measure of the FAP, but rather an upper
+limit that is shown to be quite close to the true FAP even
+with highly structured window functons.
+
+%\begin{itemize}
+%  \item Peak Width and why it's a poor proxy
+%  \item False Alarm Probability and its pitfalls
+%  \item Baluev method and why it's not enough
+%  \item Motivation and example of Bootstrap
+%  \item Is my data periodic?
+%\end{itemize}
 

### 2017-01-16 PracticalLombScargle.tex
Finish false alarm probability section

@@
 both analytically and via simulations
-\citep[see, e.g.][]{Horne86,Schwarzenberg-Czerny98,Cumming04,Fresura2008},
+\citep[see, e.g.][]{Horne86,Schwarzenberg-Czerny98,Cumming04,Frescura08},
 but all such approaches are necessarily only approximations.
@@
 with highly structured window functons.
-For the PSD-normalized periodogram, \citet{Baluev08} showed that the following
+For the PSD-normalized periodogram, \citet{Baluev2008} showed that the following
 formula for the FAP provides a close upper-limit even in the case of
@@
 \end{equation}
-where, for the PSD-normalized periodogram (\eq{lomb-scargle-periodogram}),
-\begin{equation}
-  \tau(z) = W e^{-z}\sqrt{z}
+where, for the normalized periodogram (\eq{lomb-scargle-normalized}),
+\begin{equation}
+  \tau(z) \approx W (1 - z)^{(N - 4)/2}\sqrt{z}
 \end{equation}
@@
 where the window function has very little structure.
-Nonetheless, \citet{Baluev08} showed that this approximation holds quite well
+Nonetheless, \citet{Baluev2008} showed that this approximation holds quite well
 even in the case of realistic survey windows.
@@
 the true distribution for the case with no periodic signal present.
+The bootstrap produces the most robust estimate of the FAP because it makes
+no assumptions about statistical distributions, and fully accounts for survey
+window effects.
+
+Unfortunately, the computational costs of the bootstrap can be quite
+prohibitive: to accurately measure small levels of FAP requires constructing
+a large number of full periodograms.
+If you are probing a false positive rate of $r$ among $N$ bootstrap reseamples,
+you would roughly expect to find $rN \pm \sqrt{rN}$ peaks in your sample.
+More concretely, for 1,000 bootstrap samples, you'll find only $\sim 10\pm 3$
+peaks reflecting a 1\% false positive rate.
+The random fluctuations in that count translate directly to an inability to
+accurately estimate the false positive rate at that level.
+A good rule-of-thumb is that to accurately characterize a FPR $r$ requires
+something like $50 / r$ individual periodograms to be computed---this grows
+computationally intractable quite quickly.
 
@@
   \includegraphics[width=\textwidth]{fig27_FAP_bootstrap}
-  \caption{
+  \caption{Comparison of various approaches to computing the false alarm
+    probability (FAP) for simulated observations with both structured and
+    unstructured survey windows. Note that the approximate $N_{eff}$ method
+    in its most na{\"i}ve form tends to underestimate the bootstrapped FAP,
+    while the \citet{Baluev2008} method tends to overestimate the bootstrapped
+    FAP, particularly for data with a highly-structured survey window.
     \figlabel{FAP-bootstrap}}
@@
 
-\fig{FAP-bootstrap} compares
-
-A nice description of the use of the bootstrap to estimate false alarm
-probabilities in a periodogram is gien by \citet{Suveges12}.
-
-he correlations are the result of the convolution with the su; nonetheless, a common approach has been to assume that the
-distribution can be modeled based on some effective number of independent
-
-This strong correlation between frequencies means that the simple formul
-There have been various attempts to estimate this effective number of
-frequencies and/or quantify the distribution of the maximum peak,
-both analytically and via simulations
-\citep[see, e.g.][]{Horne86,Schwarzenberg-Czerny98,Cumming04,
-  Fresura2008,Baluev2008,Suveges12}.
-Perhaps the most sophisticated approach is that of \citet{Baluev2008},
-who derived an analytic result based on the theory of stochastic processes.
-The Baluev result is not an exact measure of the FAP, but rather an upper
-limit that is shown to be quite close to the true FAP even
-with highly structured window functons.
-
-%\begin{itemize}
-%  \item Peak Width and why it's a poor proxy
-%  \item False Alarm Probability and its pitfalls
-%  \item Baluev method and why it's not enough
-%  \item Motivation and example of Bootstrap
-%  \item Is my data periodic?
-%\end{itemize}
+\fig{FAP-bootstrap} compares these methods of estimating the false alarm
+probability, for both an unstructured survey window, and a structured window
+which produces the kinds of aliasing we have discussed above.
+The naive approximation in this case under-estimates the FAP at nearly all
+levels---so, for example, it might lead you to think a peak has a FAP of 10\%
+when in fact it is closer to 30\%.
+The Baluev method, by design, over-estimates the FAP, and is quite close to
+the bootstraped distribution in the case of an unstructured window.
+For a highly structured window, the Baluev method does not perform as well,
+but still tends to over-estimate the FAP---so, for example, it might lead you
+to think a peak as a FAP of 10\% when in fact it is closer to 5\%.
+
+Finally, I will note that \citet{Suveges12} has shown the promise of a
+hybrid approach of the bootstrap method and the extreme value statistics
+of \citet{Baluev2008}; this has the ability to compute accurate FAPs
+without the need to compute thousands of full bootstrapped periodograms.
+
+\subsubsubsection{What Does the False Alarm Probability Measure?}
+
+While the False Alarm Probability can be a useful concept, you should be
+careful to keep in mind that it answers a very specific question:
+{\it what is the probability that a signal with no periodic component would
+lead to a peak of this magnitude?}
+In other words, it is a measure of the false positive rate: the rate at which
+we would expect to incorrectly identify a periodic signal where none
+actually exists.
+It tells us nothing about the false negative rate; {i.e.} the rate at which
+we would expect to miss a periodic signal that is, in fact, present.
+It also tells us nothing about the error rate; {i.e.} the rate at which we
+would expect to identify an incorrect alias of a signal present in our data
+(as we saw in \sect{failure-modes}).
+Too often users are tempted to interpret the FAP too broadly, with the hope that
+it would speak to these questions as well.
+
+To the best of my knowledge, there is currently no silver bullet for answering
+these broader questions within the Lomb-Scargle approach.
+Perhaps the most fruitful path toward understanding of such effects for a
+particular survey---with particular noise characteristics and a particular
+observing window---is via simulated data.
+In fact, this type of simulation is a vital component of most current and
+future time-domain surveys that seek to catalog periodic objects
+\citep[see, e.g.][]{opsim1, opsim2, Ivezic08LSST}
+\todo{refs for SDSSII, Kepler, PanSTARRS, Gaia}.
 
@@
 
-(Summarize key recommendations here)
+\todo{Summarize key recommendations}
+
+\begin{itemize}
+\item always use the floating-mean model---it has many benefits with virtually zero cost
+\item always explore the window power first, so that you know what alias structures to expect
+\item the peak of the periodogram rarely tells the entire story. Always check
+  for aliases that are expected from window structure ($f = f_0 \pm n\delta f$)
+  and from non-sinusoidal signals ($f = f_0 / m$).
+\item avoid Bayesian methods unless you are very comfortable with the
+  assumptions they make about your signal
+\item avoid peak widths as a proxy for uncertainty unless you are quite
+  certain that your analysis has found the correct peak.
+\item avoid false alarm probabilities as a proxy for uncertainty
+  unless you make very clear {\it exactly} what kind of error they measure.
+  Estimate the FAP via bootstrap if computationally feasible, and otherwise
+  keep in mind the failures of available estimates.
+\end{itemize}
+
+\todo{wrap up with some witty \& pithy lines of encouragement.}
+
+\todo{acknowledge folks who helped with this.}
 

### 2017-01-31 PracticalLombScargle.tex
Add CLEAN discussion & some missing references

@@
 
-The left panel of \fig{LINEAR-power} shows the Lomb-Scargle periodogram computed for the data in \fig{LINEAR-data}.
+The left panel of \fig{LINEAR-power} and its inset show the Lomb-Scargle periodogram computed for the data in \fig{LINEAR-data}.
 Computing the Lomb-Scargle periodogram for the data gives us a measure of the
@@
 
-\todo{Add simulated RR Lyrae data and show the Shuster periodogram}
-
 \section{Non-uniform Sampling}
@@
 unique solution.
+
+\subsection{A Note on Window Deconvolution}
+
+There have been a few attempts in the literature to use procedural algorithms
+to solve this under-constrained deconvolution problem, perhaps most notably
+by adapting the iterative CLEAN algorithm developed for deconvolution in the
+context of radio astronomy \citep{Roberts87}.
+For cleaning of Lomb-Scargle periodograms, the CLEAN approach is hindered by
+three main deficiencies: first and most importantly, the CLEAN algorithm at each
+iteration assumes that the highest peak is the location of the primary signal;
+this is not always borne out for realistic observations of faint objects where
+the cleaning is most necessary (see the discussion in \sect{failure-modes}).
+Second, the convolution takes place at the level of the Fourier transform
+rather than the PSD; trying to clean a PSD directly ignores important phase
+information. Third, the CLEAN algorithm assumes a Schuster-type FFT analysis:
+while the Lomb-Scargle periodogram is equivalent to Schuster in the limit of
+equally-spaced observations, it is not equivalent in the relevant case of
+unequal observations (see \sect{schuster-to-lomb-scargle}),
+and so any attempt to apply CLEAN to Lomb-Scargle
+analysis directly is fundamentally flawed.
+
+My feeling is that there is room for a more principled approach to the
+unconstrained deconvolution problem for periodograms derived from non-uniform
+fast Fourier transforms (perhaps through some sort of regularization),
+but this does not appear to have been explored anywhere in the astronomy
+literature.
 
@@
 Too often users are tempted to interpret the FAP too broadly, with the hope that
-it would speak to these questions as well.
+it would speak to these additional questions as well.
 
@@
 particular survey---with particular noise characteristics and a particular
-observing window---is via simulated data.
-In fact, this type of simulation is a vital component of most current and
-future time-domain surveys that seek to catalog periodic objects
-\citep[see, e.g.][]{opsim1, opsim2, Ivezic08LSST}
-\todo{refs for SDSSII, Kepler, PanSTARRS, Gaia}.
+observing window---is via simulated data injected into the detection
+pipeline.
+In fact, this type of simulation is a vital component of most (if not all)
+current and future time-domain surveys that seek to catalog periodic objects,
+regardless of the particular approach used to detect periodicity
+\citep[see, e.g.][]{opsim1, opsim2, Ivezic08LSST, Sesar2010, Oluseyi12, McQuillan2012, Drake2014}
 

### 2017-02-01 PracticalLombScargle.tex
Minor edits + acknowledgments

@@
 \begin{abstract}
-This paper gives an introduction to the practical aspects of the use of Lomb-Scargle periodograms and related algorithms to detect periodic signals.
-Rather than a rigorous mathematical treatment, the goal of this paper is to build intuition about what assumptions are baked into any periodogram computation
-and to help develop the intuition required to interpret the periodogram
-correctly and use it effectively.
+This paper gives an introduction to the practical aspects of the use of
+Lomb-Scargle periodograms and related algorithms to detect and characterize 
+periodicity within a signal.
+Rather than a rigorous mathematical treatment, the goal of this paper is to
+build intuition about what assumptions are baked into any periodogram
+computation and to help develop the intuition required to interpret the
+periodogram correctly and use it effectively.
 \end{abstract}
@@
 in unevenly-sampled time-series, particularly within the astronomy community.
-For example, consider the data shown in \fig{LINEAR-data}.
-This is an irregularly-sampled timeseries showing a single object from the
+For example, consider the data shown in \fig{LINEAR-data}:
+this is an irregularly-sampled timeseries showing a single object from the
 LINEAR survey \citep{LINEAR1, LINEAR3}, with magnitude measured 280 times over
@@
   \url{http://github.com/jakevdp/PracticalLombScargle/}}.
-
 By eye, it is clear that the brightness of the object varies in time with a range spanning approximately 0.8 magnitudes, but what is not immediately clear is that this variation is periodic in time.
@@
   \item What is the source of the multiple peaks revealed by the Lomb-Scargle Periodogram?
-  \item What is the Nyquist frequency for unevenly-sampled data?
+  \item What is the largest frequency (i.e. Nyquist-like limit) that such an analysis is sensitive to?
   \item How should we choose the spacing of the frequency grid for our periodogram?
@@
 \end{enumerate}
-The answers to these sorts of questions are discussed in various textbooks and review papers, but I have not come across any single concise reference that gives a good intuition for how to think about such questions.
+Quantitative answers to these sorts of questions are discussed in various textbooks and review papers, but I have not come across any single concise reference that gives a good intuition for how to think about such questions.
 This paper seeks to fill that gap, and provide a practical, just-technical-enough guide to the effective use of the Lomb-Scargle method for periodic analysis.
@@
 
+\subsection{Outline}
 The paper is organized as follows:
@@
 In \sect{window-functions}, we use these properties to understand how different
-observation patterns (i.e. window functions) transform the power spectrum, and
-discuss a conceptual understanding of the well-known Nyquist limit.
-In \sect{non-uniform-sampling}, we consider non-uniformly sampled signals, and
-show that the Nyquist limit for this case is quite different than what many
-practitioners in the field seem to assume.
-In \sect{schuster-to-lomb-scargle}, we discuss the Lomb-Scargle periodogram, a
-modified version of the classical periodogram for unevenly-sampled data, as
-well as the motivation behind these modifications.
+observation patterns (i.e. window functions) affect the power spectrum, and
+discuss a conceptual understanding of the well-known Nyquist limiting frequency.
+In \sect{non-uniform-sampling}, we consider non-uniformly sampled signals as
+a special case of such windowing, and show that the Nyquist-like limit for
+this case is quite different than what many practitioners in the field
+assume.
+In \sect{schuster-to-lomb-scargle}, we discuss the properties of the
+Lomb-Scargle periodogram, a modified version of the classical periodogram
+for unevenly-sampled data, as well as the motivation behind these modifications.
+In \sect{lomb-scargle-extensions}, we discuss a complementary view of the
+periodogram as the result of least-squares model fitting, as well as some
+extensions that allows (including the ``Bayesian'' periodogram).
 In \sect{practical-considerations}, we discuss important practical
 considerations for the use of the Lomb-Scargle periodogram, including
-the choice of frequency grid and the various normalizations of the periodogram.
-In \sect{uncertainties}, we discuss how to quantify uncertainties in periods
-derived from the periodogram, including various notions of the
-``False Alarm Probability''.
-
-All figures in this work were produced using Python, and in particular the
+the choice of frequency grid, false alarm probabilities, and various modes
+of failure that should be expected when working with real-world data.
+Finally in \sect{conclusion}, we conclude and summarize the key recommendations
+for practical use of the Lomb-Scargle periodogram.
+
+\subsection{Code and Figures}
+All figures in this paper were produced using Python, and in particular the
 Numpy \citep{numpy, numpybook},
@@
 Consider a continuously-defined signal $g(t)$.
-Its Fourier transform is given by the following integral:
+Its Fourier transform is given by the following integral, where $i\equiv\sqrt{-1}$ denotes the imaginary unit:
 \begin{equation}
@@
      \end{eqnarray}
-     which follows from the linearity of the Fourier integral.
+     Both identities follow from the linearity of the Fourier integral.
 
@@
 of scales between a function and its Fourier transform:
-that is, a function with a characteristic scale $T$ will in general
-have a Fourier transform with characteristic scale of $1/T$.
-This feature will turn out to be very important as we push further in
-understanding the Lomb-Scargle Periodogram.
+a narrow function will have a broad transform, and vice versa.
+More quantitatively, a function with a characteristic scale $T$ will in
+general have a Fourier transform with characteristic scale of $1/T$.
+Such reciprocity comes up in many diverse applications of Fourier analysis,
+from wavelet transforms to music to quantum mechanics.
+For our purposes, this feature will turn out to be quite important as we
+push further in understanding the Lomb-Scargle Periodogram.
 
@@
 \begin{equation}
-  [f \ast g](t) = \int_{-\infty}^\infty f(\tau)g(t - \tau) d\tau
+  [f \ast g](t) \equiv \int_{-\infty}^\infty f(\tau)g(t - \tau) d\tau
   \eqlabel{convolution-definition}
@@
 From the definition, it is clear that a convolution amounts to ``sliding'' one
-function past the other.
+function past the other, integrating at each step.
 Such an operation is commonly used, for example, in smoothing a function,
@@
 \end{figure}
-When thinking about frequency components of time-domain measurements, this
-property of the Fourier transform becomes essential, as we will see.
-
 An important corrollary is that the Fourier transform of a product is a convolution of the two transforms:
@@
 \end{equation}
-We will explore the impact of this in the next section.
+We will see that these properties of the Fourier transform become essential
+when thinking about frequency components of time-domain measurements.
 
@@
 \sectlabel{window-functions}
+
+\todo{discuss data modes}
 
@@
 convolved with the transform of the obverving window}.
+Symbolically, if the signal is $g(t)$ and the window is $W(t)$, the observed
+function is
+\begin{equation}
+  g_{obs}(t) = g(t)W(t),
+\end{equation}
+and its Fourier transform is a convolution:
+\begin{equation}
+  \mathcal{F}\{g_{obs}\} = \mathcal{F}\{g\} \ast \mathcal{F}\{W\}.
+\end{equation}
 This has some interesting consequences, as we shall see.
@@
 a limited span of time, as illustrated in \Fig{rectangular-window}.
-Here we consider a periodic function made up of three frequency components, observed within a 10-unit time window.
+Here we consider a periodic function observed only within the window $[-3, 3]$.
 The observed signal in this case can be thought of as a pointwise product of an infinite periodic signal, and a rectangular window function.
@@
     rectangular observing window (i.e., a continuous signal observed in its
-    entirety within a finite range of time). The observed Fourier
+    entirety within a finite range of time). The function used here is
+    $g(t) = 1.2\sin(2\pi t) + 0.8\sin(4\pi t) + 0.4\sin(6\pi t) + 0.1\sin(8\pi t)$; The observed Fourier
     transform is a convolution of the true transform (here a series of Delta
-    functions) and the window transform (here a narrow sinc function).
+    functions indicating the component frequencies) and the window transform
+    (here a narrow sinc function).
     \figlabel{rectangular-window}}
@@
 have a situation similar to that in \fig{comb-window-2}.
-With a lower sampling rate, the signal transform no longer fits within the
-frequency window, and the true Fourier transform {\it cannot be recovered}
+A sparser sampling of the signal leads to a narrower sampling in frequency,
+and the signal transform no longer fits within the frequency window.
+The result is that the true Fourier transform {\it cannot be recovered}
 from the observed transform.
@@
 This implies that if we have a regularly-sampled function with a sampling
-rate of of $f_0 = 1/T$, we can only fully recover the freuqency information
+rate of $f_0 = 1/T$, we can only fully recover the freuqency information
 if the signal is {\it band-limited} between frequencies $\pm f_0/2$.
@@
 whose Fourier transform is zero outside the range $\pm B$,
-we must sample the data with a rate of at least $2B$.
+we must sample the data with a rate of at least $f_{Ny} = 2B$.
 
@@
     aliasing of signals, such that not all frequency information can be
-    recovered.
+    recovered. This is the reason for the famous Nyquist sampling theorem,
+    which conceptually says that only a function whose Fourier transform can
+    fit entirely between the ``teeth'' of the comb is able to be fully
+    recovered for regularly-spaced observations.
     \figlabel{comb-window-2}}
@@
 Suppose we have a true (infinitely long and continuous) signal $g(t)$, but
-we observe it only at a regular grid with spacing $T$. In this case, our
+we observe it only at a regular grid with spacing $\Delta t$. In this case, our
 observed signal is $g_{obs} = g(t) \III_T(t)$ and its Fourier transform is
 \begin{equation}
-  \hat{g}_{obs}(f) = \sum_{n=-\infty}^\infty g(nT) e^{-2\pi i f n T},
+  \hat{g}_{obs}(f) = \sum_{n=-\infty}^\infty g(n\Delta t) e^{-2\pi i f n \Delta t},
+  \eqlabel{DFT-f-inf}
 \end{equation}
@@
 We can choose the coordinate system appropriately and define
-$g_n \equiv g(nT)$ to write
-\begin{equation}
-  \hat{g}_{obs}(f) = \sum_{n=0}^N g_n e^{-2\pi i f n T}
+$g_n \equiv g(n\Delta t)$ to write
+\begin{equation}
+  \hat{g}_{obs}(f) = \sum_{n=0}^N g_n e^{-2\pi i f n \Delta t}
   \eqlabel{DFT-f}
@@
 From the arguments around Nyquist aliasing, we know that the only relevant
-frequency range is from $0 \le f \le 1/T$, and so we can define $N$
-evenly-spaced frequencies with $\Delta f = 1 / (NT)$ covering this range.
+frequency range is from $0 \le f \le 1/\Delta t$, and so we can define $N$
+evenly-spaced frequencies with $\Delta f = 1 / (N\Delta t)$ covering this range.
 Denoting the sampled transform as
@@
 from an infinite number of samples to a finite number of samples.
-In moving from $\hat{g}^{(T)}$ to $\hat{g}^{(N,T)}$, we have effectively applied
+In moving from \eq{DFT-f-inf} to \eq{DFT-f}, we have effectively applied
 to our data a rectangular window function of width $NT$.
@@
 
+\todo{mention difference between astronomy \& math treatment: astro is definitely {\it not} sampled as a well-defined random process...
+e.g.
+new observations made based on implications of previous observations.}
+
+\todo{how to compute empirical window functions}
+
 In the real world, particularly in fields like Astronomy where observations are
@@
     structure in its transform, causing the observed transform to be ``noisy''.
+    Here black and gray lines represent the real and imaginary parts of the
+    transform, respectively.
     \figlabel{random-window}}
@@
     Even with very dense sampling of the function, the Fourier transform
-    cannot be exactly recovered due to the 
-    These samples have the same average spacing as those in \fig{comb-window-1},
-    but the lack of structure in observing window translates to a lack of
-    structure in its transform, causing the observed transform to be a
-    ``noisy'' representation of the true transform
+    cannot be exactly recovered due to the imperfect aliasing present in
+    the window transform.
     \figlabel{random-window-2}}
 \end{figure}
+
+\subsection{A Note on Window Deconvolution}
 
@@
 
-\subsection{A Note on Window Deconvolution}
+\todo{Move this section?}
 
@@
 As we have seen in \sect{nyquist}, the Nyquist limit is a direct consequence
-of the window function created by evenly-sampled data,
+of the Dirac comb window function created by evenly-sampled data,
 and uneven sampling destroys the symmetry that underlies its definition.
 Nevertheless, the idea of the ``Nyquist frequency'' seems to have taken hold
-in the scientific psyche to the extent that it's often mis-applied in areas
-where it is utterly irrelevant.
-For unevenly-sampled data, the ``Nyquist limit'' might or might not exist,
-and even in cases where it does exist it tends to be far less relevant than
-in the evenly-sampled case.
+in the scientific psyche to the extent that the idea is often mis-applied in
+areas where it is mathematically irrelevant.
+For unevenly-sampled data, the truth is that the ``Nyquist limit'' might or
+might not exist, and even in cases where it does exist it tends to be far
+larger (and thus far less relevant) than in the evenly-sampled case.
 
@@
 
-As a simple example of where such logic fails,
+As a simple example of where such logic can fail spectacularly,
 consider the data from \fig{LINEAR-data}: though the mean sample
@@
 seconds, but it would be irresponsible to claim that this single pair of
-observations itself defines some limit beyond which frequency information is
-unattainable.
+observations by itself defines some limit beyond which frequency information
+is unattainable.
 
@@
     Nevertheless, the periodogram (lower right) clearly identifies the correct
-    period, though it's orders of magnitude larger than pseudo-Nyquist
+    period, though it is orders of magnitude larger than pseudo-Nyquist
     estimates based on average or minimum sampling rate.
@@
 this results in a Nyquist frequency $f_{Ny} = 50$, and we see the expected
-behavior beyond this frequency: the signal at $f > f_{Ny}$ consists of repeated
-aliases of the signal at $f < f_{Ny}$.
+behavior beyond this frequency: the signal at $f > f_{Ny}$ consists of a series
+of exact aliases of the signal at $f < f_{Ny}$.
 
@@
 To realize this situation in practice, however, 
-requires infinitely precise measurements of the
-times $t_i$; more often in the real-world, the Nyquist frequency will be
-defined by the precision of the time measurements.
-That is, if times are recorded to $D$ decimal places, the Nyquist frequency
-will be, at most
+would require infinitely precise measurements of the
+times $t_i$; finite precision of time measurements means the Nyquist
+frequency can be large, but not infinite.
+Quantitatively, if the times are recorded to $D$ decimal places,
+the Nyquist frequency will be at most
 \begin{equation}
@@
 \end{equation}
-as in \fig{nyquist-eyer99}.
+(cf. \fig{nyquist-eyer99}).
 In other words, absent other relevant patterns in the observations,
@@
 
-\subsubsection{Limit due to Windowing}
+\subsubsection{Frequency Limit due to Windowing}
 
@@
 beyond which all signal is zero or nearly zero.
+
+\subsubsection{Which Limiting Frequency is Useful?}
+
+\todo{Computational constraints; use of prior knowledge}
 
@@
 (see \sect{uncertainties}), it would be preferable if the non-equispaced
-periodogram, had uncertainties that were uncorrelated, and could be
+periodogram had a noise distribution that could be 
 straightforwardly computed from the uncertainties of the data.
+
+\todo{Mention requirement of uncorrelated noise, and correction for this explored by \citet{Vio2010}}
 
@@
 of a multiplicative weight $1/\sigma_n$ within each of the summations.
-Early versions of this sort of modification appeared in \citet{Gilliland87}
-and \citet{Irwin89}; \citet{Zechmeister09} showed that such a modification
+Early versions of this sort of modification appeared in \citet{Gilliland87},
+\citet{Scargle89}, and \citet{Irwin89};
+\citet{Zechmeister09} showed that such a modification
 does not change the desirable statistical properties of the resulting
@@
 coverage can not always be guaranteed, which can potentially lead to situations
-where the periodogram entirely misses peaks of interest.
+where the periodogram suppresses peaks of interest.
 \Fig{standard-vs-floatingmean} shows a simulated example of this: the data
@@
 
-\citet{Zechmeister09} provides a detailed study of this
+\citet{Zechmeister09} provide a detailed study of this
 generalized/floating-mean periodogram, and shows that the beneficial
 statistical properties of the standard periodogram apply here as well;
-namely, periodogram power is independent and distributed according to a
-simple, known function.
+namely, the periodogram noise at each frequency is given by the sum of
+two independent Gaussian random variables.
 
@@
 analysis to more arbitrary models.
-For example, \citet{Bretthorst88} studies ``periodograms'' which measure
+For example, \citet{Bretthorst88} studies generalized 
+Bayesian ``periodograms'' that can measure
 decaying signals, non-stationary signals, multiple signals, and other models,
@@
 medical imaging and astronomy.
-
-On the astronomy side, several examples of such analyses exist.
+The crux of such extensions is the fact that you often need to use some prior
+knowledge of your data to decide when it indicates a need for a more
+complicated model, and additionally what form of additional variation to add.
+In practice this comes up only when searching for very specific signals for
+which a more complicated model has some physical motivation.
+
+On the astronomy side, several examples of such extensions exist.
 For example, the Supersmoother approach to detecting periodicity
@@
 regularization to control model complexity \citep{VanderPlas2015, Long2016}.
-With many of these types of extensions, computational complexity quickly
-becomes an issue, because fast $\mathcal{O}(N\log N)$ approaches along the
-lines of \citet{Press89} are not available for more general functional forms
+With many of these least-squares and/or Bayesian extensions,
+computational complexity quickly becomes an issue, because fast
+$\mathcal{O}(N\log N)$ approaches along the lines of \citet{Press89}
+are not available for more general functional forms
 (see \sect{algorithmic-considerations}).
@@
 The problem, however, is that the Bayesian periodogram does {\it not} return
-a simple probability that the data oscillates with a given frequency, but rather
-a probability of each frequency conditioned on a relatively strong assumption:
+a simple probability that the data is periodic with a given frequency, but rather
+a probability at each frequency conditioned on a relatively strong assumption:
 that the data are drawn from a sinusoidal model.
@@
 
+\subsection{Computing the Window Power}
+\sectlabel{computing-window-power}
+The discussion of \sect{window-functions} highlighted the importance of
+examining the window function
+
+\todo{outline this}
+
 \subsection{Failure Modes}
@@
 Panel (c) of \fig{failure-modes} shows the final pattern of failure in
-Lomb-Scarlge results, which has an opposite trend with true Period.
+Lomb-Scargle results, which has an opposite trend with true period.
 This is closely related to the effects shown in panels (a) and (b),
@@
 \end{equation}
-Panel (c) shows the 5\% of objects that fall along this failure mode for
-$m=1$, $n=-2$.
-After accounting for all these known sources of failure, only roughly
-1\% of points are misclassified in an ``unexplained'' way, seen in panel (d).
+Panel (c) shows the 5\% of objects that fall along this reflected
+failure mode for $m=1$, $n=-2$.
+After accounting for all these known sources of periodogram failure,
+only roughly 1\% of points are misclassified in an ``unexplained'' way,
+seen in panel (d).
 
@@
 
-One point of confusion in periodogram implementations is the choice of
-normalization.
+One point of confusion in periodogram implementations is the interpretation
+of the periodogram values and their meaning.
 Recall that there are two equally valid perspectives of what the periodogram
 measures: the Fourier view and the least-squares view.
-Each of these perspectives lends itself to a different approach to
-periodogram normalization.
+Each of these suggests a different natural approach to normalizing and interpreting the periodogram.
 
@@
 \subsubsection{Least-squares normalization}
-In the least-squares view of the periodogram, it is more common to normalize
-the periodogram so that the values lie between 0 and 1.
+In the least-squares view of the periodogram, the periodogram is interpreted
+as an inverse measure of the goodness-of-fit for a model.
 When we express the periodogram as a function of $\chi^2$ model fits as in
@@
 
-The main practical consequence of the normalization choice is that the
-resulting periodogram has slightly different statistical properties, and
-this needs to be taken into account when computing uncertainties and false
-alarm probabilities.
+While the normalization does not affect the shape of the periodogram,
+its main practical consequence is that the statistics of the resulting
+periodogram differ slightly, and this needs to be taken into account when
+computing analytic estimates of uncertainties and false alarm probabilities.
 Other normalizations exist as well, but seem to be rarely used in practice.
-For a summary of these alternative normalizations and their statistical
+For a concise summary of periodogram normalizations and their statistical
 properties, refer to the introduction of \citet{Baluev2008}.
@@
 This is why in general, peak widths and Gaussian error bars are not all that
-useful for reporting uncertainties in the context of periodogram analysis.
+useful for reporting useful or accurate uncertainties in the context of
+periodogram analysis.
 
 \subsubsection{False Alarm Probability}
-
-A much more relevant quantity for uncertainty in periodogram results is the
-height of the peak, and in particular the height compared to the background
-noise.
+A much more relevant quantity for expressing uncertainty of periodogram results
+is the height of the peak, and in particular the height compared to the
+background ``noise'' displayed in the periodogram.
 \fig{peak-width-height} indicates that this property {\it does} depend on
@@
 
-But we are generally not interested in the distribution of one particular
-frequency, but rather the distribution of {\it the highest peak} of the
+We are generally not interested in the distribution of one particular
+randomly-chosen frequency,
+but rather the distribution of {\it the highest peak} of the
 periodogram, which is a quite different situation.
@@
 than $1/3$; it is $p(max(r,10) > 4) = 1 - (1 - 1/3)^{10} \approx 0.98$.
-The case for the periodogram is analogous: the chance of getting a high peak
-at any random location (\eq{LS-prob-z}) is relatively small, but if you search
-among a number of large frequencies it is much larger.
-
-With the dice, the calculation is easy because the number of independent
-rolls is clear.
+The case for the periodogram is analogous: the probability of seeing a high peak
+simply by chance at any random location (\eq{LS-prob-z}) is relatively small,
+but the probability of seeing a peak when scanning a large number of
+frequencies is much higher.
+
+With the dice, the calculation is easy because the rolls are independent: the
+result of one roll does not affect the result of the next.
 With the periodogram, though, the value at one frequency is correlated with
@@
 \begin{equation}
-  FAP(z) \approx 1 - \big(P_{single}(z)\big)^{N_{eff}}
+  FAP(z) \approx 1 - \big[P_{single}(z)\big]^{N_{eff}}
   \eqlabel{FAP-neff}
@@
 
-\subsubsubsection{Baluev (2008) Method}
+\subsubsubsection{\citet{Baluev2008} Method}
 
@@
 computationally intractable quite quickly.
+The bootstrap is also not universally applicable: for example, it does not
+correctly account for cases where noise in observations is correlated;
+for more discussion of the bootstrap approach and its caveats, see
+\citet{ICVG2014} and references therein.
 
@@
 \section{Conclusion and Summary}
+\sectlabel{conclusion}
+
+\subsection{Summary of Recommendations}
 
@@
 
+\subsection{Postscript: Why Lomb-Scargle?}
+
+\begin{itemize}
+  \item Advantages: equivalence to least squares, 
+  \item Myths: not faster than Schuster, does not correct for irregular sampling, frequencies are not independent, no savings over full model fit at each frequency \citep{ICVG2014}, straight Least Squares treatment not useful \sect{failure-modes}.
+\end{itemize}
+
 \todo{wrap up with some witty \& pithy lines of encouragement.}
 
-\todo{acknowledge folks who helped with this.}
+{\it Acknowledgments:} I am grateful to Jeff Scargle, Max Mahlke,
+Zeljko Ivezic, and Andy Connolly for helpful feedback on early drafts
+of this paper.
+This work was supported by the University of Washington eScience Institute,
+with funding from the Gordon and Betty Moore Foundation, the Alfred P. Sloan
+Foundation, and the Washington Research Foundation.
+
+% Edelson-Krolik + FFT -> Lomb-Scargle?
 

### 2017-02-01 PracticalLombScargle.tex
edits and rewordings

@@
 The Lomb-Scargle periodogram \citep{Lomb76, Scargle82}
-is a well-known algorithm for detecting periodicity
-in unevenly-sampled time-series, particularly within the astronomy community.
-For example, consider the data shown in \fig{LINEAR-data}:
+is a well-known algorithm for detecting periodicity in unevenly-sampled
+time-series, and has seen particularly wide use within the astronomy community.
+As an example of a typical application of this method,
+consider the data shown in \fig{LINEAR-data}:
 this is an irregularly-sampled timeseries showing a single object from the
@@
   \url{http://github.com/jakevdp/PracticalLombScargle/}}.
-By eye, it is clear that the brightness of the object varies in time with a range spanning approximately 0.8 magnitudes, but what is not immediately clear is that this variation is periodic in time.
-The Lomb-Scargle periodogram is a method that allows efficient computation of a Fourier-like power spectrum from such unevenly-sampled data, resulting in an intuitive means of determining the period of oscillation.
-
-The left panel of \fig{LINEAR-power} and its inset show the Lomb-Scargle periodogram computed for the data in \fig{LINEAR-data}.
+By eye, it is clear that the brightness of the object varies in time with a
+range spanning approximately 0.8 magnitudes, but what is not immediately clear
+is that this variation is periodic in time.
+The Lomb-Scargle periodogram is a method that allows efficient computation of
+a Fourier-like power spectrum from such unevenly-sampled data, resulting in
+an intuitive means of determining the period of oscillation.
+
+The left panel of \fig{LINEAR-power} and its inset show the Lomb-Scargle
+periodogram computed for the data in \fig{LINEAR-data}.
 Computing the Lomb-Scargle periodogram for the data gives us a measure of the
@@
 
-Often this is exactly how the Lomb-Scargle periodogram is presented: as a clean, well-defined procedure to detect the periodic component in an unevenly-sampled dataset.
-In practice, however, there are a number of subtle issues that must be considered when applying a Lomb-Scargle analysis to real-world datasets, and I have found that these are rarely presented together at an introductory level.
-Here are a few questions in particular that we might wish to ask about the results in \fig{LINEAR-power}:
+Often this is exactly how the Lomb-Scargle periodogram is presented: as a
+clean, well-defined procedure to detect the periodic component in an
+unevenly-sampled dataset.
+In practice, however, there are a number of subtle issues that must be
+considered when applying a Lomb-Scargle analysis to real-world datasets.
+Here are a few questions in particular that we might wish to ask about
+the results in \fig{LINEAR-power}:
 \begin{enumerate}
-  \item What is the source of the multiple peaks revealed by the Lomb-Scargle Periodogram?
-  \item What is the largest frequency (i.e. Nyquist-like limit) that such an analysis is sensitive to?
-  \item How should we choose the spacing of the frequency grid for our periodogram?
-  \item What assumptions is the Lomb-Scargle periodogram making about the signal?
-  \item How should we understand the uncertainty of the computed frequency?
+  \item What is the source of the pattern of multiple peaks revealed by the
+    Lomb-Scargle Periodogram?
+  \item What is the largest frequency (i.e. Nyquist-like limit) that such
+    an analysis is sensitive to?
+  \item How should we choose the spacing of the frequency grid for our
+    periodogram?
+  \item What assumptions is the Lomb-Scargle periodogram making about the
+    signal?
+  \item How should we understand the uncertainty of the peak frequency?
 \end{enumerate}
-Quantitative answers to these sorts of questions are discussed in various textbooks and review papers, but I have not come across any single concise reference that gives a good intuition for how to think about such questions.
-This paper seeks to fill that gap, and provide a practical, just-technical-enough guide to the effective use of the Lomb-Scargle method for periodic analysis.
-This paper does not seek a complete or rigorous treatment of the mathematics involved, but rather seeks to develop the intuition of {\it how} to think about the issues involved.
+Quantitative answers to these sorts of questions are discussed in various
+textbooks and review papers, but I have not come across any single concise
+reference that gives a good intuition for how to think about such questions.
+This paper seeks to fill that gap, and provide a practical,
+just-technical-enough guide to the effective use of the Lomb-Scargle method
+for detection and characterization of periodic signals.
+This paper does not seek a complete or rigorous treatment of the mathematics
+involved, but rather seeks to develop the intuition of {\it how} to think
+about these questions.
 
@@
      \end{eqnarray}
-     In other words, a sinusoidal signal with frequency $f_0$ has a Fourier transform consisting of a weighted sum of delta functions at $\pm f_0$.
+     In other words, a sinusoidal signal with frequency $f_0$ has a Fourier
+     transform consisting of a weighted sum of delta functions at $\pm f_0$.
 
@@
 These properties taken together make the Fourier transform quite useful for the study of periodic signals.
-The linearity of the transform means that any signal made up of a sum of periodic components will have a Fourier transform consisting of a sum of delta functions at those frequencies: that is, the Fourier transform is a {\it direct} measure of the periodic content of a function.
-
-Further, if we compute the squared amplitude of the resulting transform, we can both do away with complex components and remove the phase imparted by any time offsets; this squared amplitude is usually known as the {\it Power Spectrum}:
-\begin{equation}
-  \mathcal{P}\{g\} \equiv \left|\mathcal{F}\{g\}\right|^2
+The linearity of the transform means that any signal made up of a sum of
+sinusoidal components will have a Fourier transform consisting of a sum
+of delta functions marking the frequencies of those sinusoids: that is,
+the Fourier transform is a {\it direct} measure of the periodic
+content of a function.
+
+Further, if we compute the squared amplitude of the resulting transform, we
+can both do away with complex components and remove the phase imparted by any
+time offsets; this squared amplitude is usually known as the
+{\it Power Spectrum}:
+\begin{equation}
+  \mathcal{P}_g \equiv \left|\mathcal{F}\{g\}\right|^2
   \eqlabel{power-spectrum}
 \end{equation}
-The power spectrum of a function is a positive real-valued function of $f$ that quantifies the contribution of each frequency $f$ to the total signal.
+The power spectrum of a function is a positive real-valued function of the
+frequency $f$ that quantifies the contribution of each frequency $f$ to
+the total signal.
+Note that if $g$ is real-valued, it can be shown that
+$\mathcal{P}_g(f) = \mathcal{P}_g(-f)$.
 
@@
 general have a Fourier transform with characteristic scale of $1/T$.
-Such reciprocity comes up in many diverse applications of Fourier analysis,
-from wavelet transforms to music to quantum mechanics.
-For our purposes, this feature will turn out to be quite important as we
-push further in understanding the Lomb-Scargle Periodogram.
+Such reciprocity is central to many diverse applications of Fourier analysis,
+from electrodynamics to music theory to quantum mechanics.
+For our purposes, this property of the transform will turn out to be quite
+important as we push further in understanding the Lomb-Scargle Periodogram.
 
@@
 
-Given this definition of a convolution, it can be shown that the Fourier transform of a convolution is the point-wise product of the Fourier transforms:
+Given this definition of a convolution, it can be shown that the Fourier transform of a convolution is the point-wise product of the Fourier transforms of
+the two functions:
 \begin{equation}
@@
 
-\todo{discuss data modes}
-
-In a typical observational setting, we will be observing some small part of
-a larger time-varying signal.
-For example, if you have a very long signal that you measure for a finite
-amount of time, your measurement is a point-wise product between the true
-signal and a suitable rectangular function describing your measurement process.
-If you have a continuous signal that you measure at regular intervals, your
-measurement is a point-wise product between the true signal and an appropriate
-Dirac comb describing your measurement process.
-From what we learned about the convolution theorem in the previous section, it
-is clear that {\it the measured Fourier transform will be the true transform
-convolved with the transform of the obverving window}.
+Until now we have been discussing Fourier transforms of continuous signals
+which are well-defined for all times $-\infty < t < \infty$.
+Real-world temporal measurements of a signal, however, can only measure some
+finite span of time, at some finite sampling rate.
+In either case, the resulting data can be described by a point-wise product
+of the true underlying continuous signal with a window function describing
+the observation.
+For example, a continuous signal measured over a finite duration is described
+by a rectangular window function spanning the duration of the observation,
+and a signal measured at regular intervals is described by a Dirac comb window
+function marking those measurement times.\footnote{
+  Throughout this paper we are concerned only with the mode of measurement
+  common in optical astronomy; {i.e.} point observation. Other modes of
+  observation, such as time-tag-events, binned event data, or
+  time-to-spill events, are common in high-energy astronomy and other areas.
+  We will not consider such event-driven data modes here, but it should be
+  noted that there have been some interesting explorations of unified
+  statistical treatments of all of the above modes of observation
+  \citep[e.g.][]{Scargle98, Scargle2002}.
+}
+
+The Fourier transform we measure in these cases, then, is not the
+transform of the continuous underlying function, but rather a 
+{\it convolution of the true transform with the transform of the observing
+window}.
 Symbolically, if the signal is $g(t)$ and the window is $W(t)$, the observed
@@
 \end{equation}
-and its Fourier transform is a convolution:
+and its Fourier transform is a convolution of two transforms:
 \begin{equation}
@@
 \end{equation}
-This has some interesting consequences, as we shall see.
+In the general case, such a convolution is not an invertible operation, meaning
+that, except in special circumstances, it is not possible to exactly recover
+the transform of the underlying signal.
+This has some interesting consequences for the use and interpretation of
+periodograms, as we shell see.
 
@@
 a limited span of time, as illustrated in \Fig{rectangular-window}.
-Here we consider a periodic function observed only within the window $[-3, 3]$.
-The observed signal in this case can be thought of as a pointwise product of an infinite periodic signal, and a rectangular window function.
+Here we consider a periodic function observed only within the window $-3<t<3$.
+The observed signal in this case can be thought of as a pointwise product of
+the underlying infinite periodic signal, and a rectangular window function.
 Using \eq{convolution-theorem-inverse}, we see that the Fourier transform of
-this function is equivalent to the convolution of the Fourier transforms of each component, which are a set of delta functions and a sinc function, respectively.
-For a purely periodic signal, this convolution has the effect of broadening the delta functions representing the signal, turning each of them into a shifted sinc function.
-Because of the inverse relationship between the width of the window and the width of its transform (see \Fig{fourier-pairs}), it follows that a wider observing window leads to proportionally less spread in the Fourier transform of the observed function.
+this function is equivalent to the convolution of the Fourier transforms of
+each component, which are a set of delta functions and a sinc function,
+respectively.
+For a purely periodic signal, this convolution has the effect of broadening
+the delta functions representing the signal, turning each of them into a
+shifted sinc function.
+Because of the inverse relationship between the width of the window and the
+width of its transform (see \Fig{fourier-pairs}), it follows that a wider
+observing window leads to proportionally less spread in the Fourier
+transform of the observed function.
 
@@
 comb, the effect of such an observing window is to create a long sequence
-of aliases with a spacing of $1/T$.
+of aliases of the underlying transform with a spacing of $1/T$.
 With this in mind, we can be assured in this case that evaluating the
@@
 The example in \fig{comb-window-1} is somewhat of a best-case scenario, because
-the true Fourier transform values lie entirely in a range of width $1/T$.
+the true Fourier transform values are non-zero only within a range of width $1/T$.
 If we change the sampling rate so that this is no longer the case, we will
@@
 This implies that if we have a regularly-sampled function with a sampling
-rate of $f_0 = 1/T$, we can only fully recover the freuqency information
+rate of $f_0 = 1/T$, we can only fully recover the frequency information
 if the signal is {\it band-limited} between frequencies $\pm f_0/2$.
@@
 For regularly-spaced observations defined by a Dirac comb, the delta functions
-serve to collapse the Fourier integral into a sum over periodic components.
+serve to collapse the Fourier integral into a sum over periodic components,
+and in this way we can arrive at the common form of the discrete Fourier
+transform.
 Suppose we have a true (infinitely long and continuous) signal $g(t)$, but
@@
 In moving from \eq{DFT-f-inf} to \eq{DFT-f}, we have effectively applied
-to our data a rectangular window function of width $NT$.
-From the discussion accompanying \fig{rectangular-window}, we know what this does:
-it gives us a Fourier transform convolved with a sinc function of width
-$1 / (NT)$, resulting in the ``smearing'' of the Fourier transform signal
+to our data a rectangular window function of width $N\Delta t$.
+From the discussion accompanying \fig{rectangular-window}, we know what this
+does: it gives us a Fourier transform convolved with a sinc function of width
+$1 / (N\Delta t)$, resulting in the ``smearing'' of the Fourier transform signal
 with this width.
 Roughly speaking, then, any two Fourier transform values at frequencies within
-$1/(NT)$ of each other will not be independent, and so we should space our
-evaluations of the frequency with $\Delta f \ge 1/(NT)$.
+$1/(N\Delta t)$ of each other will not be independent, and so we should space
+our evaluations of the frequency with $\Delta f \ge 1/(N\Delta t)$.
 Comparing to above, we see that this is {\it exactly the frequency spacing}
@@
 \end{equation}
-Apart from the $1/N$ scaling, this is precisely the Fourier power defined
-in \eq{power-spectrum} for the expression in \eq{DFT-f}, and it follows that
-for a uniformly-sampled function $g_n$, the Schuster periodogram captures
+Apart from the $1/N$ scaling, this sum is precisely the Fourier power
+spectrum in \eq{power-spectrum}, computed for a continuous signal with
+uniform sampling.
+It follows that, in the uniform sampling case the Schuster periodogram captures
 all of the relevant frequency information present in the data.

### 2017-02-02 PracticalLombScargle.tex
Add section on windows and deconvolutions

@@
 
-\todo{how to compute empirical window functions}
-
-In the real world, particularly in fields like Astronomy where observations are
-subject to interruptions from weather and diurnal cycles, the sampling rate
-is generally far from uniform.
-Using the same approach as we used to explore uniform sampling in the previous
-section, we can now explore non-uniform sampling here.
-
-In the general non-uniform case, we measure some signal at a set of $N$ times
-which we will denote $\{t_n\}$.
-Our observing window function in this case is still a sum of delta functions,
-but at non-uniform locations.
-Mathematically, the observing window will look like this:
-
-\begin{equation}
-W(t; \{t_n\}) = \sum_{n=1}^{N} \delta(t - t_n)
-\end{equation}
-
-Applying this window to our true underlying signal $g(t)$, we find an observed
-signal of the form:
-
-\begin{eqnarray}
-  g_{obs}(t) &=& g(t) W(t; \{t_n\}) \nonumber\\
-             &=& \sum_{n=1}^{N} g(t_n)\delta(t - t_n)
-  \eqlabel{g-nonuniform}
-\end{eqnarray}
-
-Just as in the uniform case, the observed signal is a convolution of the true
-signal transform and the window transform; unlike in the uniform case, the
-window transform will generally {\it not} be a nice clean function.
-\Fig{random-window} shows the effect on the Fourier transform of a
-non-uniform observing cadence with a sampling rate identical to that
-in \fig{comb-window-1}.
-
-A few things stand-out in this figure. In particular, the Fourier transform of
-the non-uniformly spaced delta functions looks like random noise, and in some
-senses it is: the Fourier space representation reflects the frequency of
-observations, and so non-structured spacing of observations will lead to
-a non-structured Fourier representation of the window function.
-This non-structured window transform, when convolved with the Fourier transform
-of the true signal, results in an observed Fourier transform with the same
-lack of useful structure.
-Unlike the uniform result in \Fig{comb-window-1}, there is no direct aliasing
-of the true signal, and no way to exactly recover any portion of the true
-Fourier transform of this data.
-
 \begin{figure}[ht]
@@
 
+In the real world, particularly in fields like Astronomy where observations are
+subject to interruptions from weather and diurnal cycles, the sampling rate
+is generally far from uniform.
+Using the same approach as we used to explore uniform sampling in the previous
+section, we can now explore non-uniform sampling here.
+
+In the general non-uniform case, we measure some signal at a set of $N$ times
+which we will denote $\{t_n\}$.
+Our observing window function in this case is still a sum of delta functions,
+but at non-uniform locations.
+Mathematically, the observing window will look like this:
+
+\begin{equation}
+  W(t; \{t_n\}) = \sum_{n=1}^{N} \delta(t - t_n)
+  \eqlabel{nonuniform-window}
+\end{equation}
+
+Applying this window to our true underlying signal $g(t)$, we find an observed
+signal of the form:
+
+\begin{eqnarray}
+  g_{obs}(t) &=& g(t) W(t; \{t_n\}) \nonumber\\
+             &=& \sum_{n=1}^{N} g(t_n)\delta(t - t_n)
+  \eqlabel{g-nonuniform}
+\end{eqnarray}
+
+Just as in the evenly-sampled case, the Fourier transform of the observed
+signal is a convolution of the transforms of the true signal and the window:
+\begin{equation}
+  \mathcal{F}\{g_{obs}\} = \mathcal{F}\{g\} \ast \mathcal{F}\{W\}
+  \eqlabel{g-obs-conv}
+\end{equation}
+Unlike in the uniform case, the window transform $\mathcal{F}\{W\}$ will
+generally {\it not} be a nice clean function; from \eq{FT-def} and
+\eq{nonuniform-window}, we can write
+\begin{equation}
+  \mathcal{F}\{W\}(f) = \sum_{n=1}^{N} e^{-2\pi i f t_n}
+  \eqlabel{nonuniform-window-FT}
+\end{equation}
+For uniformly-spaced observing times $t_n$, the symmetry leads to cancellations
+so that the window transform is a clean Dirac comb.
+For non-uniform observing times, this symmetry is broken and the window
+transform will be much more ``nosiy''.
+This can be seen in \fig{random-window}, which shows the Fourier transform of
+a non-uniform observing window with an average sampling rate identical to
+that in \fig{comb-window-1}, along with its imact on the observed Fourier
+transform.
+
+A few things stand-out in this figure. In particular, the Fourier transform of
+the non-uniformly spaced delta functions looks like random noise, and in some
+senses it is: the locations and heights of the Fourier peaks are related to
+the intervals between observations, and so randomization of observation times
+leads to a randomization of Fourier peak locations and heights.
+That is, {\it non-structured spacing of observations will lead to
+a non-structured frequency peaks in the window transform}.
+This non-structured window transform, when convolved with the Fourier transform
+of the true signal, results in an observed Fourier transform with the same
+lack of useful structure.
+Comparing to the uniformly-spaced observations in \fig{comb-window-1}, we
+see that the unstructured nature of the window transform means that there
+is no exact aliasing of the true signal, and thus no way to exactly recover
+any portion of the true Fourier transform of this data.
+
 One might hope that sampling the signal more densely might limit these problems,
@@
 underlying signal, but still contains a degree of ``noise'' rooted in the
-unstructured nature of the spacing between observations.
+randomized frequency peaks due to randomized spacing between observations.
 
@@
 \end{figure}
-
-\subsection{A Note on Window Deconvolution}
-
-Since the observed transform is a well-defined convolution of the (unknown) true
-transform and the (known) window transform, you might hope that we could
-perform some sort of deconvolution operation to recover the result, but this
-turns out not to be straightforward, as deconvolution (particularly in this
-case where delta functions are involved) is an ill-posed problem without a
-unique solution.
-
-\todo{Move this section?}
-
-There have been a few attempts in the literature to use procedural algorithms
-to solve this under-constrained deconvolution problem, perhaps most notably
-by adapting the iterative CLEAN algorithm developed for deconvolution in the
-context of radio astronomy \citep{Roberts87}.
-For cleaning of Lomb-Scargle periodograms, the CLEAN approach is hindered by
-three main deficiencies: first and most importantly, the CLEAN algorithm at each
-iteration assumes that the highest peak is the location of the primary signal;
-this is not always borne out for realistic observations of faint objects where
-the cleaning is most necessary (see the discussion in \sect{failure-modes}).
-Second, the convolution takes place at the level of the Fourier transform
-rather than the PSD; trying to clean a PSD directly ignores important phase
-information. Third, the CLEAN algorithm assumes a Schuster-type FFT analysis:
-while the Lomb-Scargle periodogram is equivalent to Schuster in the limit of
-equally-spaced observations, it is not equivalent in the relevant case of
-unequal observations (see \sect{schuster-to-lomb-scargle}),
-and so any attempt to apply CLEAN to Lomb-Scargle
-analysis directly is fundamentally flawed.
-
-My feeling is that there is room for a more principled approach to the
-unconstrained deconvolution problem for periodograms derived from non-uniform
-fast Fourier transforms (perhaps through some sort of regularization),
-but this does not appear to have been explored anywhere in the astronomy
-literature.
 
@@
 number of observed samples gets very large.
+For a discussion of nonuniform window functions and attempts to correct
+for their effect on the periodogram, see \sect{windows-and-deconvolution}.
 
@@
 for this reason it is preferable to use a regular grid in frequency.
-
-\subsection{Computing the Window Power}
-\sectlabel{computing-window-power}
-The discussion of \sect{window-functions} highlighted the importance of
-examining the window function
-
-\todo{outline this}
 
@@
 
+\subsection{Window Functions and Deconvolution}
+\sectlabel{windows-and-deconvolution}
+
+Our discussion of window functions here and in \sect{window-functions}
+has highlighted the impact of structure in the survey window on the
+resulting observed power spectrum, and the importance of examining that
+window power to understand features of the resulting periodogram.
+The window power spectrum can be computed directly from the
+delta-function representation of the window
+function; recalling \eq{nonuniform-window-FT}, the window power spectrum
+for observation times $\{t_n\}$ is
+\begin{equation}
+  \mathcal{P}_W(f;\{t_n\}) = \left|\sum_{n=1}^{N} e^{-2\pi i f t_n}\right|^2
+  \eqlabel{window-power-analytic}
+\end{equation}
+Comparing this to \eq{schuster-periodogram}, we see that this is essentially
+the Schuster periodogram for data $g_n=1$ at all times $t_n$;
+with that in mind we can compute the Lomb-Scargle window power by computing
+a standard Lomb-Scargle periodogram on a series of unit measurements,
+making sure to {\it not} pre-center the data or use a floating mean model.
+This method is how window power spectra have been computed throughout this
+paper.
+
+Recall from \eq{g-nonuniform} that the observed data is a point-wise product
+of the underlying function and the survey window:
+\begin{equation}
+  g_{obs}(t) = g_{true}(t)W_{\{t_n\}}(t),
+\end{equation}
+and the convolution theorem tells us that the observed Fourier transform is
+a convolution of the true transform and the window transform:
+\begin{equation}
+  \mathcal{F}\{g_{obs}\} = \mathcal{F}\{g_{true}\} \ast \mathcal{F}\{W_{\{t_n\}}\}
+\end{equation}
+Given this direct relationship, we might hope to be able to invert this
+convolution to recover $\mathcal{F}\{g_{true}\}$ directly.
+For example, we could write:
+\begin{eqnarray}
+  \mathcal{F}\{g_{true}\} &=& \mathcal{F}\{g_{obs}/W_{\{t_n\}}\}\nonumber\\
+                   &=& \mathcal{F}\{g_{obs}\} \ast \mathcal{F}\{1/W_{\{t_n\}}\}
+\end{eqnarray}
+Because of the localization of observations, $W_{\{t_n\}}(t)$ is zero for most
+values of $t$ and such an inversion is not well-defined.
+More precisely, it can be shown that deconvolution in this context is an
+ill-posed problem without a unique solution, and so direct deconvolution
+is not an option.
+
+There have been a few attempts in the literature to use procedural algorithms
+to solve this under-constrained deconvolution problem, perhaps most notably
+by adapting the iterative CLEAN algorithm developed for deconvolution in the
+context of radio astronomy \citep{Roberts87}.
+For cleaning of Lomb-Scargle periodograms, the CLEAN approach is hindered by
+three main deficiencies: first and most importantly, the CLEAN algorithm at each
+iteration assumes that the highest peak is the location of the primary signal;
+this is not always borne out for realistic observations of faint objects where
+the cleaning is most necessary (see the discussion in \sect{failure-modes}).
+Second, the convolution takes place at the level of the Fourier transform
+rather than the PSD; trying to clean a PSD directly ignores important phase
+information. Third, the CLEAN algorithm assumes a Schuster-type FFT analysis:
+while the Lomb-Scargle periodogram is equivalent to Schuster in the limit of
+equally-spaced observations, it is not equivalent in the relevant case of
+unequal observations (see \sect{schuster-to-lomb-scargle}),
+and so any attempt to apply CLEAN to Lomb-Scargle
+analysis directly is fundamentally flawed.
+
+My feeling is that there is room for a more principled approach to the
+unconstrained deconvolution problem for periodograms derived from non-uniform
+fast Fourier transforms (perhaps through some sort of regularization),
+but this does not appear to have been explored anywhere in the astronomy
+literature.
+
 \subsection{Algorithmic Considerations}

### 2017-02-02 PracticalLombScargle.tex
reword CLEAN section

@@
 Mathematically, the observing window will look like this:
-
-\begin{equation}
-  W(t; \{t_n\}) = \sum_{n=1}^{N} \delta(t - t_n)
+\begin{equation}
+  W_{\{t_n\}}(t) = \sum_{n=1}^{N} \delta(t - t_n)
   \eqlabel{nonuniform-window}
 \end{equation}
-
 Applying this window to our true underlying signal $g(t)$, we find an observed
 signal of the form:
-
 \begin{eqnarray}
-  g_{obs}(t) &=& g(t) W(t; \{t_n\}) \nonumber\\
+  g_{obs}(t) &=& g(t) W_{\{t_n\}}(t) \nonumber\\
              &=& \sum_{n=1}^{N} g(t_n)\delta(t - t_n)
@@
 \begin{equation}
-  \mathcal{F}\{g_{obs}\} = \mathcal{F}\{g\} \ast \mathcal{F}\{W\}
+  \mathcal{F}\{g_{obs}\} = \mathcal{F}\{g\} \ast \mathcal{F}\{W_{\{t_n\}}\}
   \eqlabel{g-obs-conv}
 \end{equation}
-Unlike in the uniform case, the window transform $\mathcal{F}\{W\}$ will
-generally {\it not} be a nice clean function; from \eq{FT-def} and
+Unlike in the uniform case, the window transform $\mathcal{F}\{W_{\{t_n\}}\}$
+will generally {\it not} be a nice clean function; from \eq{FT-def} and
 \eq{nonuniform-window}, we can write
 \begin{equation}
-  \mathcal{F}\{W\}(f) = \sum_{n=1}^{N} e^{-2\pi i f t_n}
+  \mathcal{F}\{W_{\{t_n\}}\}(f) = \sum_{n=1}^{N} e^{-2\pi i f t_n}
   \eqlabel{nonuniform-window-FT}
@@
 window power to understand features of the resulting periodogram.
+
+\subsubsection{Computing the Window Function}
 The window power spectrum can be computed directly from the
@@
 the Schuster periodogram for data $g_n=1$ at all times $t_n$;
-with that in mind we can compute the Lomb-Scargle window power by computing
+with that in mind one quick way to compute the window power is to use
 a standard Lomb-Scargle periodogram on a series of unit measurements,
-making sure to {\it not} pre-center the data or use a floating mean model.
+making sure to {\it not} pre-center the data or to use a floating mean model.
 This method is how window power spectra have been computed throughout this
@@
 
+\subsubsection{Deconvolution and CLEANing}
+With this ability to compute the window function, we might hope to be able to
+use it {\it quantitatively} to remove spurious peaks from the observed
+power spectrum
 Recall from \eq{g-nonuniform} that the observed data is a point-wise product
@@
 \end{equation}
-Given this direct relationship, we might hope to be able to invert this
+Given this relationship, we might hope to be able to invert this
 convolution to recover $\mathcal{F}\{g_{true}\}$ directly.
@@
 Because of the localization of observations, $W_{\{t_n\}}(t)$ is zero for most
-values of $t$ and such an inversion is not well-defined.
-More precisely, it can be shown that deconvolution in this context is an
-ill-posed problem without a unique solution, and so direct deconvolution
-is not an option.
+values of $t$ and so $1/W_{\{t_n\}}(t)$ and its Fourier transform are not well
+defined. For this reason, direct deconvolution is not an option in most
+cases of interest---in other words, the deconvolution problem is ill-posed
+and does not have a unique solution.
 
@@
 this is not always borne out for realistic observations of faint objects where
-the cleaning is most necessary (see the discussion in \sect{failure-modes}).
+the cleaning is most necessary (recall the discussion in \sect{failure-modes}).
 Second, the convolution takes place at the level of the Fourier transform
@@
 unequal observations (see \sect{schuster-to-lomb-scargle}),
-and so any attempt to apply CLEAN to Lomb-Scargle
-analysis directly is fundamentally flawed.
-
+and so any attempt to apply CLEAN to Lomb-Scargle analysis directly would
+be fundamentally flawed even if it were not ill-posed to begin with.
+
+The latter two issues could be addressed by focusing on the non-uniform
+Fourier transform and the Schuster periodogram rather than the Lomb-Scargle
+approach, but one would lose the advantageous statistical properties of the
+resulting transform, as well as the correspondance with least-squares fitting
+that allows useful extensions to models handling a floating mean and
+observational noise.
+Even in this case, though, the first issue remains quite problematic.
 My feeling is that there is room for a more principled approach to the
 unconstrained deconvolution problem for periodograms derived from non-uniform
-fast Fourier transforms (perhaps through some sort of regularization),
-but this does not appear to have been explored anywhere in the astronomy
-literature.
+fast Fourier transforms, perhaps through some sort of L1/lasso regularization
+that imposes assumptions of sparsity on the true periodogram, but to date this 
+does not appear to have been explored anywhere in the literature.
 

### 2017-02-02 PracticalLombScargle.tex
some rewordings

@@
 The paper is organized as follows:
-In \sect{continuous-fourier-transform}, we review the continous Fourier
-transform and some of its useful properties, including defining the notion of
-a power spectrum for measuring periodic content in a signal.
-In \sect{window-functions}, we use these properties to understand how different
-observation patterns (i.e. window functions) affect the power spectrum, and
-discuss a conceptual understanding of the well-known Nyquist limiting frequency.
-In \sect{non-uniform-sampling}, we consider non-uniformly sampled signals as
-a special case of such windowing, and show that the Nyquist-like limit for
-this case is quite different than what many practitioners in the field
-assume.
-In \sect{schuster-to-lomb-scargle}, we discuss the properties of the
-Lomb-Scargle periodogram, a modified version of the classical periodogram
-for unevenly-sampled data, as well as the motivation behind these modifications.
-In \sect{lomb-scargle-extensions}, we discuss a complementary view of the
-periodogram as the result of least-squares model fitting, as well as some
-extensions that allows (including the ``Bayesian'' periodogram).
-In \sect{practical-considerations}, we discuss important practical
-considerations for the use of the Lomb-Scargle periodogram, including
-the choice of frequency grid, false alarm probabilities, and various modes
-of failure that should be expected when working with real-world data.
-Finally in \sect{conclusion}, we conclude and summarize the key recommendations
+\begin{description}
+  \item[\sect{introduction}] is this introduction.
+  \item[\sect{continuous-fourier-transform}] presents a review of the 
+    continous Fourier transform and some of its useful properties,
+    including defining the notion of a power spectrum for measuring
+    periodic content in a signal.
+  \item[\sect{window-functions}] builds on these properties to explore how
+    different observation patterns (i.e. window functions) affect the observed
+    power spectrum, and discusses a conceptual understanding of the
+    well-known Nyquist limiting frequency.
+  \item[\sect{non-uniform-sampling}] considers non-uniformly sampled signals as
+    a special case of an observational window, and shows that the Nyquist-like
+    limit for this case is quite different than what many practitioners in 
+    the field often assume.
+  \item[\sect{schuster-to-lomb-scargle}] details the Lomb-Scargle periodogram,
+    a modified version of the classical periodogram for unevenly-sampled data,
+    as well as the motivation behind these modifications.
+  \item[\sect{lomb-scargle-extensions}] discusses a complementary view of the
+    Lomb-Scargle periodogram as the result of least-squares model fitting,
+    as well as some extensions enabled by that viewpoint
+    (including various ``Bayesian'' periodograms).
+  \item[\sect{practical-considerations}] utilizes the concepts introduced
+    in earlier sections to discuss important practical considerations for
+    the use of the Lomb-Scargle periodogram, including the choice of
+    frequency grid, uncertainties and false alarm probabilities, and modes of
+    failure that should be expected when working with real-world data.
+  \item[\sect{conclusion}] concludes and summarizes the key recommendations
     for practical use of the Lomb-Scargle periodogram.
-
-\subsection{Code and Figures}
+\end{description}
 All figures in this paper were produced using Python, and in particular the
@@
 \sectlabel{practical-considerations}
+The previous sections have given a conceptual introduction to the Lomb-Scargle
+periodogram, intertwining the Fourier and Least-squares views.
+With these principles understood, this section gets to the meat of the subject
+at hand: given this understanding of the Lomb-Scargle periodogram and related
+approaches, what practical issues are involved in its effective use?
 
@@
 All code behind this paper is available in the form of Jupyter notebooks in the
-paper's GitHub repository\footnote{\url{http://github.com/jakevdp/PracticalLombScargle}}.
+paper's GitHub repository at \url{http://github.com/jakevdp/PracticalLombScargle/}.
 
@@
 
-One point of confusion in periodogram implementations is the interpretation
-of the periodogram values and their meaning.
+Often it is useful to be able to interpret the periodogram values themselves.
 Recall that there are two equally valid perspectives of what the periodogram
-measures: the Fourier view and the least-squares view.
-Each of these suggests a different natural approach to normalizing and interpreting the periodogram.
+is measuring---the Fourier view and the least-squares view---and each of these
+lends itself to a different approach to normalizion and interpretation of the
+periodogram.
 
@@
 periodogram differ slightly, and this needs to be taken into account when
-computing analytic estimates of uncertainties and false alarm probabilities.
+computing analytic estimates of uncertainties and false alarm probabilities
+(see the next section).
 Other normalizations exist as well, but seem to be rarely used in practice.
@@
 In all cases, the width of the peaks (most clearly visible in one or another
-normalization) are essentially {\it identical}, regardless of the data
-quality or quantity!
+normalization) are essentially {\it identical}, regardless of the quality
+or quantity of data!
 Evidently, peak width ({i.e.} frequency precision) is not a particularly
@@
 careful to keep in mind that it answers a very specific question:
-{\it what is the probability that a signal with no periodic component would
-lead to a peak of this magnitude?}
-In other words, it is a measure of the false positive rate: the rate at which
-we would expect to incorrectly identify a periodic signal where none
-actually exists.
-It tells us nothing about the false negative rate; {i.e.} the rate at which
+{\it ``what is the probability that a signal with no periodic component would
+  lead to a peak of this magnitude?''}.
+In particular, it emphatically does {\it not} answer the much more relevant
+question {\it ``what is the probability that this data is periodic
+  given these observations?''.}
+This boils down to a case of $P(A~|~B) \ne P(B~|~A)$, but still it is common to
+see the false alarm probability treated as if it speaks to the second rather
+than the first question.
+In other words, the false alarm probability tells you the false positive rate
+in a frequentist, rather than Bayesian sense: that is, it is a false positive
+rate conditioned on the entire space of possible observations rather than on
+the particular data you observed.
+
+Similarly, the false alarm probability tells us nothing about the false
+negative rate; {i.e.} the rate at which
 we would expect to miss a periodic signal that is, in fact, present.
@@
 Too often users are tempted to interpret the FAP too broadly, with the hope that
-it would speak to these additional questions as well.
+it would speak to questions that are beyond its reach.
 
 To the best of my knowledge, there is currently no silver bullet for answering
-these broader questions within the Lomb-Scargle approach.
+these broader, more relevant questions of uncertainty of Lomb-Scargle results.
 Perhaps the most fruitful path toward understanding of such effects for a
-particular survey---with particular noise characteristics and a particular
-observing window---is via simulated data injected into the detection
+particular set of observations---with particular noise characteristics and a
+particular observing window---is via simulated data injected into the detection
 pipeline.
-In fact, this type of simulation is a vital component of most (if not all)
-current and future time-domain surveys that seek to catalog periodic objects,
-regardless of the particular approach used to detect periodicity
-\citep[see, e.g.][]{opsim1, opsim2, Ivezic08LSST, Sesar2010, Oluseyi12, McQuillan2012, Drake2014}
+In fact, this type of simulation is a vital component of most---I might even
+venture to say all---current and future time-domain surveys that seek to
+detect, characterize, and catalog periodic objects,
+regardless of the particular approach used to identify periodicity
+\citep[see, e.g.][]{opsim1, opsim2, Ivezic08LSST, Sesar2010,
+  Oluseyi12, McQuillan2012, Drake2014}
 

### 2017-02-02 PracticalLombScargle.tex
Edit sections 1-6

@@
 \begin{abstract}
-This paper gives an introduction to the practical aspects of the use of
-Lomb-Scargle periodograms and related algorithms to detect and characterize 
-periodicity within a signal.
+This paper presents a conceptual introduction to the Lomb-Scargle periodogram,
+and discusses important practical aspects of its use to detect and characterize
+periodic signals.
 Rather than a rigorous mathematical treatment, the goal of this paper is to
-build intuition about what assumptions are baked into any periodogram
-computation and to help develop the intuition required to interpret the
+build intuition about what assumptions are implicit in the use of the
+periodogram, and to help develop the intuition required to interpret the
 periodogram correctly and use it effectively.
@@
 \sectlabel{introduction}
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=\textwidth]{fig01_LINEAR_data}
-\caption{Observed light curve from LINEAR object ID 11375941. Uncertainties
-  are indicated by the gray errorbars on each point.
-  \figlabel{LINEAR-data}
-}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=\textwidth]{fig02_LINEAR_PSD}
-\caption{{\it Left panel:} the Lomb-Scargle periodogram computed from the data
-    shown in \fig{LINEAR-data}, with an inset detailing the region near the peak.
-    {\it Right panel:} the input data in \fig{LINEAR-data}, folded over the
-    detected 2.58-hour period to show the coherent periodic variability.
-    For more discussion of this data, see \citep{LINEAR3}.
-    \figlabel{LINEAR-power}
-}
-\end{figure}
 
@@
 LINEAR survey \citep{LINEAR1, LINEAR3}, with magnitude measured 280 times over
-the course of five and a half years\footnote{
-  Python code to reproduce \fig{LINEAR-data}, as well as all other figures
-  in this manuscript, is available at 
-  \url{http://github.com/jakevdp/PracticalLombScargle/}}.
+the course of five and a half years.
 By eye, it is clear that the brightness of the object varies in time with a
@@
 an intuitive means of determining the period of oscillation.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\textwidth]{fig01_LINEAR_data}
+\caption{Observed light curve from LINEAR object ID 11375941. Uncertainties
+  are indicated by the gray errorbars on each point.
+  Python code to reproduce this figure, as well as all other figures
+  in this manuscript, is available at {\tt http://github.com/jakevdp/PracticalLombScargle/}
+  \figlabel{LINEAR-data}
+}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\textwidth]{fig02_LINEAR_PSD}
+\caption{{\it Left panel:} the Lomb-Scargle periodogram computed from the data
+    shown in \fig{LINEAR-data}, with an inset detailing the region near the peak.
+    {\it Right panel:} the input data in \fig{LINEAR-data}, folded over the
+    detected 2.58-hour period to show the coherent periodic variability.
+    For more discussion of this particular object, see \citep{LINEAR3}.
+    \figlabel{LINEAR-power}
+}
+\end{figure}
 
@@
 Often this is exactly how the Lomb-Scargle periodogram is presented: as a
-clean, well-defined procedure to detect the periodic component in an
-unevenly-sampled dataset.
+clean, well-defined procedure to generate a power spectrum and to
+detect the periodic component in an unevenly-sampled dataset.
 In practice, however, there are a number of subtle issues that must be
@@
   \item What assumptions is the Lomb-Scargle periodogram making about the
-    signal?
+    shape of the unknown signal?
   \item How should we understand the uncertainty of the peak frequency?
@@
 \subsection{Outline}
-The paper is organized as follows:
-\begin{description}
-  \item[\sect{introduction}] is this introduction.
-  \item[\sect{continuous-fourier-transform}] presents a review of the 
+The remainder of this paper is organized as follows:
+{\bf\sect{continuous-fourier-transform}} presents a review of the 
 continous Fourier transform and some of its useful properties,
-    including defining the notion of a power spectrum for measuring
-    periodic content in a signal.
-  \item[\sect{window-functions}] builds on these properties to explore how
+including defining the notion of a power spectrum (i.e. Schuster
+periodogram) for detecting periodic content in a signal.
+{\bf\sect{window-functions}} builds on these properties to explore how
 different observation patterns (i.e. window functions) affect the observed
@@
 well-known Nyquist limiting frequency.
-  \item[\sect{non-uniform-sampling}] considers non-uniformly sampled signals as
+{\bf\sect{non-uniform-sampling}} considers non-uniformly sampled signals as
 a special case of an observational window, and shows that the Nyquist-like
@@
 the field often assume.
-  \item[\sect{schuster-to-lomb-scargle}] details the Lomb-Scargle periodogram,
+{\bf\sect{schuster-to-lomb-scargle}} introduces the Lomb-Scargle periodogram,
 a modified version of the classical periodogram for unevenly-sampled data,
 as well as the motivation behind these modifications.
-  \item[\sect{lomb-scargle-extensions}] discusses a complementary view of the
+{\bf\sect{lomb-scargle-extensions}} discusses a complementary view of the
 Lomb-Scargle periodogram as the result of least-squares model fitting,
@@
 (including various ``Bayesian'' periodograms).
-  \item[\sect{practical-considerations}] utilizes the concepts introduced
+{\bf\sect{practical-considerations}} utilizes the concepts introduced
 in earlier sections to discuss important practical considerations for
@@
 failure that should be expected when working with real-world data.
-  \item[\sect{conclusion}] concludes and summarizes the key recommendations
+{\bf\sect{conclusion}} concludes and summarizes the key recommendations
 for practical use of the Lomb-Scargle periodogram.
-\end{description}
+
 All figures in this paper were produced using Python, and in particular the
@@
    \item[The Fourier transform is a linear operation.]
-     That is, for any constant $A$ and any functions $f$ and $g$, we can write
+     That is, for any constant $A$ and any functions $f(t)$ and $g(t)$,
+     we can write:
      \begin{eqnarray}
@@
 the total signal.
-Note that if $g$ is real-valued, it can be shown that
-$\mathcal{P}_g(f) = \mathcal{P}_g(-f)$.
+Note that if $g$ is real-valued, it follows that $P_g$ is an even function;
+{i.e.} $\mathcal{P}_g(f) = \mathcal{P}_g(-f)$.
 
@@
 from electrodynamics to music theory to quantum mechanics.
-For our purposes, this property of the transform will turn out to be quite
+For our purposes, this property will turn out to be quite
 important as we push further in understanding the Lomb-Scargle Periodogram.
@@
 ability to convert convolutions into point-wise products.
-A convolution of two functions, usually denoted with a $\ast$, is defined
-as follows:
+A convolution of two functions, usually denoted by the $\ast$ symbol,
+is defined as follows:
 \begin{equation}
@@
 \end{equation}
-In practice, this is a much more efficient means of computing the convolution
+In practice, this can be a much more efficient means of computing a convolution
 than to directly solve at each time $t$ the integral over $\tau$ that appears
@@
 a limited span of time, as illustrated in \Fig{rectangular-window}.
-Here we consider a periodic function observed only within the window $-3<t<3$.
-The observed signal in this case can be thought of as a pointwise product of
+Here we consider a continuous periodic function observed only within the
+window $-3<t<3$.
+The observed signal in this case can be understood as the pointwise product of
 the underlying infinite periodic signal, and a rectangular window function.
-Using \eq{convolution-theorem-inverse}, we see that the Fourier transform of
-this function is equivalent to the convolution of the Fourier transforms of
-each component, which are a set of delta functions and a sinc function,
-respectively.
-For a purely periodic signal, this convolution has the effect of broadening
-the delta functions representing the signal, turning each of them into a
-shifted sinc function.
+By the convolution theorem, the Fourier transform will be given by the
+convolution of the transform of the underlying function---here a set of
+delta functions at the component frequencies---and the transform of the window
+function---here a sinc function.
+For the purely periodic signal like the one seen in \fig{rectangular-window},
+this convolution has the effect of replacing each delta function with a
+sinc function.
 Because of the inverse relationship between the width of the window and the
@@
 Another window function that commonly arises is when a continuous signal is
-sampled instantaneously at regular intervals.
-Such an observation is, in effect, a point-wise product between the true
+sampled (nearly) instantaneously at regular intervals.
+Such an observation can be thought of as a point-wise product between the true
 underlying signal and a Dirac comb with the $T$ parameter matching the spacing
@@
 With this in mind, we can be assured in this case that evaluating the
-transform in the range $(2T)^{-1} \le f < (2T)^{-1}$ is sufficient to capture
+transform in the range $0 \le f < 1/T$ is sufficient to capture
 all the available frequency information:
 the signal outside that range is a sequence of identical aliases of
-what lies in that range.
-In fact, due to the nature of the aliasing, it is in fact sufficient to
-compute the transform in {\it any} range of width $1/T$, and it is often
-convenient to instead use the range $0 \le f < 1/T$.
+what lies within that range.
 
@@
 The example in \fig{comb-window-1} is somewhat of a best-case scenario, because
-the true Fourier transform values are non-zero only within a range of width $1/T$.
-If we change the sampling rate so that this is no longer the case, we will
-have a situation similar to that in \fig{comb-window-2}.
-A sparser sampling of the signal leads to a narrower sampling in frequency,
-and the signal transform no longer fits within the frequency window.
-The result is that the true Fourier transform {\it cannot be recovered}
-from the observed transform.
+the true Fourier transform values are non-zero only within a range of
+width $1/T$.
+If we increase the time between observations, decreasing the spacing of the
+frequency comb, the true transform no longer ``fits'' inside the window
+transform, and we will have a situation similar to that in \fig{comb-window-2}.
+The result is a mixing of different portions of the signal, such that
+the true Fourier transform {\it cannot be recovered}
+from the observed data!
 
@@
 This is one way to motivate the famous Nyquist sampling limit, which
-approaches the question the other way around and says
+approaches the question from the other direction and states
 that to fully represent the frequency content of a ``band-limited signal''
@@
 we observe it only at a regular grid with spacing $\Delta t$. In this case, our
-observed signal is $g_{obs} = g(t) \III_T(t)$ and its Fourier transform is
+observed signal is $g_{obs} = g(t) \III_{\Delta t}(t)$ and its Fourier transform is
 \begin{equation}
@@
 \end{equation}
-Apart from the $1/N$ scaling, this sum is precisely the Fourier power
+Apart from the $1/N$ proportionality, this sum is precisely the Fourier power
 spectrum in \eq{power-spectrum}, computed for a continuous signal with
 uniform sampling.
-It follows that, in the uniform sampling case the Schuster periodogram captures
+It follows that, in the uniform sampling case, the Schuster periodogram captures
 all of the relevant frequency information present in the data.
@@
     These samples have the same average spacing as those in \fig{comb-window-1},
-    but the lack of structure in observing window translates to a lack of
-    structure in its transform, causing the observed transform to be ``noisy''.
+    but the irregular spacing within the observing window translates to 
+    irregular frequency peaks in its transform,
+    causing the observed transform to be ``noisy''.
     Here black and gray lines represent the real and imaginary parts of the
@@
 In the general non-uniform case, we measure some signal at a set of $N$ times
-which we will denote $\{t_n\}$.
-Our observing window function in this case is still a sum of delta functions,
-but at non-uniform locations.
-Mathematically, the observing window will look like this:
+which we will denote $\{t_n\}$, which leads to the following observing window:
 \begin{equation}
@@
 Unlike in the uniform case, the window transform $\mathcal{F}\{W_{\{t_n\}}\}$
-will generally {\it not} be a nice clean function; from \eq{FT-def} and
-\eq{nonuniform-window}, we can write
-\begin{equation}
-  \mathcal{F}\{W_{\{t_n\}}\}(f) = \sum_{n=1}^{N} e^{-2\pi i f t_n}
-  \eqlabel{nonuniform-window-FT}
-\end{equation}
-For uniformly-spaced observing times $t_n$, the symmetry leads to cancellations
-so that the window transform is a clean Dirac comb.
-For non-uniform observing times, this symmetry is broken and the window
-transform will be much more ``nosiy''.
+will generally {\it not} be a nice clean function; the symmetry present in
+the Dirac comb is broken by the uneven sampling,
+leading the transform to be much more ``noisy''.
 This can be seen in \fig{random-window}, which shows the Fourier transform of
@@
 the non-uniformly spaced delta functions looks like random noise, and in some
-senses it is: the locations and heights of the Fourier peaks are related to
+sense it is: the locations and heights of the Fourier peaks are related to
 the intervals between observations, and so randomization of observation times
@@
 definition of the Schuster periodogram to the non-uniform case.
-Plugging \eq{g-nonuniform} into the Fourier transform of \eq{FT-def}, in analogy to \eq{schuster-periodogram}, gives the following form of the power spectrum:
+Combining \eq{g-nonuniform} with the Fourier transform in \eq{FT-def}, in analogy to \eq{schuster-periodogram}, gives the following form of the power spectrum:
 \begin{equation}
@@
 number of observed samples gets very large.
-For a discussion of nonuniform window functions and attempts to correct
-for their effect on the periodogram, see \sect{windows-and-deconvolution}.
 
@@
 As we have seen in \sect{nyquist}, the Nyquist limit is a direct consequence
-of the Dirac comb window function created by evenly-sampled data,
+of the symmetry in the Dirac comb window function that describes evenly-sampled data,
 and uneven sampling destroys the symmetry that underlies its definition.
@@
 In the scientific literature it is quite common to come across various proposals
-for a Nyquist-like or pseudo-Nyquist limit in the case of irregular sampling.
+for a Nyquist-like limit applied in the case of irregular sampling.
 A few typical approaches include using the mean of the sampling intervals
@@
 or the minimum sample spacing \citep[e.g.][]{Percy86, Roberts87, Press89, Hilditch01}.
-All of these are tempting criteria in that they recover the classical Nyquist 
+All of these ``pseudo-Nyquist'' limits are tempting criteria in that
+they recover the classical Nyquist 
 frequency in the limit of evenly-spaced data.
@@
 unevenly-sampled data can probe frequencies far larger than any of these
-supposed limits (a fact that several of these citations do hint at
-parenthetically).
+supposed limits---a fact that several of these citations do hint at
+parenthetically.
 
@@
 pseudo-Nyquist definition based on an average or minimum sample spacings
-will be {\it far} smaller than the true frequency of 100; still, the
-the Lomb-Scargle periodogram in the lower right panel shows quite cleanly
-the true frequency.
+will be {\it far} below the true frequency of 100; still, the
+the Lomb-Scargle periodogram in the lower right panel quite cleanly
+recovers the true frequency.
 
@@
 by thinking of such data as a windowed version of uniformly sampled
-data with spacing $p$, where the window is zero at some observations.
+data with spacing $p$, where the window is zero at all points but the
+location of the observations.
 Such uniform data has a classical Nyquist limit of $1/(2p)$, and a window
-function applied to that data does not change that fact.
+function applied on top of that sampling does not change that fact.
 
@@
   \includegraphics[width=\textwidth]{fig13_nyquist_eyer99}
-  \caption{A visualization of the \citep{Eyer99} definition of the Nyquist
+  \caption{A visualization of the \citet{Eyer99} definition of the Nyquist
     frequency. Data are non-uniformly sampled at times $t_i = n_i p$, for
@@
 frequency can be large, but not infinite.
-Quantitatively, if the times are recorded to $D$ decimal places,
+For example, if your observation times are recorded to $D$ decimal places,
 the Nyquist frequency will be at most
 \begin{equation}
-  f_{Ny} = \frac{1}{2} 10^D,
+  f_{Ny} \le \frac{1}{2} 10^D,
   \eqlabel{nonuniform-nyquist}
 \end{equation}
-(cf. \fig{nyquist-eyer99}).
+with the inequality due to the fact that larger common factors may exist.
 In other words, absent other relevant patterns in the observations,
@@
 kind of limiting frequency for irregularly-sampled data.
-Such a situation means that the measurement is effectively a convolution of
+Such a situation means that the observation is effectively a convolution of
 the underlying signal with a rectangular window function of width $\delta t$,
@@
 
-\subsubsection{Which Limiting Frequency is Useful?}
-
-\todo{Computational constraints; use of prior knowledge}
+\subsubsection{Practicality of Limiting Frequencies}
+The frequency grid implied by such large Nyquist frequencies in practice can be
+much larger than is computationally feasible.
+For example, consider the LSST project, which will have roughly 30-second
+exposures taken over a 10-year baseline. With
+$f_{Ny} \approx (2\cdot 30{\rm sec})^{-1}$ and a grid spacing
+of $\Delta f = (5\cdot10 {\rm yr})^{-1}$ (see \sect{frequency-grid}),
+the size of the required frequency grid is on order 25 million bins, which is
+far too large a periodogram to be computationally feasible given the
+computational constraints of the LSST pipeline, even for extremely
+efficient periodogram methods (see \sect{algorithmic-considerations}).
+
+In practice, the limiting frequency of the periodogram should be set based on
+some prior knowledge of the types of signals you are interested in.
+For example, if you are interested in galactic RR Lyrae stars, the shortest
+applicable period is on order a few hours \citep{Sesar2010}, requiring
+evaluation of only a hundred thousand or so periodogram frequencies for
+data with a 10-year baseline.
+The key point is that for working with unevenly-sampled data in practice,
+computational constraints will mean that the maximum frequency
+should more often be set by {\it what you are looking for}
+rather than any hard limit based on Nyquist-like limits.
 
@@
 data \citep[see, e.g.][]{Deeming75}.
+In this section, we will examine two typical window functions derived from
+real-world observations: one ground-based (LINEAR) and one space-based (Kepler).
+For details on how window power spectra can be computed in practice, see
+\sect{windows-and-deconvolution}.
 
@@
 
-Thus the strong spike at a period of 1 day indicate that observations are
+With this in mind, the strong spike at a period of 1 day indicates
+that observations are
 taken near the same time of day: this is typical of a ground-based survey
-with observations recorded only at night.
+with observations recorded only during the night hours.
 The additional spikes at periods of $1/n$ days (for integer $n$) are aliases
-of this same effect.
-
-This window also shows a strong spike at a period of 1 year, indicating that
-there was a strong seasonal pattern in the observations.
+of this same feature.
+\fig{LINEAR-window} also shows a wide spike at a period of 1 year, indicating
+that there is a detectable annual pattern in the observations.
 Finally, there is a noticeable spike at approximately 14 days that is likely
-related to patterns in scheduling of observing runs.
+related to patterns of scheduling within the survey.
 
@@
 
-This behavior is illustrated in \fig{LINEAR-window-effect}.
-The upper panel shows the window power spectrum as a function of frequency (not period, as in \fig{LINEAR-window}), while lower panel shows the observed signal power spectrum as a function of frequency.
-The upper panel and its inset show clearly the 1-day and 1-year features we noted previously.
-The lower panel shows the observed power spectrum of the data: these diurnal and annual peaks in the window function are quite clearly imprinted on the observed power spectrum at relevant scales.
-This approximate aliasing is similar to the exact aliasing seen in regularly-sampled data at the Nyquist frequency; however, in this case the magnitude of the
-aliased signal fades further from the frequency driving the signal.
+This imprint of the window power is illustrated in \fig{LINEAR-window-effect}.
+The upper panel shows the window power spectrum as a function of frequency
+(not period, as in \fig{LINEAR-window}), while lower panel shows the observed
+signal power spectrum as a function of frequency.
+The upper panel and its inset show clearly the 1-day and 1-year features we
+noted previously.
+The lower panel shows the observed power spectrum of the data: these diurnal
+and annual peaks in the window function are quite clearly imprinted on the
+observed power spectrum at relevant scales.
+This approximate aliasing is similar to the exact aliasing seen in
+regularly-sampled data at the Nyquist frequency; however, in this case
+the magnitude of the aliased signal fades further from the frequency
+driving the signal.
 
@@
 shows observations of an RR-Lyrae variable star from the Kepler survey,
-measured over 4000 times over a period of three months,
+measured 4083 times over a period of three months,
 with an irregular observing cadence of just under 30 minutes
@@
 The window function is a series of very narrow evenly-spaced spikes,
-reminiscent of the Dirac comb shown in \figs{comb-window-1}{comb-window-2},
-which we used to motivate the idea of the Nyquist limit.
-Thus, in this case we can treat $f_{Ny} = 0.5/29.4$ minutes$^{-1}$ as the
-effective Nyquist limit for the data, though aliasing is imperfect due to
-the uneven spacing of samples, shown in the lower-left panels of
-\fig{kepler-data}.
-
+reminiscent of the Dirac comb shown in \figs{comb-window-1}{comb-window-2}.
+By analogy we can treat $f_{Ny} = 0.5/29.4$ minutes$^{-1}$ as the
+effective Nyquist limit for the data, keeping in mind that aliasing beyond this
+``limit'' will be imperfect due to the uneven spacing of the samples
+(see the lower-left panels of \fig{kepler-data}).
 The lower-right panel of \fig{kepler-data} shows the power spectrum of the
-observations, with gray shading indicating the ``Aliased'' region of the
-spectrum.
-The period of 13.6 hours is strongly apparent, along with smaller spikes at
-integer multiples of this frequency that indicate higher-order periodic
+observations, with gray shading indicating the (nearly) aliased region
+of the spectrum.
+The period of 13.6 hours is quite strongly apparent, along with smaller spikes
+at integer multiples of this frequency that indicate higher-order periodic
 components in the signal.
@@
 \fig{kepler-data}, are quite different, but in both cases essential features
-of the observed power spectra can be understood by recognizing it reflects
+of the observed power spectra can be understood by recognizing that the
+periodogram measures
 not the power spectrum of the underlying signal, but a power specrtrum from
@@
 the window function.
-For another example of this type of observing window analysis, see
-\citep{Deeming75}
 
@@
 
-Up until now, we've been mainly discussing the direct extension of the Schuster
-periodogram in \eq{schuster-periodogram} to non-uniform data.
+Up until now, we have been mainly discussing the direct extension of the
+Schuster periodogram in \eq{schuster-periodogram} to non-uniform data.
 Returning to this definition, we can rewrite the expression in a more
@@
 Gaussian noise).
-
 For the sake of evaluating the significance of periodogram peaks
@@
 straightforwardly computed from the uncertainties of the data.
-
-\todo{Mention requirement of uncorrelated noise, and correction for this explored by \citet{Vio2010}}
 
@@
 where $A$ and $B$ are arbitrary functions of the frequency $f$ and
-observing times $\{t_i\}$ (but not the values $\{g_n\}$, and showed
+observing times $\{t_i\}$ (but not the values $\{g_n\}$), and showed
 that you can choose one particular form of $A$ and $B$ such that
 \begin{enumerate}
-  \item The periodogram reduces to the classical form in the case of equally-spaced data,
+  \item The periodogram reduces to the classical form in the case of equally-spaced observations,
   \item The periodogram has desireable statistical properties, in the sense discussed above,
@@
 frequency--an estimator which was considered in some depth by \citet{Lomb76}.
-In this view, the $\tau$ shift defined in \eq{tau-def} serves to orthogonalize
-the normal equations used in the least squares analysis.
-
+From this perspective, the $\tau$ shift defined in \eq{tau-def} serves to
+orthogonalize the normal equations used in the least squares analysis.
 Partly due to the deep connection between Fourier analysis and least-squares
 analysis, the modified periodogram in \eq{lomb-scargle-periodogram}
-has since become commonly referred to as the {\it Lomb-Scargle Periodogram}
-\citep[though versions of this approach had been employed even earlier;
-  see, e.g.][]{Gottlieb75}.
-
+has since become commonly referred to as the {\it Lomb-Scargle Periodogram},
+although versions of this approach had been employed even earlier
+\citep[see, e.g.][]{Gottlieb75}.
 Because of this similarity between the classical and Lomb-Scargle periodograms,
-our discussion of windowing effects applies---at least qualitatively---to
-results computed with the Lomb-Scargle periodogram.
+the bulk of or previous discussion applies---at least qualitatively---to
+periodograms computed with the Lomb-Scargle method.
 In particular, reasoning about the effect of window functions on the observed
@@
 
+One important caveat of the Lomb-Scargle approach is that the statistical
+guarantees only apply when the observations have {\it uncorrelated white
+noise}; data with more complicated noise characteristics must be treated
+more carefully; see, e.g., \citep{Vio2010}.
+
 \section{The Least-Square Periodogram and its Extensions}
@@
 The equivalence of the Fourier interpretation and least squares interpretation of the Lomb-Scargle periodogram allows for some interesting and useful
-extensions, which we will explore in this section.
-First, let's look at the Least Squares periodogram itself:
+extensions, some of which we will explore in this section.
+First, let's consider the Least Squares periodogram itself.
 
@@
 \begin{equation}
-  y_{model}(t;f) = A_f \sin(2 \pi f (t - \phi_f))
+  y(t;f) = A_f \sin(2 \pi f (t - \phi_f))
 \end{equation}
 where the amplitude $A_f$ and phase $\phi_f$ can vary as a function of frequency.
-This is fit to the data in the standard least-squares sense, by constructing
-the $\chi^2$ statistic at each frequency:
-\begin{equation}
-  \chi^2(f) \equiv \sum_n \big(y_n - y_{model}(t_n;f)\big)^2
+These model parameters are fit to the data in the standard least-squares sense,
+by constructing the $\chi^2$ statistic at each frequency:
+\begin{equation}
+  \chi^2(f) \equiv \sum_n \big(y_n - y(t_n;f)\big)^2
   \eqlabel{chi2-simple}
@@
 
-\subsection{Observational Noise}
+\subsection{Measurement Errors}
 \sectlabel{extensions-observational-noise}
@@
 of a multiplicative weight $1/\sigma_n$ within each of the summations.
-Early versions of this sort of modification appeared in \citet{Gilliland87},
-\citet{Scargle89}, and \citet{Irwin89};
-\citet{Zechmeister09} showed that such a modification
-does not change the desirable statistical properties of the resulting
-periodogram.
+Early versions of this sort of modification appeared in \citet{Gilliland87}, and
+\citet{Irwin89}; \citet{Scargle89} derived this ``weighted'' form of the
+periodogram without direct reference to the least squares model,
+and \citet{Zechmeister09} showed that such a modification does not change
+the useful statistical properties of the resulting periodogram.
 
@@
 Another commonly-applied modification of the periodogram has variously been
-called the {\it Date-compensated Discrete Fourier Transform} \citep{Ferraz-Mello81}, the {\it floating-mean periodogram} \citep{Cumming99}, and the {\it generalized Lomb-Scargle Method} \citep{Zechmeister09}, and involves adding an offset term to the sinusoidal model at each frequency\footnote{We choose to follow \citet{Cumming99} and \citet{VanderPlas2015} and call this a ``Floating Mean'' model, to avoid confusion surrounding the different uses of the term ``generalized periodogram'' in, {e.g.} \citet{Bretthorst88} and \citet{Zechmeister09}.}:
+called the
+{\it Date-compensated Discrete Fourier Transform} \citep{Ferraz-Mello81},
+the {\it floating-mean periodogram} \citep{Cumming99,VanderPlas2015},
+or the {\it generalized Lomb-Scargle Method} \citep{Zechmeister09},
+and involves adding an offset term to the sinusoidal model at each
+frequency\footnote{We choose to follow \citet{Cumming99} and
+  \citet{VanderPlas2015} and call this a ``Floating Mean'' model, to
+  avoid confusion surrounding the different uses of the term
+  ``generalized periodogram'' in, {e.g.} \citet{Bretthorst88} and
+  \citet{Zechmeister09}.
+}:
 \begin{equation}
@@
 Applying the standard Lomb-Scargle periodogram to pre-centered data leads to a
-periodogram that misses the true period of 0.3 days (lower right panel).
+periodogram that suppresses the true period of 0.3 days (lower right panel).
 Using the floating-mean model of \eq{floating-mean-model} yields a periodogram
@@
 
-A further generalization involves multi-term Fourier models at each frequency:
+A further generalization involves multi-term Fourier models:
 rather than fitting just a single sinusoid at each frequency, we might fit a
-partial Fourier series at each frequency, adding $K$ additional components at
+partial Fourier series, adding $K-1$ additional components at
 integer multiples of the fundamental frequency:
@@
     14752041, an eclipsing binary. Notice that the standard periodogram
-    (top panels) finds an alias of the true 17.5-hour frequency,
+    (upper panels) finds an alias of the true 17.5-hour frequency,
     because a simple sinusoidal model cannot closely fit both the primary
-    and secondary eclipse. A six-term Fourier model (bottom panels), on
-    the other hand, does find the true period.
+    and secondary eclipse. A six-term Fourier model, on
+    the other hand, does find the true period (lower panels).
     \figlabel{binary-multiterm}}
@@
 quickly wash-out any gains from the more complicated model.
-For example, consider the periodogram in \fig{LINEAR-multiterm}: as previously
-discussed, this periodogram contains a number of aliases of the true peak due
-to the dominant 24-hour signal in the window function.
+For example, consider the multi-term periodograms for the LINEAR data
+depicted in \fig{LINEAR-multiterm}: as previously discussed,
+the LINEAR periodogram contains a number of aliases of the true peak due
+to the dominant 1-day signal in the window function.
 Adding terms to this model only compounds this problem, increasing the number of
 spurious peaks at the low-frequency end.
-For data with even moderate levels of noise, coincidences of these peaks can
-easily lead to spurious signals dominating the true peak.
+For data with even moderate levels of noise, chance coincidences of these
+peaks can easily lead to spurious detections which dominate the true peak.
 
@@
 Bayesian ``periodograms'' that can measure
-decaying signals, non-stationary signals, multiple signals, and other models,
-and shows the effectiveness of such approaches in applications from
+decaying signals, non-stationary signals, multi-frequency signals, and other models.
+He shows the effectiveness of such approaches in applications from
 medical imaging and astronomy.
 The crux of such extensions is the fact that you often need to use some prior
-knowledge of your data to decide when it indicates a need for a more
-complicated model, and additionally what form of additional variation to add.
+knowledge of the system being observed to decide whether a more complicated
+model is indicated, as well as what form of model to apply.
 In practice this comes up only when searching for very specific signals for
@@
 
-On the astronomy side, several examples of such extensions exist.
+On the astronomy side, several examples of this flavor of extension exist.
 For example, the Supersmoother approach to detecting periodicity
@@
 In fact, \citet{Jaynes87} showed that the standard form of the Lomb-Scargle
-periodogram can be derived directly from the principles of Bayesian probability
-theory outlined in his comprehensive treatment of the subject \citep{Jaynes03}.
+periodogram can be derived directly from the axioms of Bayesian probability
+theory outlined in his comprehensive treatment of the subject
+\citep{Jaynes03}\footnote{Read this. Really.}.
 In the Bayesian view, the Lomb-Scargle periodogram is the optimal statistic
@@
 
-For the standard model, the Bayesian periodogram is given by the posterior
+For the standard, simple-sinusoid model, the Bayesian periodogram is given by the posterior
 probability of frequency $f$ given the data $D$ and sinusoidal model $M$:
@@
 
-The benefit of the Bayesian approach is that it allows more principled treatment
-of nuisance parameters related to noise in the data and choices of model,
-and Bayesian periodogram extensions have been used quite effectively in a
+The benefit of the Bayesian approach is that it allows more flexible models,
+more principled treatment of nuisance parameters related to noise in the data,
+and the ability to make use of prior information in a periodic analysis.
+As mentioned above, Bayesian periodogram extensions have been used quite effectively in a
 variety of circumstances \citep[see, e.g.][]{Bretthorst88}.
@@
 a probability at each frequency conditioned on a relatively strong assumption:
-that the data are drawn from a sinusoidal model.
+that the data are drawn from a sinusoidal model (or, more generally, the
+particular form of model you have selected).
 Now, the standard Lomb-Scargle periodogram also assumes a sinusoidal model,
@@
 The window power spectrum can be computed directly from the
-delta-function representation of the window
-function; recalling \eq{nonuniform-window-FT}, the window power spectrum
+delta-function representation of the window function;
+From \eq{FT-def}, \eq{power-spectrum}, and \eq{nonuniform-window},
+we can write
+\begin{equation}
+  \mathcal{F}\{W_{\{t_n\}}\}(f) = \sum_{n=1}^{N} e^{-2\pi i f t_n}
+  \eqlabel{nonuniform-window-FT}
+\end{equation}
+For uniformly-spaced observing times $t_n$, the symmetry leads to cancellations
+so that the window transform is a clean Dirac comb.
+For non-uniform observing times, this symmetry is broken and the window
+transform will be much more ``nosiy''.
 for observation times $\{t_n\}$ is

### 2017-02-03 PracticalLombScargle.tex
Add section on correlated noise

@@
 noise}; data with more complicated noise characteristics must be treated
-more carefully; see, e.g., \citep{Vio2010}.
+more carefully; see, e.g., \citep{Vio2010} or the Least Squares approach
+discussed in \sect{extensions-observational-noise}.
 
@@
 the useful statistical properties of the resulting periodogram.
+
+This also suggests a convenient way to construct a periodogram that accounts
+for correlated noise across observations; denoting the  $N\times N$ noise
+covariance matrix for $N$ observations by $\Sigma$, and constructing the
+vectors $\vec{y} = [y_1, y_2,\cdots y_n]^T$
+and $\vec{y}_{model} = [y_{model}(t_1),y_{model}(t_1),\cdots y_{model}(t_n)]^T$,
+then the $\chi^2$ statistic can be written
+\begin{equation}
+  \chi^2(f) = (\vec{y}-\vec{y}_{model})^T\Sigma^{-1}(\vec{y}-\vec{y}_{model})
+  \eqlabel{chi2-with-correlated-errors}
+\end{equation}
+In the case of uncorrelated errors, the off-diagonal terms of $\Sigma$
+are zero, and this expression reduces to \eq{chi2-with-errors}.
+This is similar in spirit to the approach to correlated noise developed
+by \citet{Vio2010} from the Fourier perspective, and is in fact identical
+in the case of the ``zero-mean colored noise'' example considered therein.
 
@@
 the Schuster periodogram for data $g_n=1$ at all times $t_n$;
-with that in mind one quick way to compute the window power is to use
-a standard Lomb-Scargle periodogram on a series of unit measurements,
-making sure to {\it not} pre-center the data or to use a floating mean model.
-This method is how window power spectra have been computed throughout this
+with that in mind one quick way to compute a qualitatively correct window
+power is to use a standard Lomb-Scargle periodogram on a series of unit
+measurements, making sure to {\it not} pre-center the data or to use a
+floating mean model.
+As \citep{Scargle82} notes, this is {\it not} the true window---it differs
+from the true window just as the Lomb-Scargle periodogram differs from the
+Schuster periodogram---but in practice it is accurate enough to give
+useful insight into the window function's important features.
+This method is how window power spectra has been computed throughout this
 paper.
@@
 \begin{equation}
-  g_{obs}(t) = g_{true}(t)W_{\{t_n\}}(t),
+  g_{obs}(t) = g(t)W_{\{t_n\}}(t),
 \end{equation}
@@
 \begin{equation}
-  \mathcal{F}\{g_{obs}\} = \mathcal{F}\{g_{true}\} \ast \mathcal{F}\{W_{\{t_n\}}\}
+  \mathcal{F}\{g_{obs}\} = \mathcal{F}\{g\} \ast \mathcal{F}\{W_{\{t_n\}}\}
 \end{equation}
 Given this relationship, we might hope to be able to invert this
-convolution to recover $\mathcal{F}\{g_{true}\}$ directly.
+convolution to recover $\mathcal{F}\{g\}$ directly.
 For example, we could write:
 \begin{eqnarray}
-  \mathcal{F}\{g_{true}\} &=& \mathcal{F}\{g_{obs}/W_{\{t_n\}}\}\nonumber\\
+  \mathcal{F}\{g\} &=& \mathcal{F}\{g_{obs}/W_{\{t_n\}}\}\nonumber\\
                    &=& \mathcal{F}\{g_{obs}\} \ast \mathcal{F}\{1/W_{\{t_n\}}\}

### 2017-02-03 PracticalLombScargle.tex
Add Koen 2006 reference

@@
 the Nyquist frequency for irregularly-sampled data is most
-typically set by {\it the precision of the time measurements}.
+typically set by {\it the precision of the time measurements}
+\citep[see][for a more rigorous derivation of this]{Koen2006}.
 
@@
 
-One important caveat of the Lomb-Scargle approach is that the statistical
+One important caveat of the simple Lomb-Scargle formula is that the statistical
 guarantees only apply when the observations have {\it uncorrelated white
 noise}; data with more complicated noise characteristics must be treated
-more carefully; see, e.g., \citep{Vio2010} or the Least Squares approach
+more carefully; see, e.g., \citet{Vio2010} or the Least Squares approach
 discussed in \sect{extensions-observational-noise}.
@@
 
-This also suggests a convenient way to construct a periodogram that accounts
-for correlated noise across observations; denoting the  $N\times N$ noise
-covariance matrix for $N$ observations by $\Sigma$, and constructing the
-vectors $\vec{y} = [y_1, y_2,\cdots y_n]^T$
-and $\vec{y}_{model} = [y_{model}(t_1),y_{model}(t_1),\cdots y_{model}(t_n)]^T$,
-then the $\chi^2$ statistic can be written
-\begin{equation}
-  \chi^2(f) = (\vec{y}-\vec{y}_{model})^T\Sigma^{-1}(\vec{y}-\vec{y}_{model})
+This generalization of $\chi^2(f)$ also suggests a convenient way to
+construct a periodogram in the presence of correlated observational noise.
+If we let $\Sigma$ denote the  $N\times N$ noise
+covariance matrix for $N$ observations, and construct the
+vectors
+\begin{eqnarray}
+  &\vec{y} = [y_1, y_2,\cdots y_n]^T \nonumber\\
+  &\vec{y}_{model} = [y_{model}(t_1),y_{model}(t_1),\cdots y_{model}(t_n)]^T,
+\end{eqnarray}
+then the $\chi^2$ expression for correlated errors can be written
+\begin{equation}
+  \chi^2(f) = (\vec{y}-\vec{y}_{model})^T\Sigma^{-1}(\vec{y}-\vec{y}_{model}),
   \eqlabel{chi2-with-correlated-errors}
 \end{equation}
-In the case of uncorrelated errors, the off-diagonal terms of $\Sigma$
-are zero, and this expression reduces to \eq{chi2-with-errors}.
-This is similar in spirit to the approach to correlated noise developed
-by \citet{Vio2010} from the Fourier perspective, and is in fact identical
-in the case of the ``zero-mean colored noise'' example considered therein.
+which reduces to \eq{chi2-with-errors} if noise is uncorrelated (i.e., if the
+off-diagonal terms of $\Sigma$ are zero).
+This resulting periodogram is quite similar to the approach to correlated
+noise developed by \citet{Vio2010} from the Fourier perspective,
+and is in fact exactly equivalent in the case of the
+``zero-mean colored noise'' example considered therein.
 

### 2017-02-04 PracticalLombScargle.tex
Proof-read section 7

@@
 typically set by {\it the precision of the time measurements}
-\citep[see][for a more rigorous derivation of this]{Koen2006}.
+\citep[see][for a more rigorous treatment of this result]{Koen2006}.
 
@@
 
-\subsubsection{Practicality of Limiting Frequencies}
-The frequency grid implied by such large Nyquist frequencies in practice can be
-much larger than is computationally feasible.
-For example, consider the LSST project, which will have roughly 30-second
-exposures taken over a 10-year baseline. With
-$f_{Ny} \approx (2\cdot 30{\rm sec})^{-1}$ and a grid spacing
-of $\Delta f = (5\cdot10 {\rm yr})^{-1}$ (see \sect{frequency-grid}),
-the size of the required frequency grid is on order 25 million bins, which is
-far too large a periodogram to be computationally feasible given the
-computational constraints of the LSST pipeline, even for extremely
-efficient periodogram methods (see \sect{algorithmic-considerations}).
-
-In practice, the limiting frequency of the periodogram should be set based on
-some prior knowledge of the types of signals you are interested in.
-For example, if you are interested in galactic RR Lyrae stars, the shortest
-applicable period is on order a few hours \citep{Sesar2010}, requiring
-evaluation of only a hundred thousand or so periodogram frequencies for
-data with a 10-year baseline.
-The key point is that for working with unevenly-sampled data in practice,
-computational constraints will mean that the maximum frequency
-should more often be set by {\it what you are looking for}
-rather than any hard limit based on Nyquist-like limits.
+%\subsubsection{Practicality of Limiting Frequencies}
+%The frequency grid implied by such large Nyquist frequencies in practice can be
+%much larger than is computationally feasible.
+%For example, consider the LSST project, which will have roughly 30-second
+%exposures taken over a 10-year baseline. With
+%$f_{Ny} \approx (2\cdot 30{\rm sec})^{-1}$ and a grid spacing
+%of $\Delta f = (5\cdot10 {\rm yr})^{-1}$ (see \sect{frequency-grid}),
+%the size of the required frequency grid is on order 25 million bins, which is
+%far too large a periodogram to be computationally feasible given the
+%computational constraints of the LSST pipeline, even for extremely
+%efficient periodogram methods (see \sect{algorithmic-considerations}).
+
+%In practice, the limiting frequency of the periodogram should be set based on
+%some prior knowledge of the types of signals you are interested in.
+%For example, if you are interested in galactic RR Lyrae stars, the shortest
+%applicable period is on order a few hours \citep{Sesar2010}, requiring
+%evaluation of only a hundred thousand or so periodogram frequencies for
+%data with a 10-year baseline.
+%The key point is that for working with unevenly-sampled data in practice,
+%computational constraints will mean that the maximum frequency
+%should more often be set by {\it what you are looking for}
+%rather than any hard limit based on Nyquist-like limits.
 
@@
 \begin{eqnarray}
-  P(f) =
+  P_{LS}(f) =
   \frac{1}{2} \Bigg\{ &
@@
 the mean estimated from the data points themselves.
-
 When the data provide full phase coverage of the observed signal,
@@
 where the periodogram suppresses peaks of interest.
+
 \Fig{standard-vs-floatingmean} shows a simulated example of this: the data
 consist of noisy observations of a sinusoidal signal in which the faintest
-observations are omitted from the dataset (due to, {e.g.} a detection threshold).
+observations are omitted from the dataset (due to, e.g., a detection threshold).
 Applying the standard Lomb-Scargle periodogram to pre-centered data leads to a
@@
 Using the floating-mean model of \eq{floating-mean-model} yields a periodogram
-that detects this true period (upper right panel).
+that identifies this true period (upper right panel).
 
@@
 
-\citet{Zechmeister09} provide a detailed study of this
-generalized/floating-mean periodogram, and shows that the beneficial
-statistical properties of the standard periodogram apply here as well;
-namely, the periodogram noise at each frequency is given by the sum of
-two independent Gaussian random variables.
+\citet{Zechmeister09} provide a detailed study of such a model,
+and show that the addition of the floating mean term does not change
+the advantageous statistical properties of the periodogram.
 
@@
 how well the model fits the data.
-For Linear models, adding additional terms will {\it always} provide an
-equal or better fit to data than the simpler model, and so a periodogram based
-on a more complex model will be higher {\it at all frequencies}, not just
-at frequencies of interest!
+For linear models, adding additional terms will {\it always} provide an
+equal or better fit to data than the simpler model, and so it follows that
+a periodogram based on a more complex model will be higher
+{\it at all frequencies}, not just at frequencies of interest!
 
@@
 third harmonic.
-In general, you can expect an $N$-term periodogram to contain $N$ copies of
-every feature present in the standard periodogram, in addition to any
-new or strengthened features due to the more flexible model.
+In general, you can expect an $N$-term periodogram to contain $N$ aliases of
+every feature present in the standard periodogram; any strong peak revealed
+by the multi-term periodogram is due to two or more of these aliases coinciding.
 
@@
 For data with even moderate levels of noise, chance coincidences of these
-peaks can easily lead to spurious detections which dominate the true peak.
+peaks can lead to spurious detections which dominate the true peak, particularly
+for models with many Fourier terms.
 
@@
 In practice this comes up only when searching for very specific signals for
-which a more complicated model has some physical motivation.
+which a more complicated model has some {\it a priori} physical motivation.
 
@@
 involves fitting a flexible non-parametric smoothing
-function to the data at each frequency: the flexibility of the model leads
-to fewer issues of aliasing \citep{Reimann94}.
+function to the data at each frequency \citep{Reimann94}: the flexibility
+of the model leads to fewer aliasing issues when compared to the more
+constrained sinusoidal model.
 Another approach is to use empirically-derived templates as a functional fit
@@
 computational complexity quickly becomes an issue, because fast
-$\mathcal{O}(N\log N)$ approaches along the lines of \citet{Press89}
-are not available for more general functional forms
-(see \sect{algorithmic-considerations}).
+$\mathcal{O}(N\log N)$ approaches which can be used for sinusoidal fits
+(see \sect{algorithmic-considerations}) are not available for more general
+functional forms.
 
@@
 In the Bayesian view, the Lomb-Scargle periodogram is the optimal statistic
-for detecting a stationary sinusoidal signal in the presence of Gaussian
+for detecting a stationary {\it sinusoidal} signal in the presence of Gaussian
 noise.
@@
 \begin{equation}
-  p(f~|~D, M) \propto e^{-N \chi^2 / 2} \propto e^{P_{LS}(f)}
+  p(f\mid D, M) \propto e^{-N \chi^2 / 2} \propto e^{P_{LS}(f)}
   \eqlabel{bayesian-periodogram}
@@
 The problem, however, is that the Bayesian periodogram does {\it not} return
-a simple probability that the data is periodic with a given frequency, but rather
-a probability at each frequency conditioned on a relatively strong assumption:
-that the data are drawn from a sinusoidal model (or, more generally, the
-particular form of model you have selected).
-Now, the standard Lomb-Scargle periodogram also assumes a sinusoidal model,
-and is thus subject to the same criticism.
-But unlike the simple Bayesian approach, the standard periodogram affords
-interpretation in light of Fourier transforms, window functions, and the
-like: in most cases of interest, this turns out to be {\it vital} to the
+the probability that the data is periodic with a given frequency, but rather
+the probability conditioned on a relatively strong assumption:
+that {\it the data are drawn from a sinusoidal model} (or, more generally, the
+particular form of model used to compute the likelihood).
+As such, you should not treat the Bayesian periodogram as a measure of
+generic periodicity, but rather as a measure of how close the data is to
+a sinusoid.
+Now, the standard Lomb-Scargle periodogram can also be viewed as a sinusoidal
+fit, and is thus subject to the same criticism in the least-squares
+interpretation.
+But unlike the Bayesian periodogram, the standard periodogram affords
+interpretation in light of Fourier analysis and window functions:
+in most cases of interest, such analysis turns out to be {\it vital} to the
 application and interpretation of the periodogram in most cases of interest
@@
 approach, they tend to be far more expensive computationally.
-Bayesian approaches based on MCMC also tend to run into stability problems,
+Bayesian approaches based on Markov Chain Monte Carlo (MCMC) also tend to
+run into stability problems,
 particularly for multimodal or other complicated posterior distributions
@@
 periodogram, intertwining the Fourier and Least-squares views.
-With these principles understood, this section gets to the meat of the subject
-at hand: given this understanding of the Lomb-Scargle periodogram and related
+This section gets to the meat of the subject at hand:
+given this understanding of the Lomb-Scargle periodogram and related
 approaches, what practical issues are involved in its effective use?
+The following sections will identify several of the important issues and
+questions that are often ignored in the literature, but are nevertheless
+vital to using the periodogram in practice.
 
@@
 \sectlabel{frequency-grid}
-The choice of frequency grid for a Lomb-Scargle analysis is an important one
+The frequency grid used for a Lomb-Scargle analysis is an important choice
 that is too-often glossed-over, probably because the choice is so
@@
 periodogram up to some well-motivated limiting frequency $f_{max}$. This could
-be a true Nyquist limit based on the \citep{Eyer99} definition, a pseudo-Nyquist
+be a true Nyquist limit based on the \citet{Eyer99} definition, a pseudo-Nyquist
 limit based on careful scrutiny of the window function (cf. \fig{kepler-data}),
-or a limiting frequency based on the integration time of individual
-observations.
+a limiting frequency based on the integration time of individual
+observations, or a limit based on prior knowledge of the kinds of signals
+you expect to detect.
 
@@
 the frequencies between the limits.
-This turns out to be quite important: if you choose too fine a grid, it can
+This choice turns out to be vital: if you choose too fine a grid, it can
 lead to unnecessarily long computation times that can add up quickly in the
@@
 Because the spacing of the 10,000-point grid is much larger than the width of
-the periodogram peaks, the computation {\it entirely misses} the most
+the periodogram peaks, the analysis {\it entirely misses} the most
 important peaks in the periodogram!
@@
 ${\sim}1/T$.
-To ensure that our grid sufficiently covers each peak, it is prudent to
-over-sample by some factor---say $n_{over}$ samples per peak---and
-use a grid of size $\Delta f = \frac{1}{n_{over} T}$.
+To ensure that our grid sufficiently samples each peak, it is prudent to
+over-sample by some factor---say $n_o$ samples per peak---and
+use a grid of size $\Delta f = \frac{1}{n_o T}$.
 This pushes the total number of required periodogram evaluations to
@@
 evaluations of the periodogram.
-
 On the other hand, the LINEAR data shown in \fig{LINEAR-data} does not have
@@
 the 10-year data up to a limiting frequency defined by the 30-second
-exposure time for each observation will require roughly 2.5 million
-periodogram evaluations per object, which for fast Lomb-Scargle implementations
-(see the next section) can be accomplished in seconds on a single core.
+exposure time for each observation would require roughly 25 million
+periodogram evaluations per object, which for fast implementations
+(see the next section) could be accomplished in several seconds on
+a single core.
 
@@
     Each has 60 irregular observations over 180 nights,
-    with a cadence typical of ground-based surveys ({i.e.} showing a strong
+    with a cadence typical of ground-based surveys (i.e., showing a strong
     diurnal window pattern similar to \fig{LINEAR-window}).
@@
   P_{obs} = \left(\frac{1}{P_{true}} + n\ \delta f\right)^{-1}
+  \eqlabel{alias-periods}
 \end{equation}
 where, to be clear, $n$ is a positive or negative integer and $\delta f$ is
-a strong feature in the window function (here, 1 day).
-For this data, close to 36\% of the objects are mis-characterized along these
-failure modes.
+the frequency of a strong feature in the window  (here, $1\ {\rm day}^{-1}$).
+For the simulated data, close to 36\% of the objects are mis-characterized
+along these failure modes.
 
@@
 $f_0$ isolates a higher harmonic of that fundamental frequency, such as $2f_0$.
-This may occur for signals that are not strictly sinusoidal, and this harmonic
-frequency peak is subject to the same windowing effects as above.
-Thus we can describe the failure modes depicted in panel (b) of \fig{failure-modes} with the following equation, for positive integers $m > 0$:
+This may occur for periodic signals that are not strictly sinusoidal, and so
+has power at a higher harmonics.
+Additionally, this higher harmonic peak can be subject to the same aliasing
+effects as in \eq{alias-periods}.
+Thus we can adapt \eq{alias-periods} and describe the failure modes depicted
+in panel (b) of \fig{failure-modes} with the following equation,
+for positive integers $m > 0$:
 \begin{equation}
   P_{obs} = \left(\frac{m}{P_{true}} + n\delta f\right)^{-1}
+  \eqlabel{alias-harmonic-periods}
 \end{equation}
@@
 This is closely related to the effects shown in panels (a) and (b),
-but comes from the fact that
-the Lomb-Scargle periodogram at frequency $f_0$ contains contributions from
-signals at both $+f_0$ and $-f_0$.
-When diurnal aliasing leads to power at negative frequencies, this is reflected
-at the absolute value of that frequency, so it is more accurate to take the
-absolute value of the above formula
+but comes from the even symmetry of the Lomb-Scargle periodogram.
+Every periodogram peak at frequency $f_0$ has a corresponding peak at $-f_0$,
+and this is true of aliases as well.
+The aliases of negative-frequency peaks can cross into the positive-frequency
+range, and this can be accounted for by a further modification of
+\eq{alias-harmonic-periods}---taking its absolute value:
 \begin{equation}
@@
   \item Check for peaks at $|f_{peak} \pm n\delta f|$ for at least
-    $n \in \{1, 2\}$. If these aliases exist, then it is likely that you have
+    $n \in \{1, 2\}$. If these aliases exist, then it is possible that you have
     found a peak on the sequence of expected aliases---though keep in mind that
@@
 For noisy observations, this procedure cannot generally guarantee that you
-have found the correct peak, but it is far preferable than making the mistake
-of blindly trusting the highest peak in the periodogram!
+have found the correct peak, but it is far preferable to the simplistic
+approach of blindly trusting the highest peak in the periodogram!
 We will come back to the question of uncertainty in the periodogram result
@@
 window power to understand features of the resulting periodogram.
+Here we will consider the computation of the window function, as well
+as the possibility of recovering the true periodogram by deconvolving
+the window.
 
@@
 \begin{equation}
-  \mathcal{F}\{W_{\{t_n\}}\}(f) = \sum_{n=1}^{N} e^{-2\pi i f t_n}
-  \eqlabel{nonuniform-window-FT}
-\end{equation}
-For uniformly-spaced observing times $t_n$, the symmetry leads to cancellations
-so that the window transform is a clean Dirac comb.
-For non-uniform observing times, this symmetry is broken and the window
-transform will be much more ``nosiy''.
-for observation times $\{t_n\}$ is
-\begin{equation}
   \mathcal{P}_W(f;\{t_n\}) = \left|\sum_{n=1}^{N} e^{-2\pi i f t_n}\right|^2
@@
 Comparing this to \eq{schuster-periodogram}, we see that this is essentially
-the Schuster periodogram for data $g_n=1$ at all times $t_n$;
-with that in mind one quick way to compute a qualitatively correct window
-power is to use a standard Lomb-Scargle periodogram on a series of unit
-measurements, making sure to {\it not} pre-center the data or to use a
+the Schuster periodogram for data $g_n=1$ at all times $t_n$.
+With this fact in mind, one convenient way to estimate the form of the
+window power is to compute a standard Lomb-Scargle periodogram on a series of
+unit measurements, making sure to {\it not} pre-center the data or to use a
 floating mean model.
-As \citep{Scargle82} notes, this is {\it not} the true window---it differs
+As \citet{Scargle82} notes, this computation does {\it not} give the true
+window---it differs
 from the true window just as the Lomb-Scargle periodogram differs from the
@@
 use it {\it quantitatively} to remove spurious peaks from the observed
-power spectrum
+power spectrum.
 Recall from \eq{g-nonuniform} that the observed data is a point-wise product
@@
 \begin{eqnarray}
+  g(t) &=& g_{obs}(t) / W_{\{t_n\}}(t)\nonumber\\
   \mathcal{F}\{g\} &=& \mathcal{F}\{g_{obs}/W_{\{t_n\}}\}\nonumber\\
@@
 
-The latter two issues could be addressed by focusing on the non-uniform
+The latter two issues could be remedied by focusing on the non-uniform
 Fourier transform and the Schuster periodogram rather than the Lomb-Scargle
-approach, but one would lose the advantageous statistical properties of the
-resulting transform, as well as the correspondance with least-squares fitting
+modification, but doing so would jettison the useful statistical properties
+of Lomb-Scargle, as well as the correspondance with least-squares fitting
 that allows useful extensions to models handling a floating mean and
-observational noise.
-Even in this case, though, the first issue remains quite problematic.
+observational noise, and would still not address the far more problematic
+first issue.
 My feeling is that there is room for a more principled approach to the
@@
 cadence of $\overline{\delta t} = N/T$.
-From \eq{n-eval} we see that the number of frequencies evaluated is
+From \eq{n-eval} we see that the number of frequencies we need to evaluate is
 $N_f \propto T f_{max} = N f_{max} / \overline{\delta t}$.
@@
 Due to the trigonometric functions involved, this turns out to be a rather
-``expensive'' $\mathcal{O}(N^2)$, which makes the straightforward
+``expensive'' $\mathcal{O}(N^2)$, which makes the direct
 implementation impractical for even modestly-sized datasets.
@@
 
-All periodograms in this paper were computed using the Astropy
+All periodograms in this paper were computed using the
+Astropy\footnote{\url{http://docs.astropy.org/en/stable/stats/lombscargle.html}}
 implementations of the \citet{Press89}, \citet{Zechmeister09}, and

### 2017-02-06 PracticalLombScargle.tex
Update uncertainty discussion

@@
 typically set by {\it the precision of the time measurements}
-\citep[see][for a more rigorous treatment of this result]{Koen2006}.
+\citep[see][for more rigorous treatments
+  of this result]{Bretthorst2003, Koen2006}.
 
@@
 \begin{equation}
-  P(f) = \frac{N}{2}\big[\hat{\chi}^2_0 - \hat{\chi}^2(f)\big]
+  P(f) = \frac{1}{2}\big[\hat{\chi}^2_0 - \hat{\chi}^2(f)\big]
   \eqlabel{lomb-scargle-chi2}
@@
 
+\subsection{Uncertainties in Periodogram Results}
+\sectlabel{uncertainties}
+An important piece of reporting results from the Lomb-Scargle periodogram is
+the uncertainty of the resulting period estimate.
+In most areas of science we are used to being able to report uncertainties in
+terms of errorbars, {e.g.} ``the period is $16.3 \pm 0.6$ hours''.
+For periods derived from the Lomb-Scargle periodogram, however, uncertainties
+generally cannot be meaningfully expressed in this way:
+as we saw in the discussion of failure modes in \sect{failure-modes},
+the concern for periodograms is more often inaccuracy associated with
+false peaks and aliases, rather than with imprecision in peak location.
+
+\subsubsection{Peak Width and Frequency Precision}
+
+A periodic signal will be reflected in the periodogram by the presence of a
+peak of a certain width and height.
+In the Fourier view, the precision with which a peak's frequency can be
+identified is directly related to the width of the peak; often the
+width at half-maximum is used.
+This can be formalized more precisely in the least squares interpretation of
+the periodogram, in which the inverse of the curvature of the peak is
+identified with the uncertainty \citep{ICVG2014}---which in the Bayesian
+view amounts to fitting a Gaussian curve to the (exponentiated)
+peak \citep{Jaynes1987, Bretthorst88}---and this introduces some first-order
+dependence on the number of samples and their signal-to-noise ratio rather
+than the peak width alone.
+
+We have motivated from the Fourier window discussions that peak width
+scales with the observational baseline; the surprising result is that to
+first order the peak width in the periodogram {\it does not depend} on
+either the number of observations or their signal-to-noise ratio!
+This can be seen visually in \fig{peak-width-height},
+which shows periodograms for simulated data with a fixed 4-period baseline,
+with varying sample sizes and signal-to-noise ratios.
+In all cases, the widths of the primary peak are essentially {\it identical},
+regardless of the quality or quantity of data!
+Data quality and quantity are reflected in the height of the peak in relation
+to the ``background noise'', which speaks to the peak {\it significance} rather
+than the precision of frequency estimation.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig26_peak_width_height}
+  \caption{The effect of the number of points $N$ and the signal-to-noise
+    ratio $S/N$ on the expected width and height of the periodogram peak.
+    Top panels show the normalized periodogram (\eq{lomb-scargle-normalized}),
+    while bottom panels show the PSD-normalized periodogram
+    (\eq{lomb-scargle-periodogram})
+    Somewhat surprisingly, neither the number of points nor the signal-to-noise
+    ratio affects the peak width \figlabel{peak-width-height}}
+\end{figure}
+
+Because of this, if you insist on reporting frequency errorbars derived from
+peak width, the results can be pretty silly for observations with
+long baselines.
+For example, going by the peak width in \fig{LINEAR-power}, the
+periodogram reveals a period of $2.58014 \pm 0.00004$ hours.
+While this is an accurate characterization of the period precision
+{\it assuming that peak is the correct one}, it does not capture the
+much more relevant uncertainties demonstrated in \sect{failure-modes},
+in which we might find ourselves on the wrong peak entirely!
+This is why in general, peak widths and Gaussian error bars are not all that
+useful for reporting useful or accurate uncertainties in the context of
+periodogram analysis.
+
+\subsubsection{False Alarm Probability}
+A much more relevant quantity for expressing uncertainty of periodogram results
+is the height of the peak, and in particular the height compared to the
+background ``noise'' displayed in the periodogram.
+\fig{peak-width-height} indicates that this property {\it does} depend on
+both the number of observations and their signal-to-noise ratio: for the
+small-$N$/low signal-to-noise cases, the spurious peaks in the background
+become comparable to the size of the true peak.
+In fact, as we saw in \sect{schuster-to-lomb-scargle}, the ability to
+analytically define and quantify the relationship between peak height
+and significance is one of the primary considerations that led to the
+standard form of the Lomb-Scargle periodogram.
+
+The typical approach to quantifying the significance of a peak is the {\it
+False Alarm Probability} (FAP), which measures the probability that a dataset
+with no signal would---due to coincidental alignment among the random
+errors---lead to a peak of a similar magnitude.
+\citet{Scargle82} showed that for Gaussian noise, the values of the unnormalized
+periodogram in \eq{lomb-scargle-periodogram} are exponentially distributed;
+that is, for {\it one single frequency} $f_0$, if $Z = P(f_0)$ is the
+periodogram value from \eq{lomb-scargle-periodogram}, then
+\begin{equation}
+  P_{single}(Z) = 1 - \exp(-Z)
+  \eqlabel{LS-prob-z}
+\end{equation}
+is the cumulative probability of observing a periodogram value less than $Z$,
+in data with no periodic signal
+(be aware that for different periodogram normalizations,
+the form of this distribution changes; see \citet{Cumming99} or
+\citet{Baluev2008} for a summary for various normalizations).
+
+\subsubsubsection{Independent Frequency Method}
+
+We are generally not interested in the distribution of one particular
+randomly-chosen frequency,
+but rather the distribution of {\it the highest peak} of the
+periodogram, which is a quite different situation.
+If this is confusing, consider rolling a standard 6-sided die.
+The probability, in a single roll, of rolling a number, say $r > 4$ is easy
+to compute: it's 2 sides out of six, or $p(r>4) = 1/3$.
+If, on the other hand, you roll 10 dice and choose the largest number among
+them, the probability that it will be greater than 4 is {\it far} larger
+than $1/3$; it is $p(max(r,10) > 4) = 1 - (1 - 1/3)^{10} \approx 0.98$.
+The case for the periodogram is analogous: the probability of seeing a high peak
+simply by chance at any random location (\eq{LS-prob-z}) is relatively small,
+but the probability of seeing a peak when scanning a large number of
+frequencies is much higher.
+
+With the dice, the calculation is easy because the rolls are independent: the
+result of one roll does not affect the result of the next.
+With the periodogram, though, the value at one frequency is correlated with
+the value at other frequencies in a way that is quite difficult to analytically
+quantify---these correlations come from the convolution with the survey
+window.
+Nonetheless, one common approach to estimating the distribution has been to
+assume it can be modeled on some ``effective number'' of independent
+frequencies $N_{eff}$, so that the FAP can be estimated as
+\begin{equation}
+  FAP(z) \approx 1 - \big[P_{single}(z)\big]^{N_{eff}}
+  \eqlabel{FAP-neff}
+\end{equation}
+A very simple estimate for $N_{eff}$ is based on our arguments of the
+expected peak width, $\delta f = 1/T$.
+In this approximation, the number of independent peaks in a range
+$0 \le f \le f_{max}$ is assumed to be $N_{eff} = f_{max} T$.
+There have been various attempts to estimate this value more carefully,
+both analytically and via simulations
+\citep[see, e.g.][]{Horne86,Schwarzenberg-Czerny98,Cumming04,Frescura08},
+but all such approaches are necessarily only approximations.
+
+\subsubsubsection{\citet{Baluev2008} Method}
+
+A more sophisticated treatment of the problem is that of \citet{Baluev2008},
+who derived an analytic result based on theory regarding extreme values
+of stochastic processes.
+The Baluev result is not an exact measure of the FAP, but rather an upper
+limit that is shown to be quite close to the true FAP even
+with highly structured window functons.
+For the PSD-normalized periodogram, \citet{Baluev2008} showed that the following
+formula for the FAP provides a close upper-limit even in the case of
+highly structured window functions:
+\begin{equation}
+  FAP(z) \approx 1 - P_{single}(z)e^{-\tau(z)}
+  \eqlabel{FAP-baluev}
+\end{equation}
+where, for the normalized periodogram (\eq{lomb-scargle-normalized}),
+\begin{equation}
+  \tau(z) \approx W (1 - z)^{(N - 4)/2}\sqrt{z}
+\end{equation}
+and $W = f_{max} \sqrt{4\pi\ {\rm var}(t)}$ measures the effective width of the
+observing window in units of the maximum frequency chosen for the analysis.
+This result is not an exact measure of the FAP, but rather an
+upper limit that is valid for alias-free periodograms, {i.e.} cases
+where the window function has very little structure,
+but which holds quite well even in the case of more realistic survey windows.
+
+\subsubsubsection{Bootstrap Method}
+In the absence of a true analytic solution to the False Alarm Probability, we
+can turn to computational methods such as the bootstrap.
+The bootstrap method is a technique in which the statistic in question is
+computed repeatedly on many random resamplings of the data in order to
+approximate the distribution of that statistic
+\citep[see][for a useful general discussion of this technique]{ICVG2014}.
+For the periodogram, in each resampling we keep the temporal coordinates
+the same, draw observations randomly with replacement from the observed
+values, and then compute the maximum of the resulting periodogram.
+For enough resamplings, the distribution of these maxima will approximate
+the true distribution for the case with no periodic signal present.
+The bootstrap produces the most robust estimate of the FAP because it makes
+no assumptions about statistical distributions, and fully accounts for survey
+window effects.
+
+Unfortunately, the computational costs of the bootstrap can be quite
+prohibitive: to accurately measure small levels of FAP requires constructing
+a large number of full periodograms.
+If you are probing a false positive rate of $r$ among $N$ bootstrap reseamples,
+you would roughly expect to find $rN \pm \sqrt{rN}$ relevant peaks in your
+bootstrap sample.
+More concretely, for 1,000 bootstrap samples, you'll find only $\sim 10\pm 3$
+peaks reflecting a 1\% false positive rate.
+The random fluctuations in that count translate directly to an inability to
+accurately estimate the false positive rate at that level.
+A good rule-of-thumb is that to accurately characterize a false positive rate
+$r$ requires
+something like $\sim 10 / r$ individual periodograms to be computed---this grows
+computationally intractable quite quickly.
+The bootstrap is also not universally applicable: for example, it does not
+correctly account for cases where noise in observations is correlated;
+for more discussion of the bootstrap approach and its caveats, see
+\citet{ICVG2014} and references therein.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig27_FAP_bootstrap}
+  \caption{Comparison of various approaches to computing the false alarm
+    probability (FAP) for simulated observations with both structured and
+    unstructured survey windows. Note that the approximate $N_{eff}$ method
+    in its most na{\"i}ve form tends to underestimate the bootstrapped FAP,
+    while the \citet{Baluev2008} method tends to overestimate the bootstrapped
+    FAP, particularly for data with a highly-structured survey window.
+    \figlabel{FAP-bootstrap}}
+\end{figure}
+
+\fig{FAP-bootstrap} compares these methods of estimating the false alarm
+probability, for both an unstructured survey window, and a structured window
+which produces the kinds of aliasing we have discussed above.
+The na{\"i}ve approximation in this case under-estimates the FAP at nearly all
+levels---so, for example, it might lead you to think a peak has a FAP of 10\%
+when in fact it is closer to 30\%.
+The Baluev method, by design, over-estimates the FAP, and is quite close to
+the bootstraped distribution in the case of an unstructured window.
+For a highly structured window, the Baluev method does not perform as well,
+but still tends to over-estimate the FAP---so, for example, it might lead you
+to think a peak as a FAP of 10\% when in fact it is closer to 5\%.
+
+Finally, I will note that \citet{Suveges12} has shown the promise of a
+hybrid approach of the bootstrap method and the extreme value statistics
+of \citet{Baluev2008}; this has the ability to compute accurate FAPs
+without the need to compute thousands of full bootstrapped periodograms.
+
+\subsubsubsection{What Does the False Alarm Probability Measure?}
+
+While the False Alarm Probability can be a useful concept, you should be
+careful to keep in mind that it answers a very specific question:
+{\it ``what is the probability that a signal with no periodic component would
+  lead to a peak of this magnitude?''}.
+In particular, it emphatically does {\it not} answer the much more relevant
+question {\it ``what is the probability that this data is periodic
+  given these observations?''.}
+This boils down to a case of $P(A~|~B) \ne P(B~|~A)$, but still it is common to
+see the false alarm probability treated as if it speaks to the second rather
+than the first question.
+In other words, the false alarm probability tells you the false positive rate
+in a frequentist, rather than Bayesian sense: that is, it is a false positive
+rate conditioned on the entire space of possible observations rather than on
+the particular data you observed.
+
+Similarly, the false alarm probability tells us nothing about the false
+negative rate; {i.e.} the rate at which
+we would expect to miss a periodic signal that is, in fact, present.
+It also tells us nothing about the error rate; {i.e.} the rate at which we
+would expect to identify an incorrect alias of a signal present in our data
+(as we saw in \sect{failure-modes}).
+Too often users are tempted to interpret the FAP too broadly, with the hope that
+it would speak to questions that are beyond its reach.
+
+To the best of my knowledge, there is currently no silver bullet for answering
+these broader, more relevant questions of uncertainty of Lomb-Scargle results.
+Perhaps the most fruitful path toward understanding of such effects for a
+particular set of observations---with particular noise characteristics and a
+particular observing window---is via simulated data injected into the detection
+pipeline.
+In fact, this type of simulation is a vital component of most---I might even
+venture to say all---current and future time-domain surveys that seek to
+detect, characterize, and catalog periodic objects,
+regardless of the particular approach used to identify periodicity
+\citep[see, e.g.][etc.]{opsim1, opsim2, Ivezic08LSST, Sesar2010,
+  Oluseyi12, McQuillan2012, Drake2014}
+
+\subsection{Periodogram Normalization and Interpretation}
+\sectlabel{normalizations}
+
+Often it is useful to be able to interpret the periodogram values themselves.
+Recall that there are two equally valid perspectives of what the periodogram
+is measuring---the Fourier view and the least-squares view---and each of these
+lends itself to a different approach to normalizion and interpretation of the
+periodogram.
+
+\subsubsection{PSD normalization}
+When considering the periodogram from the Fourier perspective,
+it is useful to normalize the periodogram such that in the special case
+of equal spaced data, it recovers the standard Fourier power spectrum.
+This is the normalization used in \eq{lomb-scargle-periodogram} and the
+equivalent least-squares expression seen in \eq{lomb-scargle-chi2}:
+\begin{equation}
+  P(f) = \frac{1}{2}\big[\hat{\chi}^2_0 - \hat{\chi}^2(f)\big]
+  \eqlabel{lomb-scargle-chi2-2}
+\end{equation}
+For equally-spaced data, this periodogram is identical to the standard
+definition based on the fast Fourier transform:
+\begin{equation}
+  P(f) = \frac{1}{N} \left| FFT(y_n) \right|^2
+\end{equation}
+in particular, this means that the units of the periodogram are $unit(y)^2$,
+and can be interpreted as squared-amplitudes of the Fourier component at
+each frequency.
+Note that this changes if data uncertainties are incorporated into
+the periodogram as in \sect{extensions-observational-noise},
+the periodogram in this normalization becomes unitless: it is essentially
+a measure of periodic content in signal-to-noise ratio
+rather than in signal itself.
+
+\subsubsection{Least-squares normalization}
+In the least-squares view of the periodogram, the periodogram is interpreted
+as an inverse measure of the goodness-of-fit for a model.
+When we express the periodogram as a function of $\chi^2$ model fits as in
+\eq{lomb-scargle-chi2-2}, it becomes clear that the periodogram has a
+mathematical maximum value: if the sinusoidal model {\it perfectly} fits
+the data at some frequency $f_0$, then $\hat{\chi}^2(f_0) = 0$ and the
+periodogram is maximized at a value of $N\hat{\chi}^2_0 / 2$.
+On the other end, it is mathematically impossible for a sinusoidal model to
+do worse than the simple constant, non-varying fit reflected in
+$\hat{\chi}^2_0$, and so the minimum value of \eq{lomb-scargle-chi2-2} is
+exactly zero.
+
+This suggests a different normalization of the periodogram, where the values
+are dimensionless and lie between 0 and 1:
+\begin{equation}
+  P_{norm}(f) = 1 - \frac{\hat{\chi}^2(f)}{\hat{\chi}^2_0}
+  \eqlabel{lomb-scargle-normalized}
+\end{equation}
+This ``normalized periodogram'' is unitless, and
+is directly proportional to the unnormalized
+(or PSD-normalized) periodogram in \eq{lomb-scargle-chi2-2}.
+
+While the normalization does not affect the shape of the periodogram,
+its main practical consequence is that the statistics of the resulting
+periodogram differ slightly, and this needs to be taken into account when
+computing analytic estimates of uncertainties and false alarm probabilities
+(see the next section).
+Other normalizations exist as well, but seem to be rarely used in practice.
+For a concise summary of periodogram normalizations and their statistical
+properties, refer to the introduction of \citet{Baluev2008}.
+
 \subsection{Algorithmic Considerations}
@@
 All code behind this paper is available in the form of Jupyter notebooks in the
-paper's GitHub repository at \url{http://github.com/jakevdp/PracticalLombScargle/}.
-
-\subsection{Periodogram Normalization and Interpretation}
-\sectlabel{normalizations}
-
-Often it is useful to be able to interpret the periodogram values themselves.
-Recall that there are two equally valid perspectives of what the periodogram
-is measuring---the Fourier view and the least-squares view---and each of these
-lends itself to a different approach to normalizion and interpretation of the
-periodogram.
-
-\subsubsection{PSD normalization}
-When considering the periodogram from the Fourier perspective,
-it is useful to normalize the periodogram such that in the special case
-of equal spaced data, it recovers the standard Fourier power spectrum.
-This is the normalization used in \eq{lomb-scargle-periodogram} and the
-equivalent least-squares expression seen in \eq{lomb-scargle-chi2}:
-\begin{equation}
-  P(f) = \frac{N}{2}\big[\hat{\chi}^2_0 - \hat{\chi}^2(f)\big]
-  \eqlabel{lomb-scargle-chi2-2}
-\end{equation}
-For equally-spaced data, this periodogram is identical to the standard
-definition based on the fast Fourier transform:
-\begin{equation}
-  P(f) = \frac{1}{N} \left| FFT(y_n) \right|^2
-\end{equation}
-in particular, this means that the units of the periodogram are $unit(y)^2$,
-and can be interpreted as squared-amplitudes of the Fourier component at
-each frequency.
-Note that this changes if data uncertainties are incorporated into
-the periodogram as in \sect{extensions-observational-noise},
-the periodogram in this normalization becomes unitless: it is essentially
-a measure of periodic content in signal-to-noise ratio
-rather than in signal itself.
-
-\subsubsection{Least-squares normalization}
-In the least-squares view of the periodogram, the periodogram is interpreted
-as an inverse measure of the goodness-of-fit for a model.
-When we express the periodogram as a function of $\chi^2$ model fits as in
-\eq{lomb-scargle-chi2-2}, it becomes clear that the periodogram has a
-mathematical maximum value: if the sinusoidal model {\it perfectly} fits
-the data at some frequency $f_0$, then $\hat{\chi}^2(f_0) = 0$ and the
-periodogram is maximized at a value of $N\hat{\chi}^2_0 / 2$.
-On the other end, it is mathematically impossible for a sinusoidal model to
-do worse than the simple constant, non-varying fit reflected in
-$\hat{\chi}^2_0$, and so the minimum value of \eq{lomb-scargle-chi2-2} is
-exactly zero.
-
-This suggests a different normalization of the periodogram, where the values
-are dimensionless and lie between 0 and 1:
-\begin{equation}
-  P_{norm}(f) = 1 - \frac{\hat{\chi}^2(f)}{\hat{\chi}^2_0}
-  \eqlabel{lomb-scargle-normalized}
-\end{equation}
-This ``normalized periodogram'' is unitless, and
-is directly proportional to the unnormalized
-(or PSD-normalized) periodogram in \eq{lomb-scargle-chi2-2}.
-
-While the normalization does not affect the shape of the periodogram,
-its main practical consequence is that the statistics of the resulting
-periodogram differ slightly, and this needs to be taken into account when
-computing analytic estimates of uncertainties and false alarm probabilities
-(see the next section).
-Other normalizations exist as well, but seem to be rarely used in practice.
-For a concise summary of periodogram normalizations and their statistical
-properties, refer to the introduction of \citet{Baluev2008}.
-
-\subsection{Uncertainties in Periodogram Results}
-\sectlabel{uncertainties}
-An important piece of reporting results from the Lomb-Scargle periodogram is
-the uncertainty of the resulting period estimate.
-In most areas of science we are used to being able to report uncertainties in
-terms of errorbars, {e.g.} ``the period is $16.3 \pm 0.6$ hours''.
-For periods derived from the Lomb-Scargle periodogram, however, uncertainties
-generally cannot be meaningfully expressed in this way:
-as we saw in the discussion of failure modes in \sect{failure-modes},
-the concern for periodograms is more often errors associated with
-false peaks and aliases, rather than with imprecision in peak location.
-
-\subsubsection{Peak Width and Frequency Precision}
-
-A periodic signal will be reflected in a periodogram by the presence of a
-peak of a certain height and width.
-The width of the peak is directly related to the precision with which that
-peak's frequency is measured; the height of the peak 
-is related to the significance of the detection.
-We have motivated from Fourier window discussions that the peak width scales
-with the observational baseline; the surprising result is that the peak width
-{\it does not depend} on either the number of observations or the noise
-in those observations!
-
-This can be proven mathematically \citep[see, e.g.][]{Bretthorst88, ICVG2014},
-but it is perhaps more intuitive to see this visually.
-\fig{peak-width-height} shows periodograms for simulated data with a fixed
-baseline $T=4$ units covering four cycles, with varying sample sizes and
-signal-to-noise ratios.
-In all cases, the width of the peaks (most clearly visible in one or another
-normalization) are essentially {\it identical}, regardless of the quality
-or quantity of data!
-Evidently, peak width ({i.e.} frequency precision) is not a particularly
-useful measure of uncertainty in the detection of periods.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{fig26_peak_width_height}
-  \caption{The effect of the number of points $N$ and the signal-to-noise
-    ratio $S/N$ on the expected width and height of the periodogram peak.
-    Top panels show the normalized periodogram (\eq{lomb-scargle-normalized}),
-    while bottom panels show the PSD-normalized periodogram
-    (\eq{lomb-scargle-periodogram})
-    Somewhat surprisingly, neither the number of points nor the signal-to-noise
-    ratio affects the peak width; nor do they affect the peak height under
-    the appropriate normalization.
-    \figlabel{peak-width-height}}
-\end{figure}
-
-If you insist on reporting precision derived from a peak width
-in terms of an errorbar, the results can be pretty silly for observations
-with long baselines.
-For example, going by the peak width in \fig{LINEAR-power}, the
-periodogram reveals a period of $2.58024 \pm 0.00005$ hours.
-While this is an accurate characterization of the period precision
-{\it assuming that peak is the correct one}, it does not capture the
-much more relevant uncertainties demonstrated in \sect{failure-modes},
-in which we might find ourselves on the wrong peak entirely!
-This is why in general, peak widths and Gaussian error bars are not all that
-useful for reporting useful or accurate uncertainties in the context of
-periodogram analysis.
-
-\subsubsection{False Alarm Probability}
-A much more relevant quantity for expressing uncertainty of periodogram results
-is the height of the peak, and in particular the height compared to the
-background ``noise'' displayed in the periodogram.
-\fig{peak-width-height} indicates that this property {\it does} depend on
-both the number of observations and their signal-to-noise ratio: for the
-small-$N$/low signal-to-noise cases, the spurious peaks in the background
-become comparable to the true peak.
-In fact, as we saw in \sect{schuster-to-lomb-scargle}, the ability to
-analytically define and quantify the relationship between peak height
-and significance is one of the considerations that led to the standard
-form of the Lomb-Scargle periodogram.
-
-The typical approach to quantifying the significance of a peak is the {\it
-False Alarm Probability} (FAP), which measures the probability that a dataset
-with no signal would---due to coincidental alignment among the random
-errors---lead to a peak of a similar magnitude.
-\citet{Scargle82} showed that for Gaussian noise, the values of the unnormalized
-periodogram in \eq{lomb-scargle-periodogram} are exponentially distributed;
-that is, for {\it one single frequency} $f_0$, if $Z = P(f_0)$ is the
-periodogram value from \eq{lomb-scargle-periodogram}, then
-\begin{equation}
-  P_{single}(Z) = 1 - \exp(-Z)
-  \eqlabel{LS-prob-z}
-\end{equation}
-is the cumulative probability of observing a periodogram value less than $Z$,
-in data with no periodic signal
-(be aware that for different periodogram normalizations,
-the form of this distribution changes; see \citet{Cumming99} or
-\citet{Baluev2008} for a summary for various normalizations).
-
-\subsubsubsection{Independent Frequency Method}
-
-We are generally not interested in the distribution of one particular
-randomly-chosen frequency,
-but rather the distribution of {\it the highest peak} of the
-periodogram, which is a quite different situation.
-If this is confusing, consider rolling a standard 6-sided die.
-The probability, in a single roll, of rolling a number, say $r > 4$ is easy
-to compute: it's 2 sides out of six, or $p(r>4) = 1/3$.
-If, on the other hand, you roll 10 dice and choose the largest number among
-them, the probability that it will be greater than 4 is {\it far} larger
-than $1/3$; it is $p(max(r,10) > 4) = 1 - (1 - 1/3)^{10} \approx 0.98$.
-The case for the periodogram is analogous: the probability of seeing a high peak
-simply by chance at any random location (\eq{LS-prob-z}) is relatively small,
-but the probability of seeing a peak when scanning a large number of
-frequencies is much higher.
-
-With the dice, the calculation is easy because the rolls are independent: the
-result of one roll does not affect the result of the next.
-With the periodogram, though, the value at one frequency is correlated with
-the value at other frequencies in a way that is quite difficult to analytically
-quantify---these correlations come from the convolution with the survey
-window.
-Nonetheless, one common approach to estimating the distribution has been to
-assume it can be modeled on some ``effective number'' of independent
-frequencies $N_{eff}$, so that the FAP can be estimated as
-\begin{equation}
-  FAP(z) \approx 1 - \big[P_{single}(z)\big]^{N_{eff}}
-  \eqlabel{FAP-neff}
-\end{equation}
-A very simple estimate for $N_{eff}$ is based on our arguments of the
-expected peak width, $\delta f = 1/T$.
-In this approximation, the number of independent peaks in a range
-$0 \le f \le f_{max}$ is assumed to be $N_{eff} = f_{max} T$.
-There have been various attempts to estimate this value more carefully,
-both analytically and via simulations
-\citep[see, e.g.][]{Horne86,Schwarzenberg-Czerny98,Cumming04,Frescura08},
-but all such approaches are necessarily only approximations.
-
-\subsubsubsection{\citet{Baluev2008} Method}
-
-A more sophisticated treatment of the problem is that of \citet{Baluev2008},
-who derived an analytic result based on the theory of stochastic processes.
-The Baluev result is not an exact measure of the FAP, but rather an upper
-limit that is shown to be quite close to the true FAP even
-with highly structured window functons.
-For the PSD-normalized periodogram, \citet{Baluev2008} showed that the following
-formula for the FAP provides a close upper-limit even in the case of
-highly structured window functions:
-\begin{equation}
-  FAP(z) \approx 1 - P_{single}(z)e^{-\tau(z)}
-  \eqlabel{FAP-baluev}
-\end{equation}
-where, for the normalized periodogram (\eq{lomb-scargle-normalized}),
-\begin{equation}
-  \tau(z) \approx W (1 - z)^{(N - 4)/2}\sqrt{z}
-\end{equation}
-and $W = f_{max} \sqrt{4\pi\ {\rm var}(t)}$ measures the effective width of the
-observing window in units of the maximum frequency chosen for the analysis.
-This result is not an exact measure of the FAP, but rather an approximation
-to the upper limit that is valid for alias-free periodograms, {i.e.} cases
-where the window function has very little structure.
-Nonetheless, \citet{Baluev2008} showed that this approximation holds quite well
-even in the case of realistic survey windows.
-
-\subsubsubsection{Bootstrap Method}
-In the absence of a true analytic solution to the False Alarm Probability, we
-can turn to computational methods such as the bootstrap.
-The bootstrap method is a technique in which the statistic in question is
-computed repeatedly on many random resamplings of the data in order to
-approximate the distribution of that statistic
-\citep[see][for a useful general discussion of this technique]{ICVG2014}.
-For the periodogram, in each resampling we keep the temporal coordinates
-the same, draw observations randomly with replacement from the observed
-values, and then compute the maximum of the resulting periodogram.
-For enough resamplings, the distribution of these maxima will approximate
-the true distribution for the case with no periodic signal present.
-The bootstrap produces the most robust estimate of the FAP because it makes
-no assumptions about statistical distributions, and fully accounts for survey
-window effects.
-
-Unfortunately, the computational costs of the bootstrap can be quite
-prohibitive: to accurately measure small levels of FAP requires constructing
-a large number of full periodograms.
-If you are probing a false positive rate of $r$ among $N$ bootstrap reseamples,
-you would roughly expect to find $rN \pm \sqrt{rN}$ peaks in your sample.
-More concretely, for 1,000 bootstrap samples, you'll find only $\sim 10\pm 3$
-peaks reflecting a 1\% false positive rate.
-The random fluctuations in that count translate directly to an inability to
-accurately estimate the false positive rate at that level.
-A good rule-of-thumb is that to accurately characterize a FPR $r$ requires
-something like $50 / r$ individual periodograms to be computed---this grows
-computationally intractable quite quickly.
-The bootstrap is also not universally applicable: for example, it does not
-correctly account for cases where noise in observations is correlated;
-for more discussion of the bootstrap approach and its caveats, see
-\citet{ICVG2014} and references therein.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{fig27_FAP_bootstrap}
-  \caption{Comparison of various approaches to computing the false alarm
-    probability (FAP) for simulated observations with both structured and
-    unstructured survey windows. Note that the approximate $N_{eff}$ method
-    in its most na{\"i}ve form tends to underestimate the bootstrapped FAP,
-    while the \citet{Baluev2008} method tends to overestimate the bootstrapped
-    FAP, particularly for data with a highly-structured survey window.
-    \figlabel{FAP-bootstrap}}
-\end{figure}
-
-\fig{FAP-bootstrap} compares these methods of estimating the false alarm
-probability, for both an unstructured survey window, and a structured window
-which produces the kinds of aliasing we have discussed above.
-The naive approximation in this case under-estimates the FAP at nearly all
-levels---so, for example, it might lead you to think a peak has a FAP of 10\%
-when in fact it is closer to 30\%.
-The Baluev method, by design, over-estimates the FAP, and is quite close to
-the bootstraped distribution in the case of an unstructured window.
-For a highly structured window, the Baluev method does not perform as well,
-but still tends to over-estimate the FAP---so, for example, it might lead you
-to think a peak as a FAP of 10\% when in fact it is closer to 5\%.
-
-Finally, I will note that \citet{Suveges12} has shown the promise of a
-hybrid approach of the bootstrap method and the extreme value statistics
-of \citet{Baluev2008}; this has the ability to compute accurate FAPs
-without the need to compute thousands of full bootstrapped periodograms.
-
-\subsubsubsection{What Does the False Alarm Probability Measure?}
-
-While the False Alarm Probability can be a useful concept, you should be
-careful to keep in mind that it answers a very specific question:
-{\it ``what is the probability that a signal with no periodic component would
-  lead to a peak of this magnitude?''}.
-In particular, it emphatically does {\it not} answer the much more relevant
-question {\it ``what is the probability that this data is periodic
-  given these observations?''.}
-This boils down to a case of $P(A~|~B) \ne P(B~|~A)$, but still it is common to
-see the false alarm probability treated as if it speaks to the second rather
-than the first question.
-In other words, the false alarm probability tells you the false positive rate
-in a frequentist, rather than Bayesian sense: that is, it is a false positive
-rate conditioned on the entire space of possible observations rather than on
-the particular data you observed.
-
-Similarly, the false alarm probability tells us nothing about the false
-negative rate; {i.e.} the rate at which
-we would expect to miss a periodic signal that is, in fact, present.
-It also tells us nothing about the error rate; {i.e.} the rate at which we
-would expect to identify an incorrect alias of a signal present in our data
-(as we saw in \sect{failure-modes}).
-Too often users are tempted to interpret the FAP too broadly, with the hope that
-it would speak to questions that are beyond its reach.
-
-To the best of my knowledge, there is currently no silver bullet for answering
-these broader, more relevant questions of uncertainty of Lomb-Scargle results.
-Perhaps the most fruitful path toward understanding of such effects for a
-particular set of observations---with particular noise characteristics and a
-particular observing window---is via simulated data injected into the detection
-pipeline.
-In fact, this type of simulation is a vital component of most---I might even
-venture to say all---current and future time-domain surveys that seek to
-detect, characterize, and catalog periodic objects,
-regardless of the particular approach used to identify periodicity
-\citep[see, e.g.][]{opsim1, opsim2, Ivezic08LSST, Sesar2010,
-  Oluseyi12, McQuillan2012, Drake2014}
+paper's GitHub repository\footnote{\url{http://github.com/jakevdp/PracticalLombScargle/}}.
 

### 2017-02-06 PracticalLombScargle.tex
Add postscript

@@
 
-All figures in this paper were produced using Python, and in particular the
-Numpy \citep{numpy, numpybook},
-Pandas \citep{pandas},
-Astropy \citep{Astropy2013},
-and Matplotlib \citep{matplotlib} packages.
-Code to reproduce these figures is available in Jupyter notebooks at
-\url{http://github.com/jakevdp/PracticalLombScargle/}.
-
 \section{Background: the Continuous Fourier Transform}
@@
 \subsubsection{Deconvolution and CLEANing}
+\sectlabel{CLEAN}
 With this ability to compute the window function, we might hope to be able to
@@
 view amounts to fitting a Gaussian curve to the (exponentiated)
-peak \citep{Jaynes1987, Bretthorst88}---and this introduces some first-order
-dependence on the number of samples and their signal-to-noise ratio rather
-than the peak width alone.
-
-We have motivated from the Fourier window discussions that peak width
-scales with the observational baseline; the surprising result is that to
+peak \citep{Jaynes87, Bretthorst88}---and this introduces first-order
+dependence on the number of samples and their signal-to-noise ratio,
+though the uncertainty remains proportional to the peak width.
+\todo{explain this difference in terms of FWHM of exponentiated power?}
+
+We have motivated from the Fourier window discussions that peak width is
+the inverse of the observational baseline; the surprising result is that to
 first order the peak width in the periodogram {\it does not depend} on
@@
 to the ``background noise'', which speaks to the peak {\it significance} rather
-than the precision of frequency estimation.
+than the precision of frequency detection.
 
@@
     (\eq{lomb-scargle-periodogram})
-    Somewhat surprisingly, neither the number of points nor the signal-to-noise
-    ratio affects the peak width \figlabel{peak-width-height}}
-\end{figure}
-
-Because of this, if you insist on reporting frequency errorbars derived from
+    Perhaps surprisingly, neither the number of points nor the signal-to-noise
+    ratio affects the peak width. \figlabel{peak-width-height}}
+\end{figure}
+
+For this reason, if you insist on reporting frequency errorbars derived from
 peak width, the results can be pretty silly for observations with
@@
 For example, going by the peak width in \fig{LINEAR-power}, the
-periodogram reveals a period of $2.58014 \pm 0.00004$ hours.
+periodogram reveals a period of $2.58014 \pm 0.00004$ hours, a relative
+precision of about 1/1000 percent.
 While this is an accurate characterization of the period precision
@@
 in which we might find ourselves on the wrong peak entirely!
-This is why in general, peak widths and Gaussian error bars are not all that
-useful for reporting useful or accurate uncertainties in the context of
-periodogram analysis.
+This is why in general, peak widths and Gaussian error bars should generally
+be avoided when reporting uncertainties in the context of a periodogram
+analysis.
 
@@
 
+This paper has been a conceptual tour of the Lomb-Scargle periodogram, from
+its roots in Fourier analysis, to its equivalence with special cases of
+periodic analysis based on least squares model fitting and Bayesian
+analysis. I will finish here with a brief summary of the practical
+recommendations contained in these pages, along with a somewhat iconoclastic
+post-script for further thought.
+
 \subsection{Summary of Recommendations}
@@
 \subsection{Postscript: Why Lomb-Scargle?}
-
+After considering all of these practical aspects, I think it is worth
+revisiting the question of {\it why} astronomers lean toward the
+Lomb-Scargle approach rather than the (in many ways simpler) Schuster
+periodogram.
+In particular, after three decades of use, there seems to be an unrealistic
+mythology about the efficacy of the Lomb-Scargle approach among astronomers.
+It's common to see statements or implications like the following:
 \begin{itemize}
-  \item Advantages: equivalence to least squares, 
-  \item Myths: not faster than Schuster, does not correct for irregular sampling, frequencies are not independent, no savings over full model fit at each frequency \citep{ICVG2014}, straight Least Squares treatment not useful \sect{failure-modes}.
+  \item {\it Myth: The Lomb-Scargle periodogram can be computed more efficiently
+    than the Schuster periodogram.} Reality: computationally, the two
+    are quite similar, and in fact the fastest Lomb-Scargle algorithm currently
+    available is based on the Schuster periodogram computed via the
+    the NFFT algorithm (see \sect{algorithmic-considerations})
+  \item {\it Myth: The Lomb-Scargle periodogram is faster than the least squares
+    periodogram because it avoids explicitly solving for model coefficients.}
+    Reality: model coefficients can be determined with not much extra
+    computation \citep[see discussion in][]{ICVG2014}.
+  \item {\it Myth: The Lomb-Scargle periodogram allows analytic computation of
+    statistics for periodogram peaks.} Reality: while this is true at
+    individual frequencies, it is not true for the more relevant issue of
+    peak-value statistics, which must be either approximated or computed by
+    bootstrap resampling.
+  \item {\it Myth: The Lomb-Scargle periodogram corrects for aliasing due to
+    sampling and leads to independent noise at each frequency.}
+    Reality: for structured window functions, the Lomb-Scargle
+    periodogram has the same partial aliasing issues as the Schuster
+    periodogram (see \sect{failure-modes}).
+  \item {\it Myth: Bayesian analysis shows that Lomb-Scargle is the optimal
+    statistic for detecting periodic signals in data.} Reality: it rather
+    shows that Lomb-Scargle is the optimal statistic for
+    {\it fitting a sinusoid to data}, which is not the same as saying it
+    is optimal for finding a generic, non-sinusoidal signal.
 \end{itemize}
-
-\todo{wrap up with some witty \& pithy lines of encouragement.}
+With these myths debunked, what advantage remains for Lomb-Scargle?
+What advantages would we lose by instead using a Schuster periodogram,
+estimating uncertainty, significance, and false alarm probability by
+resampling and simulation, as we must for Lomb-Scargle itself?
+I think we only lose the advantage of being able to generalize to more
+complicated models in the Bayesian sense, in which case you're not really
+using Lomb-Scargle at all, but rather a generative Bayesian model for your
+data.
+The point of correspondance for a sinusoidal model is perhaps useful trivia,
+but not a reason to use that model if your data is not known to be sinusoidal.
+
+Conversely, if you use a Schuster approach rather than a Lomb-Scargle approach,
+what you gain is the ability to reason quantitatively about the effects of the
+survey window function on the resulting periodogram (see \sect{CLEAN}).
+While the deconvolution problem is well-known to be ill-posed, there may be a
+way forward for cleaned periodograms based on regularized deconvolutions.
+
+In any case, I would contend that there is ample room for practitioners to
+question the prevailing folk wisdom of the unspoken advantage of Lomb-Scargle
+over approaches based directly on the Fourier transform, such as the Schuster
+periodogram.
+
+% Edelson-Krolik + FFT -> Lomb-Scargle?
+
+\subsection{Figures and Code}
+
+All computations and figures in this paper were produced using Python,
+and in particular the
+Numpy \citep{numpy, numpybook},
+Pandas \citep{pandas},
+Astropy \citep{Astropy2013},
+and Matplotlib \citep{matplotlib} packages.
+Code to reproduce these figures is available in Jupyter notebooks at
+\url{http://github.com/jakevdp/PracticalLombScargle/}.
 
@@
 
-% Edelson-Krolik + FFT -> Lomb-Scargle?
-
 \bibliographystyle{apj}

### 2017-02-06 PracticalLombScargle.tex
add to post-script discussion

@@
 \subsubsection{False Alarm Probability}
+\sectlabel{false-alarm-probability}
+
 A much more relevant quantity for expressing uncertainty of periodogram results
@@
 
-\todo{Summarize key recommendations}
+\todo{Flesh these out}
 
 \begin{itemize}
-\item always use the floating-mean model---it has many benefits with virtually zero cost
-\item always explore the window power first, so that you know what alias structures to expect
-\item the peak of the periodogram rarely tells the entire story. Always check
-  for aliases that are expected from window structure ($f = f_0 \pm n\delta f$)
-  and from non-sinusoidal signals ($f = f_0 / m$).
+\item choose your frequency range \& grid spacing carefully, so as to not
+  miss relevant peaks.
+\item always compute the window periodogram first, so that you know what alias
+  structures to expect in the periodogram for your data.
+\item always use the floating-mean model---it has many benefits with
+  virtually zero down-sides.
+\item the highest peak of the periodogram rarely tells the entire story.
+  Always check for aliases that are expected from window structure
+  ($f = f_{max} \pm n\delta f$) and from non-sinusoidal signals
+  ($f = f_{max} / m$).
 \item avoid Bayesian methods unless you are very comfortable with the
-  assumptions they make about your signal
+  assumptions they make about the form of your signal.
 \item avoid peak widths as a proxy for uncertainty unless you are quite
@@
   keep in mind the failures of available estimates.
+\item in the context of a time-domain survey, always use simulated observations
+  to evaluate your rates of false positives, false negatives, and failures to
+  identify the correct frequency.
+\item in cases of strong aliasing, consider using Lomb-Scargle as a first pass,
+  and then select among candidate frequencies with a more detailed model of
+  the signals you expect.
 \end{itemize}
@@
 \subsection{Postscript: Why Lomb-Scargle?}
-After considering all of these practical aspects, I think it is worth
-revisiting the question of {\it why} astronomers lean toward the
+After considering all of these practical aspects of the periodogram,
+I think it is worth stepping back to
+revisit the question of {\it why} astronomers lean toward the
 Lomb-Scargle approach rather than the (in many ways simpler) Schuster
 periodogram.
-In particular, after three decades of use, there seems to be an unrealistic
-mythology about the efficacy of the Lomb-Scargle approach among astronomers.
+In particular, three decades after the entry of ``Lomb-Scargle'' into the
+astronomy community's vernacular, there seems to be a well-developed
+mythology surrounding the efficacy and efficiency of the Lomb-Scargle approach.
 It's common to see statements or implications like the following:
@@
     available is based on the Schuster periodogram computed via the
-    the NFFT algorithm (see \sect{algorithmic-considerations})
-  \item {\it Myth: The Lomb-Scargle periodogram is faster than the least squares
-    periodogram because it avoids explicitly solving for model coefficients.}
+    the NFFT algorithm (see \sect{algorithmic-considerations}).
+  \item {\it Myth: The Lomb-Scargle periodogram is faster than a direct
+    least squares periodogram because it avoids explicitly solving for
+    model coefficients.}
     Reality: model coefficients can be determined with not much extra
@@
     peak-value statistics, which must be either approximated or computed by
-    bootstrap resampling.
+    bootstrap resampling (see \sect{false-alarm-probability})
   \item {\it Myth: The Lomb-Scargle periodogram corrects for aliasing due to
@@
     periodogram has the same partial aliasing issues as the Schuster
-    periodogram (see \sect{failure-modes}).
+    periodogram, including spurious peaks due to partial aliasing, and
+    highly correlated periodogram errors (see \sect{failure-modes}).
   \item {\it Myth: Bayesian analysis shows that Lomb-Scargle is the optimal
-    statistic for detecting periodic signals in data.} Reality: it rather
+    statistic for detecting periodic signals in data.} Reality: it
     shows that Lomb-Scargle is the optimal statistic for
     {\it fitting a sinusoid to data}, which is not the same as saying it
-    is optimal for finding a generic, non-sinusoidal signal.
+    is optimal for finding a generic, non-sinusoidal signal
+    (see \sect{bayesian-periodograms}).
 \end{itemize}
 With these myths debunked, what advantage remains for Lomb-Scargle?
-What advantages would we lose by instead using a Schuster periodogram,
+What would we lose if we instead used a Schuster periodogram,
 estimating uncertainty, significance, and false alarm probability by
 resampling and simulation, as we must for Lomb-Scargle itself?
+
 I think we only lose the advantage of being able to generalize to more
@@
 using Lomb-Scargle at all, but rather a generative Bayesian model for your
-data.
-The point of correspondance for a sinusoidal model is perhaps useful trivia,
-but not a reason to use that model if your data is not known to be sinusoidal.
-
-Conversely, if you use a Schuster approach rather than a Lomb-Scargle approach,
+data based on some strong prior information about the form of your signal.
+The equivalence of Lomb-Scargle to a Bayesian sinusoidal model is perhaps
+interesting trivia, but not itself a reason to use that model if your data
+is not known {\it a priori} to be sinusoidal.
+
+Conversely, if you replace your Lomb-Scargle approach with a Schuster approach,
 what you gain is the ability to reason quantitatively about the effects of the
 survey window function on the resulting periodogram (see \sect{CLEAN}).
-While the deconvolution problem is well-known to be ill-posed, there may be a
-way forward for cleaned periodograms based on regularized deconvolutions.
+While the deconvolution problem is ill-posed, there is no reason to assume
+this is a fatal defect: ill-posed linear models are solved routinely in
+other areas of computational research, particularly by using priors or
+various forms of regularization.
+
+One clear advantage of both Lomb-Scargle and Schuster periodograms is that
+they are quite fast to compute compared to more sophisticated approaches to
+periodic analysis based on more constrained generative models or Bayesian
+non-parametric models, and in the context of large time-domain surveys can
+provide rough periodicity detections and catalog parameters in near real-time.
 

### 2017-02-06 PracticalLombScargle.tex
Add summary of recommendations

@@
     periodogram?
-  \item What assumptions is the Lomb-Scargle periodogram making about the
+  \item What assumptions does the Lomb-Scargle approach make regarding the
     shape of the unknown signal?
-  \item How should we understand the uncertainty of the peak frequency?
+  \item How should we understand and report the uncertainty of the
+    determined frequency?
 \end{enumerate}
-Quantitative answers to these sorts of questions are discussed in various
+Quantitative treatments of these sorts of questions are presented in various
 textbooks and review papers, but I have not come across any single concise
@@
 analysis. I will finish here with a brief summary of the practical
-recommendations contained in these pages, along with a somewhat iconoclastic
-post-script for further thought.
+recommendations contained in these pages, along with a somewhat opinionated
+and iconoclastic post-script for further thought.
 
 \subsection{Summary of Recommendations}
-
-\todo{Flesh these out}
-
-\begin{itemize}
-\item choose your frequency range \& grid spacing carefully, so as to not
-  miss relevant peaks.
-\item always compute the window periodogram first, so that you know what alias
-  structures to expect in the periodogram for your data.
-\item always use the floating-mean model---it has many benefits with
-  virtually zero down-sides.
-\item the highest peak of the periodogram rarely tells the entire story.
-  Always check for aliases that are expected from window structure
-  ($f = f_{max} \pm n\delta f$) and from non-sinusoidal signals
-  ($f = f_{max} / m$).
-\item avoid Bayesian methods unless you are very comfortable with the
-  assumptions they make about the form of your signal.
-\item avoid peak widths as a proxy for uncertainty unless you are quite
-  certain that your analysis has found the correct peak.
-\item avoid false alarm probabilities as a proxy for uncertainty
-  unless you make very clear {\it exactly} what kind of error they measure.
-  Estimate the FAP via bootstrap if computationally feasible, and otherwise
-  keep in mind the failures of available estimates.
-\item in the context of a time-domain survey, always use simulated observations
-  to evaluate your rates of false positives, false negatives, and failures to
-  identify the correct frequency.
-\item in cases of strong aliasing, consider using Lomb-Scargle as a first pass,
-  and then select among candidate frequencies with a more detailed model of
-  the signals you expect.
-\end{itemize}
+\sectlabel{summary}
+The previous pages contain a large amount of background and advice for working
+with the Lomb-Scargle periodogram.
+Here I will quickly summarize the steps you should take when applying this
+algorithm to a dataset:
+
+\begin{enumerate}
+  \item Choose an appropriate frequency grid: the minimum can be set to zero,
+    the maximum set based on the precision of the time measurements
+    (\sect{}), and the grid spacing set based on the temporal baseline of
+    the data (\sect{}) so as to not sample too coarsely around peaks.
+    If this grid size is computationally intractable, reduce the maximum
+    frequency based on what kinds of signals you are looking for (\sect{}).
+  \item Compute the window transform using the Lomb-Scargle periodogram,
+    by substituting $g_n=1$ for each $t_n$ and making sure not to pre-center
+    the data or use a floating-mean model (\sect{}). Examine this window
+    function for dominant features, such as daily or annual aliases (cf.
+    \fig{}) or Nyquist-like limits (cf. \fig{}).
+  \item Compute the periodogram for your data. You should always use the
+    floating-mean model (\sect{}), as it produces more robust periodograms
+    and has few if any disadvantages.
+  \item Plot the periodogram, and identify any patterns that may be caused
+    by features you observed in the window function power. Plot reference
+    lines showing several False Alarm Probability levels to understand whether
+    your periodogram peaks are significant enough to be labeled detections:
+    use the Baluev method (\sect{})
+    or the bootstrap method (\sect{}) 
+    if it is computationally feasible.
+    Keep in mind exactly what the False Alarm Probability measures, and
+    avoid the temptation to misinterpret it (\sect{}).
+  \item If the window function shows strong aliasing, locate the expected
+    multiple maxima and plot the phased light curve at each
+    (cf. \fig{}).If there is indication
+    that the sinusoidal model under-fits the data (cf. \fig{}) then consider
+    re-fitting with a multi-term Fourier model (\sect{}).
+  \item If you have prior knowledge of the shape of light curves you are
+    trying to detect, consider using more complex models to choose between
+    multiple peaks in the periodogram (\sect{}). This type of refinement can
+    be quite useful in building automated pipelines for period fitting,
+    especially in cases where the window aliasing is strong.
+  \item If you are building an automated pipeline based on Lomb-Scargle for
+    use in a survey, consider injecting known signals into the pipeline to
+    measure your detection efficiency as a function of object type, brightness,
+    and other relevant characteristics (\sect{})
+\end{enumerate}
+
+This list is certainly not comprehensive for all uses of the periodogram, but
+it should serve as a gentle reminder of the kinds of issues you should keep
+in mind when using the method in practice.
+
+\subsection{Other Measures of Periodicity}
+\todo{fill this in}
 
 \subsection{Postscript: Why Lomb-Scargle?}
+\sectlabel{postscript}
 After considering all of these practical aspects of the periodogram,
@@
 periodogram.
+
+%One clear advantage retained by both the Lomb-Scargle and Schuster periodograms
+%is that they are quite efficient compared with more sophisticated approaches
+%to periodic analysis based on more constrained generative models or Bayesian
+%non-parametric models, and in the context of large time-domain surveys can
+%provide rough periodicity detections and catalog parameters in near real-time.
+
 In particular, three decades after the entry of ``Lomb-Scargle'' into the
@@
     model coefficients.}
-    Reality: model coefficients can be determined with not much extra
-    computation \citep[see discussion in][]{ICVG2014}.
+    Reality: model coefficients can be determined with little extra
+    computation \citep[see the discussion in][]{ICVG2014}.
   \item {\it Myth: The Lomb-Scargle periodogram allows analytic computation of
@@
     Reality: for structured window functions, the Lomb-Scargle
-    periodogram has the same partial aliasing issues as the Schuster
+    periodogram has the same window-driven issues as the Schuster
     periodogram, including spurious peaks due to partial aliasing, and
@@
   \item {\it Myth: Bayesian analysis shows that Lomb-Scargle is the optimal
-    statistic for detecting periodic signals in data.} Reality: it
-    shows that Lomb-Scargle is the optimal statistic for
+    statistic for detecting periodic signals in data.} Reality: Bayesian
+    analysis shows that Lomb-Scargle is the optimal statistic for
     {\it fitting a sinusoid to data}, which is not the same as saying it
-    is optimal for finding a generic, non-sinusoidal signal
-    (see \sect{bayesian-periodograms}).
+    is optimal for finding the frequency of a generic, potentially
+    non-sinusoidal signal (see \sect{bayesian-periodograms}).
 \end{itemize}
@@
 various forms of regularization.
-
-One clear advantage of both Lomb-Scargle and Schuster periodograms is that
-they are quite fast to compute compared to more sophisticated approaches to
-periodic analysis based on more constrained generative models or Bayesian
-non-parametric models, and in the context of large time-domain surveys can
-provide rough periodicity detections and catalog parameters in near real-time.
 
@@
 
+\acknowledgments
 {\it Acknowledgments:} I am grateful to Jeff Scargle, Max Mahlke,

### 2017-02-09 PracticalLombScargle.tex
Add some intro paragraphs

@@
   \item What assumptions does the Lomb-Scargle approach make regarding the
-    shape of the unknown signal?
+    form of the unknown signal?
   \item How should we understand and report the uncertainty of the
@@
 about these questions.
+
+\subsection{Why Lomb-Scargle?}
+\sectlabel{why-lomb-scargle}
+
+Before we begin exploring the Lomb-Scargle periodogram in more depth,
+it is worth briefly considering the broader context of methods for
+detecting and characterizing periodicity in time series.
+
+First, it is important to note that there are many different modes of
+time series observation.
+Point observation like that shown in \fig{LINEAR-data} is typical of
+optical astronomy: values (often with uncertainties) are measured at
+discrete point in time, which may be equally or unequally-spaced.
+Other modes of observation---e.g., time-tag-events, binned event data,
+time-to-spill events, etc.---are common in high-energy astronomy and other
+areas. We will not consider such event-driven data modes here, but note
+that there have been some interesting explorations of unified statistical
+treatments of all of the above modes of observation
+\citep[e.g.][]{Scargle98, Scargle2002}.
+
+Even limiting our focus to point observations, there are a large number of
+complementary techniques for periodic analysis, which generally can be
+categorized into a few broad categories:
+\begin{description}
+  \item[Fourier Methods] are based on the Fourier transform, power spectra,
+    and closely related correlation functions. These methods include the
+    Schuster periodogram \citep{Schuster98}, the Lomb-Scargle periodogram
+    \citep{Lomb76, Scargle82}, the correlation-based method of
+    \citet{Edelson88}, and related approaches
+    \citep[see also][who discusses wavelet transforms in this
+      context]{Foster96}.
+  \item[Phase-folding Methods], as the name implies, depend on folding
+    observations as a function of phase, computing a cost function across
+    the phased data (often within bins constructed across phase space)
+    and optimizing this cost function across candidate frequencies.
+    Some examples are
+    String Length \citep{Dworetsky83},
+    Analysis of Variance \citep{Schwarzenberg-Czerny89},
+    Phase Dispersion Minimization \citep{Stellingwerf78},
+    the Gregory-Laredo method \citep{Gregory92},
+    and the conditional entropy method \citep{Graham2013}.
+    Methods based on correntropy are similar in spirit, but do not always
+    require explicit phase folding \citep{Huijse2011,Huijse2012}.
+  \item[Least Squares Methods] involve fitting a model to the data at each
+    candidate frequency, and selecting the model which maximizes the likelihood.
+    The Lomb-Scargle periodogram also falls in this category
+    (see \sect{schuster-to-lomb-scargle}), as does the Supersmoother approach
+    \citep{Reimann94}. Other studies recommend statistics other than
+    least square residuals; see, e.g., the orthogonal polynomial fits of
+    \citet{Schwarzenberg-Czerny96}.
+  \item[Bayesian Approaches] apply Bayesian probability theory to the problem,
+    often in a similar manner to the phase-folding and/or least-squares
+    approaches.
+    Examples are the generalized Lomb-Scargle models of \citet{Bretthorst88},
+    the phase-binning model of \citet{Gregory92}, Gaussian process models
+    \citep[e.g.][]{Wang2012}, and models based on stochastic processes
+    \citep[e.g.][]{Kelly14}.
+\end{description}
+Various reviews have been undertaken to compare the efficiency and effectiveness
+of the various available methods; for example, \citet{Schwarzenberg-Czerny99}
+focuses on the statistical properties of methods and recommends methods
+based on smooth model fits over methods based on phase binning,
+while \citet{Graham2013b} instead takes an empirical approach and finds that
+for several realistic data sets, no algorithm provides universally optimal
+detection efficiencies.
+
+In light of this breadth of available methods, why focus here on the
+Lomb-Scargle periodogram?
+One reason is cultural: the Lomb-Scargle periodogram is perhaps the best-known
+technique to compute periodicity of unequally-spaced data in astronomy and
+other fields, and so is often the first technique a researcher will reach for
+when searching for periodic content in a signal.
+But there is a deeper reason as well;
+it turns out that Lomb-Scargle occupies a unique niche:
+it is motivated by Fourier analysis,
+but it can also be viewed as a least squares method
+(see \sect{schuster-to-lomb-scargle}), can be derived from the principles
+of Bayesian probability theory (see \sect{bayesian-periodograms}), and has
+been shown to be closely related to bin-based phase-folding techniques under
+some circumstances \citep[see][]{Swingler89}.
+Thus, as a unique point of correspondance between many classes of methods,
+an intuitive understanding of the Lomb-Scargle method and the practical
+issues involved in its application end up being highly relevant to
+understanding and using the other methods listed above.
 
@@
 and a signal measured at regular intervals is described by a Dirac comb window
-function marking those measurement times.\footnote{
-  Throughout this paper we are concerned only with the mode of measurement
-  common in optical astronomy; {i.e.} point observation. Other modes of
-  observation, such as time-tag-events, binned event data, or
-  time-to-spill events, are common in high-energy astronomy and other areas.
-  We will not consider such event-driven data modes here, but it should be
-  noted that there have been some interesting explorations of unified
-  statistical treatments of all of the above modes of observation
-  \citep[e.g.][]{Scargle98, Scargle2002}.
-}
+function marking those measurement times.
 
@@
 \citep[e.g.][]{Scargle82, Horne86, NumRec},
+\todo{median? Schwarzenberg-Czerny 1996, ApJ 460}
 the harmonic mean of the sampling intervals \citep[e.g.][]{Debosscher07},
@@
 \end{equation}
-
-Depending on the characteristics of the dataset, this number can vary greatly.
+So what is a good choice for $n_o$?
+Values ranging from $n_o=5$ \citep{Schwarzenberg-Czerny96, VanderPlas2015}
+to $n_o=10$ \citep{Desbosscher07, Richards12}
+seem to be common in the literature;
+for periodograms computed in this paper, we use $n_o=5$.
+% \citep{Schwarzenberg-Czerny96} discusses how $n_o$ can be selected
+% more carefully based on factors such as oversampling, number of harmonics
+% in the model, and other considerations.
+
+Depending on the characteristics of the dataset, the size of the resulting
+frequency grid can vary greatly.
 For example, the Kepler data shown in \fig{kepler-data} has a pseudo-Nyquist
@@
 a limiting frequency of $f_{max} = (1\ {\rm hour})^{-1}$ based on typical
-oscillation periods of SX Phe-type stars. This leads to a much more manageable
+oscillation periods expected for SX Phe-type stars.
+This leads to a much more manageable
 $N_{eval} \approx 240,000$ periodogram evaluations.
 
-By comparison, data from the upcoming LSST survey \citep{Ivezic08LSST}
+By comparison, data from the LSST survey \citep{Ivezic08LSST}
 will fall somewhere in-between: full frequency coverage of
 the 10-year data up to a limiting frequency defined by the 30-second
-exposure time for each observation would require roughly 25 million
+integration time for each visit would require roughly 25 million
 periodogram evaluations per object, which for fast implementations
 (see the next section) could be accomplished in several seconds on
-a single core.
+a modern CPU.
 
@@
 frequency, the peak widths are not constant in period.
-Thus regular period grids tend to over-sample at large periods and under-sample
+Regular period grids tend to over-sample at large periods and under-sample
 at small periods;
@@
 periodic analysis based on least squares model fitting and Bayesian
-analysis. I will finish here with a brief summary of the practical
-recommendations contained in these pages, along with a somewhat opinionated
-and iconoclastic post-script for further thought.
+analysis.
+From this conceptual understanding, we considered a list of challenges
+and issues to be considered when applying the method in practice.
+We will finish here with a brief summary of these practical
+recommendations, a brief discussion of other methods for detecting
+periodicity, and a somewhat opinionated
+post-script for further thought.
 
@@
 algorithm to a dataset:
+
+\todo{fill-in references}
 
@@
 This list is certainly not comprehensive for all uses of the periodogram, but
-it should serve as a gentle reminder of the kinds of issues you should keep
-in mind when using the method in practice.
-
-\subsection{Other Measures of Periodicity}
-\todo{fill this in}
+it should serve as a brief reminder of the kinds of issues you should keep
+in mind when using the method to detect periodic signals.
 
@@
 
-%One clear advantage retained by both the Lomb-Scargle and Schuster periodograms
-%is that they are quite efficient compared with more sophisticated approaches
-%to periodic analysis based on more constrained generative models or Bayesian
-%non-parametric models, and in the context of large time-domain surveys can
-%provide rough periodicity detections and catalog parameters in near real-time.
-
-In particular, three decades after the entry of ``Lomb-Scargle'' into the
-astronomy community's vernacular, there seems to be a well-developed
-mythology surrounding the efficacy and efficiency of the Lomb-Scargle approach.
-It's common to see statements or implications like the following:
+As discussed in \sect{schuster-to-lomb-scargle}, the periodogram has two
+benefits over the Schuster approach:
+the noise distribution at each individual frequency is the sum
+of two Gaussians, and the result is equivalent to a periodogram derived
+from a least squares analysis.
+But somehow along the way, a mythology seems to have developed surrounding
+the efficiency and efficacy of the Lomb-Scargle approach. In particular,
+it's common to see statements or implications along the lines of the following:
 \begin{itemize}
@@
     individual frequencies, it is not true for the more relevant issue of
-    peak-value statistics, which must be either approximated or computed by
-    bootstrap resampling (see \sect{false-alarm-probability})
+    peak heights across multiple frequencies, which must be either approximated
+    or computed by bootstrap resampling (see \sect{false-alarm-probability})
   \item {\it Myth: The Lomb-Scargle periodogram corrects for aliasing due to
@@
 \end{itemize}
-With these myths debunked, what advantage remains for Lomb-Scargle?
+With these myths set aside, what is the practical advantage of Lomb-Scargle
+over a Schuster approach?
 What would we lose if we instead used a Schuster periodogram,
@@
 
-I think we only lose the advantage of being able to generalize to more
-complicated models in the Bayesian sense, in which case you're not really
-using Lomb-Scargle at all, but rather a generative Bayesian model for your
+The advantage of analytic statistics for Lomb-Scargle evaporates in light of the
+need to account for multiple frequencies, so the only advantage left is the
+correspondance to least squares and Bayesian models, and in particular the
+ability to generalize to more complicated models where appropriate---but in
+this case you're not really using Lomb-Scargle at all,
+but rather a generative Bayesian model for your
 data based on some strong prior information about the form of your signal.
@@
 interesting trivia, but not itself a reason to use that model if your data
-is not known {\it a priori} to be sinusoidal.
+is not known {\it a priori} to be sinusoidal---it could even be construed as
+an argument {\it against} Lomb-Scargle in the general case where the assumption
+of a sinusoid is not well-founded.
 
@@
 various forms of regularization.
-
 In any case, I would contend that there is ample room for practitioners to
-question the prevailing folk wisdom of the unspoken advantage of Lomb-Scargle
-over approaches based directly on the Fourier transform, such as the Schuster
+question the prevailing folk wisdom of the advantage of Lomb-Scargle
+over approaches based directly on the Fourier transform and Schuster
 periodogram.
-
-% Edelson-Krolik + FFT -> Lomb-Scargle?
 

### 2017-02-10 PracticalLombScargle.tex
Add draft footer

@@
 \usepackage{graphicx}
+\usepackage{fancyhdr}
 \graphicspath{{figures/}}
@@
 
-\title{Understanding the Lomb-Scargle Periodogram}
-
-\newcommand{\escience}{*}
-\newcommand{\uwastro}{+}
+\title{Understanding the Lomb-Scargle Periodogram\\
+       {\it (Draft Version \today)}}
+
+% Add ``Draft version'' footer
+\pagestyle{fancy}
+\fancyhf{}\renewcommand{\headrulewidth}{0pt} % remove header titles
+\cfoot{\it \small Draft Version \today}
+
+\newcommand{\escience}{1}
 \author{Jacob T. VanderPlas\altaffilmark{\escience}}
 \altaffiltext{\escience}{eScience Institute, University of Washington}
+%\author{Jacob T. VanderPlas}
+%\affil{eScience Institute, University of Washington}
 
@@
 the same data as \fig{LINEAR-data} -- i.e.{} plotted as a function of phase
-rather than time.\footnote{
-    The periodogram in \fig{LINEAR-power} was computed using
-    {\tt astropy.stats.LombScargle}
-    from the AstroPy project; see \url{http://astropy.org/}.
-}
+rather than time.
 
@@
     \citet{Edelson88}, and related approaches
-    \citep[see also][who discusses wavelet transforms in this
+    \citep[see also][for a discussion of wavelet transforms in this
       context]{Foster96}.
-  \item[Phase-folding Methods], as the name implies, depend on folding
+  \item[Phase-folding Methods] depend on folding
     observations as a function of phase, computing a cost function across
@@
 Various reviews have been undertaken to compare the efficiency and effectiveness
-of the various available methods; for example, \citet{Schwarzenberg-Czerny99}
-focuses on the statistical properties of methods and recommends methods
+of the available methods; for example, \citet{Schwarzenberg-Czerny99}
+focuses on the statistical properties of methods and recommends those
 based on smooth model fits over methods based on phase binning,
-while \citet{Graham2013b} instead takes an empirical approach and finds that
-for several realistic data sets, no algorithm provides universally optimal
-detection efficiencies.
-
-In light of this breadth of available methods, why focus here on the
+while \citet{Graham2013b} instead take an empirical approach and find that
+when considering detection efficiency in real datasets, no algorithm
+universally outperforms the others.
+
+In light of this breadth of available methods, why limit our focus here to the
 Lomb-Scargle periodogram?
@@
 technique to compute periodicity of unequally-spaced data in astronomy and
-other fields, and so is often the first technique a researcher will reach for
+other fields, and so is the first technique many will reach for
 when searching for periodic content in a signal.
-But there is a deeper reason as well;
-it turns out that Lomb-Scargle occupies a unique niche:
-it is motivated by Fourier analysis,
+But there is a deeper reason as well; it turns out that Lomb-Scargle occupies
+a unique niche: it is motivated by Fourier analysis,
 but it can also be viewed as a least squares method
@@
 some circumstances \citep[see][]{Swingler89}.
-Thus, as a unique point of correspondance between many classes of methods,
-an intuitive understanding of the Lomb-Scargle method and the practical
-issues involved in its application end up being highly relevant to
-understanding and using the other methods listed above.
+Thus, the Lomb-Scargle periodogram occupies a unique point of correspondance
+between many classes of methods, and so provides a focus for discussion that
+can speak to important considerations for all of these methods.
 
@@
 power spectrum, and discusses a conceptual understanding of the
-well-known Nyquist limiting frequency.
+Nyquist limiting frequency.
 {\bf\sect{non-uniform-sampling}} considers non-uniformly sampled signals as
@@
 Lomb-Scargle periodogram as the result of least-squares model fitting,
-as well as some extensions enabled by that viewpoint
-(including various ``Bayesian'' periodograms).
+as well as some extensions enabled by that viewpoint.
 {\bf\sect{practical-considerations}} utilizes the concepts introduced
@@
 frequency grid, uncertainties and false alarm probabilities, and modes of
-failure that should be expected when working with real-world data.
+failure that should be accounted for when working with real-world data.
 {\bf\sect{conclusion}} concludes and summarizes the key recommendations
@@
 In order to understand how we should interpret the Lomb-Scargle periodogram, we will first briefly step back and review the subject of Fourier analysis of continuous signals.
-Consider a continuously-defined signal $g(t)$.
-Its Fourier transform is given by the following integral, where $i\equiv\sqrt{-1}$ denotes the imaginary unit:
+Consider a continuous signal $g(t)$.
+its Fourier transform is given by the following integral, where $i\equiv\sqrt{-1}$ denotes the imaginary unit:
 \begin{equation}
@@
 \end{eqnarray}
-As a further shorthand, we'll sometimes use a double arrow to denote a Fourier
-pair; for example
-\begin{equation}
-  g(x) \Longleftrightarrow \hat{g}(f).
-\end{equation}
+The functions $g$ and $\hat{g}$ are known as a {\it Fourier pair}, which we
+will sometimes denote as $g \Longleftrightarrow \hat{g}$.
 
 \subsection{Properties of the Fourier Transform}
-
+\sectlabel{ft-properties}
 The continuous Fourier transform has a number of useful properties that we will make use of in our discussion.
@@
    \item[The Fourier transform of sinusoid with frequency $f_0$ is a sum of delta functions at $\pm f_0$.]
-     Given the integral definition of the Dirac delta function, it is straightforward to show that
+     From the integral definition of the Dirac delta function, we can write
      \begin{equation}
@@
      \end{equation}
-     Euler's formula shows how such a complex exponential
-     can be expressed in terms of sines and cosines:
-     \begin{equation}
-       e^{2\pi f_0 t} = \cos(2\pi f_0 t) + i\sin(2\pi f_0 t).
-       \eqlabel{Euler-formula}
-     \end{equation}
-     Using the linearity of the Fourier transform, we can take these identites
-     and derive the following identities:
+     Using Euler's formula for the complex exponential along with the linearity
+     of the Fourier transform leads to the following identities:
      \begin{eqnarray}
@@
      \end{equation}
-     The key observation here is that the time-shift does not change the
+     Notice that the time-shift does not change the
      amplitude of the resulting transform, but only the phase.
@@
 of delta functions marking the frequencies of those sinusoids: that is,
-the Fourier transform is a {\it direct} measure of the periodic
-content of a function.
+the Fourier transform directly measures periodic content in a function.
 
@@
 \end{description}
-The key observation in each of these Fourier pairs is the reciprocity
+Notice in each of these Fourier pairs the reciprocity
 of scales between a function and its Fourier transform:
@@
 
-Given this definition of a convolution, it can be shown that the Fourier transform of a convolution is the point-wise product of the Fourier transforms of
-the two functions:
+Given this definition of a convolution, it can be shown that the Fourier transform of a convolution is the point-wise product of the individual
+Fourier transforms:
 \begin{equation}
@@
 in \eq{convolution-definition}.
-This result in \eq{convolution-theorem} is known as the
+The identity in \eq{convolution-theorem} is known as the
 {\it convolution theorem}, and is illustrated in \fig{convolution-theorem}.
@@
   \caption{Visualization of the convolution theorem (\eq{convolution-theorem}).
-    The key is that the Fourier transform of
+    Recall that the Fourier transform of
     a convolution is the pointwise product of the two Fourier transforms.
@@
 which are well-defined for all times $-\infty < t < \infty$.
-Real-world temporal measurements of a signal, however, can only measure some
+Real-world temporal measurements of a signal, however, only involve some
 finite span of time, at some finite sampling rate.
@@
 
-You may have noticed we glossed over something here: the effect of switching
+Notice, though, that we glossed over one thing: the effect of switching
 from an infinite number of samples to a finite number of samples.
@@
 \citep[see, e.g.][]{FoundationsOfSignalProcessing}.
-Despite the semi-qualitative nature of our discussion,
-I find it to be a helpful approach
+Despite our lack of rigor here, I find this to be a helpful approach
 in developing intuition regarding the relationship between the
@@
 
-\todo{mention difference between astronomy \& math treatment: astro is definitely {\it not} sampled as a well-defined random process...
-e.g.
-new observations made based on implications of previous observations.}
+In the real world, particularly in fields like Astronomy where observations are
+subject to interruptions from weather and diurnal cycles, the sampling rate
+is generally far from uniform.
+Using the same approach as we used to explore uniform sampling in the previous
+section, we can now explore non-uniform sampling here.
+
+In the general non-uniform case, we measure some signal at a set of $N$ times
+which we will denote $\{t_n\}$, which leads to the following observing window:
+\begin{equation}
+  W_{\{t_n\}}(t) = \sum_{n=1}^{N} \delta(t - t_n)
+  \eqlabel{nonuniform-window}
+\end{equation}
+Applying this window to our true underlying signal $g(t)$, we find an observed
+signal of the form:
+\begin{eqnarray}
+  g_{obs}(t) &=& g(t) W_{\{t_n\}}(t) \nonumber\\
+             &=& \sum_{n=1}^{N} g(t_n)\delta(t - t_n)
+  \eqlabel{g-nonuniform}
+\end{eqnarray}
+
+Just as in the evenly-sampled case, the Fourier transform of the observed
+signal is a convolution of the transforms of the true signal and the window:
+\begin{equation}
+  \mathcal{F}\{g_{obs}\} = \mathcal{F}\{g\} \ast \mathcal{F}\{W_{\{t_n\}}\}
+  \eqlabel{g-obs-conv}
+\end{equation}
+Unlike in the uniform case, the window transform $\mathcal{F}\{W_{\{t_n\}}\}$
+will generally {\it not} be a clean sequence of delta functions; the
+symmetry present in the Dirac comb is broken by the uneven sampling,
+leading the transform to be much more ``noisy''.
+This can be seen in \fig{random-window}, which shows the Fourier transform of
+a non-uniform observing window with an average sampling rate identical to
+that in \fig{comb-window-1}, along with its imact on the observed Fourier
+transform.
 
@@
 \end{figure}
-
-In the real world, particularly in fields like Astronomy where observations are
-subject to interruptions from weather and diurnal cycles, the sampling rate
-is generally far from uniform.
-Using the same approach as we used to explore uniform sampling in the previous
-section, we can now explore non-uniform sampling here.
-
-In the general non-uniform case, we measure some signal at a set of $N$ times
-which we will denote $\{t_n\}$, which leads to the following observing window:
-\begin{equation}
-  W_{\{t_n\}}(t) = \sum_{n=1}^{N} \delta(t - t_n)
-  \eqlabel{nonuniform-window}
-\end{equation}
-Applying this window to our true underlying signal $g(t)$, we find an observed
-signal of the form:
-\begin{eqnarray}
-  g_{obs}(t) &=& g(t) W_{\{t_n\}}(t) \nonumber\\
-             &=& \sum_{n=1}^{N} g(t_n)\delta(t - t_n)
-  \eqlabel{g-nonuniform}
-\end{eqnarray}
-
-Just as in the evenly-sampled case, the Fourier transform of the observed
-signal is a convolution of the transforms of the true signal and the window:
-\begin{equation}
-  \mathcal{F}\{g_{obs}\} = \mathcal{F}\{g\} \ast \mathcal{F}\{W_{\{t_n\}}\}
-  \eqlabel{g-obs-conv}
-\end{equation}
-Unlike in the uniform case, the window transform $\mathcal{F}\{W_{\{t_n\}}\}$
-will generally {\it not} be a nice clean function; the symmetry present in
-the Dirac comb is broken by the uneven sampling,
-leading the transform to be much more ``noisy''.
-This can be seen in \fig{random-window}, which shows the Fourier transform of
-a non-uniform observing window with an average sampling rate identical to
-that in \fig{comb-window-1}, along with its imact on the observed Fourier
-transform.
 
@@
 is no exact aliasing of the true signal, and thus no way to exactly recover
-any portion of the true Fourier transform of this data.
+any portion of the true Fourier transform for the underlying function.
 
@@
 and it does, but only to a degree.
-\Fig{random-window-2} increases the density of observations by a factor of 10,
+In \Fig{random-window-2} we increase the density of observations by a factor of 10,
 such that there are 200 total observations over the length-10 observing window.
@@
 with a non-structured window function.
-The resulting periodogram, then, can only be used as a qualitative approximation
+The resulting periodogram, then, can only be used as an approximation
 of the true periodogram, and will only approach the true periodogram as the
-number of observed samples gets very large.
+number of observed samples becomes very large.
 
@@
 \citep[e.g.][]{Scargle82, Horne86, NumRec},
-\todo{median? Schwarzenberg-Czerny 1996, ApJ 460}
 the harmonic mean of the sampling intervals \citep[e.g.][]{Debosscher07},
+the median of the sampling intervals \citep[e.g.][]{Graham2013b},
 or the minimum sample spacing \citep[e.g.][]{Percy86, Roberts87, Press89, Hilditch01}.
@@
 unevenly-sampled data can probe frequencies far larger than any of these
-supposed limits---a fact that several of these citations do hint at
-parenthetically.
+supposed limits (a fact that several of these citations do hint at
+parenthetically).
 
@@
 \subsubsection{The Non-uniform Nyquist Limit}
+\sectlabel{non-uniform-nyquist}
 
@@
 \Fig{nyquist-eyer99} shows an example of such a Nyquist frequency.
-The data are non-uniformly sampled at times $t_i = n_i * p$, with $p=0.01$
+The data are non-uniformly sampled at times $t_i = n_i \cdot p$, with $p=0.01$
 and $n_i$ drawn randomly from positive integers less than 10,000.
@@
 behavior beyond this frequency: the signal at $f > f_{Ny}$ consists of a series
-of exact aliases of the signal at $f < f_{Ny}$.
+of exact aliases of the signal at $|f| < f_{Ny}$.
 
@@
 typically set by {\it the precision of the time measurements}
-\citep[see][for more rigorous treatments
+\citep[see also][for more rigorous treatments
   of this result]{Bretthorst2003, Koen2006}.
@@
 is aliased into the Nyquist range; the windowing limit is the frequency
-beyond which all signal is zero or nearly zero.
-
-%\subsubsection{Practicality of Limiting Frequencies}
-%The frequency grid implied by such large Nyquist frequencies in practice can be
-%much larger than is computationally feasible.
-%For example, consider the LSST project, which will have roughly 30-second
-%exposures taken over a 10-year baseline. With
-%$f_{Ny} \approx (2\cdot 30{\rm sec})^{-1}$ and a grid spacing
-%of $\Delta f = (5\cdot10 {\rm yr})^{-1}$ (see \sect{frequency-grid}),
-%the size of the required frequency grid is on order 25 million bins, which is
-%far too large a periodogram to be computationally feasible given the
-%computational constraints of the LSST pipeline, even for extremely
-%efficient periodogram methods (see \sect{algorithmic-considerations}).
-
-%In practice, the limiting frequency of the periodogram should be set based on
-%some prior knowledge of the types of signals you are interested in.
-%For example, if you are interested in galactic RR Lyrae stars, the shortest
-%applicable period is on order a few hours \citep{Sesar2010}, requiring
-%evaluation of only a hundred thousand or so periodogram frequencies for
-%data with a 10-year baseline.
-%The key point is that for working with unevenly-sampled data in practice,
-%computational constraints will mean that the maximum frequency
-%should more often be set by {\it what you are looking for}
-%rather than any hard limit based on Nyquist-like limits.
+beyond which all signal is attenuated to zero.
+In practice, the limit implied by either the temporal resolution or windowing
+of individual observations may be too large to be computationally feasible;
+for discussion of frequency limits in practice, see \sect{frequency-grid}.
 
@@
 Values ranging from $n_o=5$ \citep{Schwarzenberg-Czerny96, VanderPlas2015}
-to $n_o=10$ \citep{Desbosscher07, Richards12}
+to $n_o=10$ \citep{Debosscher07, Richards12}
 seem to be common in the literature;
@@
 \subsubsection{Computing the Window Function}
+\sectlabel{computing-window-function}
 The window power spectrum can be computed directly from the
@@
 In the Fourier view, the precision with which a peak's frequency can be
-identified is directly related to the width of the peak; often the
-width at half-maximum is used.
+identified is directly related to the width of this peak; often the
+half-width at half-maximum $f_{1/2} \approx 1/T$ is used.
 This can be formalized more precisely in the least squares interpretation of
@@
 view amounts to fitting a Gaussian curve to the (exponentiated)
-peak \citep{Jaynes87, Bretthorst88}---and this introduces first-order
-dependence on the number of samples and their signal-to-noise ratio,
-though the uncertainty remains proportional to the peak width.
-\todo{explain this difference in terms of FWHM of exponentiated power?}
+peak \citep{Jaynes87, Bretthorst88}. This introduces first-order
+dependence on the number of samples $N$ and their average signal-to-noise
+ratio $\Sigma$; the scaling is approximately \citep[see, e.g.][]{Gregory2001}:
+\begin{eqnarray}
+  \sigma_f &\approx& f_{1/2} / \sqrt{P_{max}}\nonumber\\
+           &\approx& f_{1/2} \sqrt{\frac{2}{N\Sigma^2}}.
+  \eqlabel{bayes-error}
+\end{eqnarray}
+This dependence comes from the fact that the Bayesian uncertainty is related
+to the width of the {\it exponentiated} periodogram, which
+depends on $P_{max}$, the height of the peak\footnote{
+To see why, consider a periodogram with maximum value
+$P_{max} = P(f_{max})$, so that $P(f_{max} \pm f_{1/2}) = P_{max}/2$.
+The Bayesian uncertainty comes from approximating the exponentiated peak as a
+Gaussian; i.e., $\exp[P(f_{max} \pm \delta f)] \propto \exp[-\delta f^2/(2\sigma_f^2)]$.
+From this we can write $P_{max}/2 \approx P_{max} -f_{1/2}^2 / (2 \sigma_f^2)$ or
+$\sigma_f \approx f_{1/2} / \sqrt{P_{max}}$.
+In terms of signal-to-noise ratio
+$\Sigma = {\rm rms}[(y_n - \mu) / \sigma_n]$,
+a well-fit model gives
+$P_{max} \approx \hat{\chi}_0^2/2 \approx \Sigma^2 N/2$,
+which leads to the expression in \eq{bayes-error}.
+}.
 
@@
 \subsubsubsection{What Does the False Alarm Probability Measure?}
+\sectlabel{what-FAP-measures}
 
@@
 the data at some frequency $f_0$, then $\hat{\chi}^2(f_0) = 0$ and the
-periodogram is maximized at a value of $N\hat{\chi}^2_0 / 2$.
+periodogram is maximized at a value of $\hat{\chi}^2_0 / 2$.
 On the other end, it is mathematically impossible for a sinusoidal model to
@@
 with the Lomb-Scargle periodogram.
-Here I will quickly summarize the steps you should take when applying this
-algorithm to a dataset:
-
-\todo{fill-in references}
+Following is a brief summary of the considerations to keep in mind when you
+apply this algorithm to a dataset:
 
@@
     the maximum set based on the precision of the time measurements
-    (\sect{}), and the grid spacing set based on the temporal baseline of
-    the data (\sect{}) so as to not sample too coarsely around peaks.
+    (\sect{non-uniform-nyquist}), and the grid spacing set based on the
+    temporal baseline of the data (\sect{frequency-grid})
+    so as to not sample too coarsely around peaks.
     If this grid size is computationally intractable, reduce the maximum
-    frequency based on what kinds of signals you are looking for (\sect{}).
+    frequency based on what kinds of signals you are looking for.
   \item Compute the window transform using the Lomb-Scargle periodogram,
     by substituting $g_n=1$ for each $t_n$ and making sure not to pre-center
-    the data or use a floating-mean model (\sect{}). Examine this window
-    function for dominant features, such as daily or annual aliases (cf.
-    \fig{}) or Nyquist-like limits (cf. \fig{}).
+    the data or use a floating-mean model (\sect{computing-window-function}).
+    Examine this window function for dominant features, such as daily or
+    annual aliases (cf. \fig{LINEAR-window})
+    or Nyquist-like limits (cf. \fig{kepler-data}).
   \item Compute the periodogram for your data. You should always use the
-    floating-mean model (\sect{}), as it produces more robust periodograms
-    and has few if any disadvantages.
+    floating-mean model (\sect{floating-mean}), as it produces more robust
+    periodograms and has few if any disadvantages. Avoid multi-term Fourier
+    extensions (\sect{multiterm}) when the signal is unknown, because its
+    main effect is to increase periodogram noise
+    (cf. \figs{kepler-multiterm}{LINEAR-multiterm}).
   \item Plot the periodogram, and identify any patterns that may be caused
@@
     your periodogram peaks are significant enough to be labeled detections:
-    use the Baluev method (\sect{})
-    or the bootstrap method (\sect{}) 
-    if it is computationally feasible.
+    use the Baluev method or the bootstrap method if it is computationally
+    feasible (\sect{false-alarm-probability}).
     Keep in mind exactly what the False Alarm Probability measures, and
-    avoid the temptation to misinterpret it (\sect{}).
+    avoid the temptation to misinterpret it (\sect{what-FAP-measures}).
   \item If the window function shows strong aliasing, locate the expected
-    multiple maxima and plot the phased light curve at each
-    (cf. \fig{}).If there is indication
-    that the sinusoidal model under-fits the data (cf. \fig{}) then consider
-    re-fitting with a multi-term Fourier model (\sect{}).
+    multiple maxima and plot the phased light curve at each.
+    If there is indication that the sinusoidal model under-fits the data
+    (cf. \fig{binary-multiterm}) then consider
+    re-fitting with a multi-term Fourier model (\sect{multiterm}).
   \item If you have prior knowledge of the shape of light curves you are
     trying to detect, consider using more complex models to choose between
-    multiple peaks in the periodogram (\sect{}). This type of refinement can
-    be quite useful in building automated pipelines for period fitting,
-    especially in cases where the window aliasing is strong.
+    multiple peaks in the periodogram (\sect{bayesian-periodograms}).
+    This type of refinement can be quite useful in building automated
+    pipelines for period fitting, especially in cases where the window
+    aliasing is strong.
   \item If you are building an automated pipeline based on Lomb-Scargle for
@@
     measure your detection efficiency as a function of object type, brightness,
-    and other relevant characteristics (\sect{})
+    and other relevant characteristics (\sect{failure-modes})
 \end{enumerate}

### 2017-02-10 PracticalLombScargle.tex
Read-through and edit

@@
 \fancyhf{}\renewcommand{\headrulewidth}{0pt} % remove header titles
-\cfoot{\it \small Draft Version \today}
+\cfoot{\it \footnotesize $\sim$ Draft Version, \today $\ \sim$}
 
@@
    \item[The Fourier transform of sinusoid with frequency $f_0$ is a sum of delta functions at $\pm f_0$.]
-     From the integral definition of the Dirac delta function, we can write
+     From the integral definition of the Dirac delta function\footnote{
+       $\delta(f)\equiv\int_{-\infty}^\infty e^{-2\pi i x f}df$},
+     we can write
      \begin{equation}
@@
      \end{equation}
-     Using Euler's formula for the complex exponential along with the linearity
+     Using Euler's formula\footnote{$e^{ix} = \cos x + i\sin x$}
+     for the complex exponential along with the linearity
      of the Fourier transform leads to the following identities:
@@
 Real-world temporal measurements of a signal, however, only involve some
-finite span of time, at some finite sampling rate.
+finite span of time, at some finite rate of sampling.
 In either case, the resulting data can be described by a point-wise product
@@
 
-The Fourier transform we measure in these cases, then, is not the
-transform of the continuous underlying function, but rather a 
-{\it convolution of the true transform with the transform of the observing
-window}.
+The Fourier transform of measured data in these cases, then, is not the
+transform of the continuous underlying function, but rather the transform
+of the {\it point-wise product of the signal and the observing window}
 Symbolically, if the signal is $g(t)$ and the window is $W(t)$, the observed
@@
 \end{equation}
-and its Fourier transform is a convolution of two transforms:
+and by the convolution theorem, its transform is a convolution
+of the signal transform and the window transform:
 \begin{equation}
@@
 \end{equation}
-In the general case, such a convolution is not an invertible operation, meaning
-that, except in special circumstances, it is not possible to exactly recover
-the transform of the underlying signal.
 This has some interesting consequences for the use and interpretation of
@@
 \subsection{Effect of a Rectangular Window}
-
-First, let's consider the case of observing a continuous periodic signal over
-a limited span of time, as illustrated in \Fig{rectangular-window}.
-Here we consider a continuous periodic function observed only within the
-window $-3<t<3$.
-The observed signal in this case can be understood as the pointwise product of
-the underlying infinite periodic signal, and a rectangular window function.
-By the convolution theorem, the Fourier transform will be given by the
-convolution of the transform of the underlying function---here a set of
-delta functions at the component frequencies---and the transform of the window
-function---here a sinc function.
-For the purely periodic signal like the one seen in \fig{rectangular-window},
-this convolution has the effect of replacing each delta function with a
-sinc function.
-Because of the inverse relationship between the width of the window and the
-width of its transform (see \Fig{fourier-pairs}), it follows that a wider
-observing window leads to proportionally less spread in the Fourier
-transform of the observed function.
 
@@
 
+First, let's consider the case of observing a continuous periodic signal over
+a limited span of time: \fig{rectangular-window} shows
+a continuous periodic function observed only within the window $-3<t<3$.
+The observed signal in this case can be understood as the pointwise product of
+the underlying infinite periodic signal, and a rectangular window function.
+By the convolution theorem, the Fourier transform will be given by the
+convolution of the transform of the underlying function (here a set of
+delta functions at the component frequencies) and the transform of the window
+function (here a sinc function).
+For the purely periodic signal like the one seen in \fig{rectangular-window},
+this convolution has the effect of replacing each delta function with a
+sinc function.
+Because of the inverse relationship between the width of the window and the
+width of its transform (see \Fig{fourier-pairs}), it follows that a wider
+observing window leads to proportionally less spread in the Fourier
+transform of the observed function.
+
 \subsection{The Dirac Comb and the Discrete Fourier Transform}
-Another window function that commonly arises is when a continuous signal is
-sampled (nearly) instantaneously at regular intervals.
-Such an observation can be thought of as a point-wise product between the true
-underlying signal and a Dirac comb with the $T$ parameter matching the spacing
-of the observations; this is illustrated in \fig{comb-window-1}.
 
@@
 
+Another window function that commonly arises is when a continuous signal is
+sampled (nearly) instantaneously at regular intervals.
+Such an observation can be thought of as a point-wise product between the true
+underlying signal and a Dirac comb with the $T$ parameter matching the spacing
+of the observations; this is illustrated in \fig{comb-window-1}.
 Interestingly, because the Fourier transform of a Dirac comb is another Dirac
@@
 With this in mind, we can be assured in this case that evaluating the
-transform in the range $0 \le f < 1/T$ is sufficient to capture
+observed transform in the range $0 \le f < 1/T$ is sufficient to capture
 all the available frequency information:
@@
 the true Fourier transform {\it cannot be recovered}
-from the observed data!
-
-This implies that if we have a regularly-sampled function with a sampling
-rate of $f_0 = 1/T$, we can only fully recover the frequency information
-if the signal is {\it band-limited} between frequencies $\pm f_0/2$.
-This is one way to motivate the famous Nyquist sampling limit, which
-approaches the question from the other direction and states
-that to fully represent the frequency content of a ``band-limited signal''
-whose Fourier transform is zero outside the range $\pm B$,
-we must sample the data with a rate of at least $f_{Ny} = 2B$.
+from the transform of the observed data!
 
@@
 
+This implies that if we have a regularly-sampled function with a sampling
+rate of $f_0 = 1/T$, we can only fully recover the frequency information
+if the signal is {\it band-limited} between frequencies $\pm f_0/2$.
+This is one way to motivate the famous Nyquist sampling limit, which
+approaches the question from the other direction and states
+that to fully represent the frequency content of a ``band-limited signal''
+whose Fourier transform is zero outside the range $\pm B$,
+we must sample the data with a rate of at least $f_{Ny} = 2B$.
+
 \subsubsection{The Discrete Fourier Transform}
@@
 
-Notice, though, that we glossed over one thing: the effect of switching
+Notice, though, that we glossed over one important thing: the effect of switching
 from an infinite number of samples to a finite number of samples.
@@
 
-\subsubsection{The Schuster Periodogram}
+\subsection{The Schuster Periodogram}
 
@@
 uniform sampling.
-It follows that, in the uniform sampling case, the Schuster periodogram captures
-all of the relevant frequency information present in the data.
+It follows that, in the uniform sampling case, the Schuster periodogram
+captures all of the relevant frequency information present in the data.
+This definition readily generalizes to the non-uniform case, which we
+will explore in the following section.
 
@@
 In the real world, particularly in fields like Astronomy where observations are
-subject to interruptions from weather and diurnal cycles, the sampling rate
-is generally far from uniform.
+subject to influences of weather and diurnal, lunar, or seasonal cycles, the
+sampling rate is generally far from uniform.
 Using the same approach as we used to explore uniform sampling in the previous
@@
 Unlike in the uniform case, the window transform $\mathcal{F}\{W_{\{t_n\}}\}$
-will generally {\it not} be a clean sequence of delta functions; the
+will generally {\it not} be a straightforward sequence of delta functions; the
 symmetry present in the Dirac comb is broken by the uneven sampling,
@@
 This non-structured window transform, when convolved with the Fourier transform
-of the true signal, results in an observed Fourier transform with the same
-lack of useful structure.
+of the true signal, results in an observed Fourier transform reflecting the
+same random noise.
 Comparing to the uniformly-spaced observations in \fig{comb-window-1}, we
@@
 any portion of the true Fourier transform for the underlying function.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\textwidth]{fig11_random_window_2}
+  \caption{The effect of non-uniform sampling on the observed Fourier transform,
+    with a factor of 10 more samples than \fig{random-window}.
+    Even with very dense sampling of the function, the Fourier transform
+    cannot be exactly recovered due to the imperfect aliasing present in
+    the window transform.
+    \figlabel{random-window-2}}
+\end{figure}
 
@@
 randomized frequency peaks due to randomized spacing between observations.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth]{fig11_random_window_2}
-  \caption{The effect of non-uniform sampling on the observed Fourier transform,
-    with a factor of 10 more samples than \fig{random-window}.
-    Even with very dense sampling of the function, the Fourier transform
-    cannot be exactly recovered due to the imperfect aliasing present in
-    the window transform.
-    \figlabel{random-window-2}}
-\end{figure}
-
-\subsection{Non-Uniform Periodogram}
-
-From this non-uniform Fourier transform, it is straightforward to extend the
-definition of the Schuster periodogram to the non-uniform case.
-Combining \eq{g-nonuniform} with the Fourier transform in \eq{FT-def}, in analogy to \eq{schuster-periodogram}, gives the following form of the power spectrum:
-\begin{equation}
-  P(f) = \frac{1}{N}\left|\sum_{n=1}^N g(t_n)e^{-2\pi i t_n} \right|^2
-\end{equation}
-Becuase it is derived from the Fourier transform of a non-uniformly sampled
-function, this periodogram is subject to all the caveats mentioned here: in
-particular, the periodogram measures the periodic content of the
-{\it windowed signal} and thus reflects a convolution of the true signal
-with a non-structured window function.
-The resulting periodogram, then, can only be used as an approximation
-of the true periodogram, and will only approach the true periodogram as the
-number of observed samples becomes very large.
 
@@
 
-While Nyquist arguments based on average or minimum sampling fail spectacularly,
-there is a sense in which the Nyquist limit can be applied to unevenly-spaced
-data. \citet{Eyer99} explore this issue in some detail, and in particular
-prove the following:
+While pseudo-Nyquist arguments based on average or minimum sampling
+fail spectacularly, there is a sense in which the Nyquist limit can
+be applied to unevenly-spaced data. \citet{Eyer99} explore this issue
+in some detail, and in particular prove the following:
 \begin{quote}
-Let $p$ be the largest value such that each $t_i$ can be written $t_i = t_0 + n_i p$, for integers $n_i$. The Nyquist Frequency then is $f_{Ny} = 1 / (2p)$.
+{\it Let $p$ be the largest value such that each $t_i$ can be written
+  $t_i = t_0 + n_i p$, for integers $n_i$. The Nyquist Frequency then
+  is $f_{Ny} = 1 / (2p)$.}
 \end{quote}
-In other words, computing the Nyquist rate for unevenly-spaced data requires
+In other words, computing the Nyquist limit for unevenly-spaced data requires
 finding some common factor, such that each spacing $\Delta t_i$ is {\it exactly}
@@
 kind of frequency limit exists.
-This is typical in, e.g., optical astronomy, where a single observation typically consists of an integration of observed flux over a finite duration $\delta t$.
+This is typical in, e.g., optical astronomy, where a single observation typically consists of an integration of observed photons over a finite duration $\delta t$.
 As noted by \citet{ICVG2014}, this time-scale of integration represents another
@@
 a convolution of the true spectrum and the window spectrum
-(cf.\ \fig{random-window}), and so we would expect this structure to be
-imprinted on the power spectrum measured from the data.
+(cf.\ \fig{random-window}), and so we would expect the structure in the window
+to be imprinted on the power spectrum measured from the data.
 
@@
 The upper panel shows the window power spectrum as a function of frequency
-(not period, as in \fig{LINEAR-window}), while lower panel shows the observed
-signal power spectrum as a function of frequency.
+(rather than period, as in \fig{LINEAR-window}),
+while lower panel shows the observed
+signal power spectrum as a function of frequency
+(rather than period, as in \fig{LINEAR-power}).
 The upper panel and its inset show clearly the 1-day and 1-year features we
@@
   P(f)
-  &=& \frac{1}{N}\left|\sum_{n=1}^N g(t_n)e^{-2\pi i t_n} \right|^2 \nonumber\\
+  &=& \frac{1}{N}\left|\sum_{n=1}^N g_n e^{-2\pi i f t_n} \right|^2 \nonumber\\
   &=& \frac{1}{N}\left[
@@
 For the sake of evaluating the significance of periodogram peaks
-(see \sect{uncertainties}), it would be preferable if the non-equispaced
+(see \sect{uncertainties}), it would be preferable if the nonuniform
 periodogram had a noise distribution that could be 
@@
 \eq{classical-periodogram} has such desireable statistical properties
-in the case of uniform sampling, this does not hold for
-\eq{classical-periodogram} when the sampling becomes non-uniform.
+in the case of uniform sampling, this does not hold
+when the sampling becomes non-uniform.
 To address this, he considered a generalized form of the periodogram,
@@
 \end{equation}
-where $A$ and $B$ are arbitrary functions of the frequency $f$ and
+where $A$, $B$, and $\tau$ are arbitrary functions of the frequency $f$ and
 observing times $\{t_i\}$ (but not the values $\{g_n\}$), and showed
-that you can choose one particular form of $A$ and $B$ such that
+that you can choose a unique form of $A$, $B$, and $\tau$ such that
 \begin{enumerate}
@@
 constructing a ``periodogram'' from the $\chi^2$ goodness-of-fit at each
-frequency--an estimator which was considered in some depth by \citet{Lomb76}.
+frequency---an estimator which was considered in some depth by \citet{Lomb76}.
 From this perspective, the $\tau$ shift defined in \eq{tau-def} serves to
@@
 \citep[see, e.g.][]{Gottlieb75}.
-Because of this similarity between the classical and Lomb-Scargle periodograms,
+Because of this similarity between the Schuster and Lomb-Scargle periodograms,
 the bulk of or previous discussion applies---at least qualitatively---to
@@
 In particular, reasoning about the effect of window functions on the observed
-Lomb-Scargle power spectrum remains useful even if it is true in only a
-qualitative sense.
+Lomb-Scargle power spectrum remains useful even if it is not precisely
+accurate.
 
@@
 and \citet{Zechmeister09} showed that such a modification does not change
-the useful statistical properties of the resulting periodogram.
+the statistical properties of the resulting periodogram.
 
@@
 frequency\footnote{We choose to follow \citet{Cumming99} and
-  \citet{VanderPlas2015} and call this a ``Floating Mean'' model, to
-  avoid confusion surrounding the different uses of the term
-  ``generalized periodogram'' in, {e.g.} \citet{Bretthorst88} and
+  \citet{VanderPlas2015} and call this a ``Floating Mean'' model,
+  to avoid confusion of the different uses of the term
+  ``generalized periodogram'' in, {e.g.} \citet{Bretthorst2001} and
   \citet{Zechmeister09}.
@@
 
-\Fig{standard-vs-floatingmean} shows a simulated example of this: the data
-consist of noisy observations of a sinusoidal signal in which the faintest
-observations are omitted from the dataset (due to, e.g., a detection threshold).
-Applying the standard Lomb-Scargle periodogram to pre-centered data leads to a
-periodogram that suppresses the true period of 0.3 days (lower right panel).
-Using the floating-mean model of \eq{floating-mean-model} yields a periodogram
-that identifies this true period (upper right panel).
-
 \begin{figure}[ht]
@@
 
-\citet{Zechmeister09} provide a detailed study of such a model,
+\Fig{standard-vs-floatingmean} shows a simulated example of this: the data
+consist of noisy observations of a sinusoidal signal in which the faintest
+observations are omitted from the dataset (due to, e.g., a detection threshold).
+Applying the standard Lomb-Scargle periodogram to pre-centered data leads to a
+periodogram that suppresses the true period of 0.3 days (lower right panel).
+Using the floating-mean model of \eq{floating-mean-model} yields a periodogram
+that identifies this true period (upper right panel).
+\citet{Zechmeister09} provide a detailed study of this model,
 and show that the addition of the floating mean term does not change
-the advantageous statistical properties of the periodogram.
+the useful statistical properties of the periodogram.
 
@@
 rather than fitting just a single sinusoid at each frequency, we might fit a
-partial Fourier series, adding $K-1$ additional components at
+partial Fourier series, adding $K-1$ additional sinusoidal components at
 integer multiples of the fundamental frequency:
@@
 
-This background noise is easy to understand: in the least-squares view of the
-periodogram, the periodogram height at any frequency is directly related to
-how well the model fits the data.
-For linear models, adding additional terms will {\it always} provide an
+This additional periodogram noise is easy to understand: in the least-squares
+view of the periodogram, the periodogram height at any frequency is directly
+related to how well the model fits the data.
+For nested linear models, adding additional terms will {\it always} provide an
 equal or better fit to data than the simpler model, and so it follows that
@@
 Kepler data from \fig{kepler-data}.
-Given a standard periodogram peak at $f_0$, a 2-term periodogram will have an
-additional peak at $f_0 / 2$, in which the original peak is the second
-harmonic. Similarly, a 3-term periodogram will add additional peaks both at
+Given a standard periodogram with a peak or sub-peak at $f_0$,
+a 2-term periodogram will duplicate this peak at $f_0 / 2$,
+with the second harmonic driving the fit.
+Similarly, a 3-term periodogram will add additional peaks both at
 $f_0 / 2$ and $f_0 / 3$, due to the original peak falling in the second and
@@
 analysis to more arbitrary models.
-For example, \citet{Bretthorst88} studies generalized 
-Bayesian ``periodograms'' that can measure
-decaying signals, non-stationary signals, multi-frequency signals, and other models.
-He shows the effectiveness of such approaches in applications from
-medical imaging and astronomy.
-The crux of such extensions is the fact that you often need to use some prior
-knowledge of the system being observed to decide whether a more complicated
-model is indicated, as well as what form of model to apply.
+For example, periodogram models can be extended to measure
+decaying signals, non-stationary signals, multi-frequency signals, chirps,
+and other signal types \citep[see, e.g.][]{Jaynes87, Bretthorst88, Gregory2001}.
+In particular, \citet{Bretthorst88} demonstrates the effectiveness of such
+approaches in applications ranging from medical imaging to astronomy.
+The challenge of such extensions is the fact that you often need to use some
+prior knowledge of the system being observed to decide whether a more
+complicated model is indicated, as well as what form of model to apply.
 In practice this comes up only when searching for very specific signals for
@@
 \citep{Jaynes03}\footnote{Read this. Really.}.
-In the Bayesian view, the Lomb-Scargle periodogram is the optimal statistic
-for detecting a stationary {\it sinusoidal} signal in the presence of Gaussian
-noise.
+In the Bayesian view, the Lomb-Scargle periodogram is in fact the optimal
+statistic for detecting a stationary sinusoidal signal in the presence of
+Gaussian noise.
 
@@
 \begin{equation}
-  p(f\mid D, M) \propto e^{-N \chi^2 / 2} \propto e^{P_{LS}(f)}
+  p(f\mid D, M) \propto e^{P_{LS}(f)}
   \eqlabel{bayesian-periodogram}
@@
 The effect of this exponentiation, as seen in \fig{bayesian-periodogram},
-is to suppress all peaks in relation to the largest one of the spectrum.
+is to suppress side-lobes and alias peaks in relation to the largest one
+of the spectrum.
 
@@
 and the ability to make use of prior information in a periodic analysis.
-As mentioned above, Bayesian periodogram extensions have been used quite effectively in a
-variety of circumstances \citep[see, e.g.][]{Bretthorst88}.
-Additionally, the Bayesian result explicitly returns probability density as
-a function of frequency, and it is tempting to think of this as the
+It explicitly returns probability density as a function of frequency,
+and it is tempting to think of this as the
 ``natural'' way to interpret the periodogram.
 
-The problem, however, is that the Bayesian periodogram does {\it not} return
+The problem, however, is that the Bayesian periodogram does {\it not} compute
 the probability that the data is periodic with a given frequency, but rather
 the probability conditioned on a relatively strong assumption:
-that {\it the data are drawn from a sinusoidal model} (or, more generally, the
-particular form of model used to compute the likelihood).
-As such, you should not treat the Bayesian periodogram as a measure of
-generic periodicity, but rather as a measure of how close the data is to
-a sinusoid.
-Now, the standard Lomb-Scargle periodogram can also be viewed as a sinusoidal
-fit, and is thus subject to the same criticism in the least-squares
-interpretation.
+that {\it the data are drawn from a sinusoidal model}.
+As such, the standard Bayesian periodogram is not a useful measure of
+generic periodocity!
+In fact, the Baysian periodogram's derivation under the
+assumption of a sinusoidal signal is perhaps the
+best argument against its use for unknown signals:
+the result of a Bayesian analysis is only ever as good as the assumptions that
+go into it, and for a general, not-necessarily sinusoidal signal, the
+assumptions are incorrect, and the periodogram should not be trusted.
+
+Now, the standard Lomb-Scargle periodogram can also be viewed as derived
+from a sinusoidal fit, and thus might appear subject to the same criticism.
 But unlike the Bayesian periodogram, the standard periodogram affords
-interpretation in light of Fourier analysis and window functions:
-in most cases of interest, such analysis turns out to be {\it vital} to the
-application and interpretation of the periodogram in most cases of interest
-(see \sect{failure-modes}).
-
-In other words, the Bayesian approach goes ``all-in'' on the least squares
+interpretation in light of Fourier analysis and window functions: here
+the incorrect sinusoidal model manifests itself in terms of frequency
+aliases, which can be understood through the analysis of window functions.
+In most cases of interest, such analysis turns out to be {\it vital} to the
+application and interpretation of the periodogram (see \sect{failure-modes}).
+
+In other words, the Bayesian approach essentially
+goes ``all-in'' on the least squares
 interpretation of the periodogram, exponentially suppressing the information
@@
 to Fourier analysis.
-In contrived cases with structureless window functions, exponential attenuation
-of side-peaks and side-lobes may seem appealing \citep[see, e.g.][]{Mortier15},
+In contrived cases with clean sinusoidal data and unstructured window
+functions, exponential attenuation
+of side-lobes and aliases may seem appealing \citep[see, e.g.][]{Mortier15},
 but that appeal extends to the real world only in the most favorable of
-cases---{i.e.} high signal-to-noise measurements with a very well-behaved
-survey window.
+cases---i.e., high signal-to-noise measurements of near-sinusoidal data with
+a very well-behaved survey window.
+In short, you shiould be wary of placing too much trust in a Bayesian
+periodogram, unless you're certain of the type of signal you're looking for.
 
@@
 instrument-dependent parametric models \citep[e.g.][]{Angus16},
-flexible non-parametric functions \citep[e.g.][]{Gregory92}, and
+flexible non-parametric functions \citep[e.g.][]{Gregory92},
+Gaussian Process models \citep[e.g.][]{Wang2012}, and
 specially-designed stochastic models \citep[e.g.][]{Kelly14}.
-Though these can often be more accurate and powerful than the Lomb-Scargle
+Though these are often be more accurate and powerful than the Lomb-Scargle
 approach, they tend to be far more expensive computationally.
@@
 The previous sections have given a conceptual introduction to the Lomb-Scargle
-periodogram, intertwining the Fourier and Least-squares views.
+periodogram and its roots in both Fourier and Least-squares analysis.
 This section gets to the meat of the subject at hand:
 given this understanding of the Lomb-Scargle periodogram and related
-approaches, what practical issues are involved in its effective use?
+approaches, how should we use it most effectively in practice?
 The following sections will identify several of the important issues and
@@
 For non-uniform data, it is not so simple, and there are two important
-considerations: the {\bf frequency limits} and the {\bf grid spacing}.
+considerations: the {\it frequency limits} and the {\it grid spacing}.
 
@@
 the frequencies between the limits.
-This choice turns out to be vital: if you choose too fine a grid, it can
+This choice turns out to be quite important as well: too fine a grid can
 lead to unnecessarily long computation times that can add up quickly in the
-case of large surveys. If you choose too coarse a grid, you
-risk entirely missing narrow peaks that fall between grid points.
+case of large surveys, while too coarse a grid risks
+entirely missing narrow peaks that fall between grid points.
 For example, \fig{LINEAR-coarse-grid} shows the true, well-sampled periodogram
@@
 frequency, the peak widths are not constant in period.
-Regular period grids tend to over-sample at large periods and under-sample
-at small periods;
+Regular grids in period rather than frequency tend to over-sample
+at large periods and under-sample at small periods;
 for this reason it is preferable to use a regular grid in frequency.
@@
 vital to keep in mind the failure modes of the periodogram approach, which
-are rooted in the aliasing and pseudo-aliasing effects discussed in
-\sect{window-functions}.
+are rooted in the aliasing and pseudo-aliasing effects rooted in the structure
+of the window function (recall \sect{window-functions}).
 Due to the interaction of the signal, the convolution due to the survey window,
@@
 \fig{failure-modes} demonstrates this for some simulated data.
-The data consist of 60 noisy observations of each object within a span of
-180 days, with each observation recorded within a few hours of midnight
-each night.
+The data consist of 60 noisy observations each of 1,000 simulated light
+curves within a span of 180 days.
+The window is typical of ground-based data, with each observation recorded
+within a few hours before or after midnight each night.
 The left panel shows the results: the peridogram peak coincides with the true
@@
 signature $f_0$ to be partially aliased at $f_0 + n \delta f$, for integers
-$n$ and $\delta f = 1\ {\rm day}^{-1}$
+$n$ and $\delta f = 1$ cycle/day.
 (Recall \figs{LINEAR-window}{LINEAR-window-effect}).
@@
 where, to be clear, $n$ is a positive or negative integer and $\delta f$ is
-the frequency of a strong feature in the window  (here, $1\ {\rm day}^{-1}$).
+the frequency of a strong feature in the window  (here, $1$ cycle/day).
 For the simulated data, close to 36\% of the objects are mis-characterized
@@
 This may occur for periodic signals that are not strictly sinusoidal, and so
-has power at a higher harmonics.
-Additionally, this higher harmonic peak can be subject to the same aliasing
+have power at a higher harmonics.
+This higher harmonic peak is also subject to the same aliasing
 effects as in \eq{alias-periods}.
-Thus we can adapt \eq{alias-periods} and describe the failure modes depicted
+Thus we can extend \eq{alias-periods} and describe the failure modes depicted
 in panel (b) of \fig{failure-modes} with the following equation,
@@
 \end{equation}
-In \fig{failure-modes}, the $m=2$ harmonic and its aliases account for roughly
-15\% of the results.
+Panel (b) of \fig{failure-modes} illustrates this for $m=2$; this harmonic and
+its aliases account for roughly 15\% of the results.
 
@@
 and this is true of aliases as well.
-The aliases of negative-frequency peaks can cross into the positive-frequency
-range, and this can be accounted for by a further modification of
+If a peak's aliases cross into the negative frequency regime, they are
+effectively reflected into the positive-frequency range.
+This reflection can be accounted for by a further modification of
 \eq{alias-harmonic-periods}---taking its absolute value:
@@
     If a significant peak is found, then $f_{peak}$ is probably an order-$m$
-    harmonic of the true frequency.
+    harmonic of the true frequency.\footnote{Notice that this step can detect
+      aliases like those encoundered in \fig{binary-multiterm},
+      without resorting to a problematic multiterm model.}
   \item Check for peaks at $|f_{peak} \pm n\delta f|$ for at least
-    $n \in \{1, 2\}$. If these aliases exist, then it is possible that you have
+    $n \in \{1, 2\}$ (where $\delta f$ is determined from plotting the window
+    power). If these aliases exist, then it is possible that you have
     found a peak on the sequence of expected aliases---though keep in mind that
@@
 Fourier transform and the Schuster periodogram rather than the Lomb-Scargle
-modification, but doing so would jettison the useful statistical properties
-of Lomb-Scargle, as well as the correspondance with least-squares fitting
-that allows useful extensions to models handling a floating mean and
-observational noise, and would still not address the far more problematic
-first issue.
+modification, but doing so would jettison the benefits of Lomb-Scargle---i.e.,
+its useful statistical properties and extensibility via least squares---and
+this would still not address the far more problematic first issue.
 My feeling is that there is room for a more principled approach to the
@@
 as we saw in the discussion of failure modes in \sect{failure-modes},
-the concern for periodograms is more often inaccuracy associated with
-false peaks and aliases, rather than with imprecision in peak location.
+the concern for periodograms is more often a disjointed inaccuracy associated
+with false peaks and aliases, rather than a more smooth imprecision in
+location of a particular peak.
 
@@
 ratio $\Sigma$; the scaling is approximately \citep[see, e.g.][]{Gregory2001}:
-\begin{eqnarray}
-  \sigma_f &\approx& f_{1/2} / \sqrt{P_{max}}\nonumber\\
-           &\approx& f_{1/2} \sqrt{\frac{2}{N\Sigma^2}}.
+\begin{equation}
+  \sigma_f \approx f_{1/2} \sqrt{\frac{2}{N\Sigma^2}}.
   \eqlabel{bayes-error}
-\end{eqnarray}
+\end{equation}
 This dependence comes from the fact that the Bayesian uncertainty is related
@@
 is the height of the peak, and in particular the height compared to the
-background ``noise'' displayed in the periodogram.
+spurious background peaks that arise in the periodogram.
 \fig{peak-width-height} indicates that this property {\it does} depend on
@@
 errors---lead to a peak of a similar magnitude.
-\citet{Scargle82} showed that for Gaussian noise, the values of the unnormalized
-periodogram in \eq{lomb-scargle-periodogram} are exponentially distributed;
-that is, for {\it one single frequency} $f_0$, if $Z = P(f_0)$ is the
-periodogram value from \eq{lomb-scargle-periodogram}, then
+\citet{Scargle82} showed that for data consisting of pure Gaussian noise,
+the values of the unnormalized
+periodogram in \eq{lomb-scargle-periodogram} follows a $\chi^2$ distribution
+with two degrees of freedom; that is, at any given frequeny $f_0$,
+if $Z = P(f_0)$ is the periodogram value from \eq{lomb-scargle-periodogram},
+then
 \begin{equation}
@@
 is the cumulative probability of observing a periodogram value less than $Z$,
-in data with no periodic signal
-(be aware that for different periodogram normalizations,
-the form of this distribution changes; see \citet{Cumming99} or
-\citet{Baluev2008} for a summary for various normalizations).
+in data consisting only of Gaussian noise.\footnote{
+  Be aware that for different periodogram normalizations
+  (See \sect{normalizations}), the form of this distribution changes;
+  see \citet{Cumming99} or
+  \citet{Baluev2008} for a good summary of the statistical properties of
+  various periodogram normalizations.
+}
 
@@
 them, the probability that it will be greater than 4 is {\it far} larger
-than $1/3$; it is $p(max(r,10) > 4) = 1 - (1 - 1/3)^{10} \approx 0.98$.
-The case for the periodogram is analogous: the probability of seeing a high peak
-simply by chance at any random location (\eq{LS-prob-z}) is relatively small,
-but the probability of seeing a peak when scanning a large number of
-frequencies is much higher.
-
-With the dice, the calculation is easy because the rolls are independent: the
+than $1/3$; it is $p({\rm max}_{10}(r) > 4) = 1 - (1 - 1/3)^{10} \approx 0.98$.
+The case for the periodogram is analogous: the probability of seeing a 
+spurious peak at any single location (\eq{LS-prob-z}) is relatively small,
+but the probability of seeing a single spurious peak among a large number
+of frequencies is much higher.
+
+With the dice, this calculation is easy because the rolls are independent: the
 result of one roll does not affect the result of the next.
@@
 the value at other frequencies in a way that is quite difficult to analytically
-quantify---these correlations come from the convolution with the survey
+express---these correlations come from the convolution with the survey
 window.
@@
 A more sophisticated treatment of the problem is that of \citet{Baluev2008},
-who derived an analytic result based on theory regarding extreme values
-of stochastic processes.
-The Baluev result is not an exact measure of the FAP, but rather an upper
-limit that is shown to be quite close to the true FAP even
-with highly structured window functons.
+who derived an analytic result based on theory of extreme values
+for stochastic processes.
 For the PSD-normalized periodogram, \citet{Baluev2008} showed that the following
@@
 \end{equation}
-and $W = f_{max} \sqrt{4\pi\ {\rm var}(t)}$ measures the effective width of the
+and $W = f_{max} \sqrt{4\pi\ {\rm var}(t)}$ is an effective width of the
 observing window in units of the maximum frequency chosen for the analysis.
@@
 The bootstrap produces the most robust estimate of the FAP because it makes
-no assumptions about statistical distributions, and fully accounts for survey
-window effects.
+few assumptions about the form of the periodogram distribution,
+and fully accounts for survey window effects.
 
@@
   given these observations?''.}
-This boils down to a case of $P(A~|~B) \ne P(B~|~A)$, but still it is common to
-see the false alarm probability treated as if it speaks to the second rather
-than the first question.
-In other words, the false alarm probability tells you the false positive rate
-in a frequentist, rather than Bayesian sense: that is, it is a false positive
-rate conditioned on the entire space of possible observations rather than on
-the particular data you observed.
+This boils down to a case of $P(A \mid B) \ne P(B \mid A)$,
+but still it is common to see the false alarm probability
+treated as if it speaks to the second rather than the first question.
+%In other words, the false alarm probability tells you the false positive rate
+%in a frequentist, rather than Bayesian sense: that is, it is a false positive
+%rate conditioned on the entire space of possible observations rather than on
+%the particular data you observed.
 
@@
 
-To the best of my knowledge, there is currently no silver bullet for answering
+Unfortunately, there is no silver bullet for answering
 these broader, more relevant questions of uncertainty of Lomb-Scargle results.
@@
 pipeline.
-In fact, this type of simulation is a vital component of most---I might even
-venture to say all---current and future time-domain surveys that seek to
+In fact, this type of simulation is a vital component of most---we might even
+venture all---current and future time-domain surveys that seek to
 detect, characterize, and catalog periodic objects,
@@
 computing analytic estimates of uncertainties and false alarm probabilities
-(see the next section).
+explored in \sect{false-alarm-probability}.
 Other normalizations exist as well, but seem to be rarely used in practice.
-For a concise summary of periodogram normalizations and their statistical
-properties, refer to the introduction of \citet{Baluev2008}.
+For a concise summary of several periodogram normalizations and their
+statistical properties, refer to the introduction of \citet{Baluev2008}.
 
@@
 the periodogram. We will see that the na{\"i}ve implementation of
-\eq{lomb-scargle-periodogram} scales poorly with the size of the data, but that
+the standard Lomb-Scargle formula (\eq{lomb-scargle-periodogram})
+scales poorly with the size of the data, but that
 fast alternatives are available.
@@
 $N_f \propto T f_{max} = N f_{max} / \overline{\delta t}$.
-Holding constant the average survey cadence $\overline{\delta t}$ and individual
-observation properties (which control $f_{max}$), we find that the number of
+Holding constant the average survey cadence $\overline{\delta t}$ and
+$f_{max}$, we find that the number of
 frequencies required is directly proportional to the number of data points.
@@
 $K$ Fourier components.
-
-All periodograms in this paper were computed using the
-Astropy\footnote{\url{http://docs.astropy.org/en/stable/stats/lombscargle.html}}
-implementations of the \citet{Press89}, \citet{Zechmeister09}, and
-\citet{Palmer09} algorithms, which were adapted from Python code originally
-published by \citet{ICVG2014} and \citet{VanderPlas2015}.
-All code behind this paper is available in the form of Jupyter notebooks in the
-paper's GitHub repository\footnote{\url{http://github.com/jakevdp/PracticalLombScargle/}}.
 
@@
 apply this algorithm to a dataset:
-
 \begin{enumerate}
@@
 \end{enumerate}
-
 This list is certainly not comprehensive for all uses of the periodogram, but
@@
     individual frequencies, it is not true for the more relevant issue of
-    peak heights across multiple frequencies, which must be either approximated
+    maximum peak heights across multiple frequencies,
+    which must be either approximated
     or computed by bootstrap resampling (see \sect{false-alarm-probability})
@@
 All computations and figures in this paper were produced using Python,
-and in particular the
+and in particular used the
 Numpy \citep{numpy, numpybook},
@@
 and Matplotlib \citep{matplotlib} packages.
-Code to reproduce these figures is available in Jupyter notebooks at
-\url{http://github.com/jakevdp/PracticalLombScargle/}.
+Periodograms where computed using the
+Astropy\footnote{\url{http://docs.astropy.org/en/stable/stats/lombscargle.html}}
+implementations of the \citet{Press89}, \citet{Zechmeister09}, and
+\citet{Palmer09} algorithms, which were adapted from Python code originally
+published by \citet{ICVG2014} and \citet{VanderPlas2015}.
+All code behind this paper, including code to reproduce all figures,
+is available in the form of Jupyter notebooks in the paper's GitHub
+repository\footnote{\url{http://github.com/jakevdp/PracticalLombScargle/}}.
 

### 2017-02-22 PracticalLombScargle.tex
Add reference to Fast Template Method

@@
 Sesar2013} and related work.
+
 Finally, there has been some exploration of extensions to the Lomb-Scargle
@@
 (see \sect{algorithmic-considerations}) are not available for more general
-functional forms.
+functional forms, though there is some promising work in this area: see,
+for example, the {\it Fast Template Periodogram} (Hoffman et al. 2017,
+{\it in prep}) which can quickly construct periodograms from Fourier
+approximations to templates.
 

### 2017-02-22 PracticalLombScargle.tex
More tweaks

@@
 \begin{abstract}
-This paper presents a conceptual introduction to the Lomb-Scargle periodogram,
-and discusses important practical aspects of its use to detect and characterize
-periodic signals.
+The Lomb-Scargle periodogram is a well-known algorithm for detecting and
+characterizing periodic signals in unevenly-sampled data.
+This paper presents a conceptual introduction to the Lomb-Scargle periodogram
+and important practical considerations for its use.
 Rather than a rigorous mathematical treatment, the goal of this paper is to
 build intuition about what assumptions are implicit in the use of the
-periodogram, and to help develop the intuition required to interpret the
-periodogram correctly and use it effectively.
+periodogram, so as to motivate important practical considerations required
+in its application and interpretation.
 \end{abstract}
@@
 \subsection{Outline}
-The remainder of this paper is organized as follows:
+The remainder of this paper is organized as follows:\\
 {\bf\sect{continuous-fourier-transform}} presents a review of the
@@
 including defining the notion of a power spectrum (i.e. Schuster
-periodogram) for detecting periodic content in a signal.
+periodogram) for detecting periodic content in a signal.\\
 {\bf\sect{window-functions}} builds on these properties to explore how
@@
 power spectrum, and discusses a conceptual understanding of the
-Nyquist limiting frequency.
+Nyquist limiting frequency.\\
 {\bf\sect{non-uniform-sampling}} considers non-uniformly sampled signals as
@@
 limit for this case is quite different than what many practitioners in
-the field often assume.
+the field often assume.\\
 {\bf\sect{schuster-to-lomb-scargle}} introduces the Lomb-Scargle periodogram,
 a modified version of the classical periodogram for unevenly-sampled data,
-as well as the motivation behind these modifications.
+as well as the motivation behind these modifications.\\
 {\bf\sect{lomb-scargle-extensions}} discusses a complementary view of the
 Lomb-Scargle periodogram as the result of least-squares model fitting,
-as well as some extensions enabled by that viewpoint.
-{\bf\sect{practical-considerations}} utilizes the concepts introduced
+as well as some extensions enabled by that viewpoint.\\
+{\bf\sect{practical-considerations}} builds on the concepts introduced
 in earlier sections to discuss important practical considerations for
@@
 frequency grid, uncertainties and false alarm probabilities, and modes of
-failure that should be accounted for when working with real-world data.
+failure that should be accounted for when working with real-world data.\\
 {\bf\sect{conclusion}} concludes and summarizes the key recommendations
@@
 for example, the {\it Fast Template Periodogram} (Hoffman et al. 2017,
-{\it in prep}) which can quickly construct periodograms from Fourier
-approximations to templates.
+{\it in prep}\footnote{\url{http://ascl.net/code/v/1559}})
+which can quickly construct periodograms from Fourier approximations to templates.
 
@@
 I think it is worth stepping back to
-revisit the question of {\it why} astronomers lean toward the
+revisit the question of {\it why} astronomers tend to gravitate toward the
 Lomb-Scargle approach rather than the (in many ways simpler) Schuster
@@
     statistics for periodogram peaks.} Reality: while this is true at
-    individual frequencies, it is not true for the more relevant issue of
+    individual frequencies, it is not true for the more relevant question of
     maximum peak heights across multiple frequencies,
     which must be either approximated
-    or computed by bootstrap resampling (see \sect{false-alarm-probability})
+    or computed by bootstrap resampling (see \sect{uncertainties})
   \item {\it Myth: The Lomb-Scargle periodogram corrects for aliasing due to
     sampling and leads to independent noise at each frequency.}
-    Reality: for structured window functions, the Lomb-Scargle
+    Reality: for structured window functions common to most astronomical
+    applications, the Lomb-Scargle
     periodogram has the same window-driven issues as the Schuster
@@
 Pandas \citep{pandas},
-Astropy \citep{Astropy2013},
+AstroPy \citep{Astropy2013},
 and Matplotlib \citep{matplotlib} packages.
 Periodograms where computed using the
-Astropy\footnote{\url{http://docs.astropy.org/en/stable/stats/lombscargle.html}}
-implementations of the \citet{Press89}, \citet{Zechmeister09}, and
+AstroPy implementations\footnote{The AstroPy Lomb-Scargle documentation is at
+\url{http://docs.astropy.org/en/stable/stats/lombscargle.html}}
+of the \citet{Press89}, \citet{Zechmeister09}, and
 \citet{Palmer09} algorithms, which were adapted from Python code originally
@@
 is available in the form of Jupyter notebooks in the paper's GitHub
-repository\footnote{\url{http://github.com/jakevdp/PracticalLombScargle/}}.
+repository\footnote{Source code \& notebooks available at
+\url{http://github.com/jakevdp/PracticalLombScargle/}}.
 
@@
 {\it Acknowledgments:} I am grateful to Jeff Scargle, Max Mahlke,
-Zeljko Ivezic, and Andy Connolly for helpful feedback on early drafts
+and Zeljko Ivezic for helpful feedback on early drafts
 of this paper.

### 2017-03-21 PracticalLombScargle.tex
Add acknowledgment

@@
 \acknowledgments
-{\it Acknowledgments:} I am grateful to Jeff Scargle, Max Mahlke,
-and Zeljko Ivezic for helpful feedback on early drafts
-of this paper.
+{\it Acknowledgments:} I am grateful to Jeff Scargle, Max Mahlke, Dan
+Foreman-Mackey, and Zeljko Ivezic for helpful feedback
+on early drafts of this paper.
 This work was supported by the University of Washington eScience Institute,

### 2017-03-22 PracticalLombScargle.tex
Minor adjustments to text

@@
 this is an irregularly-sampled timeseries showing a single object from the
-LINEAR survey \citep{LINEAR1, LINEAR3}, with magnitude measured 280 times over
-the course of five and a half years.
+LINEAR survey \citep{LINEAR1, LINEAR3}, with un-filtered magnitude measured
+280 times over the course of five and a half years.
 By eye, it is clear that the brightness of the object varies in time with a
@@
 \begin{enumerate}
+  \item How does the Lomb-Scargle periodogram relate to the classical
+    Fourier power spectrum?
   \item What is the source of the pattern of multiple peaks revealed by the
@@
 detecting and characterizing periodicity in time series.
-
 First, it is important to note that there are many different modes of
@@
   \item[Least Squares Methods] involve fitting a model to the data at each
-    candidate frequency, and selecting the model which maximizes the likelihood.
+    candidate frequency, and selecting the frequency which maximizes the likelihood.
     The Lomb-Scargle periodogram also falls in this category
@@
 while \citet{Graham2013b} instead take an empirical approach and find that
-when considering detection efficiency in real datasets, no algorithm
-universally outperforms the others.
+when considering detection efficiency in real datasets, no suitably efficient
+algorithm universally outperforms the others.
 
@@
 technique to compute periodicity of unequally-spaced data in astronomy and
-other fields, and so is the first technique many will reach for
+other fields, and so is the first tool many will reach for
 when searching for periodic content in a signal.
 But there is a deeper reason as well; it turns out that Lomb-Scargle occupies
-a unique niche: it is motivated by Fourier analysis,
+a unique niche: it is motivated by Fourier analysis (see \sect{schuster-to-lomb-scargle}),
 but it can also be viewed as a least squares method
-(see \sect{schuster-to-lomb-scargle}), can be derived from the principles
+(see \sect{lomb-scargle-extensions}).
+It can be derived from the principles
 of Bayesian probability theory (see \sect{bayesian-periodograms}), and has
@@
 some circumstances \citep[see][]{Swingler89}.
-Thus, the Lomb-Scargle periodogram occupies a unique point of correspondance
+Thus, the Lomb-Scargle periodogram occupies a unique point of correspondence
 between many classes of methods, and so provides a focus for discussion that
-can speak to important considerations for all of these methods.
+has bearing on considerations involved in {\it all} of these methods.
 
@@
 continous Fourier transform and some of its useful properties,
-including defining the notion of a power spectrum (i.e. Schuster
+including defining the notion of a power spectrum (i.e. classical/Schuster
 periodogram) for detecting periodic content in a signal.\\
@@
 Consider a continuous signal $g(t)$.
-its Fourier transform is given by the following integral, where $i\equiv\sqrt{-1}$ denotes the imaginary unit:
+Its Fourier transform is given by the following integral, where $i\equiv\sqrt{-1}$ denotes the imaginary unit:
 \begin{equation}
@@
 of delta functions marking the frequencies of those sinusoids: that is,
-the Fourier transform directly measures periodic content in a function.
+the Fourier transform directly measures periodic content in a continuous function.
 
 Further, if we compute the squared amplitude of the resulting transform, we
-can both do away with complex components and remove the phase imparted by any
-time offsets; this squared amplitude is usually known as the
+can both do away with complex components and remove the phase imparted by
+the choice of temporal baseline; this squared amplitude is usually known as the
 {\it Power Spectrum}:
@@
 We have already discussed that the Fourier transform of a complex exponential is a single delta function.
-This is just one of many ``Fourier Pairs'' to keep in mind as we progress to understanding the Lomb-Scargle Periodogram.
+This is just one of many Fourier pairs to keep in mind as we progress to understanding the Lomb-Scargle periodogram.
 We list a few more important pairs here (see \Fig{fourier-pairs} for a visual representation of the following pairs):
@@
 For our purposes, this property will turn out to be quite
-important as we push further in understanding the Lomb-Scargle Periodogram.
+important as we push further in understanding the Lomb-Scargle periodogram.
 
@@
 transform of the continuous underlying function, but rather the transform
-of the {\it point-wise product of the signal and the observing window}
+of the {\it point-wise product of the signal and the observing window.}
 Symbolically, if the signal is $g(t)$ and the window is $W(t)$, the observed
@@
 The observed signal in this case can be understood as the pointwise product of
-the underlying infinite periodic signal, and a rectangular window function.
+the underlying infinite periodic signal with a rectangular window function.
 By the convolution theorem, the Fourier transform will be given by the
@@
 \subsubsection{The Discrete Fourier Transform}
-For regularly-spaced observations defined by a Dirac comb, the delta functions
-serve to collapse the Fourier integral into a sum over periodic components,
-and in this way we can arrive at the common form of the discrete Fourier
+When a continuous function is sampled at regular intervals, the delta functions
+in the Dirac comb window serve to collapse the Fourier integral into a Fourier
+sum, and in this manner we can arrive at the common form of the discrete Fourier
 transform.
@@
 
-\subsection{The Schuster Periodogram}
+\subsection{The Classical Periodogram}
 
@@
 apply the definition of the Fourier power spectrum from \eq{power-spectrum}
-to compute the {\it Schuster Periodogram}, first proposed by \citet{Schuster98}:
+to compute the {\it Classical Periodogram}, sometimes called the
+{\it Schuster Periodogram} after \citet{Schuster98} who first proposed it:
 \begin{equation}
@@
 spectrum in \eq{power-spectrum}, computed for a continuous signal with
-uniform sampling.
+uniform sampling defined by a Dirac comb.
 It follows that, in the uniform sampling case, the Schuster periodogram
@@
 
-One might hope that sampling the signal more densely might limit these problems,
-and it does, but only to a degree.
+One might hope that sampling the signal more densely might alleviate
+these problems, and it does, but only to a degree.
 In \Fig{random-window-2} we increase the density of observations by a factor of 10,
@@
 
-As we have seen in \sect{nyquist}, the Nyquist limit is a direct consequence
+We saw in \sect{nyquist} that the Nyquist limit is a direct consequence
 of the symmetry in the Dirac comb window function that describes evenly-sampled data,
@@
 All of these ``pseudo-Nyquist'' limits are tempting criteria in that
-they recover the classical Nyquist
+they are easy to compute, and reduce to the classical Nyquist
 frequency in the limit of evenly-spaced data.
@@
 In other words, computing the Nyquist limit for unevenly-spaced data requires
-finding some common factor, such that each spacing $\Delta t_i$ is {\it exactly}
-an integer multiple of this factor.
+finding the largest factor $p$, such that each spacing $\Delta t_i$ is
+{\it exactly} an integer multiple of this factor.
 \citet{Eyer99} prove this formally, but the result can be understood
@@
 real-world observations: one ground-based (LINEAR) and one space-based (Kepler).
-For details on how window power spectra can be computed in practice, see
+For details on how window power spectra can be estimated in practice, see
 \sect{windows-and-deconvolution}.
@@
 taken near the same time of day: this is typical of a ground-based survey
-with observations recorded only during the night hours.
+with observations recorded only during the nighttime hours.
 The additional spikes at periods of $1/n$ days (for integer $n$) are aliases
@@
 \fig{LINEAR-window} also shows a wide spike at a period of 1 year, indicating
-that there is a detectable annual pattern in the observations.
+a detectable annual pattern in the observations.
 Finally, there is a noticeable spike at approximately 14 days that is likely
@@
 measured 4083 times over a period of three months,
-with an irregular observing cadence of just under 30 minutes
+with an irregular observing cadence of around 30 minutes
  \citep[For deeper discussion of these observations, see][]{Kolenberg2010}.
@@
 
-\section{From Schuster to Lomb-Scargle}
+\section{From Classical to Lomb-Scargle Periodograms}
 \sectlabel{schuster-to-lomb-scargle}
@@
 Up until now, we have been mainly discussing the direct extension of the
-Schuster periodogram in \eq{schuster-periodogram} to non-uniform data.
+classical periodogram in \eq{schuster-periodogram} to non-uniform data.
 Returning to this definition, we can rewrite the expression in a more
@@
 periodic signals, its statistical properties are not as straightforward as in
-the uniform case: the periodogram has noise properties that are nontrivial
-to quantify, even when applied to ideal data (data with uncorrelated
-Gaussian noise).
-For the sake of evaluating the significance of periodogram peaks
-(see \sect{uncertainties}), it would be preferable if the nonuniform
-periodogram had a noise distribution that could be
-straightforwardly computed from the uncertainties of the data.
-
-\citet{Scargle82} showed that while the classical periodogram of
-\eq{classical-periodogram} has such desireable statistical properties
-in the case of uniform sampling, this does not hold
-when the sampling becomes non-uniform.
-To address this, he considered a generalized form of the periodogram,
+the uniform case.
+When the classical periodogram is applied to uniformly-sampled Gaussian noise,
+the values of the resulting periodogram is chi-square distributed.
+This property becomes quite useful in practice when the periodogram is used in the
+context of a classical hypothesis test to distinguish between periodic and
+non-periodic objects---see \sect{false-alarm-probability}.
+Unfortunately, when the sampling becomes nonuniform these properties no longer
+hold and the periodogram distribution cannot in general be analytically expressed.
+
+\citet{Scargle82} addressed this by considering a generalized form of the periodogram,
 \begin{equation}
@@
   \item The periodogram reduces to the classical form in the case of equally-spaced observations,
-  \item The periodogram has desireable statistical properties, in the sense discussed above,
+  \item The periodogram's statistics are analytically computable,
   \item The periodogram is insensitive to global time-shifts in the data.
 \end{enumerate}
-The values of $A$ and $B$ leading to these properties lead to the following
-modification of the classical expression for the periodogram:
+The values of $A$ and $B$ leading to these properties result in the following
+form of the generalized periodogram:
 \begin{eqnarray}
@@
     Though the two periodogram estimates differ quantitatively, the essential
-    qualitative features---namely the position of significant peaks---generally
+    qualitative features---namely the position of significant peaks---typically
     remain the same.
@@
 orthogonalize the normal equations used in the least squares analysis.
-Partly due to the deep connection between Fourier analysis and least-squares
+Partly due to this deep connection between Fourier analysis and least-squares
 analysis, the modified periodogram in \eq{lomb-scargle-periodogram}
@@
 \citep[see, e.g.][]{Gottlieb75}.
-Because of this similarity between the Schuster and Lomb-Scargle periodograms,
+
+Because of the close similarity between the classical and Lomb-Scargle periodograms,
 the bulk of or previous discussion applies---at least qualitatively---to
@@
 In particular, reasoning about the effect of window functions on the observed
-Lomb-Scargle power spectrum remains useful even if it is not precisely
-accurate.
+Lomb-Scargle power spectrum remains qualitatively useful even if it is not
+quantitatively precise.
 
@@
 \sectlabel{extensions-observational-noise}
-Perhaps the most important modification of the periodogram is to
-construct a periodogram that correctly handles measurement error in the data.
-This involves changing the expression for the $\chi^2$ statistic in a manner
-analogous to the change for a regression problem.
-If there are Gaussian errors $\sigma_n$ on each observed $y_n$, we can re-write \eq{chi2-simple} in the following (standard) way:
+Perhaps the most important modification to the periodogram is to
+construct it such that it correctly handles measurement error in the data.
+This can be done through the standard change to the $\chi^2$ expression:
+i.e., if there are Gaussian errors $\sigma_n$ on each observed $y_n$,
+we can re-write \eq{chi2-simple} in the following (standard) way:
 \begin{equation}
@@
 of a multiplicative weight $1/\sigma_n$ within each of the summations.
-Early versions of this sort of modification appeared in \citet{Gilliland87}, and
+Early versions of this sort of modification appeared in \citet{Gilliland87} and
 \citet{Irwin89}; \citet{Scargle89} derived this ``weighted'' form of the
@@
 data are pre-centered around the mean value of the (unknown) signal.
-In many analyses, this requirement is accomplished by pre-centering using
-the mean estimated from the data points themselves.
-When the data provide full phase coverage of the observed signal,
-this approach is generally robust.
-However, due to selection effects and survey cadence, full phase
-coverage can not always be guaranteed, which can potentially lead to situations
-where the periodogram suppresses peaks of interest.
+In many analyses, this requirement is accomplished by pre-centering data
+about the sample mean: this approach is generally robust when
+the data provide full phase coverage of the observed signal;
+however, due to selection effects and survey cadence, full phase
+coverage can not always be guaranteed.
+Using the sample mean in such cases can potentially lead to suppression of
+peaks of interest.
 
@@
 
-\Fig{standard-vs-floatingmean} shows a simulated example of this: the data
+A simulated example of this is shown in \Fig{standard-vs-floatingmean}: the data
 consist of noisy observations of a sinusoidal signal in which the faintest
@@
 that identifies this true period (upper right panel).
-\citet{Zechmeister09} provide a detailed study of this model,
-and show that the addition of the floating mean term does not change
-the useful statistical properties of the periodogram.
+A detailed study of the floating-mean model is given by
+\citet{Zechmeister09} who show that the addition of the floating mean term
+does not change the useful statistical properties outlined in \sect{schuster-to-lomb-scargle}.
 
@@
 
-A further generalization involves multi-term Fourier models:
+A further generalization of the least squares periodogram involves multi-term Fourier models:
 rather than fitting just a single sinusoid at each frequency, we might fit a
@@
 \citet{Bretthorst88} takes a comprehensive look at this type of multi-term
-extension to the periodogram, as well as extensions to decaying signals and
-other more complex models.
+extension to the periodogram, as well as related extensions to decaying signals
+and other more complex models.
 
@@
 {\it at all frequencies}, not just at frequencies of interest!
+Indeed, this effect can be readily observed in \fig{binary-multiterm}.
 
@@
 
-This background noise can also be understood in terms of aliasing.
+This background noise in the multi-term periodogram can also be understood
+in terms of aliasing.
 Consider \fig{kepler-multiterm}, which shows several multi-term fits to the
@@
 The problem, however, is that the Bayesian periodogram does {\it not} compute
-the probability that the data is periodic with a given frequency, but rather
-the probability conditioned on a relatively strong assumption:
+the probability that the data are periodic with a given frequency, but rather
+that probability conditioned on a relatively strong assumption:
 that {\it the data are drawn from a sinusoidal model}.
@@
 the result of a Bayesian analysis is only ever as good as the assumptions that
-go into it, and for a general, not-necessarily sinusoidal signal, the
+go into it, and for a general (not-necessarily sinusoidal) signal, those
 assumptions are incorrect, and the periodogram should not be trusted.
@@
 The following sections will identify several of the important issues and
-questions that are often ignored in the literature, but are nevertheless
-vital to using the periodogram in practice.
+questions that are not often addressed in the literature on the subject,
+but are nevertheless vital to consider when using the periodogram in practice.
 
@@
 by its temporal resolution, which is 6 digits beyond the decimal point in days.
-From \eq{nonuniform-nyquist}, we can write $f_{Ny} = 500,000\ {\rm days}^{-1}$.
-Given the observing window of $T = 1962\ {\rm days}$, we find that five
+From \eq{nonuniform-nyquist}, we can write $f_{Ny} = 500,000\ {\rm days}^{-1}$,
+and given the observing window of $T = 1962\ {\rm days}$, we find that five
 evaluations per peak across the entire detectable frequency range
@@
   \item Check for peaks at $|f_{peak} \pm n\delta f|$ for at least
-    $n \in \{1, 2\}$ (where $\delta f$ is determined from plotting the window
-    power). If these aliases exist, then it is possible that you have
+    $n \in \{1, 2\}$ (where $\delta f$ is determined from plotting the survey
+    window power, and is generally $(1\ {\rm day})^{-1}$ for ground-based surveys).
+    If these aliases exist, then it is possible that you have
     found a peak on the sequence of expected aliases---though keep in mind that
@@
 Comparing this to \eq{schuster-periodogram}, we see that this is essentially
-the Schuster periodogram for data $g_n=1$ at all times $t_n$.
+the classical periodogram for data $g_n=1$ at all times $t_n$.
 With this fact in mind, one convenient way to estimate the form of the
@@
 from the true window just as the Lomb-Scargle periodogram differs from the
-Schuster periodogram---but in practice it is accurate enough to give
+classical periodogram---but in practice it is accurate enough to give
 useful insight into the window function's important features.
@@
 power spectrum.
-Recall from \eq{g-nonuniform} that the observed data is a point-wise product
+Recall from \eq{g-nonuniform} that the observed data are a point-wise product
 of the underlying function and the survey window:
@@
 values of $t$ and so $1/W_{\{t_n\}}(t)$ and its Fourier transform are not well
-defined. For this reason, direct deconvolution is not an option in most
+defined. Because of this, direct deconvolution is not an option in most
 cases of interest---in other words, the deconvolution problem is ill-posed
@@
 rather than the PSD; trying to clean a PSD directly ignores important phase
-information. Third, the CLEAN algorithm assumes a Schuster-type FFT analysis:
-while the Lomb-Scargle periodogram is equivalent to Schuster in the limit of
+information. Third, the CLEAN algorithm assumes a classical FFT analysis:
+while the Lomb-Scargle periodogram is equivalent to classical periodogram in the limit of
 equally-spaced observations, it is not equivalent in the relevant case of
@@
 The latter two issues could be remedied by focusing on the non-uniform
-Fourier transform and the Schuster periodogram rather than the Lomb-Scargle
+Fourier transform and the classical periodogram rather than the Lomb-Scargle
 modification, but doing so would jettison the benefits of Lomb-Scargle---i.e.,
@@
 unconstrained deconvolution problem for periodograms derived from non-uniform
-fast Fourier transforms, perhaps through some sort of L1/lasso regularization
-that imposes assumptions of sparsity on the true periodogram, but to date this
+fast Fourier transforms (perhaps through some sort of L1/lasso regularization
+that imposes assumptions of sparsity on the true periodogram) but to date this
 does not appear to have been explored anywhere in the literature.
@@
 \sectlabel{uncertainties}
-An important piece of reporting results from the Lomb-Scargle periodogram is
-the uncertainty of the resulting period estimate.
-In most areas of science we are used to being able to report uncertainties in
+An important aspect of reporting results from the Lomb-Scargle periodogram is
+the uncertainty of the estimated period.
+In many areas of science we are used to being able to report uncertainties in
 terms of errorbars, {e.g.} ``the period is $16.3 \pm 0.6$ hours''.
@@
 
-We have motivated from the Fourier window discussions that peak width is
-the inverse of the observational baseline; the surprising result is that to
+We have motivated from the Fourier window discussions that peak width $f_{1/2}$
+is the inverse of the observational baseline; the surprising result is that to
 first order the peak width in the periodogram {\it does not depend} on
@@
 much more relevant uncertainties demonstrated in \sect{failure-modes},
-in which we might find ourselves on the wrong peak entirely!
+in which we might find ourselves on the wrong peak entirely.
 This is why in general, peak widths and Gaussian error bars should generally
@@
 In particular, it emphatically does {\it not} answer the much more relevant
-question {\it ``what is the probability that this data is periodic
+question {\it ``what is the probability that this dataset is periodic
   given these observations?''.}

### 2017-03-23 PracticalLombScargle.tex
minor edits and rewordings

@@
 periodogram, which is a quite different situation.
-If this is confusing, consider rolling a standard 6-sided die.
+
+By analogy, consider rolling a standard 6-sided die.
 The probability, in a single roll, of rolling a number, say $r > 4$ is easy
@@
 for stochastic processes.
-For the PSD-normalized periodogram, \citet{Baluev2008} showed that the following
+For the standard periodogram in \eq{lomb-scargle-periodogram},
+\citet{Baluev2008} showed that the following
 formula for the FAP provides a close upper-limit even in the case of
@@
 \subsubsubsection{Bootstrap Method}
-In the absence of a true analytic solution to the False Alarm Probability, we
-can turn to computational methods such as the bootstrap.
+In the absence of a true analytic solution to the false alarm probability,
+we can turn to computational methods such as the bootstrap.
 The bootstrap method is a technique in which the statistic in question is
@@
 pipeline.
-In fact, this type of simulation is a vital component of most---we might even
-venture all---current and future time-domain surveys that seek to
+In fact, this type of simulation is a vital component of most (if not all)
+current and future time-domain surveys that seek to
 detect, characterize, and catalog periodic objects,
@@
 each frequency.
-Note that this changes if data uncertainties are incorporated into
-the periodogram as in \sect{extensions-observational-noise},
+Note, however, that the units change if data uncertainties are incorporated
+into the periodogram as in \sect{extensions-observational-noise};
 the periodogram in this normalization becomes unitless: it is essentially
@@
 periodogram is maximized at a value of $\hat{\chi}^2_0 / 2$.
-On the other end, it is mathematically impossible for a sinusoidal model to
-do worse than the simple constant, non-varying fit reflected in
-$\hat{\chi}^2_0$, and so the minimum value of \eq{lomb-scargle-chi2-2} is
-exactly zero.
+On the other end, it is mathematically impossible for a best-fit sinusoidal
+model at any frequency to do worse than the simple constant, non-varying
+fit reflected in $\hat{\chi}^2_0$, and so the minimum value of
+\eq{lomb-scargle-chi2-2} is exactly zero.
 
@@
 We will finish here with a brief summary of these practical
-recommendations, a brief discussion of other methods for detecting
-periodicity, and a somewhat opinionated
+recommendations, along with a somewhat opinionated
 post-script for further thought.
@@
 revisit the question of {\it why} astronomers tend to gravitate toward the
-Lomb-Scargle approach rather than the (in many ways simpler) Schuster
+Lomb-Scargle approach rather than the (in many ways simpler) classical
 periodogram.
 
-As discussed in \sect{schuster-to-lomb-scargle}, the periodogram has two
-benefits over the Schuster approach:
-the noise distribution at each individual frequency is the sum
-of two Gaussians, and the result is equivalent to a periodogram derived
+As discussed in \sect{schuster-to-lomb-scargle}, the Lomb-Scargle approach has
+two distinct benefits over the classical periodogram:
+the noise distribution at each individual frequency is chi-square distributed
+under the null hypothesis,
+and the result is equivalent to a periodogram derived
 from a least squares analysis.
@@
   \item {\it Myth: The Lomb-Scargle periodogram can be computed more efficiently
-    than the Schuster periodogram.} Reality: computationally, the two
+    than the classical periodogram.} Reality: computationally, the two
     are quite similar, and in fact the fastest Lomb-Scargle algorithm currently
-    available is based on the Schuster periodogram computed via the
+    available is based on the classical periodogram computed via the
     the NFFT algorithm (see \sect{algorithmic-considerations}).
@@
     applications, the Lomb-Scargle
-    periodogram has the same window-driven issues as the Schuster
+    periodogram has the same window-driven issues as the classical
     periodogram, including spurious peaks due to partial aliasing, and
@@
 With these myths set aside, what is the practical advantage of Lomb-Scargle
-over a Schuster approach?
-What would we lose if we instead used a Schuster periodogram,
+over a classical periodogram?
+What would we lose if we instead used a simple Fourier periodogram,
 estimating uncertainty, significance, and false alarm probability by
@@
 need to account for multiple frequencies, so the only advantage left is the
-correspondance to least squares and Bayesian models, and in particular the
+correspondence to least squares and Bayesian models, and in particular the
 ability to generalize to more complicated models where appropriate---but in
@@
 The equivalence of Lomb-Scargle to a Bayesian sinusoidal model is perhaps
-interesting trivia, but not itself a reason to use that model if your data
-is not known {\it a priori} to be sinusoidal---it could even be construed as
+an interesting bit of trivia, but not itself a reason to use that model if your
+data is not known {\it a priori} to be sinusoidal---it could even be construed as
 an argument {\it against} Lomb-Scargle in the general case where the assumption
@@
 
-Conversely, if you replace your Lomb-Scargle approach with a Schuster approach,
+Conversely, if you replace your Lomb-Scargle approach with a classical periodogram,
 what you gain is the ability to reason quantitatively about the effects of the
@@
 question the prevailing folk wisdom of the advantage of Lomb-Scargle
-over approaches based directly on the Fourier transform and Schuster
+over approaches based directly on the Fourier transform and classical
 periodogram.

### 2017-03-23 PracticalLombScargle.tex
Add discussion of periodogram as estimator of power spectrum

@@
 {\bf\sect{window-functions}} builds on these properties to explore how
-different observation patterns (i.e. window functions) affect the observed
-power spectrum, and discusses a conceptual understanding of the
+different observation patterns (i.e. window functions) affect the
+periodogram, and discusses a conceptual understanding of the
 Nyquist limiting frequency.\\
@@
 the choice of temporal baseline; this squared amplitude is usually known as the
-{\it Power Spectrum}:
+{\it power spectral density} or the {\it power spectrum}:
 \begin{equation}
@@
 Apart from the $1/N$ proportionality, this sum is precisely the Fourier power
-spectrum in \eq{power-spectrum}, computed for a continuous signal with
-uniform sampling defined by a Dirac comb.
+spectrum in \eq{power-spectrum}, computed for a continuous signal observed
+with uniform sampling defined by a Dirac comb.
 It follows that, in the uniform sampling case, the Schuster periodogram
@@
 will explore in the following section.
+
+One point that should be emphasized is that the {\it periodogram} in
+\eq{schuster-periodogram} and the {\it power spectrum} in \eq{power-spectrum}
+are conceptually different things.
+As noted in \citet{Scargle82}, the astronomy community tends to use these
+terms interchangeably, but to be precise
+the periodogram---i.e., the statistic we compute from our data---is an
+{\it estimator} of the power spectrum---i.e., the underlying continuous
+function of interest.
+In fact, the classical periodogram and its extensions (including the
+Lomb-Scargle we will discuss momentarily)
+are not consistent estimators of the power spectrum: that is, the periodogram
+has unavoidable intrinsic variance, even in the limit of an infinite number
+of observations
+\citep[for a detailed discussion, see Chp 8.4 of][]{Anderson1971}.
 

### 2017-03-28 PracticalLombScargle.tex
More figure and text tweaks

@@
 build intuition about what assumptions are implicit in the use of the
-periodogram, so as to motivate important practical considerations required
-in its application and interpretation.
+Lomb-Scargle periodogram and related estimators of periodicity,
+so as to motivate important practical considerations required
+in its proper application and interpretation.
 \end{abstract}
@@
 The Lomb-Scargle periodogram \citep{Lomb76, Scargle82}
-is a well-known algorithm for detecting periodicity in unevenly-sampled
+is a well-known algorithm for detecting and characterizing
+periodicity in unevenly-sampled
 time-series, and has seen particularly wide use within the astronomy community.
@@
 The Lomb-Scargle periodogram is a method that allows efficient computation of
-a Fourier-like power spectrum from such unevenly-sampled data, resulting in
-an intuitive means of determining the period of oscillation.
+a Fourier-like power spectrum estimator from such unevenly-sampled data,
+resulting in an intuitive means of determining the period of oscillation.
 
@@
 
-The left panel of \fig{LINEAR-power} and its inset show the Lomb-Scargle
-periodogram computed for the data in \fig{LINEAR-data}.
-Computing the Lomb-Scargle periodogram for the data gives us a measure of the
+The Lomb-Scargle periodogram computed from this data is shown in the left
+panel  of \fig{LINEAR-data}.
+The Lomb-Scargle periodogram here yields an estimate of the Fourier
 power as a function of period of oscillation (\fig{LINEAR-power}, left), from
-which we can determine the period of oscillation of approximately 2.58 hours.
+which we can read-off the period of oscillation of approximately 2.58 hours.
 The right panel of \fig{LINEAR-power} shows a folded visualization of
@@
     periodogram?
-  \item What assumptions does the Lomb-Scargle approach make regarding the
-    form of the unknown signal?
+  \item What assumptions, if any, does the Lomb-Scargle approach make
+    regarding the form of the unknown signal?
   \item How should we understand and report the uncertainty of the
@@
 involved, but rather seeks to develop the intuition of {\it how} to think
-about these questions.
+about these questions, with references to relevant technical treatments.
 
@@
     and closely related correlation functions. These methods include the
-    Schuster periodogram \citep{Schuster98}, the Lomb-Scargle periodogram
+    classical or Schuster periodogram \citep{Schuster98},
+    the Lomb-Scargle periodogram
     \citep{Lomb76, Scargle82}, the correlation-based method of
@@
 the choice of temporal baseline; this squared amplitude is usually known as the
-{\it power spectral density} or the {\it power spectrum}:
+{\it power spectral density} or simply the {\it power spectrum}:
 \begin{equation}
@@
     a lower sampling rate. The result is that the Fourier transform of the
-    window function (lower left) has spacing narrower than the Fourier transform
-    of the signal (middle left), meaning the observed Fourier transform has
+    window function (middle right) has spacing narrower than the Fourier
+    transform of the signal (upper right), meaning the observed Fourier
+    transform (lower right) has
     aliasing of signals, such that not all frequency information can be
@@
 apply the definition of the Fourier power spectrum from \eq{power-spectrum}
-to compute the {\it Classical Periodogram}, sometimes called the
-{\it Schuster Periodogram} after \citet{Schuster98} who first proposed it:
+to compute the {\it classical periodogram}, sometimes called the
+{\it Schuster periodogram} after \citet{Schuster98} who first proposed it:
 \begin{equation}
@@
 functional forms, though there is some promising work in this area: see,
-for example, the {\it Fast Template Periodogram} (Hoffman et al. 2017,
-{\it in prep}\footnote{\url{http://ascl.net/code/v/1559}})
+for example, the {\it Fast Template Periodogram}\footnote{\url{http://ascl.net/code/v/1559}}
+(Hoffman et al. 2017, {\it in prep})
 which can quickly construct periodograms from Fourier approximations to templates.
@@
     while bottom panels show the PSD-normalized periodogram
-    (\eq{lomb-scargle-periodogram})
+    (\eq{lomb-scargle-periodogram}) scaled by noise variance.
     Perhaps surprisingly, neither the number of points nor the signal-to-noise
@@
 \end{itemize}
-With these myths set aside, what is the practical advantage of Lomb-Scargle
-over a classical periodogram?
-What would we lose if we instead used a simple Fourier periodogram,
+With these misconceptions corrected, what is the practical advantage of
+Lomb-Scargle over a classical periodogram?
+What would we lose if we instead used the simple classical Fourier periodogram,
 estimating uncertainty, significance, and false alarm probability by
@@
 what you gain is the ability to reason quantitatively about the effects of the
-survey window function on the resulting periodogram (see \sect{CLEAN}).
+survey window function on the resulting periodogram (cf. \sect{CLEAN}).
 While the deconvolution problem is ill-posed, there is no reason to assume
 this is a fatal defect: ill-posed linear models are solved routinely in
-other areas of computational research, particularly by using priors or
-various forms of regularization.
+other areas of computational research, particularly by using sparsity
+priors or various forms of regularization.
 In any case, I would contend that there is ample room for practitioners to
@@
 {\it Acknowledgments:} I am grateful to Jeff Scargle, Max Mahlke, Dan
-Foreman-Mackey, and Zeljko Ivezic for helpful feedback
+Foreman-Mackey, Zeljko Ivezic, and David Hogg for helpful feedback
 on early drafts of this paper.

### 2017-03-28 PracticalLombScargle.tex
Final tweaks for arXiv submission
fix some typos

@@
 %\documentclass{emulateapj}
+\pdfoutput=1
 
@@
 \usepackage{fancyhdr}
-\graphicspath{{figures/}}
+\graphicspath{{./}{figures/}}
 
@@
 
-\title{Understanding the Lomb-Scargle Periodogram\\
-       {\it (Draft Version \today)}}
+\title{Understanding the Lomb-Scargle Periodogram}
+
+%\title{Understanding the Lomb-Scargle Periodogram\\
+%       {\it (Draft Version \today)}}
 
 % Add ``Draft version'' footer
-\pagestyle{fancy}
-\fancyhf{}\renewcommand{\headrulewidth}{0pt} % remove header titles
-\cfoot{\it \footnotesize $\sim$ Draft Version, \today $\ \sim$}
+%\pagestyle{fancy}
+%\fancyhf{}\renewcommand{\headrulewidth}{0pt} % remove header titles
+%\cfoot{\it \footnotesize $\sim$ Draft Version, \today $\ \sim$}
 
@@
 The Lomb-Scargle periodogram computed from this data is shown in the left
-panel  of \fig{LINEAR-data}.
+panel  of \fig{LINEAR-power}.
 The Lomb-Scargle periodogram here yields an estimate of the Fourier
-power as a function of period of oscillation (\fig{LINEAR-power}, left), from
+power as a function of period of oscillation, from
 which we can read-off the period of oscillation of approximately 2.58 hours.

### 2017-09-13 PracticalLombScargle.tex

@@
   \caption{Multi-term models (left) and their corresponding periodograms (right)
-    for the kepler data shown in \fig{kepler-data}
+    for the Kepler data shown in \fig{kepler-data}
     \figlabel{kepler-multiterm}}
@@
 within a few hours before or after midnight each night.
-The left panel shows the results: the peridogram peak coincides with the true
+The left panel shows the results: the periodogram peak coincides with the true
 period in under 50\% of cases, and the modes of failure lead to noticeable

### 2017-12-06 PracticalLombScargle.tex
Respond to reviewers

@@
 of delta functions marking the frequencies of those sinusoids: that is,
-the Fourier transform directly measures periodic content in a continuous function.
+the Fourier transform directly measures additive periodic content in a
+continuous function.
 
@@
 \end{equation}
-From the definition, it is clear that a convolution amounts to ``sliding'' one
-function past the other, integrating at each step.
+The convolution can be thought of as an operation that ``slides'' one
+function past the other, integrating the product at each step.
 Such an operation is commonly used, for example, in smoothing a function,
@@
 \end{equation}
-In practice, this can be a much more efficient means of computing a convolution
+In practice, this can be a much more efficient means of numerically computing a convolution
 than to directly solve at each time $t$ the integral over $\tau$ that appears
@@
 width of its transform (see \Fig{fourier-pairs}), it follows that a wider
-observing window leads to proportionally less spread in the Fourier
-transform of the observed function.
+observing window leads to proportionally less spread in each peak of
+the observed Fourier transform.
 
@@
 to our data a rectangular window function of width $N\Delta t$.
-From the discussion accompanying \fig{rectangular-window}, we know what this
-does: it gives us a Fourier transform convolved with a sinc function of width
+From the discussion accompanying \fig{rectangular-window}, we know the result: it yields a Fourier transform convolved with a sinc function of width
 $1 / (N\Delta t)$, resulting in the ``smearing'' of the Fourier transform signal
@@
 In fact, the classical periodogram and its extensions (including the
-Lomb-Scargle we will discuss momentarily)
+Lomb-Scargle periodogram that we will discuss momentarily)
 are not consistent estimators of the power spectrum: that is, the periodogram
@@
 of observations
-\citep[for a detailed discussion, see Chp 8.4 of][]{Anderson1971}.
+\citep[for a detailed discussion, see Chapter 8.4 of][]{Anderson1971}.
 
@@
 leads to a randomization of Fourier peak locations and heights.
-That is, {\it non-structured spacing of observations will lead to
-a non-structured frequency peaks in the window transform}.
+In other words, {\it non-structured spacing of observations lead directly to
+non-structured frequency peaks in the window transform}.
 This non-structured window transform, when convolved with the Fourier transform
@@
 the uniform case.
-When the classical periodogram is applied to uniformly-sampled Gaussian noise,
+When the classical periodogram is applied to uniformly-sampled Gaussian
+white noise,
 the values of the resulting periodogram is chi-square distributed.
@@
   \item The periodogram reduces to the classical form in the case of equally-spaced observations,
-  \item The periodogram's statistics are analytically computable,
+  \item The periodogram's distribution is analytically computable,
   \item The periodogram is insensitive to global time-shifts in the data.
@@
 
-\subsection{Measurement Errors}
+\subsection{Measurement Uncertainties}
 \sectlabel{extensions-observational-noise}
 Perhaps the most important modification to the periodogram is to
-construct it such that it correctly handles measurement error in the data.
+construct it such that it correctly handles uncertainties in the observed data.
 This can be done through the standard change to the $\chi^2$ expression:
 i.e., if there are Gaussian errors $\sigma_n$ on each observed $y_n$,
-we can re-write \eq{chi2-simple} in the following (standard) way:
+we can re-write \eq{chi2-simple} in the following way:
 \begin{equation}
@@
 In short, you shiould be wary of placing too much trust in a Bayesian
-periodogram, unless you're certain of the type of signal you're looking for.
+periodogram, unless you're certain of the precise form of the signal
+you're looking for.
 
@@
 periodogram and its roots in both Fourier and Least-squares analysis.
-This section gets to the meat of the subject at hand:
+We now come to the meat of the subject at hand:
 given this understanding of the Lomb-Scargle periodogram and related
@@
 be a true Nyquist limit based on the \citet{Eyer99} definition, a pseudo-Nyquist
-limit based on careful scrutiny of the window function (cf. \fig{kepler-data}),
+limit based on careful scrutiny of the window function (cf.\ \fig{kepler-data}),
 a limiting frequency based on the integration time of individual
@@
 vital to keep in mind the failure modes of the periodogram approach, which
-are rooted in the aliasing and pseudo-aliasing effects rooted in the structure
+are related to the aliasing and pseudo-aliasing effects rooted in the structure
 of the window function (recall \sect{window-functions}).
@@
 $r$ requires
-something like $\sim 10 / r$ individual periodograms to be computed---this grows
-computationally intractable quite quickly.
+something like $\sim 10 / r$ individual periodograms to be computed---this
+can grow computationally intractable quite quickly.
 The bootstrap is also not universally applicable: for example, it does not
@@
 without the need to compute thousands of full bootstrapped periodograms.
+
+\subsubsubsection{Detection Significance with Red Noise}
+It is important to recognize that the above methods of measuring peak significance
+or false alarm probability generally depend on the assumption of white noise:
+that is, that the height of spurious ``background'' peaks in the periodogram
+is independent of frequency.
+In many situations of interest, however, this assumption does not hold: for
+example, compact accreting objects often display ``red noise'' -- that is,
+noise that increases in power toward lower frequencies \citep{Vaughan2003}.
+The presence of red noise complicates both the detection of peaks and
+estimation of their significance or false alarm probability; for detailed
+discussion of various approaches to periodic analysis in this context,
+see \citet{Vaughan2005, Vaughan2010}.
 
@@
 
-Unfortunately, there is no silver bullet for answering
+Unfortunately, there is no single, universally applicable answer to
 these broader, more relevant questions of uncertainty of Lomb-Scargle results.
@@
 a large number of frequencies.
-\citet{Zechmeister09} showed how this approach can be easily extended to
-the floating mean periodogram discussed in \sect{floating-mean}.
+\citet{Zechmeister09} showed how this approach can be straightforwardly
+extended to the floating mean periodogram discussed in \sect{floating-mean}.
 
@@
     periodograms and has few if any disadvantages. Avoid multi-term Fourier
-    extensions (\sect{multiterm}) when the signal is unknown, because its
-    main effect is to increase periodogram noise
+    extensions (\sect{multiterm}) when the signal has an unknown form, because
+    its main effect is to increase periodogram noise
     (cf. \figs{kepler-multiterm}{LINEAR-multiterm}).
@@
   \item If you have prior knowledge of the shape of light curves you are
-    trying to detect, consider using more complex models to choose between
+    trying to detect, consider using more complex models, such as multi-term
+    models or physically-derived templates, to choose between
     multiple peaks in the periodogram (\sect{bayesian-periodograms}).
@@
 {\it Acknowledgments:} I am grateful to Jeff Scargle, Max Mahlke, Dan
-Foreman-Mackey, Zeljko Ivezic, and David Hogg for helpful feedback
-on early drafts of this paper.
+Foreman-Mackey, Zeljko Ivezic, David Hogg, and the anonymous reviewer
+for helpful feedback on early drafts of this paper.
 This work was supported by the University of Washington eScience Institute,

### 2021-09-19 PracticalLombScargle.tex
Update PracticalLombScargle.tex
Duplicate word 'the' fix

@@
 will be {\it far} below the true frequency of 100; still, the
-the Lomb-Scargle periodogram in the lower right panel quite cleanly
+Lomb-Scargle periodogram in the lower right panel quite cleanly
 recovers the true frequency.

### 2021-09-19 PracticalLombScargle.tex
Update PracticalLombScargle.tex
Small word fix, I think this is what you meant to write.

--- 2017-12-06 PracticalLombScargle.tex
@@
 Because of the close similarity between the classical and Lomb-Scargle periodograms,
-the bulk of or previous discussion applies---at least qualitatively---to
+the bulk of our previous discussion applies---at least qualitatively---to
 periodograms computed with the Lomb-Scargle method.

### 2021-09-19 PracticalLombScargle.tex
Update PracticalLombScargle.tex
Fix typo

--- 2017-12-06 PracticalLombScargle.tex
@@
 We can find the ``best'' model $\hat{y}(t;f)$ by minimizing $\chi^2(f)$ at
-each freuqnecy with respect to $A_f$ and $\phi_f$;
+each frequency with respect to $A_f$ and $\phi_f$;
 we will denote this minimum value as $\hat{\chi}^2(f)$.
